### 2020-01-07 main.tex
Initial Overleaf Import

\documentclass{article}
\usepackage[utf8]{inputenc}

\title{path-independence}
\author{ank55 }
\date{January 2020}

\begin{document}

\maketitle

\section{Introduction}

\end{document}

### 2020-02-12 path-checker/src/lib.hs
idk current directory state

@@ -1,0 +1,13 @@
+module Lib
+    ( someFunc
+    ) where
+
+someFunc :: IO ()
+someFunc = putStrLn "someFunc"
+
+-- Extract neighbours from adjacency graph.
+{- neighbour_one_d lst = [ x | x <- [0..length lst - 1], lst !! x == 1]
+neighbours node graph = neighbour_one_d (graph !! node)
+neighbours_filtered node graph visited = 
+    let lst = graph !! node in
+    [x | x <- [0..length lst-1], lst !! x == 1, not (elem x visited)]  -}

### 2020-02-12 path-finding/flipTolerantPathFinding.py

@@ -1,47 +1,0 @@
-"""Find all paths between two nodes in a directed graph."""
-
-"""Graph is encoded as an adjacency matrix with i,j = 1 indicating that there is an edge from i to j"""
-_GRAPH = [[ 0, 1, 1, 0],
-	  [ 1, 0, 1, 1],
-	  [ 0, 0, 0, 1],
-	  [ 0, 0, 0, 0]]
-
-def findPaths(s, t, flips=0, graph = _GRAPH):
-	"""Print the paths from s to t in graph."""
-	print(getList(s, t, graph, {}, flips))
-
-def getList(s, t, graph, visited, flips=0):
-	"""Return a {number of flips -> list of paths from s to t}."""
-	if s==t:
-            return {0 :[[t]]}
-	neighbours = getNeighbours(s, graph, visited)
-	sameflipsAns = {} 
-	for neighbour in neighbours:
-	    visited[s]=''
-            sameflipsAns={ flip:sameflipsAns.get(flip, [])+[[s]+lst for lst in getList(neighbour, t, graph, visited, flip).get(flip, [])] for flip in range(flips+1) }
-	    visited.pop(neighbour, '')
-	print("{} has forward paths {}".format(s, str(sameflipsAns)))
-        if flips==0:
-            return sameflipsAns
-	preds = getPredecessors(s, graph, visited)
-	for neighbour in preds:
-	    visited[s]=''
-            sameflipsAns={ flip:sameflipsAns.get(flip, [])+[[s]+lst for lst in getList(neighbour, t, graph, visited, flip-1).get(flip-1, [])] for flip in range(flips+1) }
-	    visited.pop(neighbour, '')
-	print("{} has paths {}".format(s, str(sameflipsAns)))
-        return sameflipsAns
-
-def getNeighbours(s, graph, visited):
-	"""return a list of the neighbours of s that haven't been visited"""
-	lst = filter(lambda x : x not in visited and graph[s][x]==1, range(len(graph[s])))
-	print("{} has neighbours {}".format(s, str(lst)))
-	return lst
-
-def getPredecessors(s, graph, visited):
-        """return a list of the (immediate) predecessors of s that haven't been visited"""
-        lst = [i for i in range(len(graph)) if i not in visited and graph[i][s]==1]
-        print("{} has predecessors {}".format(s, str(lst)))
-        return lst
-
-
-findPaths(0, 3, 1)

### 2020-02-12 path-finding/pathFinding.py

@@ -1,32 +1,0 @@
-"""Find all paths between two nodes in a directed graph."""
-
-"""Graph is encoded as an adjacency matrix with i,j = 1 indicating that there is an edge from i to j"""
-_GRAPH = [[ 0, 1, 1, 0],
-	  [ 1, 0, 1, 1],
-	  [ 0, 0, 0, 1],
-	  [ 0, 0, 0, 0]]
-
-def findPaths(s, t, graph = _GRAPH):
-	"""Print the paths from s to t in graph."""
-	print(getList(s, t, graph, {}))
-
-def getList(s, t, graph, visited):
-	"""Return a list of paths from s to t."""
-	if s==t:
-		return [[t]]
-	neighbours = getNeighbours(s, graph, visited)
-	ans = []
-	for neighbour in neighbours:
-		visited[s]=''
-		ans+=[[s]+lst for lst in getList(neighbour, t, graph, visited)]
-		visited.pop(neighbour, '')
-	print("{} has paths {}".format(s, str(ans)))
-	return ans
-
-def getNeighbours(s, graph, visited):
-	"""return a list of the neighbours of s that haven't been visited"""
-	lst = filter(lambda x : x not in visited and graph[s][x]==1, range(len(graph[s])))
-	print("{} has neighbours {}".format(s, str(lst)))
-	return lst
-
-findPaths(0, 3)

### 2020-02-12 pathfinding/src/Lib.hs

@@ -1,0 +1,12 @@
+module Lib
+    ( someFunc
+    ) where
+
+someFunc :: IO ()
+someFunc = putStrLn "someFunc"
+
+neighbour_one_d lst = [ x | x <- [0..length lst - 1], lst !! x == 1]
+neighbours node graph = neighbour_one_d (graph !! node)
+neighbours_filtered node graph visited =
+    let lst = graph !! node in
+    [x | x <- [0..length lst-1], lst !! x == 1, not (elem x visited)] 

### 2020-02-12 main.tex

@@
 
+\section{Definitions}
+\begin{enumerate}
+    \item Path graph:
+    \item Conflicting paths: Paths with same source and destination node
+    \item Path independence:
+\end{enumerate}
+
+\section{Identifying the set of new conflictiong paths on addition of edge}
+\subsection{Theorem 1: The result of the two-flip tolerant path search algorithm corresponds bijectively to the set of new conflicting paths that arise in the graph dues to the addition of an edge.}
+
+Proof:\\
+Let the new edge added to the graph be (s,t).
+The two flip tolerant path algorithm returns all paths from s to t with upto two direction flips.
+
+*This proof is straightforward*
+
+\subsection{Implementation of two-flip tolerant path search}
+Normal recursive DFS but with direction and number of flips parameters to maintain state.
+
+\subsection{Verification using the new paths}
+The appropriate function and inverse compositions.
+
+\subsubsection{If any of the paths don't check out then the graph cannot be path independent.
+If all the paths do check out and the graph was originally path independent then the new graph is path independent.}
+
+\subsection{Theorem 2: If the two flip tolerant path search returns two paths P and Q that have the same flipping points then they provide the same information and one can be removed as redundant}
+
+\subsubsection{Theorem: If two paths have one flip-point in common and there exist some path connecting the second flip points, then there exists a partial ordering between the two paths}
+
 \end{document}

### 2020-02-13 main.tex
Updates from Overleaf

@@
 \usepackage[utf8]{inputenc}
+\usepackage{tikz}
 
 \title{path-independence}
-\author{ank55 }
+\author{ }
 \date{January 2020}
@@
 
+Many systems can be understood as a a category, or a collection of sets, with some transformation functions between the sets- for example, in programming languages, type systems with coercion.
+If a graph is constructed with the sets as nodes and functions as edges between the nodes, then by construction the new graph has the property of "path independence", the idea being that each object has a unique representation in a given set in the system.
+
 \section{Definitions}
 \begin{enumerate}
-    \item Path graph:
-    \item Conflicting paths: Paths with same source and destination node
-    \item Path independence:
+    \item Path graph: A graph with vertices representing sets, and directed edges representing transformations functions between the sets.
+A path represents the composition of the transformation functions in the nodes.
+    \item Conflicting paths: Paths with same source and destination node.
+    \item Path independence: The property that all conflicting paths in a graph are equal. Equality can be thought of as evaluated by some function equality checking oracle.
 \end{enumerate}
@@
 \subsection{Implementation of two-flip tolerant path search}
-Normal recursive DFS but with direction and number of flips parameters to maintain state.
+Normal recursive DFS with visited set also getting popped on backtrack (to get \textit{all} paths) but with direction and number of flips parameters to maintain state.
 
 \subsection{Verification using the new paths}
-The appropriate function and inverse compositions.
+The appropriate function and function inverse compositions.
 
@@
 
-\subsection{Theorem 2: If the two flip tolerant path search returns two paths P and Q that have the same flipping points then they provide the same information and one can be removed as redundant}
+\subsection{Theorem: If the two flip tolerant path search returns two paths P and Q that have the same flipping points then they provide the same information and one can be removed as redundant}
+
+\begin{center}
+\begin{tikzpicture}[scale=0.2]
+\tikzstyle{every node}+=[inner sep=0pt]
+\draw [black] (25,-31.9) circle (3);
+\draw (25,-31.9) node {$S$};
+\draw [black] (48,-31.9) circle (3);
+\draw (48,-31.9) node {$T$};
+\draw [black] (28.9,-19.5) circle (3);
+\draw (28.9,-19.5) node {$F_1$};
+\draw [black] (43.9,-19.5) circle (3);
+\draw (43.9,-19.5) node {$F_2$};
+\draw [black] (30.687,-17.11) arc (132.96162:47.03838:8.383);
+\fill [black] (42.11,-17.11) -- (41.87,-16.2) -- (41.19,-16.93);
+\draw (36.4,-14.36) node [above] {$g_1$};
+\draw [black] (41.409,-21.156) arc (-63.93383:-116.06617:11.399);
+\fill [black] (41.41,-21.16) -- (40.47,-21.06) -- (40.91,-21.96);
+\draw (36.4,-22.82) node [below] {$g_2$};
+\draw [black] (23.657,-29.237) arc (-164.08097:-230.83744:7.929);
+\fill [black] (23.66,-29.24) -- (23.92,-28.33) -- (22.96,-28.61);
+\draw (22.95,-24.04) node [left] {$f_1$};
+\draw [black] (30.007,-22.272) arc (11.75211:-46.67052:8.58);
+\fill [black] (27.49,-30.26) -- (28.42,-30.08) -- (27.73,-29.35);
+\draw (30.56,-27.24) node [right] {$f_2$};
+\draw [black] (45.566,-30.169) arc (-134.8387:-188.56886:9.123);
+\fill [black] (42.98,-22.34) -- (42.36,-23.06) -- (43.35,-23.21);
+\draw (42.57,-27.26) node [left] {$h_2$};
+\draw [black] (46.175,-21.439) arc (41.169:-4.57656:10.227);
+\fill [black] (46.18,-21.44) -- (46.33,-22.37) -- (47.08,-21.71);
+\draw (48.95,-24.27) node [right] {$h_1$};
+\end{tikzpicture}
+\end{center}
+
+Intuitive sketch:
+Let path $P_1$ be $f_1\circ g_1 \circ h_1$. Let path $P_2$ be $f_2\circ g_2 \circ h_2$. Then the equality constraint checking leads to identical results.
 
@@
 
+Two possible cases:
+
+\begin{center}
+\begin{tikzpicture}[scale=0.2]
+\tikzstyle{every node}+=[inner sep=0pt]
+\draw [black] (25,-31.9) circle (3);
+\draw (25,-31.9) node {$S$};
+\draw [black] (48,-31.9) circle (3);
+\draw (48,-31.9) node {$T$};
+\draw [black] (26.6,-13.6) circle (3);
+\draw (26.6,-13.6) node {$F_1$};
+\draw [black] (39,-21) circle (3);
+\draw (39,-21) node {$F_2$};
+\draw [black] (45.1,-12.9) circle (3);
+\draw (45.1,-12.9) node {$F_3$};
+\draw [black] (29.18,-15.14) -- (36.42,-19.46);
+\fill [black] (36.42,-19.46) -- (35.99,-18.62) -- (35.48,-19.48);
+\draw (31.38,-17.8) node [below] {$g_2$};
+\draw [black] (23.529,-29.292) arc (-156.64712:-213.3464:14.141);
+\fill [black] (23.53,-29.29) -- (23.67,-28.36) -- (22.75,-28.76);
+\draw (21.8,-22.39) node [left] {$f_1$};
+\draw [black] (27.889,-16.304) arc (19.98764:-29.98116:15.634);
+\fill [black] (26.74,-29.46) -- (27.57,-29.02) -- (26.71,-28.52);
+\draw (29.4,-23.08) node [right] {$f_2$};
+\draw [black] (46.09,-29.59) -- (40.91,-23.31);
+\fill [black] (40.91,-23.31) -- (41.03,-24.25) -- (41.81,-23.61);
+\draw (42.95,-27.88) node [left] {$h_2$};
+\draw [black] (29.6,-13.49) -- (42.1,-13.01);
+\fill [black] (42.1,-13.01) -- (41.28,-12.54) -- (41.32,-13.54);
+\draw (35.81,-12.7) node [above] {$g_1$};
+\draw [black] (47.55,-28.93) -- (45.55,-15.87);
+\fill [black] (45.55,-15.87) -- (45.18,-16.73) -- (46.17,-16.58);
+\draw (47.25,-22.21) node [right] {$h_1$};
+\draw [black] (40.8,-18.6) -- (43.3,-15.3);
+\fill [black] (43.3,-15.3) -- (42.41,-15.63) -- (43.21,-16.24);
+\draw (42.63,-18.35) node [right] {$l$};
+\end{tikzpicture}
+\end{center}
+
+$P_2$ is better, intuitively because l can only lose- not add information.
+
+\begin{center}
+\begin{tikzpicture}[scale=0.2]
+\tikzstyle{every node}+=[inner sep=0pt]
+\draw [black] (25,-31.9) circle (3);
+\draw (25,-31.9) node {$S$};
+\draw [black] (48,-31.9) circle (3);
+\draw (48,-31.9) node {$T$};
+\draw [black] (26.7,-7.7) circle (3);
+\draw (26.7,-7.7) node {$F_1$};
+\draw [black] (43.1,-21.8) circle (3);
+\draw (43.1,-21.8) node {$F_2$};
+\draw [black] (29.9,-16.8) circle (3);
+\draw (29.9,-16.8) node {$F_3$};
+\draw [black] (29.408,-6.43) arc (107.51449:-8.88948:11.282);
+\fill [black] (43.95,-18.93) -- (44.57,-18.22) -- (43.58,-18.06);
+\draw (41.59,-8.14) node [above] {$g_1$};
+\draw [black] (22.914,-29.751) arc (-141.57162:-226.46499:15.013);
+\fill [black] (22.91,-29.75) -- (22.81,-28.81) -- (22.02,-29.44);
+\draw (19.1,-19.32) node [left] {$f_1$};
+\draw [black] (46.06,-22.008) arc (71.20036:-19.43982:5.814);
+\fill [black] (46.06,-22.01) -- (46.66,-22.74) -- (46.98,-21.79);
+\draw (50.12,-23.9) node [right] {$h_2$};
+\draw [black] (45.319,-30.601) arc (-127.57654:-180.66292:7.327);
+\fill [black] (42.46,-24.71) -- (41.95,-25.5) -- (42.95,-25.52);
+\draw (42.49,-29.07) node [left] {$h_1$};
+\draw [black] (28.97,-19.65) -- (25.93,-29.05);
+\fill [black] (25.93,-29.05) -- (26.65,-28.44) -- (25.7,-28.13);
+\draw (26.68,-23.67) node [left] {$f_2$};
+\draw [black] (32.71,-17.86) -- (40.29,-20.74);
+\fill [black] (40.29,-20.74) -- (39.72,-19.99) -- (39.37,-20.92);
+\draw (35.16,-19.83) node [below] {$g_2$};
+\draw [black] (27.7,-10.53) -- (28.9,-13.97);
+\fill [black] (28.9,-13.97) -- (29.11,-13.05) -- (28.17,-13.38);
+\end{tikzpicture}
+\end{center}
+
+$P_1$ is better for a similar reason.
+
+\subsubsection{Theorem: In general, if there is an ordering to be derived from a paths between both flipping points of two paths, then there might exists a partial ordering between the two paths}
+
+I think the only case we know for sure is this one:
+
+\begin{center}
+\begin{tikzpicture}[scale=0.2]
+\tikzstyle{every node}+=[inner sep=0pt]
+\draw [black] (25,-31.9) circle (3);
+\draw (25,-31.9) node {$S$};
+\draw [black] (48,-31.9) circle (3);
+\draw (48,-31.9) node {$T$};
+\draw [black] (26.7,-7.7) circle (3);
+\draw (26.7,-7.7) node {$F_1$};
+\draw [black] (44.9,-17.5) circle (3);
+\draw (44.9,-17.5) node {$F_2$};
+\draw [black] (29.9,-16.8) circle (3);
+\draw (29.9,-16.8) node {$F_3$};
+\draw [black] (46.3,-7.7) circle (3);
+\draw (46.3,-7.7) node {$F_4$};
+\draw [black] (22.914,-29.751) arc (-141.57162:-226.46499:15.013);
+\fill [black] (22.91,-29.75) -- (22.81,-28.81) -- (22.02,-29.44);
+\draw (19.1,-19.32) node [left] {$f_1$};
+\draw [black] (47.37,-28.97) -- (45.53,-20.43);
+\fill [black] (45.53,-20.43) -- (45.21,-21.32) -- (46.19,-21.11);
+\draw (45.7,-25.06) node [left] {$h_2$};
+\draw [black] (28.97,-19.65) -- (25.93,-29.05);
+\fill [black] (25.93,-29.05) -- (26.65,-28.44) -- (25.7,-28.13);
+\draw (26.68,-23.67) node [left] {$f_2$};
+\draw [black] (32.9,-16.94) -- (41.9,-17.36);
+\fill [black] (41.9,-17.36) -- (41.13,-16.82) -- (41.08,-17.82);
+\draw (37.34,-17.71) node [below] {$g_2$};
+\draw [black] (27.7,-10.53) -- (28.9,-13.97);
+\fill [black] (28.9,-13.97) -- (29.11,-13.05) -- (28.17,-13.38);
+\draw [black] (29.471,-6.557) arc (108.53192:71.46808:22.115);
+\fill [black] (43.53,-6.56) -- (42.93,-5.83) -- (42.61,-6.78);
+\draw (36.5,-4.91) node [above] {$g_1$};
+\draw [black] (45.88,-10.67) -- (45.32,-14.53);
+\fill [black] (45.32,-14.53) -- (45.93,-13.81) -- (44.94,-13.67);
+\draw [black] (48.889,-9.204) arc (53.58054:-45.54392:13.727);
+\fill [black] (48.89,-9.2) -- (49.24,-10.08) -- (49.83,-9.28);
+\draw (55.03,-19.24) node [right] {$h_1$};
+\end{tikzpicture}
+\end{center}
+I think $P_1$ is better. Don't yet have a proof sketch in mind yet but this is what the intuition from the previous cases would lead to.
+
+The two other cases arise when you flip each arrow in turn. I think in these cases the result might go either way; TODO come up with the appropriate counterexamples.
+
+\section{Optimized Algorithm}
+Here, present pseudocode to do the path search with the ignoring of useless possibilities baked in. Perform runtime analysis.
+
 \end{document}

### 2020-02-13 pathfinding/flipTolerantPathFinding.py
Recovered old python stuff

@@ -1,0 +1,47 @@
+"""Find all paths between two nodes in a directed graph."""
+
+"""Graph is encoded as an adjacency matrix with i,j = 1 indicating that there is an edge from i to j"""
+_GRAPH = [[ 0, 1, 1, 0],
+	  [ 1, 0, 1, 1],
+	  [ 0, 0, 0, 1],
+	  [ 0, 0, 0, 0]]
+
+def findPaths(s, t, flips=0, graph = _GRAPH):
+	"""Print the paths from s to t in graph."""
+	print(getList(s, t, graph, {}, flips))
+
+def getList(s, t, graph, visited, flips=0):
+	"""Return a {number of flips -> list of paths from s to t}."""
+	if s==t:
+            return {0 :[[t]]}
+	neighbours = getNeighbours(s, graph, visited)
+	sameflipsAns = {} 
+	for neighbour in neighbours:
+	    visited[s]=''
+            sameflipsAns={ flip:sameflipsAns.get(flip, [])+[[s]+lst for lst in getList(neighbour, t, graph, visited, flip).get(flip, [])] for flip in range(flips+1) }
+	    visited.pop(neighbour, '')
+	print("{} has forward paths {}".format(s, str(sameflipsAns)))
+        if flips==0:
+            return sameflipsAns
+	preds = getPredecessors(s, graph, visited)
+	for neighbour in preds:
+	    visited[s]=''
+            sameflipsAns={ flip:sameflipsAns.get(flip, [])+[[s]+lst for lst in getList(neighbour, t, graph, visited, flip-1).get(flip-1, [])] for flip in range(flips+1) }
+	    visited.pop(neighbour, '')
+	print("{} has paths {}".format(s, str(sameflipsAns)))
+        return sameflipsAns
+
+def getNeighbours(s, graph, visited):
+	"""return a list of the neighbours of s that haven't been visited"""
+	lst = filter(lambda x : x not in visited and graph[s][x]==1, range(len(graph[s])))
+	print("{} has neighbours {}".format(s, str(lst)))
+	return lst
+
+def getPredecessors(s, graph, visited):
+        """return a list of the (immediate) predecessors of s that haven't been visited"""
+        lst = [i for i in range(len(graph)) if i not in visited and graph[i][s]==1]
+        print("{} has predecessors {}".format(s, str(lst)))
+        return lst
+
+
+findPaths(0, 3, 1)

### 2020-02-13 pathfinding/src/Lib.hs

@@ -1,12 +1,0 @@
-module Lib
-    ( someFunc
-    ) where
-
-someFunc :: IO ()
-someFunc = putStrLn "someFunc"
-
-neighbour_one_d lst = [ x | x <- [0..length lst - 1], lst !! x == 1]
-neighbours node graph = neighbour_one_d (graph !! node)
-neighbours_filtered node graph visited =
-    let lst = graph !! node in
-    [x | x <- [0..length lst-1], lst !! x == 1, not (elem x visited)] 

### 2020-02-21 pathfinding/pathFinding.py
Loops are working

@@ -4,8 +4,8 @@
 from itertools import combinations
 
 _GRAPH = \
-[[0, 1, 1],
- [0, 0, 0],
+[[0, 1, 0],
+ [1, 0, 0],
  [0, 0, 0]]
 
 # Find and return all paths from s to t in the given graph
@@ -46,21 +46,22 @@ def findNewConflicts(edge, graph = _GRAPH):
 
 # ---------- Helper functions ----------
 
-def getList(s, t, graph, visited):
-    """Return a list of paths from s to t."""
-    if s==t:
+def getList(s, t, graph, visited, first_node=True):
+    """Return a list of paths from s to t
+       Note that first_node allows us to get loops"""
+    if s==t and not first_node:
         return [[t]]
     neighbours = getNeighbours(s, graph, visited)
     ans = []
     for neighbour in neighbours:
-        visited.add(s)
-        ans += [[s]+lst for lst in getList(neighbour, t, graph, visited)]
+        if not first_node:
+            visited.add(s)
+        ans += [[s]+lst for lst in getList(neighbour, t, graph, visited, False)]
         visited.discard(neighbour)
-    return ans
+    return ans + ([[t]] if s==t else [])
 
 # Return a list of the neighbours of s that haven't been visited
 def getNeighbours(s, graph, visited):
-    """"""
     lst = filter(lambda x : x not in visited and graph[s][x]==1, range(len(graph[s])))
     return lst
 

### 2020-02-21 main.tex

--- 2020-01-07 main.tex
@@
 
+Many systems can be understood as a a category, or a collection of sets, with some transformation functions between the sets- for example, in programming languages, type systems with coercion.
+If a graph is constructed with the sets as nodes and functions as edges between the nodes, then by construction the new graph has the property of "path independence", the idea being that each object has a unique representation in a given set in the system.
+
+Much like conservative fields in physics, path independent graphs may arise in many systems.
+It might be useful to be able to perform some basic verification and ensure the graphs in question are indeed path independent.
+As a case study, gator, a domain specific graphics language that allows automatic conversion between geometric spaces internally creates a graph with he transformations between the spaces.
+In order for the compiler to give predictable results, gator requires that this space transformation graph be path independent.
+
+Naively checking if all paths in a given graph are path independent could require a number of function equality checks as bad as factorial in the number of nodes, since a path consists of an ordering of nodes.
+As comparing function equalities can be expensive, we seek a more refined approach to verifying path independence, specifically over the course of online addition to a graph, as would be useful in cases like the gator compiler.
+
+\section{Definitions}
+\begin{enumerate}
+    \item Path graph: A graph with vertices representing sets, and directed edges representing transformations functions between the sets.
+A path represents the composition of the transformation functions in the nodes.
+    \item Conflicting paths: Paths with same source and destination node.
+(Note that in this definition of the term we don not require paths to actually disagree, but only have the potential to give different results).
+    \item Path independence: The property that all conflicting paths in a graph are equal. Equality can be thought of as evaluated by some function equality checking oracle.
+\end{enumerate}
+
+\section{Identifying the set of new conflicting paths on addition of edge}
+
+The naive approach to verifying path independence would be to compare all conflicting paths pairwise and thus ensure that there are no instances of violation of the property.
+Over the course of online addition of an edge, given that the graph before adding the edge was path independent, it would only be necessary to check conflicting pairs involving a new path that were created as a consequence of adding the new edge to the path.
+To identify all such conflicting pairs, we present the \textit{two-flip tolerant path search}.
+
+\subsection{Two flip tolerant path search}
+In a normal directed graph path, only forward edges, i.e. edges that go outward from the current node, are considered. 
+On the other hand, the \textit{two flip path} goes through up to three phases: in the first phase, only backward edges- pointing inward to the current node- are accepted.
+In the second phase, only forward edges are accepted, and in the third phase, again only backward edges are accepted.
+As a consequence, there is a node in the path for which both path edges are oriented outwards, that lies between the first and second phase.
+This will be referred to as the \textit{first flipping point}.
+Similarly, there is a \textit{second flipping point} between the second and third phase, for which both path edges point inwards.
+
+Consider the situation where a new edge from source node S to sink T is being added to a graph.
+A two flip tolerant path from S to T would correspond to a new pair of conflicting paths- flipping point 1 would be the source of both paths, and flipping point 2, the sink.
+The first path would correspond to the second phase of the flip tolerant path search, and the second path, involving the new edge, would correspond to composing the first phase of the two flip path, the new edge, and the third phase.
+
+To present this idea diagrammatically, we represent path phases with arrows (note that these are the composition of many edges, not a single edge). The new edge is represented with a dashed arrow.
+
+\begin{center}
+\begin{tikzpicture}[scale=0.2]
+\tikzstyle{every node}+=[inner sep=0pt]
+\draw [black] (18.1,-36.1) circle (3);
+\draw (18.1,-36.1) node {$S$};
+\draw [black] (50.2,-36.1) circle (3);
+\draw (50.2,-36.1) node {$T$};
+\draw [black] (26,-22.9) circle (3);
+\draw (26,-22.9) node {$F_1$};
+\draw [black] (41.3,-22.9) circle (3);
+\draw (41.3,-22.9) node {$F_2$};
+\draw [dashed] (21.1,-36.1) -- (47.2,-36.1);
+\fill [black] (47.2,-36.1) -- (46.4,-35.6) -- (46.4,-36.6);
+\draw (34.15,-35.6) node [above] {new edge};
+\draw [black] (24.46,-25.47) -- (19.64,-33.53);
+\fill [black] (19.64,-33.53) -- (20.48,-33.1) -- (19.62,-32.58);
+\draw [black] (29,-22.9) -- (38.3,-22.9);
+\fill [black] (38.3,-22.9) -- (37.5,-22.4) -- (37.5,-23.4);
+\draw [black] (48.52,-33.61) -- (42.98,-25.39);
+\fill [black] (42.98,-25.39) -- (43.01,-26.33) -- (43.84,-25.77);
+\end{tikzpicture}
+\end{center}
+Here, (S, $F_1$, $F_2$, T) is a two flip tolerant path. ($F_1$, S, T, $F_2$) is a new path, created because of the addition of (S,T), that conflicts with ($F_1$, $F_2$).
+
+The two flip tolerant path search returns the set of all paths between a given source and sink that have upto two flips (paths that omit one or more of the three phases are also accepted).
+
+\paragraph{Theorem} The result of the two-flip tolerant path search from the source to the sink node of an edge that is to be newly added has a one-to-one* correspondence to the set of new pairs of conflicting paths that arise in the graph due to its addition.
+
+*= This excludes paths whose "area of difference" does not include the new edge since they should remain path independent.
+
+\paragraph{Proof}
+Let the new edge added to the graph be (S, T).
+It has already been shown how each path corresponds to a conflicting pair in the case where a path has two flips.
+When a path has only the first flip (which is to say, the third phase of the path is missing), but not the second flip, sink node T can be treated as the sink of the two conflicting path, while the first flipping point remains the source.
+Similarly when only the second flipping point is present then it is the sink, and S is the source.
+Finally when no flipping points are present, there are two possibilities.
+Either the path is oriented from S to T, in which case the conflicting paths are simply the edge (S, T) and the entire flip-less path.
+Or, the path is oriented from T to S. In this case, the conflicting paths are the identity on S (that implicitly exists) and the cycle obtained from composing the new edge with the path.
+
+It remains to show that every new conflicting pair corresponds to a two flip tolerant path. Let the common source be $F_1$ and common sink be $F_2$. 
+Note that we are excluding pair where both paths include the new edge, since such paths are equal from $F_1$ to S because of the path independence of the original graph, identical from S to T, and again equal from T to $F_2$ because of original path independence.
+Being the composition of equal parts, these paths must always be equal.
+
+Now, it is the case that only one path passes through (S, T), which we will name path 1.
+We construct a two flip tolerant path from S to T: phase 1 is the segment of path 1 from $F_1$ to S, phase 2 is path 2, and phase 3 is the segment of path 1 from T to $F_2$.
+It is possible that some of $F_1, F_2, S$ and $T$ coincide, in which case the corresponding segments between them disappear, and the resultant path has fewer than two flips.
+
+\subsection{Implementation of two-flip tolerant path search}
+A modified DFS with visited set also getting popped on backtrack (to get \textit{all} paths) and additionally, variables to maintain number of flips and flip state.
+
+\subsection{Verification using the new paths}
+The flip tolerant paths search returns pairs of conflicting paths.
+Verification is performed by composing edge functions to derive path functions, and then performing an equality check to verify if the path functions are equal.
+
+If any of the paths don't check out then the graph, by definition, cannot be path independent.
+If all the paths do check out and the graph was originally path independent then the new graph is path independent, since new pairs have been verified and old pairs remain unchanged.
+
+% \subsubsection{Theorem: If two paths have one flip-point in common and there exist some path connecting the second flip points, then there exists a partial ordering between the two paths}
+
+% Two possible cases:
+
+% \begin{center}
+% \begin{tikzpicture}[scale=0.2]
+% \tikzstyle{every node}+=[inner sep=0pt]
+% \draw [black] (25,-31.9) circle (3);
+% \draw (25,-31.9) node {$S$};
+% \draw [black] (48,-31.9) circle (3);
+% \draw (48,-31.9) node {$T$};
+% \draw [black] (26.6,-13.6) circle (3);
+% \draw (26.6,-13.6) node {$F_1$};
+% \draw [black] (39,-21) circle (3);
+% \draw (39,-21) node {$F_2$};
+% \draw [black] (45.1,-12.9) circle (3);
+% \draw (45.1,-12.9) node {$F_3$};
+% \draw [black] (29.18,-15.14) -- (36.42,-19.46);
+% \fill [black] (36.42,-19.46) -- (35.99,-18.62) -- (35.48,-19.48);
+% \draw (31.38,-17.8) node [below] {$g_2$};
+% \draw [black] (23.529,-29.292) arc (-156.64712:-213.3464:14.141);
+% \fill [black] (23.53,-29.29) -- (23.67,-28.36) -- (22.75,-28.76);
+% \draw (21.8,-22.39) node [left] {$f_1$};
+% \draw [black] (27.889,-16.304) arc (19.98764:-29.98116:15.634);
+% \fill [black] (26.74,-29.46) -- (27.57,-29.02) -- (26.71,-28.52);
+% \draw (29.4,-23.08) node [right] {$f_2$};
+% \draw [black] (46.09,-29.59) -- (40.91,-23.31);
+% \fill [black] (40.91,-23.31) -- (41.03,-24.25) -- (41.81,-23.61);
+% \draw (42.95,-27.88) node [left] {$h_2$};
+% \draw [black] (29.6,-13.49) -- (42.1,-13.01);
+% \fill [black] (42.1,-13.01) -- (41.28,-12.54) -- (41.32,-13.54);
+% \draw (35.81,-12.7) node [above] {$g_1$};
+% \draw [black] (47.55,-28.93) -- (45.55,-15.87);
+% \fill [black] (45.55,-15.87) -- (45.18,-16.73) -- (46.17,-16.58);
+% \draw (47.25,-22.21) node [right] {$h_1$};
+% \draw [black] (40.8,-18.6) -- (43.3,-15.3);
+% \fill [black] (43.3,-15.3) -- (42.41,-15.63) -- (43.21,-16.24);
+% \draw (42.63,-18.35) node [right] {$l$};
+% \end{tikzpicture}
+% \end{center}
+
+% $P_2$ is better, intuitively because l can only lose- not add information.
+
+% \begin{center}
+% \begin{tikzpicture}[scale=0.2]
+% \tikzstyle{every node}+=[inner sep=0pt]
+% \draw [black] (25,-31.9) circle (3);
+% \draw (25,-31.9) node {$S$};
+% \draw [black] (48,-31.9) circle (3);
+% \draw (48,-31.9) node {$T$};
+% \draw [black] (26.7,-7.7) circle (3);
+% \draw (26.7,-7.7) node {$F_1$};
+% \draw [black] (43.1,-21.8) circle (3);
+% \draw (43.1,-21.8) node {$F_2$};
+% \draw [black] (29.9,-16.8) circle (3);
+% \draw (29.9,-16.8) node {$F_3$};
+% \draw [black] (29.408,-6.43) arc (107.51449:-8.88948:11.282);
+% \fill [black] (43.95,-18.93) -- (44.57,-18.22) -- (43.58,-18.06);
+% \draw (41.59,-8.14) node [above] {$g_1$};
+% \draw [black] (22.914,-29.751) arc (-141.57162:-226.46499:15.013);
+% \fill [black] (22.91,-29.75) -- (22.81,-28.81) -- (22.02,-29.44);
+% \draw (19.1,-19.32) node [left] {$f_1$};
+% \draw [black] (46.06,-22.008) arc (71.20036:-19.43982:5.814);
+% \fill [black] (46.06,-22.01) -- (46.66,-22.74) -- (46.98,-21.79);
+% \draw (50.12,-23.9) node [right] {$h_2$};
+% \draw [black] (45.319,-30.601) arc (-127.57654:-180.66292:7.327);
+% \fill [black] (42.46,-24.71) -- (41.95,-25.5) -- (42.95,-25.52);
+% \draw (42.49,-29.07) node [left] {$h_1$};
+% \draw [black] (28.97,-19.65) -- (25.93,-29.05);
+% \fill [black] (25.93,-29.05) -- (26.65,-28.44) -- (25.7,-28.13);
+% \draw (26.68,-23.67) node [left] {$f_2$};
+% \draw [black] (32.71,-17.86) -- (40.29,-20.74);
+% \fill [black] (40.29,-20.74) -- (39.72,-19.99) -- (39.37,-20.92);
+% \draw (35.16,-19.83) node [below] {$g_2$};
+% \draw [black] (27.7,-10.53) -- (28.9,-13.97);
+% \fill [black] (28.9,-13.97) -- (29.11,-13.05) -- (28.17,-13.38);
+% \end{tikzpicture}
+% \end{center}
+
+% $P_1$ is better for a similar reason.
+
+\section{Optimization}
+
+The naive approach presented above identifies and verifies all conflicting pairs in the graph.
+However, because of the relationship between the various paths, there is some redundancy.
+It sometimes possible to conclude from verifying one pair, that a different pair must necessarily also be equal.
+We leverage such situations, since checking the equality of paths may be an expensive operation and an all-path search can have a bad worst-case run time.
+
+\subsubsection{Ordering on informativeness of paths}
+
+Given a pair of two flip tolerant paths, we look at situations where there is a path between corresponding flipping points.
+There are two cases.
+
+\paragraph{Case 1} The arrows connecting flip points of path 1 to flip points of path 2 are in opposite directions.
+
+\begin{center}
+\begin{tikzpicture}[scale=0.2]
+\tikzstyle{every node}+=[inner sep=0pt]
+\draw [black] (16.6,-34.1) circle (3);
+\draw (16.6,-34.1) node {$S$};
+\draw [black] (56.8,-33.6) circle (3);
+\draw (56.8,-33.6) node {$T$};
+\draw [black] (28,-20.1) circle (3);
+\draw (28,-20.1) node {$P_1$};
+\draw [black] (44.5,-20.1) circle (3);
+\draw (44.5,-20.1) node {$P_2$};
+\draw [black] (28,-44.9) circle (3);
+\draw (28,-44.9) node {$Q_1$};
+\draw [black] (44.5,-44.9) circle (3);
+\draw (44.5,-44.9) node {$Q_2$};
+\draw [black] (31,-20.1) -- (41.5,-20.1);
+\fill [black] (41.5,-20.1) -- (40.7,-19.6) -- (40.7,-20.6);
+\draw (36.25,-20.6) node [below] {$g_1$};
+\draw [black] (26.11,-22.43) -- (18.49,-31.77);
+\fill [black] (18.49,-31.77) -- (19.39,-31.47) -- (18.61,-30.84);
+\draw (21.74,-25.67) node [left] {$f_1$};
+\draw [black] (25.82,-42.84) -- (18.78,-36.16);
+\fill [black] (18.78,-36.16) -- (19.01,-37.08) -- (19.7,-36.35);
+\draw (23.57,-39.02) node [above] {$f_2$};
+\draw [black] (31,-44.9) -- (41.5,-44.9);
+\fill [black] (41.5,-44.9) -- (40.7,-44.4) -- (40.7,-45.4);
+\draw (36.25,-45.4) node [below] {$g_2$};
+\draw [black] (54.59,-35.63) -- (46.71,-42.87);
+\fill [black] (46.71,-42.87) -- (47.64,-42.7) -- (46.96,-41.96);
+\draw (49.21,-38.76) node [above] {$h_2$};
+\draw [black] (54.78,-31.38) -- (46.52,-22.32);
+\fill [black] (46.52,-22.32) -- (46.69,-23.25) -- (47.43,-22.57);
+\draw (51.19,-25.39) node [right] {$h_1$};
+\draw [black] (28,-23.1) -- (28,-41.9);
+\fill [black] (28,-41.9) -- (28.5,-41.1) -- (27.5,-41.1);
+\draw (27.5,-32.5) node [left] {$l$};
+\draw [black] (19.6,-34.06) -- (53.8,-33.64);
+\fill [black] (53.8,-33.64) -- (52.99,-33.15) -- (53.01,-34.15);
+\draw [black] (44.5,-41.9) -- (44.5,-23.1);
+\fill [black] (44.5,-23.1) -- (44,-23.9) -- (45,-23.9);
+\draw (45,-32.5) node [right] {$m$};
+\end{tikzpicture}
+\end{center}
+Let Path$_1=f_1\circ g_1\circ h_1$ and Path$_2$=$f_2\circ g_2\circ h_2$. We choose Path$_2$.
+\paragraph{Theorem} 
+If conflicting paths $g_2 = f_2 \circ (S,T) \circ h_2$ then it must be that $g_1 = f_1 \circ (S,T) \circ h_1$.
+\paragraph{Proof}
+We use the fact that $f_1$=$l\circ f_2$ and $h_1$=$h_2 \circ m$.
+\[g_2 = f_2 \circ (S,T) \circ h_2 \Rightarrow l \circ g_2 = l \circ f_2 \circ (S,T) \circ h_2 \Rightarrow l \circ g_2 \circ m = l \circ f_2 \circ (S,T) \circ h_2 \circ m \Rightarrow g_1 = f_1 \circ (S,T) \circ h_1\]
+
+% Notice this ordering is like the subtyping rule for a function. What is the relationship?
+
+Notice that this generalization will apply to situations where any of these paths are the identity, such as when the first flipping point of both paths is the same, or when there is no first flipping point- it is identical to S.
+
+\paragraph{Case 2} The arrows connecting flip points of path 1 to flip points of path 2 are in the same directions.
+
+\begin{center}
+\begin{tikzpicture}[scale=0.2]
+\tikzstyle{every node}+=[inner sep=0pt]
+\draw [black] (16.6,-34.1) circle (3);
+\draw (16.6,-34.1) node {$S$};
+\draw [black] (56.8,-33.6) circle (3);
+\draw (56.8,-33.6) node {$T$};
+\draw [black] (28,-20.1) circle (3);
+\draw (28,-20.1) node {$P_1$};
+\draw [black] (44.5,-20.1) circle (3);
+\draw (44.5,-20.1) node {$P_2$};
+\draw [black] (28,-44.9) circle (3);
+\draw (28,-44.9) node {$Q_1$};
+\draw [black] (44.5,-44.9) circle (3);
+\draw (44.5,-44.9) node {$Q_2$};
+\draw [black] (31,-20.1) -- (41.5,-20.1);
+\fill [black] (41.5,-20.1) -- (40.7,-19.6) -- (40.7,-20.6);
+\draw (36.25,-20.6) node [below] {$g_1$};
+\draw [black] (26.11,-22.43) -- (18.49,-31.77);
+\fill [black] (18.49,-31.77) -- (19.39,-31.47) -- (18.61,-30.84);
+\draw (21.74,-25.67) node [left] {$f_1$};
+\draw [black] (25.82,-42.84) -- (18.78,-36.16);
+\fill [black] (18.78,-36.16) -- (19.01,-37.08) -- (19.7,-36.35);
+\draw (23.57,-39.02) node [above] {$f_2$};
+\draw [black] (31,-44.9) -- (41.5,-44.9);
+\fill [black] (41.5,-44.9) -- (40.7,-44.4) -- (40.7,-45.4);
+\draw (36.25,-45.4) node [below] {$g_2$};
+\draw [black] (54.59,-35.63) -- (46.71,-42.87);
+\fill [black] (46.71,-42.87) -- (47.64,-42.7) -- (46.96,-41.96);
+\draw (49.21,-38.76) node [above] {$h_2$};
+\draw [black] (54.78,-31.38) -- (46.52,-22.32);
+\fill [black] (46.52,-22.32) -- (46.69,-23.25) -- (47.43,-22.57);
+\draw (51.19,-25.39) node [right] {$h_1$};
+\draw [black] (28,-23.1) -- (28,-41.9);
+\fill [black] (28,-41.9) -- (28.5,-41.1) -- (27.5,-41.1);
+\draw (27.5,-32.5) node [left] {$l$};
+\draw [black] (19.6,-34.06) -- (53.8,-33.64);
+\fill [black] (53.8,-33.64) -- (52.99,-33.15) -- (53.01,-34.15);
+\draw [black] (44.5,-23.1) -- (44.5,-41.9);
+\fill [black] (44.5,-41.9) -- (45,-41.1) -- (44,-41.1);
+\draw (44,-32.5) node [left] {$m$};
+\end{tikzpicture}
+\end{center}
+
+In this case, nothing can be said. As a counterexample, let $P_1$ contain element a.
+Suppose $g_1(a)=a_{g_1}$. Similarly, $l(a)=a_{l}$.
+*TODO Write. Basically there can be a disagreement at $P_2$.*
+
+\subsection{Polynomial verification bound}
+ This is a special instance of case 1, where l and m are the identity.
+\begin{center}
+\begin{tikzpicture}[scale=0.2]
+\tikzstyle{every node}+=[inner sep=0pt]
+\draw [black] (25,-31.9) circle (3);
+\draw (25,-31.9) node {$S$};
+\draw [black] (48,-31.9) circle (3);
+\draw (48,-31.9) node {$T$};
+\draw [black] (28.9,-19.5) circle (3);
+\draw (28.9,-19.5) node {$F_1$};
+\draw [black] (43.9,-19.5) circle (3);
+\draw (43.9,-19.5) node {$F_2$};
+\draw [black] (30.687,-17.11) arc (132.96162:47.03838:8.383);
+\fill [black] (42.11,-17.11) -- (41.87,-16.2) -- (41.19,-16.93);
+\draw (36.4,-14.36) node [above] {$g_1$};
+\draw [black] (41.409,-21.156) arc (-63.93383:-116.06617:11.399);
+\fill [black] (41.41,-21.16) -- (40.47,-21.06) -- (40.91,-21.96);
+\draw (36.4,-22.82) node [below] {$g_2$};
+\draw [black] (23.657,-29.237) arc (-164.08097:-230.83744:7.929);
+\fill [black] (23.66,-29.24) -- (23.92,-28.33) -- (22.96,-28.61);
+\draw (22.95,-24.04) node [left] {$f_1$};
+\draw [black] (30.007,-22.272) arc (11.75211:-46.67052:8.58);
+\fill [black] (27.49,-30.26) -- (28.42,-30.08) -- (27.73,-29.35);
+\draw (30.56,-27.24) node [right] {$f_2$};
+\draw [black] (45.566,-30.169) arc (-134.8387:-188.56886:9.123);
+\fill [black] (42.98,-22.34) -- (42.36,-23.06) -- (43.35,-23.21);
+\draw (42.57,-27.26) node [left] {$h_2$};
+\draw [black] (46.175,-21.439) arc (41.169:-4.57656:10.227);
+\fill [black] (46.18,-21.44) -- (46.33,-22.37) -- (47.08,-21.71);
+\draw (48.95,-24.27) node [right] {$h_1$};
+\end{tikzpicture}
+\end{center}
+
+Thus, when flip points of the two paths are identical, they give the same information and it is safe to verify only one and discard the other.
+This means that after our path search gives us all two flip tolerant paths, we only need to verify one path for each unique choice of flip points, and can discard the rest of the paths.
+
+In an N node graph, there are $^NC_2 = \frac{N(N-1)}{2}$ choices of flip points. Thus we can place a polynomial upper bound on the number of paths that actually need to be verified.
+
+If we allow only Case 1 above as a reducing law, this bound is tight. 
+This can be seen in the case where the graph contains 2N nodes besides S and T.
+We consider N of the nodes to be in group 1, and the other N to be in group 2.
+Every node in group 1 has a forward node to every node in group 2, as well as to S.
+T has a forward edge to every node in group 2.
+In this graph, on the addition of edge (S, T), $N^2$ paths need to be verified which is polynomial in 2N+2.
+
+\subsection{Optimized Path Finding Algorithm}
+We utilize this idea of selecting flip points to devise a polynomial time path search.
+This search does not return all two flip tolerant paths, but rather, a path for every feasible choice of flipping point.
+ *TODO Write algorithm.*
+
 \end{document}

### 2020-02-24 pathfinding/pathFinding.py
Two-flip search seems to be working

@@ -4,8 +4,10 @@
 from itertools import combinations
 
 _GRAPH = \
-[[0, 1],
- [0, 0]]
+[[0, 1, 1, 0],
+ [0, 0, 1, 0],
+ [1, 0, 0, 0],
+ [0, 0, 0, 0]]
 
 # _GRAPH = \
 # [[0, 1, 1, 1, 0, 1],
@@ -17,34 +19,57 @@ _GRAPH = \
 
 # ----------- Fast functions -----------
 
-def findPaths(s, t, graph = _GRAPH):
-    return [[0]]
+# A list of all pairs of paths that both share a source and sink
+#  and include 'edge' in the path
+# Throws an error if 'edge' is not in the graph
+def findEdgeConflicts(s, t, graph = _GRAPH):
+    if graph[s][t] != 1:
+        raise "Edge to check conflicts with does not exist"
+    return findNewConflicts(s, t, graph)
 
-def findAllConflictingPaths(graph = _GRAPH):
-    return [[(0, 0)]]
-
-def findNewConflicts(edge, graph = _GRAPH):
-    result = [(0, 0)]
-    graph[edge[0]][edge[1]] = 1
+# A list of all pairs of paths that both share a source and sink
+#  and include 'edge' in the path
+# Adds 'edge' if it is not already part of the graph
+def findNewConflicts(s, t, graph = _GRAPH):
+    graph[s][t] = 0
+    result = []
+    for path,flips in findFlippingPoints(t, s, graph, flips_allowed=2):
+        print(path)
+        # It's a cycle, so find all cycle conflicts
+        if len(flips) == 0:
+            for i in range(len(path)):
+                result.append(([path[i]] + path[i+1:] + path[:i+1], [path[i]]))
+        # Common ancestor, so check two meeting paths
+        if len(flips) == 1:
+            new_res = [s]
+            for i in range(len(path)):
+                new_res.append(path[i])
+                if path[i] == flips[0]:
+                    rev_path = path[i:]
+                    rev_path.reverse()
+                    new_res = (new_res, rev_path)
+                    break
+            result.append(new_res)
+        # Two-flip result, so compare paths between flip source and sink
+        if len(flips) == 2:
+            new_res = []
+            next_path = []
+            for i in range(len(path)):
+                next_path.append(path[i])
+                # Note that we assume the order of reported flips
+                # Flips must be ordered to occur in the path _from_ s to t
+                if path[i] == flips[0]:
+                    new_res = next_path
[…]

### 2020-02-24 pathfinding/pathFinding.py
Speed test added

@@ -1,6 +1,7 @@
 """Find all paths between two nodes in a directed graph."""
 """Graph is encoded as an adjacency matrix with i,j = 1 indicating that there is an edge from i to j"""
 
+import random
 from itertools import combinations
 
 _GRAPH = \
@@ -25,7 +26,7 @@ _GRAPH = \
 def findEdgeConflicts(s, t, graph = _GRAPH):
     if graph[s][t] != 1:
         raise "Edge to check conflicts with does not exist"
-    return findNewConflicts(s, t, graph)
+    return findNewConflicts(s, t, graph)[0]
 
 # A list of all pairs of paths that both share a source and sink
 #  and include 'edge' in the path
@@ -34,7 +35,6 @@ def findNewConflicts(s, t, graph = _GRAPH):
     graph[s][t] = 0
     result = []
     for path,flips in findFlippingPoints(t, s, graph, flips_allowed=2):
-        print(path)
         # It's a cycle, so find all cycle conflicts
         if len(flips) == 0:
             for i in range(len(path)):
@@ -67,7 +67,7 @@ def findNewConflicts(s, t, graph = _GRAPH):
                     break
             result.append(new_res)
     graph[s][t] = 1
-    return result
+    return (result, graph)
 
 # -------- Reference functions ---------
 
@@ -98,9 +98,30 @@ def findEdgeConflictsReference(s, t, graph = _GRAPH):
 # Adds 'edge' if it is not already part of the graph
 def findNewConflictsReference(s, t, graph = _GRAPH):
     graph[s][t] = 1
-    findEdgeConflictsReference(edge, graph)
+    return (findEdgeConflictsReference(s, t, graph), graph)
     
 # ---------- Testing functions ---------
+
+# Returns a size x size adjacency matrix
+#  with an average of 'density' 0/1 edge ratio
+def generateGraph(size, density=.5):
+    to_return = [[0 for _ in range(size)] for _ in range(size)]
+    for i in range(size):
+        for j in range(size):
+            if random.random() < density:
+                to_return[i][j] = 1
+    return to_return
+
+# Returns a list of edges contained in the given graph
+def graphEdges(graph):
+    to_return = []
+    for i in range(len(graph)):
+        if len(graph[i]) != len(graph):
+            raise "Graph must be a square matrix"
+        for j in range(len(graph[i])):
+            if graph[i][j] == 1:
+                to_return.append([i, j])
[…]

### 2020-02-25 pathfinding/pathFinding.py
Updated timing code

@@ -3,6 +3,7 @@
 
 import random
 from itertools import combinations
+import time
 
 _GRAPH = \
 [[0, 1, 1, 0],
@@ -186,20 +187,26 @@ def printDict(d):
 
 def main():
     # --- Accuracy Test ---
-    printDict(findAllPaths())
-    print("-------------------")
-    s, t = (0, 1)
-    print(findEdgeConflictsReference(s, t))
-    print("-------------------")
-    print(findEdgeConflicts(s, t))
+    # printDict(findAllPaths())
+    # print("-------------------")
+    # s, t = (0, 1)
+    # print(findEdgeConflictsReference(s, t))
+    # print("-------------------")
+    # print(findEdgeConflicts(s, t))
 
     # --- Speed Test ---
-    # size = 11
-    # edges = graphEdges(generateGraph(size))
-    # g = generateGraph(size, 0)
-    # for s,t in edges:
-    #     g = findNewConflicts(s, t, g)[1]
-    # print(g)
+    total = 0
+    size = 7
+    density = .35
+    iterations = 100
+    for _ in range(iterations):
+        edges = graphEdges(generateGraph(size, density))
+        g = generateGraph(size, 0)
+        tm = time.time()
+        for s,t in edges:
+            g = findNewConflictsReference(s, t, g)[1]
+        total += time.time() - tm
+    print(total / iterations)
 
 if __name__=="__main__":
     main()

### 2020-02-25 main.tex

@@
 Over the course of online addition of an edge, given that the graph before adding the edge was path independent, it would only be necessary to check conflicting pairs involving a new path that were created as a consequence of adding the new edge to the path.
-To identify all such conflicting pairs, we present the \textit{two-flip tolerant path search}.
+To identify all such conflicting pairs, we present the \textit{two-flip tolerant path search}. This naive approach may be considered a baseline solution.
 
@@
 
-\paragraph{Theorem} The result of the two-flip tolerant path search from the source to the sink node of an edge that is to be newly added has a one-to-one* correspondence to the set of new pairs of conflicting paths that arise in the graph due to its addition.
+The \textit{path extraction algorithm} then transforms the output of the two flip path search to the set of new pairs of conflicting paths that arise in the graph due to its addition.
 
 *= This excludes paths whose "area of difference" does not include the new edge since they should remain path independent.
+It also excludes paths that in their course contain cycles that exclude the new edge, since these paths must, again by path independence of the original graph, be equal to the path resulting from removing their cycles.
 
-\paragraph{Proof}
+\paragraph{Path extraction algorithm}
+The purpose of this algorithm is to derive the conflicting pairs that a given two-flip tolerant path corresponds to.
+
 Let the new edge added to the graph be (S, T).
@@
 Either the path is oriented from S to T, in which case the conflicting paths are simply the edge (S, T) and the entire flip-less path.
-Or, the path is oriented from T to S. In this case, the conflicting paths are the identity on S (that implicitly exists) and the cycle obtained from composing the new edge with the path.
+Or, the path is oriented from T to S.
+In this case, the conflicting paths are the identity on S (that implicitly exists) and the cycle obtained from composing the new edge with the path.
+The cycle conflicts with all other cycles starting from each node in the cycle.
+Notice that a single cycle can give rise to multiple conflicting pairs.
+
+\paragraph{Theorem} The result of the two-flip tolerant path search from the source to the sink node of an edge that is to be newly added, followed with the path extraction algorithm has a one-to-one* correspondence to the set of new pairs of conflicting paths that arise in the graph due to its addition.
+
+*= Same conditions as above.
+
+\paragraph{Proof}
+
+Every element in the output of the path extraction algorithm was by definition a conflicting pair.
 
@@
 
+*Insert python code here*
+
+\subsection{Implementation of path extraction}
+ *Again, insert code*
+
 \subsection{Verification using the new paths}
-The flip tolerant paths search returns pairs of conflicting paths.
+The algorithm returns pairs of conflicting paths.
 Verification is performed by composing edge functions to derive path functions, and then performing an equality check to verify if the path functions are equal.
@@
 
-% \subsubsection{Theorem: If two paths have one flip-point in common and there exist some path connecting the second flip points, then there exists a partial ordering between the two paths}
-
-% Two possible cases:
-
-% \begin{center}
-% \begin{tikzpicture}[scale=0.2]
-% \tikzstyle{every node}+=[inner sep=0pt]
-% \draw [black] (25,-31.9) circle (3);
-% \draw (25,-31.9) node {$S$};
-% \draw [black] (48,-31.9) circle (3);
-% \draw (48,-31.9) node {$T$};
-% \draw [black] (26.6,-13.6) circle (3);
-% \draw (26.6,-13.6) node {$F_1$};
-% \draw [black] (39,-21) circle (3);
-% \draw (39,-21) node {$F_2$};
-% \draw [black] (45.1,-12.9) circle (3);
-% \draw (45.1,-12.9) node {$F_3$};
-% \draw [black] (29.18,-15.14) -- (36.42,-19.46);
-% \fill [black] (36.42,-19.46) -- (35.99,-18.62) -- (35.48,-19.48);
-% \draw (31.38,-17.8) node [below] {$g_2$};
-% \draw [black] (23.529,-29.292) arc (-156.64712:-213.3464:14.141);
-% \fill [black] (23.53,-29.29) -- (23.67,-28.36) -- (22.75,-28.76);
-% \draw (21.8,-22.39) node [left] {$f_1$};
-% \draw [black] (27.889,-16.304) arc (19.98764:-29.98116:15.634);
-% \fill [black] (26.74,-29.46) -- (27.57,-29.02) -- (26.71,-28.52);
-% \draw (29.4,-23.08) node [right] {$f_2$};
-% \draw [black] (46.09,-29.59) -- (40.91,-23.31);
-% \fill [black] (40.91,-23.31) -- (41.03,-24.25) -- (41.81,-23.61);
-% \draw (42.95,-27.88) node [left] {$h_2$};
-% \draw [black] (29.6,-13.49) -- (42.1,-13.01);
-% \fill [black] (42.1,-13.01) -- (41.28,-12.54) -- (41.32,-13.54);
-% \draw (35.81,-12.7) node [above] {$g_1$};
-% \draw [black] (47.55,-28.93) -- (45.55,-15.87);
-% \fill [black] (45.55,-15.87) -- (45.18,-16.73) -- (46.17,-16.58);
-% \draw (47.25,-22.21) node [right] {$h_1$};
-% \draw [black] (40.8,-18.6) -- (43.3,-15.3);
-% \fill [black] (43.3,-15.3) -- (42.41,-15.63) -- (43.21,-16.24);
-% \draw (42.63,-18.35) node [right] {$l$};
-% \end{tikzpicture}
-% \end{center}
-
-% $P_2$ is better, intuitively because l can only lose- not add information.
-
-% \begin{center}
-% \begin{tikzpicture}[scale=0.2]
-% \tikzstyle{every node}+=[inner sep=0pt]
-% \draw [black] (25,-31.9) circle (3);
-% \draw (25,-31.9) node {$S$};
-% \draw [black] (48,-31.9) circle (3);
-% \draw (48,-31.9) node {$T$};
-% \draw [black] (26.7,-7.7) circle (3);
-% \draw (26.7,-7.7) node {$F_1$};
-% \draw [black] (43.1,-21.8) circle (3);
-% \draw (43.1,-21.8) node {$F_2$};
-% \draw [black] (29.9,-16.8) circle (3);
-% \draw (29.9,-16.8) node {$F_3$};
-% \draw [black] (29.408,-6.43) arc (107.51449:-8.88948:11.282);
-% \fill [black] (43.95,-18.93) -- (44.57,-18.22) -- (43.58,-18.06);
-% \draw (41.59,-8.14) node [above] {$g_1$};
-% \draw [black] (22.914,-29.751) arc (-141.57162:-226.46499:15.013);
-% \fill [black] (22.91,-29.75) -- (22.81,-28.81) -- (22.02,-29.44);
-% \draw (19.1,-19.32) node [left] {$f_1$};
-% \draw [black] (46.06,-22.008) arc (71.20036:-19.43982:5.814);
-% \fill [black] (46.06,-22.01) -- (46.66,-22.74) -- (46.98,-21.79);
-% \draw (50.12,-23.9) node [right] {$h_2$};
-% \draw [black] (45.319,-30.601) arc (-127.57654:-180.66292:7.327);
-% \fill [black] (42.46,-24.71) -- (41.95,-25.5) -- (42.95,-25.52);
-% \draw (42.49,-29.07) node [left] {$h_1$};
-% \draw [black] (28.97,-19.65) -- (25.93,-29.05);
-% \fill [black] (25.93,-29.05) -- (26.65,-28.44) -- (25.7,-28.13);
-% \draw (26.68,-23.67) node [left] {$f_2$};
-% \draw [black] (32.71,-17.86) -- (40.29,-20.74);
-% \fill [black] (40.29,-20.74) -- (39.72,-19.99) -- (39.37,-20.92);
-% \draw (35.16,-19.83) node [below] {$g_2$};
-% \draw [black] (27.7,-10.53) -- (28.9,-13.97);
-% \fill [black] (28.9,-13.97) -- (29.11,-13.05) -- (28.17,-13.38);
-% \end{tikzpicture}
-% \end{center}
-
-% $P_1$ is better for a similar reason.
-
-\section{Optimization}
+\subsection{Optimization of two-flip search}
 
@@
 
-\subsection{Polynomial verification bound}
- This is a special instance of case 1, where l and m are the identity.
+\section{Optimization}
+We now present a polynomial time algorithm, that is based on the observation that all conflicting paths that have the same source and same sink, but do not include the new edge, must be equal because of path independence in the pre-existing graph.
+Therefore, it suffices to check for one path for a given source and sink, and it follows that all the rest will check out.
+
+In terms of flip points, this is a special instance of case 1, where l and m are the identity.
 \begin{center}
@@
 
-In an N node graph, there are $^NC_2 = \frac{N(N-1)}{2}$ choices of flip points. Thus we can place a polynomial upper bound on the number of paths that actually need to be verified.
+In an N node graph, there are $^NP_2 = N(N-1)$ choices of flip points. Thus we can place a polynomial upper bound on the number of paths that actually need to be verified.
 
@@
 We utilize this idea of selecting flip points to devise a polynomial time path search.
-This search does not return all two flip tolerant paths, but rather, a path for every feasible choice of flipping point.
- *TODO Write algorithm.*
+This search does not return all flip tolerant paths, but rather, a path for every feasible choice of flipping point.
+
+\begin{verbatim}
+let (s,t) be the new edge.
+The graph being used for path search does not contain (s,t).
+
+# set of tuples storing conflicting pairs of paths
+set_of_pairs = {}
+
+for every node f_1 in the graph:
+    for every node f_2 in the graph:
+        try:
+            find a path p_1 from f_1 to s.
+            find a path p_2 from t to f_2.
+            let path p= p_1 + (s,t) + p_2.
+            find a path q from f_1 to f_2.
+            add (p,q) to set_of_pairs    
+        catch(path search failed):
+            print('There is no conflicting pair from f_1 and f_2 such that
+            one of the paths contains the new edge')
+
+equality_check_verify(set_of_pairs)
+
+\end{verbatim}
+
+As observed earlier, the number of paths that this algorithm returns is asymptotically tight.
+
+\subsection{Proof of correctness}
+
+*TODO fill out*
+
+- Ignoring cycle containing paths is ok
+
+- No point in doing a pair unless exactly one has new edge
+
+- Excepting the cases already dismissed, there is no conflicting pair such that its information is not covered by one of the paths in the final set (best way to write might be by cases).
 

### 2020-02-25 pathfinding/pathFinding.py
Polynomial algorithm

@@ -4,11 +4,12 @@
 import random
 from itertools import combinations
 import time
+import queue
 
 _GRAPH = \
 [[0, 1, 1, 0],
  [0, 0, 1, 0],
- [1, 0, 0, 0],
+ [0, 0, 0, 0],
  [0, 0, 0, 0]]
 
 # _GRAPH = \
@@ -69,6 +70,31 @@ def findNewConflicts(s, t, graph = _GRAPH):
             result.append(new_res)
     graph[s][t] = 1
     return (result, graph)
+
+# Finds a useful subset of conflicts in polynomial time
+def findEdgeConflictsPolynomial(s, t, graph = _GRAPH):
+    if graph[s][t] != 1:
+        raise "Edge to check conflicts with does not exist"
+    return findNewConflictsPolynomial(s, t, graph)[0]
+
+# Finds a useful subset of conflicts in polynomial time
+def findNewConflictsPolynomial(s, t, graph = _GRAPH):
+    graph[s][t] = 0
+    result = []
+    for i in range(len(graph)):
+        if len(graph[i]) != len(graph):
+            raise "Graph must be a square matrix"
+        for j in range(len(graph[i])):
+            p1 = findPath(i, s, graph)
+            p2 = findPath(t, j, graph)
+            p = p1 + p2
+            q = findPath(i, j, graph)
+            if p1 == [] or p2 == [] or q == []:
+                continue
+            result.append((p, q))
+    graph[s][t] = 1
+    return (result, graph)
+    
 
 # -------- Reference functions ---------
 
@@ -173,6 +199,29 @@ def findDirectedPathsRec(s, t, graph, visited, first_node, forward, flips_allowe
             ans += [([s]+lst[0], [s] + lst[1]) for lst in res]
     return ans + ([([t], [])] if s==t else [])
 
+# Finds a path from s to t
+def findPath(s, t, graph):
+    if s == t:
+        return [s]
+    visited = set()
+    visited.add(s)
+    toVisit = queue.Queue()
+    for neighbour in getNeighbours(s, graph, visited, True):
+        visited.add(neighbour)
+        path = [s, neighbour]
+        if neighbour == t:
+            return path
+        toVisit.put(path)
+    while toVisit.qsize() != 0:
+        next_path = toVisit.get()
+        for neighbour in getNeighbours(next_path[-1], graph, visited, True):
+            visited.add(neighbour)
+            path = next_path + [neighbour]
+            if neighbour == t:
+                return path
[…]

### 2020-02-26 main.tex
Updates from Overleaf

@@
 
+\paragraph{Theorem} This reduction rule is the most general possible.
+\paragraph{Proof} TODO write.
+Basically we create an axiom set where the only allowed action is composing with functions.
+This leads to only one inference rule to show implication between path equalities.
+We show that to derive p1>p2 using the inference rule some number of times, it is necessarily true that p1 and p2 have the relationship described in this case.
+
 \section{Optimization}
@@
 
+Also notice that if trying to optimize for path length (say, if composing functions is expensive) then "find any path" can be replaced with "find shortest path".
+
 \subsection{Proof of correctness}
@@
 
+\subsection{Minimal path set}
+While the algorithm above eliminates a large amount of redundancy, there is still potential for reducing the number of path checks.
+This is something we would want to do in situations in which equality checking is very expensive.
+By paying a cost in run-time, it is possible to derive a minimal set of paths to verify.
+
+We create a flip-point graph where each node represents a choice of flip point, so that there are $|V|^2$ nodes in all for every choice of flip point pair.
+An algorithm is run on the original graph to mark ever node with the set of its predecessors and successors.
+If node $n_1$ has a set of successors S and $n_2$ has a set of predecessors P then on the new flip-point graph we group together all the points in $S\times P$ in a set corresponding to the choice of adding $n_1, n_2$ to the minimal flipping point set.
+This is done for every node pair.
+Now the set cover problem is solved on the flip-point graph, and a minimal flip point set is obtained.
+
+The path finding algorithm is then performed with the minimal set as follows:
+
+\begin{verbatim}
+let (s,t) be the new edge.
+The graph being used for path search does not contain (s,t).
+
+# set of tuples storing conflicting pairs of paths
+set_of_pairs = {}
+
+for every node (f_1, f_2) in minimal_flipping_point_set:
+    try:
+        find a path p_1 from f_1 to s.
+        find a path p_2 from t to f_2.
+        let path p= p_1 + (s,t) + p_2.
+        find a path q from f_1 to f_2.
+        add (p,q) to set_of_pairs    
+    catch(path search failed):
+        print('There is no conflicting pair from f_1 and f_2 such that
+        one of the paths contains the new edge')
+
+equality_check_verify(set_of_pairs)
+
+\end{verbatim}
+
+\subsection{Proof of correctness}
+TODO: Write. Basically just say you're applying the reduction rule.
+
+\subsection{Proof of optimality}
+TODO: Write. Really just use the definition of the set cover problem.
+
+\section{Applications}
+
+TODO. Gator and Scala?
+
 \end{document}

### 2020-03-08 main.tex
idk

@@
 
-Every element in the output of the path extraction algorithm was by definition a conflicting pair.
+Every element in the output of the path extraction algorithm was by construction a conflicting pair.
 
@@
 Being the composition of equal parts, these paths must always be equal.
+Further, we are excluding pairs where one of the paths contains a cycle that excludes the new edge, because again the path independence of the original graph implies that the path with the cycle must be equal to the path with the cycle removed.
 
@@
 It is possible that some of $F_1, F_2, S$ and $T$ coincide, in which case the corresponding segments between them disappear, and the resultant path has fewer than two flips.
+
+The special case is that of cycles, where path 1 corresponds to some cycle in the graph that excludes the new edge and path 2 is a cycle that includes the new edge.
+Path 2 corresponds to a path with 0 flips that is caught by the two-flip tolerant algorithm. Subsequently the path extraction algorithm finds path 1 and generates the pair of path 1 and path 2.
+
+We have shown that all paths that need verification are caught by the two flip tolerant search followed with path extraction.
 
@@
 This leads to only one inference rule to show implication between path equalities.
-We show that to derive p1>p2 using the inference rule some number of times, it is necessarily true that p1 and p2 have the relationship described in this case.
+We show that to derive p1$>$p2 using the inference rule some number of times, it is necessarily true that p1 and p2 have the relationship described in this case.
 

### 2020-04-10 April_edition.tex

\documentclass{article}
\usepackage[utf8]{inputenc}
\usepackage{amsthm}
\usepackage{soul}
\usepackage{tikz}
\usepackage{amssymb}
\usepackage[options ]{algorithm2e}
\usepackage{listings}
\usepackage{xparse}
\usepackage{xcolor}

\newtheorem{lemma}{Lemma}
\newtheorem{theorem}{Theorem}
\newtheorem{definition}{Definition}

\title{Online verification of Commutativity}

\author{ }
\date{April 2020}

\begin{document}

\maketitle

\section{Introduction}

Many systems can be understood as a category, with a collection of sets, and some transformation functions between the sets- for example, in programming languages, type systems with coercion. If a graph is constructed with the sets as nodes and functions as edges between the nodes, then by construction the new graph is commutative. The idea is that each object has a unique representation in a given set in the system, so the transformation mapping between any two nodes should not depend on choice of path used to obtain the transformation.

Since commutative graphs arise in many systems, it is often useful to be able to perform verification and ensure the graphs in question are indeed path independent. 
% As a case study, gator, a domain specific graphics language that allows automatic conversion between geometric spaces internally creates a graph with he transformations between the spaces. 
Naively checking if all paths in a given graph are path independent could require a number of function equality checks as bad as factorial in the number of nodes, since a path consists of an ordering of nodes. 
We present a polynomial time algorithm to verify that commutativity of a graph is maintained over the course of online addition. It uses the minimal possible number of equality checks, since these checks can be expensive. 
To evaluate our results we use two case studies.
First, in the domain specific geometry type language \textit{Gator}, our algorithm is used to ensure that user defined transformations between spaces stay consistent, and the compiler, deterministic as it makes automatic conversion choices. As a second case study, the the algorithm is used to identify inefficiencies in a currency conversion graph.

\section{Problem setup and Terminology}
The closely related problem of identifying the minimal subset of paths to check in order to verify whether a given acyclic graph is commutative, is solved in \cite{commutative}. 
I reproduce relevant terminology and propositions from the paper here for convenience.

First, we formalize the notion of a graph that represents transformations.

We start with a directed graph G=(V,E) with spaces attached to vertices and maps to arcs. Each edge (u, v) in E corresponds to a function that maps elements of u to elements in v. 
This correspondence is stored as a mapping f:$E\rightarrow F$, where F is the semi-group that contains all the functions in the graph, with associative multiplication denoted by $\circ$. f maps each edge to the function it represents.

$P_{all}$ is the set of all paths in G. Since there may be cycles, it may be infinite.

A path is a sequence of edges. The edge-to-function mapping f can be naturally extended to paths. If path p=$e_1\circ e_2 ... e_n$ then $f(p)=f(e_1) \circ f(e_2) ... f(e_n)$.

The notation $\partial^{+}(p)$ is used to indicate that start node pf path p, and similarly $\partial^-(p)$ is the end node of path p. $\partial(p)$ maps to the tuple (start node of p, end node of p).

A pair of paths p and q is called \textit{parallel} iff their terminal nodes are the same, i.e. $\partial(p)=\partial(q)$.

$R_{all}$ is the set of all parallel paths in the graph.

The graph commutes iff f(p)=f(q) $\forall (p,q)\in R_{all}$, which is to say, the compositions of the maps along any path connecting u to v is independent of choice of path.

A bilinking is a parallel pair that is disjoint but for their terminal nodes. The set of all bilinkings is $R_0$.

The main result of \cite{commutative} can be used to verify a graph commutes in the acyclic case. The paper describes the following algorithm to find the ($V^2E$ bounded) minimal set of pairs that needs to be checked.

Define $r_1>r_2$ for bilinkings $r_1$ = $\{p_l,q_l\}$, $r_2$= $\{p_2, q_2)$ $\in$ $R_0$, if there exists a path p such that $\partial p=\partial r_1$ and p contains $p_2$.
The $\langle\rangle$ function is defined as:
$\langle r \rangle = \{ s\in R_0| r>s\}$.

For bilinking s, let F(s) be the vector in $\mathbb{F_2}^{|E|}$ representing the edges present in s (there is a dimension corresponding to each edge, the $n^{th}$ dimension of F(s) is 1 if the corresponding edge is in s, and 0 otherwise). Let this function be extended to sets, so that for some set of bilinkings S, F(S) = $\{ F(s) | s\in S \}$.

for a set of bilinkings r, the cl function is defined as:
cl(r) = $\{ s\in R_0|$ s is linearly dependent on $F(r) \}$.

Using these two function we finally define the function $\sigma$ on a set of bilinkings r as
$\sigma(r) = \{s \in R_0 | s\in cl(R\cap \langle s \rangle) \}$

Finally, an implementation of the described algorithm is

\begin{verbatim}
Find a spanning set Rs, = [r_1, ... ,r_k].

R := Rs,
    for i:=l to K do
        if r_i in sigma(R - r_i) then R:=R - r_i;
return R.

sigma(inputSet S, codomain = Rs):
    output = {}
    for bilinking in Rs:
        smallerPairs = {}
        for candidateBilinking in Rs:
            if exists path(source: bilinking.source, sink:candidateBilinking.source) and
            exists path(source: candidateBilinking.sink, sink: bilinking.sink):
                smallerPairs.add(candidateBilinking)
        # Check if there's a linear dependence between the bilinking and smallerPairs
        # vectorize(pair) returns a vector with a dimension for each edge in the graph,
        # with a 1 if pair includes that edge and 0 if it doesn't.
        matrix = [vectorize(pair) for pair in smallerPair]
        matrix = matrix+[vectorize(bilinking)]
        # if there is a linear dependence
        if determinant(matrix)%2==0:
            output.add(bilinking)
    return output

\end{verbatim}

and to initially set Rs,

\begin{verbatim}
Graph existingGraph;
R_s = {}
for each node v in V:
    # Let the subsection of the graph reachable from v be S.
    S = existingGraph.extractReachableSection(v)
    # Create a minimum spanning tree of S
    T = createMinimumSpanningTree(S)
    # Find the excluded edges
    excludedEdges = S.edges - T.edges
    # Use these edges to create bilinkings
    for each edge e in excludedEdges:
        firstPath = T.findPath(source: e.source, sink: e.sink)
        R_s.addElement(new Bilinking(firstPath, e)
return R_s
\end{verbatim}

The proof of why this works is complex and the subject of \cite{commutative}, 
% Actually I may be wrong about this
% but at a very high level the cl function captures the redundancy one gets from how "gluing together" bilinkings. For example in the graph below verifying any two of the bilinkings commute implies that the third must too. So cl(any two)=all three bilinkings.

% \begin{center}
% \begin{tikzpicture}[scale=0.2]
% \tikzstyle{every node}+=[inner sep=0pt]
% \draw [black] (15.1,-25.8) circle (3);
% \draw (15.1,-25.8) node {$A$};
% \draw [black] (31.6,-34.7) circle (3);
% \draw (31.6,-34.7) node {$C$};
% \draw [black] (31.6,-18.9) circle (3);
% \draw (31.6,-18.9) node {$B$};
% \draw [black] (47.7,-26.5) circle (3);
% \draw (47.7,-26.5) node {$D$};
% \draw [black] (17.87,-24.64) -- (28.83,-20.06);
% \fill [black] (28.83,-20.06) -- (27.9,-19.9) -- (28.29,-20.83);
% \draw [black] (31.6,-21.9) -- (31.6,-31.7);
% \fill [black] (31.6,-31.7) -- (32.1,-30.9) -- (31.1,-30.9);
% \draw [black] (17.74,-27.22) -- (28.96,-33.28);
% \fill [black] (28.96,-33.28) -- (28.49,-32.46) -- (28.02,-33.34);
% \draw [black] (34.27,-33.34) -- (45.03,-27.86);
% \fill [black] (45.03,-27.86) -- (44.09,-27.78) -- (44.54,-28.67);
% \draw [black] (34.31,-20.18) -- (44.99,-25.22);
% \fill [black] (44.99,-25.22) -- (44.48,-24.43) -- (44.05,-25.33);
% \end{tikzpicture}
% \end{center}

% The $\langle \rangle$ function captures structures like this one, which display the redundancy above. Here, for p=($p_1, p_2), q=(q_1, q_2)$, we have that $p>q$.

% \begin{center}
% \begin{tikzpicture}[scale=0.2]
% \tikzstyle{every node}+=[inner sep=0pt]
% \draw [black] (20.7,-15.7) circle (3);
% \draw (20.7,-15.7) node {$A$};
% \draw [black] (44,-15.7) circle (3);
% \draw (44,-15.7) node {$B$};
% \draw [black] (20.7,-33) circle (3);
% \draw (20.7,-33) node {$A'$};
% \draw [black] (44,-33) circle (3);
% \draw (44,-33) node {$B'$};
% \draw [black] (22.498,-13.306) arc (136.73613:43.26387:13.529);
% \fill [black] (42.2,-13.31) -- (42.02,-12.38) -- (41.29,-13.07);
% \draw (32.35,-8.55) node [above] {$p_1$};
% \draw [black] (41.785,-17.717) arc (-53.13545:-126.86455:15.728);
% \fill [black] (41.79,-17.72) -- (40.85,-17.8) -- (41.45,-18.6);
% \draw (32.35,-21.36) node [below] {$p_2$};
% \draw [black] (23.143,-31.264) arc (120.64183:59.35817:18.065);
% \fill [black] (41.56,-31.26) -- (41.12,-30.43) -- (40.61,-31.29);
% \draw (32.35,-28.24) node [above] {$q_1$};
% \draw [black] (41.52,-34.683) arc (-60.46289:-119.53711:18.601);
% \fill [black] (41.52,-34.68) -- (40.58,-34.64) -- (41.07,-35.51);
% \draw (32.35,-37.6) node [below] {$q_2$};
% \draw [black] (20.7,-18.7) -- (20.7,-30);
% \fill [black] (20.7,-30) -- (21.2,-29.2) -- (20.2,-29.2);
% \draw (20.2,-24.35) node [left] {$\alpha$};
% \draw [black] (44,-30) -- (44,-18.7);
% \fill [black] (44,-18.7) -- (43.5,-19.5) -- (44.5,-19.5);
% \draw (44.5,-24.35) node [right] {$\beta$};
% \end{tikzpicture}
% \end{center}
% $\sigma$ weaves these together to identify, given an input subset, the set of bilinkings that would be commutative should the input subset be verified. 
$\sigma$ identifies reduncancies arising from the structure of the graph and then the approach is to greedily eliminate redundancies.
\cite{commutative} solves the problem of verifying whether a given acyclic diagram is commutative.

In the next section, the distinct problem of verification that a diagram (that may contain cycles) commutes on addition of a new edge is solved. The polynomial time algorithm is found to entail at worst O($|V|^2$) checks and also can be extended to identify the minimum possible set of checks. If the algorithm is run over the course of construction of the graph, it has the same worst case asymptotic runtime O($|V|^2|E|)$ as the one already described.

\section{Solving the online addition problem}

The first step is to identify all bilinkings that involve the new edge.
Bilinkings that do not involve the new edge must already commute by assumption.
Bilinkings where both paths involve the new edge cannot exist since bilinkings are node-disjoint.
Further, for a given (source, sink) pair only a single bilinking needs to be verified. This is an implication of Theorem \ref{reductionRule}.
This fact, along with the equivalent proposition 1.1 from \cite{commutative} for graphs with cycles means that should our selected set of bilinkings along with cycles passing through the new edge be verified commutative, the entire graph must commute. %TODO prove this equivalent?
% \begin{theorem}
% Verifying that all bilinkings in the graph commute implies that the entire graph must commute.
% \end{theorem}

\begin{verbatim}
Graph existingGraph;
Edge newEdge;

Set bilinkings = new Set();
for S in existingGraph.Nodes():
    for T in existingGraph.Nodes():
        try:
            Path pathWithNewEdge = 
                FindPath(
                    graph: existingGraph, 
                    sourceNode: S,
                    sinkNode: newEdge.Source()) +
                newEdge as Path +
                FindPath(
                    graph: existingGraph, 
                    sourceNode: newEdge.Sink(), 
                    sinkNode: T)
            if (S==T):
                // Cycles are a speacial case.
                pathInOldGraph = pathWithNewEdge + pathWithNewEdge
            else:
                Path pathInOldGraph = FindPath(
                    graph: existingGraph, 
                    sourceNode: S, 
                    sinkNode: T);
            bilinkings.add((pathInOldGraph, pathWithNewEdge));
        catch (pathFindingFailed Exception):
            print(
                $"No comparable pairs from node {S} to node {T} 
                that need to be checked");

output bilinkings;
\end{verbatim}

The try block is executed at most n(n-1)/2 = $O(n^2)$ times where n is the number of nodes.

The bound is asymptotically tight.
This can be seen in the case where the graph contains 2N nodes besides S and T. We consider N of the nodes to be in group 1, and the other N to be in group 2. Every node in group 1 has a forward node to every node in group 2, as well as to S. T has a forward edge to every node in group 2. In this graph, on the addition of edge (S, T), $N^2$ paths need to be verified which is polynomial in 2N+2.

Also notice that if trying to optimize for path length (say, if composing functions is expensive) then "find any path" can be replaced with "find shortest path".

\subsection{(Optional) Optimization Step}

In the case where equality checks are very expensive, we begin by finding the minimal set of (source, sink) pairs such that checking for these pairs logically implies having checked the full graph.

We observe that there are some redundancies in the graph.

Consider the following situation:

\begin{center}
\begin{tikzpicture}[scale=0.2]
\tikzstyle{every node}+=[inner sep=0pt]
\draw [black] (16.6,-34.1) circle (3);
\draw (16.6,-34.1) node {$S$};
\draw [black] (56.8,-33.6) circle (3);
\draw (56.8,-33.6) node {$T$};
\draw [black] (28,-20.1) circle (3);
\draw (28,-20.1) node {$P_1$};
\draw [black] (44.5,-20.1) circle (3);
\draw (44.5,-20.1) node {$P_2$};
\draw [black] (28,-44.9) circle (3);
\draw (28,-44.9) node {$Q_1$};
\draw [black] (44.5,-44.9) circle (3);
\draw (44.5,-44.9) node {$Q_2$};
\draw [black] (31,-20.1) -- (41.5,-20.1);
\fill [black] (41.5,-20.1) -- (40.7,-19.6) -- (40.7,-20.6);
\draw (36.25,-20.6) node [below] {$g_1$};
\draw [black] (26.11,-22.43) -- (18.49,-31.77);
\fill [black] (18.49,-31.77) -- (19.39,-31.47) -- (18.61,-30.84);
\draw (21.74,-25.67) node [left] {$f_1$};
\draw [black] (25.82,-42.84) -- (18.78,-36.16);
\fill [black] (18.78,-36.16) -- (19.01,-37.08) -- (19.7,-36.35);
\draw (23.57,-39.02) node [above] {$f_2$};
\draw [black] (31,-44.9) -- (41.5,-44.9);
\fill [black] (41.5,-44.9) -- (40.7,-44.4) -- (40.7,-45.4);
\draw (36.25,-45.4) node [below] {$g_2$};
\draw [black] (54.59,-35.63) -- (46.71,-42.87);
\fill [black] (46.71,-42.87) -- (47.64,-42.7) -- (46.96,-41.96);
\draw (49.21,-38.76) node [above] {$h_2$};
\draw [black] (54.78,-31.38) -- (46.52,-22.32);
\fill [black] (46.52,-22.32) -- (46.69,-23.25) -- (47.43,-22.57);
\draw (51.19,-25.39) node [right] {$h_1$};
\draw [black] (28,-23.1) -- (28,-41.9);
\fill [black] (28,-41.9) -- (28.5,-41.1) -- (27.5,-41.1);
\draw (27.5,-32.5) node [left] {$l$};
\draw [black] (19.6,-34.06) -- (53.8,-33.64);
\fill [black] (53.8,-33.64) -- (52.99,-33.15) -- (53.01,-34.15);
\draw [black] (44.5,-41.9) -- (44.5,-23.1);
\fill [black] (44.5,-23.1) -- (44,-23.9) -- (45,-23.9);
\draw (45,-32.5) node [right] {$m$};
\end{tikzpicture}
Each arrow represents a path, and (S,T) is the new edge being added.
\end{center}
\paragraph{Theorem} 
\label{reductionRule}
If conflicting paths $g_2 = f_2; (S,T); h_2$ then it must be that $g_1 = f_1; (S,T); h_1$.

% monomorphisms

\paragraph{Proof}
We use the fact that $f_1$=$l; f_2$ and $h_1$=$h_2; m$.
\[g_2 = f_2; (S,T); h_2 \Rightarrow l; g_2 = l; f_2; (S,T) ; h_2 \]
\[\Rightarrow l ; g_2 ; m = l ; f_2 ; (S,T) ; h_2 ; m \Rightarrow g_1 = f_1 ; (S,T) ; h_1\]

Note that the proof is not affected if an of these paths is the identity, eg. if $f_1$ is the identity and S and $P_1$ are actually the same node.

We conclude that verifying a comparable pair of paths with end points ($P_1$, $P_2$) implies the verification of all path pairs ($Q_1$, $Q_2$) such that $Q_1$ is a successor of $P_1$ and $P_2$ is a successor of $Q_2$. A successor S to node N is any node such that there exists a path from N to S. Nodes are also their own successors and predecessors.

Under the assumption that the only path operations allowed are composition and replacement of one path by a different, equal path, as would be true when edges are generic functions and no other information is available, this reduction rule is also the only reduction rule.

That is to say, if verifying a comparable pair of paths with end points ($P_1$, $P_2$) implies the verification of a pair with endpoints ($Q_1$, $Q_2$), then it must be that $Q_1$ is a successor of $P_1$ and $P_2$ is a successor of $Q_2$. %TODO proof required?

Using this information it is possible to choose a minimal subset of path pairs to verify.

We construct a graph with a node for each possible (source, sink) pair in the graph- each node then represents a possible choice for bilinking endpoint pairs.

\begin{verbatim}
# Existng graph
Graph existingGraph;
# New edge
Edge (S, T);

predecessors = S.predecessors(existingGraph)
successors = T.successors(existingGraph)

# Construct the graph
Graph terminalPairGraph;
for q in successors:
    for p in predecessors:
        terminalPairGraph.addNode(q, p)
        # addEdge function should handle addition of edges between nodes that 
        # haven't been defined yet
        for q_pred in q.predecessors(existingGraph):
            for p_succ in p.successors(existingGraph):
                terminalPairGraph.addEdge((q_pred, p_succ))

# Now we reduce the graph.
verificationSet = {}
while (len(terminalPairGraph.nodes) > 0):
    # Choose a random node to start reduction from.
    current_node = terminalPairGraph.nodes[0]
    visited_nodes = set()
    # Climb up the graph till a "root" node is found
    while (len(currentNode.parents()) > 0):
        visited_nodes.add(currentNode)
        # Choose a parent at random.
        currentNode = currentNode.parents()[0]
        # Deal with cycles by reducing them
        if currentNode in visited_nodes:
            edges = getAllEdges(visited_nodes, terminalPairGraph)
            terminalPairGraph.removeNodes(visitedNodes)
            # All the incoming and outgoing edges from the entire visited set are now 
            # transferred to a single node that represents the class
            terminalPairGraph.addNode(currentNode, edges)
    # No predecessor! We should have reached a root!
    verificationSet.add(currentNode)
    terminalPairGraph.removeNodes(currentNode.successors(terminalPairGraph)

# This contains the minimal set of (source, sink) pairs whose verification 
# implies verification of the entire graph. 
return verificationSet

\end{verbatim}

The result of the algorithm cannot be further reduced. Further, the verification of the bilinkings returned in the algorithm implies that the output of the previous algorithm must commute, and transitively, that the graph must commute.
% Here is a naive implemenetation. Note that an important optimization in sparse graphs can be made: Source-sink pairs with no viable adjacent pair should never be added to the graph (as verified by the try-catch block from the first algorithm).

\subsection{Verification}
Ultimately verification is performed simply by calling an equality oracle for every comparable pair returned by the previous stage.

\begin{verbatim}
for (path_1, path_2) in bilinkings:
    if (path_1 != path_2):
        print($"Edge addition does not maintain path independence. 
        Counterexample: {path_1}, {path_2}")
        return False
print("Edge addition maintains path independence.")
return True
\end{verbatim}

\begin{thebibliography}{9}
\bibitem{commutative} 
Kazuo Murota. 
\textit{HOMOTOPY BASE OF ACYCLIC GRAPHS - A COMBINATORIAL ANALYSIS OF COMMUTATIVE BY MEANS OF PREORDERED MATROID}. 
Discrete Applied Mathematics, 1987.

\end{thebibliography}
\end{document}

### 2020-04-12 April_edition.tex
Made modifications to deal with cycles

@@
 
-Define $r_1>r_2$ for bilinkings $r_1$ = $\{p_l,q_l\}$, $r_2$= $\{p_2, q_2)$ $\in$ $R_0$, if there exists a path p such that $\partial p=\partial r_1$ and p contains $p_2$.
-The $\langle\rangle$ function is defined as:
-$\langle r \rangle = \{ s\in R_0| r>s\}$.
+% Define $r_1>r_2$ for bilinkings $r_1$ = $\{p_l,q_l\}$, $r_2$= $\{p_2, q_2)$ $\in$ $R_0$, if there exists a path p such that $\partial p=\partial r_1$ and p contains $p_2$.
+% The $\langle\rangle$ function is defined as:
+% $\langle r \rangle = \{ s\in R_0| r>s\}$.
 
-For bilinking s, let F(s) be the vector in $\mathbb{F_2}^{|E|}$ representing the edges present in s (there is a dimension corresponding to each edge, the $n^{th}$ dimension of F(s) is 1 if the corresponding edge is in s, and 0 otherwise).
+% For bilinking s, let F(s) be the vector in $\mathbb{F}2^{|E|}$ representing the edges present in s (there is a dimension corresponding to each edge, the $n^{th}$ dimension of F(s) is 1 if the corresponding edge is in s, and 0 otherwise).
 Let this function be extended to sets, so that for some set of bilinkings S, F(S) = $\{ F(s) | s\in S \}$.
 
-for a set of bilinkings r, the cl function is defined as:
+% for a set of bilinkings r, the cl function is defined as:
 cl(r) = $\{ s\in R_0|$ s is linearly dependent on $F(r) \}$.
 
-Using these two function we finally define the function $\sigma$ on a set of bilinkings r as
-$\sigma(r) = \{s \in R_0 | s\in cl(R\cap \langle s \rangle) \}$
+% Using these two function we finally define the function $\sigma$ on a set of bilinkings r as
+% $\sigma(r) = \{s \in R_0 | s\in cl(R\cap \langle s \rangle) \}$
 
-Finally, an implementation of the described algorithm is
+An implementation of the described algorithm is
+
+\begin{verbatim}
+Graph existingGraph;
+R_s = {}
+for each node v in V:
+    # Let the subsection of the graph reachable from v be S.
+    S = existingGraph.extractReachableSection(v)
+    # Create a minimum spanning tree of S
+    T = createMinimumSpanningTree(S)
+    # Find the excluded edges
+    excludedEdges = S.edges - T.edges
+    # Use these edges to create bilinkings
+    for each edge e in excludedEdges:
+        firstPath = T.findPath(source: e.source, sink: e.sink)
+        R_s.addElement(new Bilinking(firstPath, e)
+return R_s
+\end{verbatim}
 
@@
 
-\end{verbatim}
-
-and to initially set Rs,
-
-\begin{verbatim}
-Graph existingGraph;
-R_s = {}
-for each node v in V:
-    # Let the subsection of the graph reachable from v be S.
-    S = existingGraph.extractReachableSection(v)
-    # Create a minimum spanning tree of S
-    T = createMinimumSpanningTree(S)
-    # Find the excluded edges
-    excludedEdges = S.edges - T.edges
-    # Use these edges to create bilinkings
-    for each edge e in excludedEdges:
-        firstPath = T.findPath(source: e.source, sink: e.sink)
-        R_s.addElement(new Bilinking(firstPath, e)
-return R_s
 \end{verbatim}
@@
 % $\sigma$ weaves these together to identify, given an input subset, the set of bilinkings that would be commutative should the input subset be verified. 
-$\sigma$ identifies reduncancies arising from the structure of the graph and then the approach is to greedily eliminate redundancies.
+But at a high level, $\sigma$ identifies reduncancies arising from the structure of the graph and then the approach is to greedily eliminate redundancies.
 \cite{commutative} solves the problem of verifying whether a given acyclic diagram is commutative.
@@
 
-The first step is to identify all bilinkings that involve the new edge.
-Bilinkings that do not involve the new edge must already commute by assumption.
-Bilinkings where both paths involve the new edge cannot exist since bilinkings are node-disjoint.
-Further, for a given (source, sink) pair only a single bilinking needs to be verified. This is an implication of Theorem \ref{reductionRule}.
-This fact, along with the equivalent proposition 1.1 from \cite{commutative} for graphs with cycles means that should our selected set of bilinkings along with cycles passing through the new edge be verified commutative, the entire graph must commute.
+The solution presented here assumes that there is an implicit identity function on every node, as would be the case in most applications.
+However, a modified version of the algorithm for graphs without this assumption is presented in the appendix.
+
+The first step is to identify all pairs that involve the new edge.
+Pairs that do not involve the new edge must already commute by assumption.
+Further, for a given (source, sink) pair only a single parallel pair needs to be verified. This is an implication of Theorem \ref{reductionRule}.
+Theorem \ref{verifyingSet} shows that should our selected set of pairs along with cycles passing through the new edge be verified commutative, the entire graph must commute.
 %TODO prove this equivalent?
 % \begin{theorem}
-% Verifying that all bilinkings in the graph commute implies that the entire graph must commute.
+% Verifying that all parallel pairs in the graph commute implies that the entire graph must commute.
 % \end{theorem}
@@
 
-Set bilinkings = new Set();
+Set parallelPairs = new Set();
 for S in existingGraph.Nodes():
@@
                 // Cycles are a speacial case.
-                pathInOldGraph = pathWithNewEdge + pathWithNewEdge
+                pathInOldGraph = identity(S)
             else:
@@
                     sinkNode: T);
-            bilinkings.add((pathInOldGraph, pathWithNewEdge));
+            parallelPairs.add((pathInOldGraph, pathWithNewEdge));
         catch (pathFindingFailed Exception):
-            print(
-                $"No comparable pairs from node {S} to node {T} 
-                that need to be checked");
+            // No comparable pairs from node {S} to node {T} 
+            // that need to be checked
+            continue
 
-output bilinkings;
+output parallelPairs;
 \end{verbatim}
@@
 \end{center}
-\paragraph{Theorem} 
+\begin{theorem}
 \label{reductionRule}
 If conflicting paths $g_2 = f_2; (S,T); h_2$ then it must be that $g_1 = f_1; (S,T); h_1$.
+\end{theorem}
 
@@
 
-\paragraph{Proof}
+\begin{proof}
 We use the fact that $f_1$=$l; f_2$ and $h_1$=$h_2; m$.
@@
 \[\Rightarrow l ; g_2 ; m = l ; f_2 ; (S,T) ; h_2 ; m \Rightarrow g_1 = f_1 ; (S,T) ; h_1\]
+\end{proof}
 
@@
 
-Under the assumption that the only path operations allowed are composition and replacement of one path by a different, equal path, as would be true when edges are generic functions and no other information is available, this reduction rule is also the only reduction rule.
+Under the assumption that the only path operations allowed are composition and replacement of one path by a different, equal path, as would be true when edges are generic functions, and no other information is available, so that F is a semi-group, this "reduction" rule is also the only rule to reduce the set of path pairs to check by finding implications.
 
@@
 
-We construct a graph with a node for each possible (source, sink) pair in the graph- each node then represents a possible choice for bilinking endpoint pairs.
+We construct a graph with a node for each possible (source, sink) pair in the graph- each node then represents a possible choice for parallel pair endpoint pairs, and greedily search for the smallest set of nodes from which the entire graph would be reachable.
 
@@
 The result of the algorithm cannot be further reduced.
-Further, the verification of the bilinkings returned in the algorithm implies that the output of the previous algorithm must commute, and transitively, that the graph must commute.
+Also, the verification of the parallel pairs returned in the algorithm implies that the output of the previous algorithm must commute, and transitively, that the entire graph must commute.
 % Here is a naive implemenetation.
@@
 \subsection{Verification}
-Ultimately verification is performed simply by calling an equality oracle for every comparable pair returned by the previous stage.
+Ultimately verification is performed by calling an equality oracle (that is beyond the scope of this work) for every comparable pair returned by the previous stage.
 
 \begin{verbatim}
-for (path_1, path_2) in bilinkings:
+for (path_1, path_2) in parallelPairs:
     if (path_1 != path_2):
@@
 \end{verbatim}
+
+\section{Applications}
+TODO.
+\subsection{Gator}
+\subsection{Currency Graph}
+
+\section{Appendix: Theorems and additional cases}
+This should be read as a continuation of the notation section, the definitions there will be used.
+
+%Define a cycle-pair to mean the pair (c, c;c) where c is a cycle, which is to say the pair whose first element is a cycle and second element is the cycle traversed twice.
+*tempted to name this bi-cycle*.
+
+%Let i be the identity function. Define a cycle-identity pair for cycle c to be (c, i). *should I call this unicycle?*
+
+Consider all the cycles in the graph that pass through the new edge, unique to their terminal point.
+Define a cycle-pair corresponding to cycle c to mean the pair of paths (c, identity on terminal node of c).
+Let $C$ be the set of all cycle-pairs (unique to their terminal point) corresponding to cycles that pass through the new edge.
+
+Let V, the set of pairs to verify, be $C \cup R_0$.
+
+\begin{lemma}
+A path that involves the new edge more than once is equal to a path without multiple occurrences of the edge, under the assumption that the pairs in C are verified.
+\end{lemma}
+\begin{proof}
+Consider the part of the path between the first occurrence of the new edge and the second. This forms a cycle, which has a corresponding check in $C$ and must be verified to be equal to the identity.
+This means that the loops containing the multiple occurrences of the new edge may be ignored and the path with a single occurrence of the new edge with the cycles removed is equal to this path.
+\end{proof}
+
+\begin{theorem}
+\label{verifyingSet}
+Verifying that all pairs in V commute implies that all pairs in $R_{all}$ commute.
+\end{theorem}
+\begin{proof}
+% We do a kind of "structural induction" to show that verification every pair in $R_{all}$ that aren't in V can wither be reduced down to the verification of smaller components or can be shown to be implied by the verification of V.
+
+Each pair in $R_{all}$ that is not in V would fall into one of these categories:
+
+Case A: pairs that do not involve the new edge: By assumption that the original graph commutes, these pairs must commute.
+
+Case B: Cases where there is a different pair corresponding to this choice of (source, sink) in V already.
+There are four paths to be considered, two from each pair. Those paths which do not include the new edge must already be equal by the assumption that the original graph commutes.
+Those paths which do include the new edge (S,T) can be reduced, by the case for multiple occurrences of a new edge, to a path with a single occurrence of the new edge as described in the preceding lemma.
+These new paths can be divided into three segments: the segment from the source to S, the new edge, and the segment from the last occurrence of T to the sink.
+The first segments of all paths are equal by assumption that the original graph commutes.
+This is also true of the third segment of all the paths.
+The second segment consists of only the new edge and is the same for all graphs.
+The composition of equal functions is equal, so all the paths passing through the new edge must be equal.
+Therefore it is sufficient to check one path that passes through the new edge and one that doesn't. Such a pair belongs to V by construction.
+\end{proof}
+
+Modified algorithm for graphs without the identity:
+
+\begin{verbatim}
+Graph existingGraph;
+Edge newEdge;
+
+Set parallelPairs = new Set();
+for S in existingGraph.Nodes():
+    for T in existingGraph.Nodes():
+        try:
+            Path pathWithNewEdge = 
+                FindPath(
+                    graph: existingGraph, 
+                    sourceNode: S,
+                    sinkNode: newEdge.Source()) +
+                newEdge as Path +
+                FindPath(
+                    graph: existingGraph, 
+                    sourceNode: newEdge.Sink(), 
+                    sinkNode: T)
+            if (S==T):
+                // Cycles are a special case.
+                parallelPairs.add(pathWithNewEdge, pathWithNewEdge+pathWithNewEdge)
+                for edge in (S.forwardEdges - pathWithNewEdge[0]):
+                    parallelPairs.add(edge, edge+pathWithNewEdge)
+                for edge in (S.backwardEdges - pathWithNewEdge[-1]):
+                    parallelPairs.add(edge, pathWithNewEdge+edge)
+            else:
+                Path pathInOldGraph = FindPath(
+                    graph: existingGraph, 
+                    sourceNode: S, 
+                    sinkNode: T);
+                parallelPairs.add((pathInOldGraph, pathWithNewEdge));
+        catch (pathFindingFailed Exception):
+            // No comparable pairs from node {S} to node {T} 
+            // that need to be checked
+            continue
+    for cycle in findCyclesWithTerminatNode(newEdge.Source()):
+        parallelPairs.add(newEdge, cycle+newEdge)
+    for cycle in findCyclesWithTerminatNode(newEdge.Source()):
+        parallelPairs.add(newEdge, newEdge+cycle)
+output parallelPairs;
+\end{verbatim}
+
+The optional minimization algorithm is modified to exclude nodes representing cycles, since the reduction case no longer will apply in the absence of an implicit identity on the terminal node of the cycle.
+*I think the minimality of minimized subset still holds, but not 100\% sure*
+
+\section{Case Studies}
 

### 2020-04-22 April_edition.tex
commented out an unwanted line
Idk

@@
 Many systems can be understood as a category, with a collection of sets, and some transformation functions between the sets- for example, in programming languages, type systems with coercion.
-If a graph is constructed with the sets as nodes and functions as edges between the nodes, then by construction the new graph is commutative.
+If a directed graph is constructed with the sets as nodes and functions as edges between the nodes, then by construction, in the new graph, choice of path should not matter- any choice of path should yield the same transformation between any two nodes.
 The idea is that each object has a unique representation in a given set in the system, so the transformation mapping between any two nodes should not depend on choice of path used to obtain the transformation.
+We will use the term "diagram" to refer to such graphs with their associated sets (vertices) and transformations (edges).
 
-Since commutative graphs arise in many systems, it is often useful to be able to perform verification and ensure the graphs in question are indeed path independent. 
+Since such commutative diagrams arise in many systems, it is often useful to be able to perform verification and ensure the graphs in question are indeed path independent. 
+We formalize the problem as checking the commutativity of a diagram incrementally.
+A diagram is said to commute when on selecting any two vertices, any directed path leads to the same result by composition.
 % As a case study, gator, a domain specific graphics language that allows automatic conversion between geometric spaces internally creates a graph with he transformations between the spaces. 
 Naively checking if all paths in a given graph are path independent could require a number of function equality checks as bad as factorial in the number of nodes, since a path consists of an ordering of nodes.
-We present a polynomial time algorithm to verify that commutativity of a graph is maintained over the course of online addition.
-It uses the minimal possible number of equality checks, since these checks can be expensive.
+Previous work has identified an $O(|E|^2|V|^4)$ algorithm to verify that an acyclic diagram commutes.
+
+We present a polynomial time $O(|E||V|^2+|V|^3)$ algorithm to verify that commutativity of a graph is maintained over the course of online addition, assuming an oracle to check the abstract notion of equality of functions.
+The number of calls to the equality checking oracle required by the algorithm is asymptotically tight.
+With an additional $O(|V|^4)$ optimization step, is is possible to use instead the minimal possible number of equality checks, should these checks be expensive.
+We compare our solution to three baselines: a naive checking of all path pairs, with only special handling for cycles, a check for all path pairs that involve the new edge, and with an algorithm suggested by previous work to solve the batch version of the problem for acyclic diagrams.
+
 To evaluate our results we use two case studies.
@@
 
-\section{Problem setup and Terminology}
-The closely related problem of identifying the minimal subset of paths to check in order to verify whether a given acyclic graph is commutative, is solved in \cite{commutative}. 
-I reproduce relevant terminology and propositions from the paper here for convenience.
+\section{Formal Problem Setup and Terminology}
 
-First, we formalize the notion of a graph that represents transformations.
+First, we formalize the notion of a diagram, drawing inspiration from Murota\cite{commutative}.
 
-We start with a directed graph G=(V,E) with spaces attached to vertices and maps to arcs. Each edge (u, v) in E corresponds to a function that maps elements of u to elements in v. 
-This correspondence is stored as a mapping f:$E\rightarrow F$, where F is the semi-group that contains all the functions in the graph, with associative multiplication denoted by $\circ$.
-f maps each edge to the function it represents.
+We start with a directed graph G=(V,E) with sets attached to vertices and maps to arcs. Each edge (u, v) in E corresponds to a function that maps elements of u to elements in v.
+All the functions form a semigroup F, with multiplication denoted by $\circ$.
+A semigroup is an algebraic structure with a set and an associative binary operation. We use it to represent functions in order to capture the one allowed operation- function composition.
 
-$P_{all}$ is the set of all paths in G. Since there may be cycles, it may be infinite.
+w
+The correspondence between edges and functions is stored as a mapping f:$E\rightarrow F$. f maps each edge to the function it represents.
 
@@
 
-The notation $\partial^{+}(p)$ is used to indicate that start node pf path p, and similarly $\partial^-(p)$ is the end node of path p. $\partial(p)$ maps to the tuple (start node of p, end node of p).
+Let $P_{all}$ is the set of all paths in G.
 
-A pair of paths p and q is called \textit{parallel} iff their terminal nodes are the same, i.e. $\partial(p)=\partial(q)$.
+The notation $\partial^{+}(p)$ is used to indicate the start node of path p, and similarly $\partial^-(p)$ is the end node of path p. $\partial(p)$ maps to the tuple (start node of p, end node of p).
 
-$R_{all}$ is the set of all parallel paths in the graph.
+A pair of paths p and q is called \textit{parallel} iff their terminal nodes are the same, i.e., $\partial(p)=\partial(q)$.
+The problem of checking if diagrams commute amounts to checking if the two members of every parallel pair that can be constructed is equal.
 
-The graph commutes iff f(p)=f(q) $\forall (p,q)\in R_{all}$, which is to say, the compositions of the maps along any path connecting u to v is independent of choice of path.
+Let $R_{all}$ be the set of all parallel pairs of paths in the diagram.
+
+The diagram commutes iff f(p)=f(q) $\forall (p,q)\in R_{all}$, which is to say, the compositions of the maps along any path connecting u to v is independent of choice of path.
+Here, equality of functions is defined by an equality oracle whose exact specifics will vary with the nature of the functions and the domain to which the diagram belongs.
+
+\section{Baselines}
+
+\subsection{Naive all-path check}
+The naive solution is to verify that in every parallel pair in the diagram, the two paths are equal.
+However since cycles potentially introduce an infinite number of pairs, they must be dealt with separately.
+
+Two situation can be imagined- the first, where the existence of the identity function from every node to itself is assumed, and the second, where no such identity needs to exist.
+
+When the identity is assumed, every cycle forms a parallel pair with an identity function on the terminal node.
+Therefore, the cycle must be verified to be equal to the identity on every node in the cycle.
+If this check succeeds for every elementary cycle in the diagram, then paths with multiple instance of the cycle must be equal to the version of the path without the cycle.
+It is therefore then enough to perform verification for all pairs where none of the path involves cycles.
+This is a finite set, bounded by $^{|E|!}C_2$.
+All elementary cycles in a diagram can be found in $O((|V|+|E|)(C+1))$ using Johnson's algorithm where C is the number of cycles, that can be greater than $2^{|V|}$.
+Thus the baseline is a factorial time algorithm that is nevertheless guaranteed to terminate.
+
+Without the identity assumed, a cycle c forms a parallel pair with itself repeated twice- the pair (c, c;c).
+This verification needs to be performed for all elementary cycles.
+Further, if there is any "exit" or "entrance" into an elementary cycle- an edge that is not a part of the cycle whose one end is a node that belongs to the cycle- then it must be verified that the transformation represented by the edge is equal to the transformation represented by the edge and cycle composed (the cycle is pre-pended in the case of an edge directed away from the cycle and appended in the case of an edge directed into the cycle).
+Should this verification succeed, by logic similar to the previous case, it must be that
+all paths (including complex cycles) that include the elementary cycle must be equal to version of the paths without the cycle (since a segment of the path with the cycle has already been verified to be equal to a version of that segment without the cycle).
+Once again this is a factorial time, terminating algorithm.
+
+\subsection{Naive algorithm for online addition}
+
+Since we solve the question of online addition, where the original diagram is path independent and a new edge is added, the naive baseline can be improved to verify only parallel pairs that have been affected by the addition of the new edge.
+Pairs that do not involve the new edge must continue to be path independent because they already belonged to the old diagram.
+
+The cycle checking steps can also be improved upon.
+In the case where the identity is assumed, only elementary cycles passing through the new edge need to be verified to be equal to the identity.
+When the identity is not assumed, cycles that include one of the nodes of the new edge must be checked for the new exit or entry point that the new edge becomes (if the cycle includes the source of the new edge then the cycle followed by the new edge must be equal to the new edge alone and if the cycle include the sink then the new edge followed by the cycle must be equal to the new edge).
+Additionally, all cycles that include the new edge would need full verification like in the batch algorithm- the cycle would have to be equal to itself composed with itself, and would have to be verified for all its entry and exit points.
+
+Now, the commutativity of the original diagram could be used to narrow down the set of pairs to verify.
+Pairs where both paths do not involve the new edge would remain equal, but also, pairs where both paths involved the new edge would have to be equal.
+To see why this is true, each path could be thought of as consisting of the composition of three segments, the first segment from the common source to the source of the new edge, the second, the new edge itself, and the third, from the sink of the new edge to the common sink of the parallel pair.
+The first segment of both pairs would have to be equal because they existed in as parallel pairs in the original diagram, and similarly the third segments would also have to be equal.
+The second segments, consisting of the same edge, would also have to be equal.
+Therefore the composition of these three equal components would be equal.
+Note that the new edge could only appear once because cycles have already been dealt with and none of the paths under consideration hit a node twice.
+Thus we are left only to verify the pairs where exactly one of the paths includes the new edge.
+
+a "two-flip tolerant" path search could be used to identify the pairs of paths where exactly one path includes the new edge.
+
+\subsubsection{Two flip tolerant path search}
+In a normal directed graph path, only forward edges, i.e. edges that go outward from the current node, are considered. 
+On the other hand, the \textit{two flip path} goes through up to three phases: in the first phase, only backward edges- pointing inward to the current node- are accepted.
+In the second phase, only forward edges are accepted, and in the third phase, again only backward edges are accepted.
+As a consequence, there is a node in the path for which both path edges are oriented outwards, that lies between the first and second phase.
+This will be referred to as the \textit{first flipping point}.
+Similarly, there is a \textit{second flipping point} between the second and third phase, for which both path edges point inwards.
+
+Consider the situation where a new edge from source node S to sink T is being added to a diagram.
+A two flip tolerant path from S to T would correspond to a new pair of parallel paths- flipping point 1 would be the source of both paths, and flipping point 2, the sink.
+The first path would correspond to the second phase of the flip tolerant path search, and the second path, involving the new edge, would correspond to composing the first phase of the two flip path, the new edge, and the third phase.
+
+To present this idea diagrammatically, we represent path phases with arrows (note that these are the composition of many edges, not a single edge). The new edge is represented with a dashed arrow.
+
+\begin{center}
+\begin{tikzpicture}[scale=0.2]
+\tikzstyle{every node}+=[inner sep=0pt]
+\draw [black] (18.1,-36.1) circle (3);
+\draw (18.1,-36.1) node {$S$};
+\draw [black] (50.2,-36.1) circle (3);
+\draw (50.2,-36.1) node {$T$};
+\draw [black] (26,-22.9) circle (3);
+\draw (26,-22.9) node {$F_1$};
+\draw [black] (41.3,-22.9) circle (3);
+\draw (41.3,-22.9) node {$F_2$};
+\draw [dashed] (21.1,-36.1) -- (47.2,-36.1);
+\fill [black] (47.2,-36.1) -- (46.4,-35.6) -- (46.4,-36.6);
+\draw (34.15,-35.6) node [above] {new edge};
+\draw [black] (24.46,-25.47) -- (19.64,-33.53);
+\fill [black] (19.64,-33.53) -- (20.48,-33.1) -- (19.62,-32.58);
+\draw [black] (29,-22.9) -- (38.3,-22.9);
+\fill [black] (38.3,-22.9) -- (37.5,-22.4) -- (37.5,-23.4);
+\draw [black] (48.52,-33.61) -- (42.98,-25.39);
+\fill [black] (42.98,-25.39) -- (43.01,-26.33) -- (43.84,-25.77);
+\end{tikzpicture}
+\end{center}
+Here, (S, $F_1$, $F_2$, T) is a two flip tolerant path. ($F_1$, S, T, $F_2$) is a new path, created because of the addition of (S,T), that conflicts with ($F_1$, $F_2$).
+
+The two flip tolerant path search returns the set of all paths between a given source and sink that have up to two flips (paths that omit one or more of the three phases are also accepted).
+
+The \textit{path extraction algorithm} then transforms the output of the two flip path search to the set of new pairs of parallel paths that arise in the diagram due to its addition, that need to be verified.
+
+\paragraph{Path extraction algorithm}
+The purpose of this algorithm is to derive the conflicting pairs that a given two-flip tolerant path corresponds to.
+
+Let the new edge added to the diagram be (S, T).
+It has already been shown how each path corresponds to a conflicting pair in the case where a path has two flips.
+When a path has only the first flip (which is to say, the third phase of the path is missing), but not the second flip, sink node T can be treated as the sink of the two conflicting path, while the first flipping point remains the source.
+Similarly when only the second flipping point is present then it is the sink, and S is the source.
+Finally when no flipping points are present, there are two possibilities.
+Either the path is oriented from S to T, in which case the conflicting paths are simply the edge (S, T) and the entire flip-less path, or the path is oriented from T to S.
+In this case, we have found a cycle that has already been dealt with.
+
+\paragraph{Theorem} The result of the two-flip tolerant path search from the source to the sink node of an edge that is to be newly added, followed with the path extraction algorithm has a one-to-one correspondence to the set of new parallel pairs with exactly one path passing through the new edge, with neither paths containing any cycles.
+
+\paragraph{Proof}
+
+Every element in the output of the path extraction algorithm was by construction a conflicting pair.
+
+It remains to show that every new conflicting pair corresponds to a two flip tolerant path. Let the common source be $F_1$ and common sink be $F_2$.
+
+Only one path passes through (S, T), which we will name path 1.
+We construct a two flip tolerant path from S to T: phase 1 is the segment of path 1 from $F_1$ to S, phase 2 is path 2, and phase 3 is the segment of path 1 from T to $F_2$.
+It is possible that some of $F_1, F_2, S$ and $T$ coincide, in which case the corresponding segments between them disappear, and the resultant path has fewer than two flips.
+
+We have shown that all paths that need verification are caught by the two flip tolerant search followed with path extraction.
+
+\paragraph{Implementation}
+*Copy in python code from implementation*
+
+This algorithm is also in the worst case as bad as factorial in the number of nodes.
+However, in the average case, it significantly outperforms the naive batch baseline.
+For sparse diagrams it is even competitive with our ultimate polynomial solution.
+Average case heuristic analysis will be presented in the evaluation section.
+
+\subsection{Optimal solution for batch problem}
+Murota's main result \cite{commutative} is a solution to the question of whether a given acyclic diagram commutes that returns the minimal number of equality checks.
+The paper describes the following algorithm to find the ($V^2E$ bounded) minimal set of pairs that needs to be checked.
+
+A diagram's structure can result in some redundancy, in the sense that verifying a subset of pairs can imply the verification of path pairs that aren't in the subset.
+For example, in the figure below, verifying $p_1$ and $p_2$ implies that the paths in $p_3$ are equal too.
+
+\begin{center}
+\begin{tikzpicture}[scale=0.2]
+\tikzstyle{every node}+=[inner sep=0pt]
+\draw [black] (12,-15.8) circle (3);
+\draw [black] (30.5,-30.5) circle (3);
+\draw [black] (46.4,-15.8) circle (3);
+\draw [black] (29.9,-15.8) circle (3);
+\draw [black] (15,-15.8) -- (26.9,-15.8);
+\fill [black] (26.9,-15.8) -- (26.1,-15.3) -- (26.1,-16.3);
+\draw (20.95,-16.3) node [below] {$f_1$};
+\draw [black] (14.35,-17.67) -- (28.15,-28.63);
+\fill [black] (28.15,-28.63) -- (27.84,-27.74) -- (27.21,-28.53);
+\draw (19.94,-23.64) node [below] {$g_1$};
+\draw [black] (29.063,-27.873) arc (-157.62406:-197.70132:13.638);
+\fill [black] (29.06,-27.87) -- (29.22,-26.94) -- (28.3,-27.32);
+\draw (27.49,-23.25) node [left] {$f_2$};
+\draw [black] (31.53,-18.309) arc (25.84176:-21.16715:11.992);
+\fill [black] (31.53,-18.31) -- (31.43,-19.25) -- (32.33,-18.81);
+\draw (33.28,-23.03) node [right] {$j_1$};
+\draw [black] (32.9,-15.8) -- (43.4,-15.8);
+\fill [black] (43.4,-15.8) -- (42.6,-15.3) -- (42.6,-16.3);
+\draw (38.15,-16.3) node [below] {$j_2$};
+\draw [black] (32.7,-28.46) -- (44.2,-17.84);
+\fill [black] (44.2,-17.84) -- (43.27,-18.01) -- (43.95,-18.75);
+\draw (39.77,-23.64) node [below] {$h_1$};
+\end{tikzpicture}
+
+$p_1$=($f_1;f_2$, $g_1$)\\
+$p_2$=($j_1;j_2$, $h_1$)\\
+$p_3$=($f_1;h_1$, $f_1;j_2$)
+\end{center}
+
+The approach in the algorithm, at a high level, is to define a function that takes in a subset of pairs and returns the subset of pairs whose verification is implied by the verification of the input set.
+Then, greedy elimination of redundancies is performed until a minimal set is reached.
 
 A bilinking is a parallel pair that is disjoint but for their terminal nodes. The set of all bilinkings is $R_0$.
+In an acyclic diagram, if all bilinkings are equal, all parallel pairs must also be equal since they any given pair can be expressed as a composition of one or more bilinkings.
 
-The main result of \cite{commutative} can be used to verify a graph commutes in the acyclic case.
-The paper describes the following algorithm to find the ($V^2E$ bounded) minimal set of pairs that needs to be checked.
+Define $r_1>r_2$ for bilinkings $r_1$ = $\{p_l,q_l\}$, $r_2$= $\{p_2, q_2)$ $\in$ $R_0$, if there exists a path p such that $\partial p=\partial r_1$ and p contains $p_2$.
+The $\langle\rangle$ function is defined as:
+$\langle r \rangle = \{ s\in R_0| r>s\}$.
 
-% Define $r_1>r_2$ for bilinkings $r_1$ = $\{p_l,q_l\}$, $r_2$= $\{p_2, q_2)$ $\in$ $R_0$, if there exists a path p such that $\partial p=\partial r_1$ and p contains $p_2$.
-% The $\langle\rangle$ function is defined as:
-% $\langle r \rangle = \{ s\in R_0| r>s\}$.
-
-% For bilinking s, let F(s) be the vector in $\mathbb{F}2^{|E|}$ representing the edges present in s (there is a dimension corresponding to each edge, the $n^{th}$ dimension of F(s) is 1 if the corresponding edge is in s, and 0 otherwise).
+For bilinking s, let F(s) be the vector in $\mathbb{F}2^{|E|}$ representing the edges present in s (there is a dimension corresponding to each edge, the $n^{th}$ dimension of F(s) is 1 if the corresponding edge is in s, and 0 otherwise).
 Let this function be extended to sets, so that for some set of bilinkings S, F(S) = $\{ F(s) | s\in S \}$.
 
-% for a set of bilinkings r, the cl function is defined as:
+for a set of bilinkings r, the cl function is defined as:
 cl(r) = $\{ s\in R_0|$ s is linearly dependent on $F(r) \}$.
 
-% Using these two function we finally define the function $\sigma$ on a set of bilinkings r as
-% $\sigma(r) = \{s \in R_0 | s\in cl(R\cap \langle s \rangle) \}$
+The closure function on r basically captures all the pairs that can be made by made by composing or "gluing together" the bilinkings in r.
+The $\langle\rangle$ function captures the orientation of paths, and can be used to constrain compositions to be in the right direction.
+
+Using these two function we finally define the function $\sigma$ on a set of bilinkings r as
+$\sigma(r) = \{s \in R_0 | s\in cl(R\cap \langle s \rangle) \}$
+This is the function used to capture all the pairs whose verification is implied by the verification of pairs in r.
+It is used to iteratively check if a given pair is redundant. bilinkings are eliminated until a minimum "spanning" subset is reached.
 
@@
 
-The proof of why this works is complex and the subject of \cite{commutative}, 
-% Actually I may be wrong about this
+The proof of correctness can be found in Murota\cite{commutative}, 
+The number of checks returned by the algorithm is at worst $O(|V|^2|E|)$. The overall run time of an optimized implementation is $O(|V|^4|E|^2)$.
+
 % but at a very high level the cl function captures the redundancy one gets from how "gluing together" bilinkings.
@@
 % $\sigma$ weaves these together to identify, given an input subset, the set of bilinkings that would be commutative should the input subset be verified.
-But at a high level, $\sigma$ identifies reduncancies arising from the structure of the graph and then the approach is to greedily eliminate redundancies.
-\cite{commutative} solves the problem of verifying whether a given acyclic diagram is commutative.
-
-In the next section, the distinct problem of verification that a diagram (that may contain cycles) commutes on addition of a new edge is solved.
-The polynomial time algorithm is found to entail at worst O($|V|^2$) checks and also can be extended to identify the minimum possible set of checks.
-If the algorithm is run over the course of construction of the graph, it has the same worst case asymptotic runtime O($|V|^2|E|)$ as the one already described.
 
@@
 The solution presented here assumes that there is an implicit identity function on every node, as would be the case in most applications.
-However, a modified version of the algorithm for graphs without this assumption is presented in the appendix.
+However, a modified version of the algorithm for diagrams without this assumption is presented in the appendix, with modification that is similar in spirit to what was described in the description of the naive baselines.
 
-The first step is to identify all pairs that involve the new edge.
-Pairs that do not involve the new edge must already commute by assumption.
-Further, for a given (source, sink) pair only a single parallel pair needs to be verified. This is an implication of Theorem \ref{reductionRule}.
-Theorem \ref{verifyingSet} shows that should our selected set of pairs along with cycles passing through the new edge be verified commutative, the entire graph must commute.
-%TODO prove this equivalent?
-% \begin{theorem}
-% Verifying that all parallel pairs in the graph commute implies that the entire graph must commute.
-% \end{theorem}
+% Like with the online baseline, the first step is to identify all pairs that involve the new edge.
+% Pairs that do not involve the new edge must already commute by assumption.
+Like with the online baseline, we do not concern ourselves with parallel pairs where neither or both paths pass through the new edge.
+The key observation that allows us to improve on the online baseline is that for a given (source, sink) pair only a single parallel pair needs to be verified.
+This is an implication of Theorem \ref{reductionRule}, expanded on later.
+Theorem \ref{verifyingSet} shows that should our selected set of pairs along with cycles passing through the new edge be verified commutative, the entire diagram must commute.
+The approach is to identify a parallel pair with exactly one path through the edge for each (source, sink) pair.
 
@@
 
-The try block is executed at most n(n-1)/2 = $O(n^2)$ times where n is the number of nodes.
+The try block is executed at most $O(|V|^2)$ times.
 
@@
 T has a forward edge to every node in group 2.
-In this graph, on the addition of edge (S, T), $N^2$ paths need to be verified which is polynomial in 2N+2.
+In this diagram, on the addition of edge (S, T), $N^2$ paths need to be verified which is polynomial in 2N+2.
 
@@
 
+The overall run time of the algorithm, assuming that path finding is done using depth first search, is $O((|V|+|E|)|V|^2)$.
+
 \subsection{(Optional) Optimization Step}
 
-In the case where equality checks are very expensive, we begin by finding the minimal set of (source, sink) pairs such that checking for these pairs logically implies having checked the full graph.
+In the case where equality checks are very expensive, we begin by finding the minimal set of (source, sink) pairs such that checking for these pairs logically implies having checked the full diagram.
 
-We observe that there are some redundancies in the graph.
+We observe that there are some redundancies in the diagram.
 
@@
 \end{theorem} 
-
 % monomorphisms
-
 \begin{proof}
@@
 We construct a graph with a node for each possible (source, sink) pair in the graph- each node then represents a possible choice for parallel pair endpoint pairs, and greedily search for the smallest set of nodes from which the entire graph would be reachable.
+The idea is to look for "roots" in the graph that have to be included in the ultimate verification set because they have no predecessor an cannot be verified "through" the verification of some other pair.
+Then all the successors whose verification is implied by the roots are eliminated.
 
@@
 The result of the algorithm cannot be further reduced.
-Also, the verification of the parallel pairs returned in the algorithm implies that the output of the previous algorithm must commute, and transitively, that the entire graph must commute.
-% Here is a naive implemenetation.
-Note that an important optimization in sparse graphs can be made: Source-sink pairs with no viable adjacent pair should never be added to the graph (as verified by the try-catch block from the first algorithm).
+Also, the verification of the parallel pairs returned in the algorithm implies that the output of the previous algorithm must commute, and transitively with Theorem \ref{verifyingSet} that the entire diagram must commute.
+
+The run time of the first step is $O(|V|^4)$, and that of the second step is $O(|V|)$, so that the overall bound is $O(|V|^4)$, which is still a gain of a factor of $|E|^2$ over the batch algorithm.
 
@@
 \section{Appendix: Theorems and additional cases}
-This should be read as a continuation of the notation section, the definitions there will be used.
 
@@
 \begin{lemma}
+\label{cycles_lemma}
 A path that involves the new edge more than once is equal to a path without multiple occurrences of the edge, under the assumption that the pairs in C are verified.
@@
 
-Case A: pairs that do not involve the new edge: By assumption that the original graph commutes, these pairs must commute.
+Case A: pairs that do not involve the new edge: By assumption that the original diagram commutes, these pairs must commute.
 
 Case B: Cases where there is a different pair corresponding to this choice of (source, sink) in V already.
-There are four paths to be considered, two from each pair. Those paths which do not include the new edge must already be equal by the assumption that the original graph commutes.
+There are four paths to be considered, two from each pair. Those paths which do not include the new edge must already be equal by the assumption that the original diagram commutes.
 Those paths which do include the new edge (S,T) can be reduced, by the case for multiple occurrences of a new edge, to a path with a single occurrence of the new edge as described in the preceding lemma.
 These new paths can be divided into three segments: the segment from the source to S, the new edge, and the segment from the last occurrence of T to the sink.
-The first segments of all paths are equal by assumption that the original graph commutes.
+The first segments of all paths are equal by assumption that the original diagram commutes.
 This is also true of the third segment of all the paths.
-The second segment consists of only the new edge and is the same for all graphs.
+The second segment consists of only the new edge and is the same for all paths.
 The composition of equal functions is equal, so all the paths passing through the new edge must be equal.
@@
 
-Modified algorithm for graphs without the identity:
+Modified algorithm for diagrams without the identity:
 

### 2020-04-22 main.tex

@@
 \end{center}
+
 Let Path$_1=f_1\circ g_1\circ h_1$ and Path$_2$=$f_2\circ g_2\circ h_2$. We choose Path$_2$.
@@
 
+\section{Ordering}
+We can try answering a similar question in partial function territory, where instead of asking whether adding an edge will introduce comparable paths that are not equal, we ask if adding a new edge introduces comparable paths that disagree, i.e.
+there is some element in the domain for which both functions are defined and give equal results?
+Further we can also ask if the addition of a new edge introduces a path that is less informative than all the other paths in the graph.
+
+TODO add the whole two flip tolerant schpeal.
+
+TOTHINK figure out if there's a way to prove two flip is asymptotically optimal.
+
+\section{Future Work}
+Cant think of any theory questions we haven't tried to answer.
+Think of how to apply in other areas like reflexive programming, formal verification.
+
 \end{document}

### 2020-04-22 new_structure.tex

--- 2020-04-10 April_edition.tex
@@
 
-\title{Online verification of Commutativity}
+\title{Online verification of commutativity}
 
 \author{ }
-\date{April 2020}
+\date{March 2020}
 
@@
 \section{Introduction}
+Consider the graph representing a category where every path between a common source and sink is equal. 
+% Morphisms semantically represent natural transformations between objects.
+Such \textit{path independent graphs} arise frequently in practice- from the implicit type conversion graph for a program in a language like scala, to the category of units of measure where morphisms are conversions between the units.
+Fundamentally, graphs representing systems with sets and "natural" transformations between the sets have this property by construction.
+It is often useful to be able to verify that a category is indeed stays path independent over the course of construction.
+We present an efficient way to perform such a check, and evaluate our idea through two case studies.
 
-Many systems can be understood as a category, with a collection of sets, and some transformation functions between the sets- for example, in programming languages, type systems with coercion.
-If a graph is constructed with the sets as nodes and functions as edges between the nodes, then by construction the new graph is commutative.
-The idea is that each object has a unique representation in a given set in the system, so the transformation mapping between any two nodes should not depend on choice of path used to obtain the transformation.
-
-Since commutative graphs arise in many systems, it is often useful to be able to perform verification and ensure the graphs in question are indeed path independent. 
-% As a case study, gator, a domain specific graphics language that allows automatic conversion between geometric spaces internally creates a graph with he transformations between the spaces. 
-Naively checking if all paths in a given graph are path independent could require a number of function equality checks as bad as factorial in the number of nodes, since a path consists of an ordering of nodes.
-We present a polynomial time algorithm to verify that commutativity of a graph is maintained over the course of online addition.
-It uses the minimal possible number of equality checks, since these checks can be expensive.
-To evaluate our results we use two case studies.
 First, in the domain specific geometry type language \textit{Gator}, our algorithm is used to ensure that user defined transformations between spaces stay consistent, and the compiler, deterministic as it makes automatic conversion choices.
@@
 
-\section{Problem setup and Terminology}
-The closely related problem of identifying the minimal subset of paths to check in order to verify whether a given acyclic graph is commutative, is solved in \cite{commutative}. 
-I reproduce relevant terminology and propositions from the paper here for convenience.
+\section{Algorithm}
+The objective of the algorithm, given a path independent graph and a new edge to add, is to return that either the new graph after the addition of the edge is path independent, or that addition of the edge results in a loss of path independence along with an example pair of paths that have a common source and sink (\textit{comparable paths}).
+It assumes the existence of a path equality checking oracle.
 
-First, we formalize the notion of a graph that represents transformations.
+For simplicity we will also assume that the graph has an implicit identity morphism on every node. However, with a few modifications, this algorithm can be extended to graphs without the identity.
 
-We start with a directed graph G=(V,E) with spaces attached to vertices and maps to arcs. Each edge (u, v) in E corresponds to a function that maps elements of u to elements in v. 
-This correspondence is stored as a mapping f:$E\rightarrow F$, where F is the semi-group that contains all the functions in the graph, with associative multiplication denoted by $\circ$.
-f maps each edge to the function it represents.
+\subsection{Comparable Path pair finding}
 
-$P_{all}$ is the set of all paths in G. Since there may be cycles, it may be infinite.
+A graph is path independent if there exists no pair of comparable paths that are not equal.
 
-A path is a sequence of edges. The edge-to-function mapping f can be naturally extended to paths. If path p=$e_1\circ e_2 ... e_n$ then $f(p)=f(e_1) \circ f(e_2) ... f(e_n)$.
+Thus we begin by identifying all possible common sources and sinks that have a pair of paths, noting that equality of all pairs of paths such that one path does not include the new edge and the other does, along with the assumption that the existing graph is path independent, implies that all pairs comparable paths between the source and sink in the new, augmented graph are equal.
 
-The notation $\partial^{+}(p)$ is used to indicate that start node pf path p, and similarly $\partial^-(p)$ is the end node of path p. $\partial(p)$ maps to the tuple (start node of p, end node of p).
+\begin{lemma}
+Equality of all pairs of paths such that one path does not include the new edge and the other does, and of cycles that include the new edge exactly once, along with the assumption that the existing graph is path independent, implies that the new, augmented graph is path independent.
+\end{lemma}
+\begin{proof}
+A graph is path independent if there exists no comparable pair of paths that is not equal.
 
-A pair of paths p and q is called \textit{parallel} iff their terminal nodes are the same, i.e. $\partial(p)=\partial(q)$.
+For a comparable pair where neither path contains the new edge, the pair was also present in the old, path independent graph. Since it was path independent, the two paths must be equal.
 
-$R_{all}$ is the set of all parallel paths in the graph.
+Upon verification that all cycles are equal to the identity, the path containing the new edge more than once is known to be equal to a path with the same start and end nodes that
 
-The graph commutes iff f(p)=f(q) $\forall (p,q)\in R_{all}$, which is to say, the compositions of the maps along any path connecting u to v is independent of choice of path.
-
-A bilinking is a parallel pair that is disjoint but for their terminal nodes. The set of all bilinkings is $R_0$.
-
-The main result of \cite{commutative} can be used to verify a graph commutes in the acyclic case.
-The paper describes the following algorithm to find the ($V^2E$ bounded) minimal set of pairs that needs to be checked.
-
-Define $r_1>r_2$ for bilinkings $r_1$ = $\{p_l,q_l\}$, $r_2$= $\{p_2, q_2)$ $\in$ $R_0$, if there exists a path p such that $\partial p=\partial r_1$ and p contains $p_2$.
-The $\langle\rangle$ function is defined as:
-$\langle r \rangle = \{ s\in R_0| r>s\}$.
-
-For bilinking s, let F(s) be the vector in $\mathbb{F_2}^{|E|}$ representing the edges present in s (there is a dimension corresponding to each edge, the $n^{th}$ dimension of F(s) is 1 if the corresponding edge is in s, and 0 otherwise).
-Let this function be extended to sets, so that for some set of bilinkings S, F(S) = $\{ F(s) | s\in S \}$.
-
-for a set of bilinkings r, the cl function is defined as:
-cl(r) = $\{ s\in R_0|$ s is linearly dependent on $F(r) \}$.
-
-Using these two function we finally define the function $\sigma$ on a set of bilinkings r as
-$\sigma(r) = \{s \in R_0 | s\in cl(R\cap \langle s \rangle) \}$
-
-Finally, an implementation of the described algorithm is
-
-\begin{verbatim}
-Find a spanning set Rs, = [r_1, ... ,r_k].
-
-R := Rs,
-    for i:=l to K do
-        if r_i in sigma(R - r_i) then R:=R - r_i;
-return R.
-
-sigma(inputSet S, codomain = Rs):
-    output = {}
-    for bilinking in Rs:
-        smallerPairs = {}
-        for candidateBilinking in Rs:
-            if exists path(source: bilinking.source, sink:candidateBilinking.source) and
-            exists path(source: candidateBilinking.sink, sink: bilinking.sink):
-                smallerPairs.add(candidateBilinking)
-        # Check if there's a linear dependence between the bilinking and smallerPairs
-        # vectorize(pair) returns a vector with a dimension for each edge in the graph,
-        # with a 1 if pair includes that edge and 0 if it doesn't.
-        matrix = [vectorize(pair) for pair in smallerPair]
-        matrix = matrix+[vectorize(bilinking)]
-        # if there is a linear dependence
-        if determinant(matrix)%2==0:
-            output.add(bilinking)
-    return output
-
-\end{verbatim}
-
-and to initially set Rs,
-
-\begin{verbatim}
-Graph existingGraph;
-R_s = {}
-for each node v in V:
-    # Let the subsection of the graph reachable from v be S.
-    S = existingGraph.extractReachableSection(v)
-    # Create a minimum spanning tree of S
-    T = createMinimumSpanningTree(S)
-    # Find the excluded edges
-    excludedEdges = S.edges - T.edges
-    # Use these edges to create bilinkings
-    for each edge e in excludedEdges:
-        firstPath = T.findPath(source: e.source, sink: e.sink)
-        R_s.addElement(new Bilinking(firstPath, e)
-return R_s
-\end{verbatim}
-
-The proof of why this works is complex and the subject of \cite{commutative}, 
-% Actually I may be wrong about this
-% but at a very high level the cl function captures the redundancy one gets from how "gluing together" bilinkings.
-For example in the graph below verifying any two of the bilinkings commute implies that the third must too.
-So cl(any two)=all three bilinkings.
-
-% \begin{center}
-% \begin{tikzpicture}[scale=0.2]
-% \tikzstyle{every node}+=[inner sep=0pt]
-% \draw [black] (15.1,-25.8) circle (3);
-% \draw (15.1,-25.8) node {$A$};
-% \draw [black] (31.6,-34.7) circle (3);
-% \draw (31.6,-34.7) node {$C$};
-% \draw [black] (31.6,-18.9) circle (3);
-% \draw (31.6,-18.9) node {$B$};
-% \draw [black] (47.7,-26.5) circle (3);
-% \draw (47.7,-26.5) node {$D$};
-% \draw [black] (17.87,-24.64) -- (28.83,-20.06);
-% \fill [black] (28.83,-20.06) -- (27.9,-19.9) -- (28.29,-20.83);
-% \draw [black] (31.6,-21.9) -- (31.6,-31.7);
-% \fill [black] (31.6,-31.7) -- (32.1,-30.9) -- (31.1,-30.9);
-% \draw [black] (17.74,-27.22) -- (28.96,-33.28);
-% \fill [black] (28.96,-33.28) -- (28.49,-32.46) -- (28.02,-33.34);
-% \draw [black] (34.27,-33.34) -- (45.03,-27.86);
-% \fill [black] (45.03,-27.86) -- (44.09,-27.78) -- (44.54,-28.67);
-% \draw [black] (34.31,-20.18) -- (44.99,-25.22);
-% \fill [black] (44.99,-25.22) -- (44.48,-24.43) -- (44.05,-25.33);
-% \end{tikzpicture}
-% \end{center}
-
-% The $\langle \rangle$ function captures structures like this one, which display the redundancy above. Here, for p=($p_1, p_2), q=(q_1, q_2)$, we have that $p>q$.
-
-% \begin{center}
-% \begin{tikzpicture}[scale=0.2]
-% \tikzstyle{every node}+=[inner sep=0pt]
-% \draw [black] (20.7,-15.7) circle (3);
-% \draw (20.7,-15.7) node {$A$};
-% \draw [black] (44,-15.7) circle (3);
-% \draw (44,-15.7) node {$B$};
-% \draw [black] (20.7,-33) circle (3);
-% \draw (20.7,-33) node {$A'$};
-% \draw [black] (44,-33) circle (3);
-% \draw (44,-33) node {$B'$};
-% \draw [black] (22.498,-13.306) arc (136.73613:43.26387:13.529);
-% \fill [black] (42.2,-13.31) -- (42.02,-12.38) -- (41.29,-13.07);
-% \draw (32.35,-8.55) node [above] {$p_1$};
-% \draw [black] (41.785,-17.717) arc (-53.13545:-126.86455:15.728);
-% \fill [black] (41.79,-17.72) -- (40.85,-17.8) -- (41.45,-18.6);
-% \draw (32.35,-21.36) node [below] {$p_2$};
-% \draw [black] (23.143,-31.264) arc (120.64183:59.35817:18.065);
-% \fill [black] (41.56,-31.26) -- (41.12,-30.43) -- (40.61,-31.29);
-% \draw (32.35,-28.24) node [above] {$q_1$};
-% \draw [black] (41.52,-34.683) arc (-60.46289:-119.53711:18.601);
-% \fill [black] (41.52,-34.68) -- (40.58,-34.64) -- (41.07,-35.51);
-% \draw (32.35,-37.6) node [below] {$q_2$};
-% \draw [black] (20.7,-18.7) -- (20.7,-30);
-% \fill [black] (20.7,-30) -- (21.2,-29.2) -- (20.2,-29.2);
-% \draw (20.2,-24.35) node [left] {$\alpha$};
-% \draw [black] (44,-30) -- (44,-18.7);
-% \fill [black] (44,-18.7) -- (43.5,-19.5) -- (44.5,-19.5);
-% \draw (44.5,-24.35) node [right] {$\beta$};
-% \end{tikzpicture}
-% \end{center}
-% $\sigma$ weaves these together to identify, given an input subset, the set of bilinkings that would be commutative should the input subset be verified. 
-$\sigma$ identifies reduncancies arising from the structure of the graph and then the approach is to greedily eliminate redundancies.
-\cite{commutative} solves the problem of verifying whether a given acyclic diagram is commutative.
-
-In the next section, the distinct problem of verification that a diagram (that may contain cycles) commutes on addition of a new edge is solved.
-The polynomial time algorithm is found to entail at worst O($|V|^2$) checks and also can be extended to identify the minimum possible set of checks.
-If the algorithm is run over the course of construction of the graph, it has the same worst case asymptotic runtime O($|V|^2|E|)$ as the one already described.
-
-\section{Solving the online addition problem}
-
-The first step is to identify all bilinkings that involve the new edge.
-Bilinkings that do not involve the new edge must already commute by assumption.
-Bilinkings where both paths involve the new edge cannot exist since bilinkings are node-disjoint.
-Further, for a given (source, sink) pair only a single bilinking needs to be verified. This is an implication of Theorem \ref{reductionRule}.
-This fact, along with the equivalent proposition 1.1 from \cite{commutative} for graphs with cycles means that should our selected set of bilinkings along with cycles passing through the new edge be verified commutative, the entire graph must commute.
-%TODO prove this equivalent?
-% \begin{theorem}
-% Verifying that all bilinkings in the graph commute implies that the entire graph must commute.
-% \end{theorem}
+For a comparable pair where both paths contain 
+\end{proof}
 
@@
 
-Set bilinkings = new Set();
+Set comparablePaths = new Set();
 for S in existingGraph.Nodes():
@@
                     sinkNode: T);
-            bilinkings.add((pathInOldGraph, pathWithNewEdge));
+            comparablePaths.add((pathInOldGraph, pathWithNewEdge));
         catch (pathFindingFailed Exception):
             print(
-                $"No comparable pairs from node {S} to node {T} 
-                that need to be checked");
+                $"No comparable pairs from node {S} to node {T} that need to be checked");
 
-output bilinkings;
+output comparablePaths;
 \end{verbatim}
@@
 
-The bound is asymptotically tight.
+This bound is asymptotically tight.
 This can be seen in the case where the graph contains 2N nodes besides S and T.
@@
 
-We observe that there are some redundancies in the graph.
+However, we observe that there are some redundancies.
 
@@
 \end{center}
+TODO backward compositions.
 \paragraph{Theorem} 
-\label{reductionRule}
-If conflicting paths $g_2 = f_2; (S,T); h_2$ then it must be that $g_1 = f_1; (S,T); h_1$.
+If conflicting paths $g_2 = f_2 \circ (S,T) \circ h_2$ then it must be that $g_1 = f_1 \circ (S,T) \circ h_1$.
 
@@
 \paragraph{Proof}
-We use the fact that $f_1$=$l; f_2$ and $h_1$=$h_2; m$.
-\[g_2 = f_2; (S,T); h_2 \Rightarrow l; g_2 = l; f_2; (S,T) ; h_2 \]
-\[\Rightarrow l ; g_2 ; m = l ; f_2 ; (S,T) ; h_2 ; m \Rightarrow g_1 = f_1 ; (S,T) ; h_1\]
+We use the fact that $f_1$=$l\circ f_2$ and $h_1$=$h_2 \circ m$.
+\[g_2 = f_2 \circ (S,T) \circ h_2 \Rightarrow l \circ g_2 = l \circ f_2 \circ (S,T) \circ h_2 \Rightarrow l \circ g_2 \circ m = l \circ f_2 \circ (S,T) \circ h_2 \circ m \Rightarrow g_1 = f_1 \circ (S,T) \circ h_1\]
 
@@
 
-Under the assumption that the only path operations allowed are composition and replacement of one path by a different, equal path, as would be true when edges are generic functions and no other information is available, this reduction rule is also the only reduction rule.
+Under the assumption that the only path operations allowed are composition and replacement of one path by a different, equal path, this reduction rule is also the only reduction rule.
+That is to say, if verifying a comparable pair of paths with end points ($P_1$, $P_2$) implies the verification of a pair with endpoints ($Q_1$, $Q_2$), then it must be that $Q_1$ is a successor of $P_1$ and $P_2$ is a successor of $Q_2$.
 
-That is to say, if verifying a comparable pair of paths with end points ($P_1$, $P_2$) implies the verification of a pair with endpoints ($Q_1$, $Q_2$), then it must be that $Q_1$ is a successor of $P_1$ and $P_2$ is a successor of $Q_2$.
-%TODO proof required?
+TODO add proof. Maybe reference lemma below *after* introducing path semantics.
 
-Using this information it is possible to choose a minimal subset of path pairs to verify.
+Using this information it is possible to choose a minimal subset of path pairs to verify, using a set-cover oracle.
 
-We construct a graph with a node for each possible (source, sink) pair in the graph- each node then represents a possible choice for bilinking endpoint pairs.
+We construct a graph with a node for each possible (source, sink) pair in the graph- each node then represents a possible choice for comparable path pair endpoints.
+Then we assign sets for every choice of inclusion of a (source, sink) comparable pair- the members of a set corresponding to a pair P consists of those nodes corresponding to those pairs whose verification would be implied by the verification of P.
+All nodes that do not have conflicting pairs such that one path includes the new edge and the other doesn't are removed (i.e.
+all pairs corresponding to iterations of the comparable path finding algorithm where the catch block was entered).
+
+Set cover gives us the smallest selection of sets that would allow for verification of every necessary node pair in the graph.
 
@@
 
-The result of the algorithm cannot be further reduced.
-Further, the verification of the bilinkings returned in the algorithm implies that the output of the previous algorithm must commute, and transitively, that the graph must commute.
 % Here is a naive implemenetation.
@@
 \begin{verbatim}
-for (path_1, path_2) in bilinkings:
+for (path_1, path_2) in comparablePaths:
     if (path_1 != path_2):
@@
 
-\begin{thebibliography}{9}
-\bibitem{commutative} 
-Kazuo Murota. 
-\textit{HOMOTOPY BASE OF ACYCLIC GRAPHS - A COMBINATORIAL ANALYSIS OF COMMUTATIVE BY MEANS OF PREORDERED MATROID}. 
-Discrete Applied Mathematics, 1987.
+\section{Proofs of Correctness}
+This section might be easier if I write up a reduction semantics for paths, which I need for the proof of minimality anyway.
+Not sure where the best place to introduce these semantics would be- maybe here?
+TODO.
 
-\end{thebibliography}
+\subsection{Verifying every path returned by the Comparable Path Pair finding algorithm implies that the new graph is path independent}
+*TODO fill out*
+
+- No point in doing a pair unless exactly one has new edge / it's a cycle
+
+- Excepting the cases already dismissed, there is no conflicting pair such that its information is not covered by one of the paths in the final set.
+
+\subsection{Verifying every path returned by the Optimization algorithm implies that the new graph is path independent}
+Structure optimization section into lemmas and reference those and the previous paragraph.
+
+\section{Run time and efficiency analysis}
+
+\subsection{Comparable path finding}
+TODO fill in
+
+- Polynomial
+- Asymptotically optimal number of paths in worst case, since if addition of a path is a single step then at least $n^2$ steps required in worst case so O($n^2$) is a theoretical lower bound on run time.
+- Maybe path finding subroutine can be more efficient
+
+\subsection{Minimization problem}
+This problem is NP hard.
+
+TODO Show reduction from set cover.
+
+Therefore we can't do much better than reducing to another NP complete problem.
+
+\section{Case Studies}
+
+\subsection{Gator}
+\subsection{Currencies}
+
 \end{document}
+
+// Finish write up
+// Ordering
+// Currency
+// Semantics

### 2020-04-22 paper/paper/paper.tex
Added in paper content to paper.tex

--- 2020-04-22 April_edition.tex
@@
-\documentclass{article}
-\usepackage[utf8]{inputenc}
-\usepackage{amsthm}
-\usepackage{soul}
+% This is samplepaper.tex, a sample chapter demonstrating the
+% LLNCS macro package for Springer Computer Science proceedings;
+% Version 2.20 of 2017/10/04
+%
+\documentclass[runningheads]{llncs}
+%
 \usepackage{tikz}
 \usepackage{amssymb}
-\usepackage[options ]{algorithm2e}
-\usepackage{listings}
-\usepackage{xparse}
-\usepackage{xcolor}
-
-\newtheorem{lemma}{Lemma}
-\newtheorem{theorem}{Theorem}
-\newtheorem{definition}{Definition}
-
-\title{Online verification of Commutativity}
-
-\author{ }
-\date{April 2020}
+%
+\usepackage{graphicx}
+% Used for displaying a sample figure. If possible, figure files should
+% be included in EPS format.
+%
+% If you use the hyperref package, please uncomment the following line
+% to display URLs in blue roman font according to Springer's eBook style:
+% \renewcommand\UrlFont{\color{blue}\rmfamily}
 
 \begin{document}
+%
+\title{Online verification of Commutativity}
+%
+%\titlerunning{Abbreviated paper title}
+% If the paper title is too long for the running head, you can set
+% an abbreviated paper title here
+%
+\author{First Author\inst{1}\orcidID{0000-1111-2222-3333} \and
+Second Author\inst{1}\orcidID{1111-2222-3333-4444} \and
+Third Author\inst{1}\orcidID{2222--3333-4444-5555}}
+%
+\authorrunning{F. Author et al.}
+% First names are abbreviated in the running head.
+% If there are more than two authors, 'et al.' is used.
+%
+\institute{Cornell University, Ithaca NY 14853, USA
+\email{\{abc,lncs\}@cornell.edu}}
+%
+\maketitle              % typeset the header of the contribution
+%
+\begin{abstract}
+Commutative diagrams arise in many systems, such as an implicit type conversion graph in some programming languages.
+It can be useful in program analysis to verify that a diagram representing conversions remains commutative.
+A naive approach to solving the problem would take factorial time, and previous work provides an algorithm for the special case of acyclic diagrams that runs in $O(|V|^4|E|^2)$ time, where V is the set of vertices in the diagram, and E, the set of edges.
+We present an algorithm that runs in $O(|V|^2|E| + |V|^3)$ time.
+We evaluate the algorithm gainst baselines, and see its practical application in the compiler of a domain specific language for geometry types.
 
-\maketitle
-
-\section{Introduction}
-
+\end{abstract}
+%
+%
+%
 Many systems can be understood as a category, with a collection of sets, and some transformation functions between the sets- for example, in programming languages, type systems with coercion.
@@
 A diagram is said to commute when on selecting any two vertices, any directed path leads to the same result by composition.
-% As a case study, gator, a domain specific graphics language that allows automatic conversion between geometric spaces internally creates a graph with he transformations between the spaces. 
 Naively checking if all paths in a given graph are path independent could require a number of function equality checks as bad as factorial in the number of nodes, since a path consists of an ordering of nodes.
@@
 
-% but at a very high level the cl function captures the redundancy one gets from how "gluing together" bilinkings.
-For example in the graph below verifying any two of the bilinkings commute implies that the third must too.
-So cl(any two)=all three bilinkings.
-
-% \begin{center}
-% \begin{tikzpicture}[scale=0.2]
-% \tikzstyle{every node}+=[inner sep=0pt]
-% \draw [black] (15.1,-25.8) circle (3);
-% \draw (15.1,-25.8) node {$A$};
-% \draw [black] (31.6,-34.7) circle (3);
-% \draw (31.6,-34.7) node {$C$};
-% \draw [black] (31.6,-18.9) circle (3);
-% \draw (31.6,-18.9) node {$B$};
-% \draw [black] (47.7,-26.5) circle (3);
-% \draw (47.7,-26.5) node {$D$};
-% \draw [black] (17.87,-24.64) -- (28.83,-20.06);
-% \fill [black] (28.83,-20.06) -- (27.9,-19.9) -- (28.29,-20.83);
-% \draw [black] (31.6,-21.9) -- (31.6,-31.7);
-% \fill [black] (31.6,-31.7) -- (32.1,-30.9) -- (31.1,-30.9);
-% \draw [black] (17.74,-27.22) -- (28.96,-33.28);
-% \fill [black] (28.96,-33.28) -- (28.49,-32.46) -- (28.02,-33.34);
-% \draw [black] (34.27,-33.34) -- (45.03,-27.86);
-% \fill [black] (45.03,-27.86) -- (44.09,-27.78) -- (44.54,-28.67);
-% \draw [black] (34.31,-20.18) -- (44.99,-25.22);
-% \fill [black] (44.99,-25.22) -- (44.48,-24.43) -- (44.05,-25.33);
-% \end{tikzpicture}
-% \end{center}
-
-% The $\langle \rangle$ function captures structures like this one, which display the redundancy above. Here, for p=($p_1, p_2), q=(q_1, q_2)$, we have that $p>q$.
-
-% \begin{center}
-% \begin{tikzpicture}[scale=0.2]
-% \tikzstyle{every node}+=[inner sep=0pt]
-% \draw [black] (20.7,-15.7) circle (3);
-% \draw (20.7,-15.7) node {$A$};
-% \draw [black] (44,-15.7) circle (3);
-% \draw (44,-15.7) node {$B$};
-% \draw [black] (20.7,-33) circle (3);
-% \draw (20.7,-33) node {$A'$};
-% \draw [black] (44,-33) circle (3);
-% \draw (44,-33) node {$B'$};
-% \draw [black] (22.498,-13.306) arc (136.73613:43.26387:13.529);
-% \fill [black] (42.2,-13.31) -- (42.02,-12.38) -- (41.29,-13.07);
-% \draw (32.35,-8.55) node [above] {$p_1$};
-% \draw [black] (41.785,-17.717) arc (-53.13545:-126.86455:15.728);
-% \fill [black] (41.79,-17.72) -- (40.85,-17.8) -- (41.45,-18.6);
-% \draw (32.35,-21.36) node [below] {$p_2$};
-% \draw [black] (23.143,-31.264) arc (120.64183:59.35817:18.065);
-% \fill [black] (41.56,-31.26) -- (41.12,-30.43) -- (40.61,-31.29);
-% \draw (32.35,-28.24) node [above] {$q_1$};
-% \draw [black] (41.52,-34.683) arc (-60.46289:-119.53711:18.601);
-% \fill [black] (41.52,-34.68) -- (40.58,-34.64) -- (41.07,-35.51);
-% \draw (32.35,-37.6) node [below] {$q_2$};
-% \draw [black] (20.7,-18.7) -- (20.7,-30);
-% \fill [black] (20.7,-30) -- (21.2,-29.2) -- (20.2,-29.2);
-% \draw (20.2,-24.35) node [left] {$\alpha$};
-% \draw [black] (44,-30) -- (44,-18.7);
-% \fill [black] (44,-18.7) -- (43.5,-19.5) -- (44.5,-19.5);
-% \draw (44.5,-24.35) node [right] {$\beta$};
-% \end{tikzpicture}
-% \end{center}
-% $\sigma$ weaves these together to identify, given an input subset, the set of bilinkings that would be commutative should the input subset be verified.
-
 \section{Solving the online addition problem}
@@
 
-% Like with the online baseline, the first step is to identify all pairs that involve the new edge.
-% Pairs that do not involve the new edge must already commute by assumption.
 Like with the online baseline, we do not concern ourselves with parallel pairs where neither or both paths pass through the new edge.
@@
 
-%Define a cycle-pair to mean the pair (c, c;c) where c is a cycle, which is to say the pair whose first element is a cycle and second element is the cycle traversed twice.
-*tempted to name this bi-cycle*.
-
-%Let i be the identity function. Define a cycle-identity pair for cycle c to be (c, i). *should I call this unicycle?*
-
 Consider all the cycles in the graph that pass through the new edge, unique to their terminal point.
@@
 \begin{lemma}
-\label{cycles_lemma}
+\label{one_occurence_lemma}
 A path that involves the new edge more than once is equal to a path without multiple occurrences of the edge, under the assumption that the pairs in C are verified.
@@
 \begin{proof}
-% We do a kind of "structural induction" to show that verification every pair in $R_{all}$ that aren't in V can wither be reduced down to the verification of smaller components or can be shown to be implied by the verification of V.
 
@@
 There are four paths to be considered, two from each pair. Those paths which do not include the new edge must already be equal by the assumption that the original diagram commutes.
-Those paths which do include the new edge (S,T) can be reduced, by the case for multiple occurrences of a new edge, to a path with a single occurrence of the new edge as described in the preceding lemma.
+Those paths which do include the new edge (S,T) can be reduced, by the case for multiple occurrences of a new edge, to a path with a single occurrence of the new edge as described in the preceding lemma \ref{one_occurence_lemma}.
 These new paths can be divided into three segments: the segment from the source to S, the new edge, and the segment from the last occurrence of T to the sink.
@@
 
-\begin{thebibliography}{9}
+%
+% ---- Bibliography ----
+%
+% BibTeX users should specify bibliography style 'splncs04'.
+% References will then be sorted and formatted in the correct style.
+%
+% \bibliographystyle{splncs04}
+% \bibliography{mybibliography}
+%
+\begin{thebibliography}{8}
+
 \bibitem{commutative} 
-Kazuo Murota. 
-\textit{HOMOTOPY BASE OF ACYCLIC GRAPHS - A COMBINATORIAL ANALYSIS OF COMMUTATIVE BY MEANS OF PREORDERED MATROID}. 
-Discrete Applied Mathematics, 1987.
+Kazuo Murota: {HOMOTOPY BASE OF ACYCLIC GRAPHS - A COMBINATORIAL ANALYSIS OF COMMUTATIVE BY MEANS OF PREORDERED MATROID}. Discrete Applied Mathematics, vol. 17, Issues 1--2, 135--155 (1987).
 

### 2020-05-19 currency-exchange/noIdentityOptimalSetPathChecker.py
A couple of fixes

@@ -35,6 +35,6 @@ def testGetPathsToCheck():
     assert(checker.timeTaken > 0)
 
 def runAllTests():
-    print('\033[0m' + "Running optimalSetPathChecker Tests")
+    print('\033[0m' + "Running NoIdentityOptimalSetPathChecker Tests")
     testGetPathsToCheck()
     print(Fore.GREEN + 'Run Completed')   

### 2020-05-19 currency-exchange/noIdentityPolynomialPathChecker.py

@@ -51,7 +51,7 @@ def testGetPathsToCheck():
     assert(checker.timeTaken > 0)
 
 def runAllTests():
-    print('\033[0m' + "Running optimalSetPathChecker Tests")
+    print('\033[0m' + "Running NoIdentityPolynomialPathChecker Tests")
     testGetPathsToCheck()
     print(Fore.GREEN + 'Run Completed')    
 

### 2020-05-19 currency-exchange/nonIdentityUtilities.py

@@ -3,7 +3,7 @@ from typing import List, Tuple
 
 class NonIdentityPathChecker(BaseOnlineChecker):
 
-    def handleCycle(self, terminalNode: int) -> List[Tuple[int, int]]:
+    def handleCycle(self, terminalNode: int) -> List[Tuple[List[int], List[int]]]:
         pathsToCheck = []
         cycle = self.pathsToNewEdgeSource[terminalNode] +\
             self.pathsFromNewEdgeSink[terminalNode] 
@@ -14,32 +14,15 @@ class NonIdentityPathChecker(BaseOnlineChecker):
             pathsToCheck += [(cycle, [cycleEntry] + cycle)]
         return pathsToCheck
     
-    def getTerminalCycles(self, node: int) -> List[List[int]]:
-        """Find all the cycles that include node (but not the new edge)."""
-        # Modified DFS.
-        path = []
-        blocked = set()
-        cycles = []
-        stack = [node]
-        while (len(stack)>0):
-            currentNode = stack.pop()
-            if currentNode == self._invalid_node:
-                blocked.add(path.pop())
-            elif currentNode not in blocked and currentNode not in path:
-                path = path+[currentNode]
-                if currentNode == node and len(path)>1:
-                    cycles += [path]
-                stack += [self._invalid_node] + self.compactFwdGraph[currentNode]
-        return cycles
+    def getSinkCycles(self) -> List[Tuple[List[int], List[int]]]:
+        (cycleExists, cycle) = self.findPath(self.newEdgeSink, self.newEdgeSink)
+        if cycleExists:
+            return [(cycle, [self.newEdgeSource] + cycle)]
+        return []
     
-    def getSinkCycles(self) -> List[Tuple[int, int]]:
-        sinkCycles = self.getTerminalCycles(self.newEdgeSink)
-        sinkCyclePairs = [(cycle, [self.newEdgeSource] + cycle) 
-            for cycle in sinkCycles]
-        return sinkCyclePairs
-    
-    def getSourceCycles(self) -> List[Tuple[int, int]]:
-        sourceCycles = self.getTerminalCycles(self.newEdgeSource)
-        sourceCyclePairs = [(cycle, cycle+[self.newEdgeSink]) 
-            for cycle in sourceCycles]
-        return sourceCyclePairs
+    def getSourceCycles(self) -> List[Tuple[List[int], List[int]]]:
+        (cycleExists, cycle) = self.findPath(
[…]

### 2020-05-19 paper/paper.tex
went over paper

--- 2020-04-22 paper/paper/paper.tex
@@
 Commutative diagrams arise in many systems, such as an implicit type conversion graph in some programming languages.
-It can be useful in program analysis to verify that a diagram representing conversions remains commutative.
-A naive approach to solving the problem would take factorial time, and previous work provides an algorithm for the special case of acyclic diagrams that runs in $O(|V|^4|E|^2)$ time, where V is the set of vertices in the diagram, and E, the set of edges.
-We present an algorithm that runs in $O(|V|^2|E| + |V|^3)$ time.
-We evaluate the algorithm gainst baselines, and see its practical application in the compiler of a domain specific language for geometry types.
+It can be useful in program analysis to verify that a diagram representing conversions remains commutative over the course of its construction.
+A naive approach to solving the problem would take factorial time, and the presence of cycles could potentially lead to an infinite number of paths to check.
+Previous work provides an algorithm for the special case of acyclic diagrams that verifies that a given diagram commutes in $O(|V|^4|E|^2)$ time, where V is the set of vertices in the diagram, and E, the set of edges.
+We present an algorithm that given a diagram and an edge to add to it, verifies that the resultant diagram commutes in $O(|V|(|E| + |V|)$ time.
+For the case when checking the equality of paths is expensive, we present a second algorithm that runs in $O(V^4)$ time but uses the minimum possible number of equality checks.
+We evaluate the algorithms against baselines, and see its practical application in the compiler of a domain specific language for geometry types.
+As a demonstration of the more general applicability of the algorithm, we use it to identify discrepancies in a  currency conversion graph.
 
@@
 If a directed graph is constructed with the sets as nodes and functions as edges between the nodes, then by construction, in the new graph, choice of path should not matter- any choice of path should yield the same transformation between any two nodes.
-The idea is that each object has a unique representation in a given set in the system, so the transformation mapping between any two nodes should not depend on choice of path used to obtain the transformation.
+Each object has a unique representation in a given set in the system, so the transformation mapping between any two nodes should not depend on choice of path used to obtain the transformation.
 We will use the term "diagram" to refer to such graphs with their associated sets (vertices) and transformations (edges).
@@
 Naively checking if all paths in a given graph are path independent could require a number of function equality checks as bad as factorial in the number of nodes, since a path consists of an ordering of nodes.
-Previous work has identified an $O(|E|^2|V|^4)$ algorithm to verify that an acyclic diagram commutes.
+Further, the presence of cycles implies a potentially infinite number of paths.
+Previous work ~\cite{commutative} has identified an $O(|E|^2|V|^4)$ algorithm to verify that an acyclic diagram commutes.
 
-We present a polynomial time $O(|E||V|^2+|V|^3)$ algorithm to verify that commutativity of a graph is maintained over the course of online addition, assuming an oracle to check the abstract notion of equality of functions.
+We present a polynomial time $O(|V|(|E|+|V|)$ algorithm to verify that commutativity of a graph is maintained over the course of online addition, assuming an oracle to check the abstract notion of equality of functions.
 The number of calls to the equality checking oracle required by the algorithm is asymptotically tight.
@@
 
-First, we formalize the notion of a diagram, drawing inspiration from Murota\cite{commutative}.
+First, we formalize the notion of a diagram, drawing inspiration from Murota ~\cite{commutative}.
 
-We start with a directed graph G=(V,E) with sets attached to vertices and maps to arcs. Each edge (u, v) in E corresponds to a function that maps elements of u to elements in v.
+% A lot of this exposition is from the cited paper, I feel like I need to make this clearer to give proper credit, but not sure how.
+
+We start with a directed graph G=(V,E) with sets attached to vertices and maps, to arcs. Each edge (u, v) in E corresponds to a function that maps elements of u to elements in v.
 All the functions form a semigroup F, with multiplication denoted by $\circ$.
@@
 
-w
 The correspondence between edges and functions is stored as a mapping f:$E\rightarrow F$. f maps each edge to the function it represents.
@@
 
-Two situation can be imagined- the first, where the existence of the identity function from every node to itself is assumed, and the second, where no such identity needs to exist.
+Two situation can be imagined- the first, where the existence of the identity function from every node to itself is assumed as would describe most real world situations, and the second, where no such identity needs to exist.
 
@@
 Further, if there is any "exit" or "entrance" into an elementary cycle- an edge that is not a part of the cycle whose one end is a node that belongs to the cycle- then it must be verified that the transformation represented by the edge is equal to the transformation represented by the edge and cycle composed (the cycle is pre-pended in the case of an edge directed away from the cycle and appended in the case of an edge directed into the cycle).
-Should this verification succeed, by logic similar to the previous case, it must be that
-all paths (including complex cycles) that include the elementary cycle must be equal to version of the paths without the cycle (since a segment of the path with the cycle has already been verified to be equal to a version of that segment without the cycle).
+Should this verification succeed, by logic similar to the previous case, it must be that all paths (including complex cycles) that include the elementary cycle must be equal to version of the paths without the cycle (since a segment of the path with the cycle has already been verified to be equal to a version of that segment without the cycle).
 Once again this is a factorial time, terminating algorithm.
@@
 \paragraph{Implementation}
-*Copy in python code from implementation*
+*Copy in python code from implementation? Appendix?*
 
@@
 \subsection{Optimal solution for batch problem}
-Murota's main result \cite{commutative} is a solution to the question of whether a given acyclic diagram commutes that returns the minimal number of equality checks.
+Murota's main result ~\cite{commutative} is a solution to the question of whether a given acyclic diagram commutes that returns the minimal number of equality checks.
 The paper describes the following algorithm to find the ($V^2E$ bounded) minimal set of pairs that needs to be checked.
@@
 
-A bilinking is a parallel pair that is disjoint but for their terminal nodes. The set of all bilinkings is $R_0$.
-In an acyclic diagram, if all bilinkings are equal, all parallel pairs must also be equal since they any given pair can be expressed as a composition of one or more bilinkings.
+A bilinking is defined to be a parallel pair that is disjoint but for their terminal nodes. The set of all bilinkings is $R_0$.
+In an acyclic diagram, if all bilinkings are equal, all parallel pairs must also be equal since they any given pair can be expressed as a composition of bilinkings.
 
@@
 
-for a set of bilinkings r, the cl function is defined as:
+for a set of bilinkings r, the closure function cl is defined as:
 cl(r) = $\{ s\in R_0|$ s is linearly dependent on $F(r) \}$.
@@
 The closure function on r basically captures all the pairs that can be made by made by composing or "gluing together" the bilinkings in r. 
-The $\langle\rangle$ function captures the orientation of paths, and can be used to constrain compositions to be in the right direction.
+The aforementioned $\langle\rangle$ function captures the orientation of paths, and can be used to constrain compositions to be in the right direction.
 
 Using these two function we finally define the function $\sigma$ on a set of bilinkings r as
-$\sigma(r) = \{s \in R_0 | s\in cl(R\cap \langle s \rangle) \}$
+$\sigma(r) = \{s \in R_0 | s\in cl(R\cap \langle s \rangle) \}$.
 This is the function used to capture all the pairs whose verification is implied by the verification of pairs in r.
-It is used to iteratively check if a given pair is redundant. bilinkings are eliminated until a minimum "spanning" subset is reached.
+
+$\sigma$ is used to iteratively check if a given pair is redundant. bilinkings are eliminated until a minimum "spanning" subset is reached.
 
@@
 The solution presented here assumes that there is an implicit identity function on every node, as would be the case in most applications.
-However, a modified version of the algorithm for diagrams without this assumption is presented in the appendix, with modification that is similar in spirit to what was described in the description of the naive baselines.
+However, a modified version of the algorithm for diagrams without this assumption is presented in the appendix.
+The modification is similar in spirit to what was described in the description of the naive baselines.
 
@@
 
-The try block is executed at most $O(|V|^2)$ times.
+The try block is executed at most $O(|V|^2)$ times, so that this is the bound on the number of pairs verified.
 
@@
 
-The overall run time of the algorithm, assuming that path finding is done using depth first search, is $O((|V|+|E|)|V|^2)$.
+An efficient implementation of the algorithm can run in $O(|V|(|V|+|E|))$ time at a $O(|V|^3)$ space cost.
+In such an implementation, path finding from a given source node to all potential sink nodes would be done in a single $O(|V|+|E|)$ breadth first search and the results memoized in $O(|V|^2)$ space.
 
@@
 
-The run time of the first step is $O(|V|^4)$, and that of the second step is $O(|V|)$, so that the overall bound is $O(|V|^4)$, which is still a gain of a factor of $|E|^2$ over the batch algorithm.
+The run time of the first step is $O(|V|^4)$, and that of the second step is $O(|V|)$, so that the overall bound is $O(|V|^4)$.
 
@@
 \subsection{Currency Graph}
+
+\section{Evaluation}
 
@@
             continue
-    for cycle in findCyclesWithTerminatNode(newEdge.Source()):
+    cycle = findCycleWithTerminalNode(newEdge.Source()):
+    if cycle:    
         parallelPairs.add(newEdge, cycle+newEdge)
-    for cycle in findCyclesWithTerminatNode(newEdge.Source()):
+    cycle = findCycleWithTerminalNode(newEdge.Source()):
+    if cycle:    
         parallelPairs.add(newEdge, newEdge+cycle)
@@
 
-The optional minimization algorithm is modified to exclude nodes representing cycles, since the reduction case no longer will apply in the absence of an implicit identity on the terminal node of the cycle.
-*I think the minimality of minimized subset still holds, but not 100\% sure*
+The optional minimization algorithm is modified to exclude nodes representing cycles where the source and sink are equal, since the reduction case no longer will apply in the absence of an implicit identity on the terminal node of the cycle.
 

### 2020-05-19 paper/paper.tex
changed a couple of wards

@@
 We present an algorithm that given a diagram and an edge to add to it, verifies that the resultant diagram commutes in $O(|V|(|E| + |V|)$ time.
-For the case when checking the equality of paths is expensive, we present a second algorithm that runs in $O(V^4)$ time but uses the minimum possible number of equality checks.
+For the case when checking the equality of paths is expensive, we present an optimization that runs in $O(V^4)$ time but reduces to the minimum possible number of equality checks.
 We evaluate the algorithms against baselines, and see its practical application in the compiler of a domain specific language for geometry types.
@@
 
-\subsection{(Optional) Optimization Step}
+\subsection{Optimization Step}
 

### 2020-05-22 acmPaper/paper.tex
Added paper in ACMConf format for TAPAS

--- 2020-05-19 paper/paper.tex
@@
-% This is samplepaper.tex, a sample chapter demonstrating the
-% LLNCS macro package for Springer Computer Science proceedings;
-% Version 2.20 of 2017/10/04
-%
-\documentclass[runningheads]{llncs}
+\documentclass[sigconf,authordraft]{acmart}
 %
@@
 \usepackage{amssymb}
+\usepackage{algorithm}
 %
@@
 % \renewcommand\UrlFont{\color{blue}\rmfamily}
+\AtBeginDocument{%
+  \providecommand\BibTeX{{%
+    \normalfont B\kern-0.5em{\scshape i\kern-0.25em b}\kern-0.8em\TeX}}}
 
 \begin{document}
-%
+\settopmatter{printacmref=false}
 \title{Online verification of Commutativity}
@@
 %
-\author{First Author\inst{1}\orcidID{0000-1111-2222-3333} \and
-Second Author\inst{1}\orcidID{1111-2222-3333-4444} \and
-Third Author\inst{1}\orcidID{2222--3333-4444-5555}}
-%
-\authorrunning{F. Author et al.}
-% First names are abbreviated in the running head.
-% If there are more than two authors, 'et al.' is used.
-%
-\institute{Cornell University, Ithaca NY 14853, USA
-\email{\{abc,lncs\}@cornell.edu}}
-%
-\maketitle              % typeset the header of the contribution
-%
+\author{First Author}
+\email{email}
+\author{Second Author}
+\authornotemark[1]
+\email{email 2}
+\affiliation{%
+  \institution{Cornell University}
+  \streetaddress{address}
+  \city{Ithaca}
+  \state{New York}
+  \postcode{14850}
+}
+
+\author{Third Author}
+\email{email 3}
+
 \begin{abstract}
@@
 As a demonstration of the more general applicability of the algorithm, we use it to identify discrepancies in a  currency conversion graph.
-
 \end{abstract}
@@
 %
+
+\maketitle
+
+\section{Introduction}
 Many systems can be understood as a category, with a collection of sets, and some transformation functions between the sets- for example, in programming languages, type systems with coercion.

### 2020-05-25 currency-exchange/baseOnlineChecker.py
Added new edge case, transitioned some of the more testing to TestDefinition.

@@ -3,7 +3,7 @@ import copy
 from iPathChecker import IPathChecker
 from testUtilities import assertEqual, test_graph, test_s, test_t,\
     expected_paths_to_s, expected_paths_from_t, expected_fwd_graph,\
-        expected_bkd_graph, NO_EDGE
+        expected_bkd_graph, NO_EDGE, TestDefinition, defaultTestSuite
 from typing import List, Set, Tuple
 
 _default_no_edge = -1
@@ -18,7 +18,7 @@ class BaseOnlineChecker(IPathChecker):
 
     def getForwardEdges(self, source: int) -> List[int]:
         return [i for i in range(len(self.graph)) \
-            if self.graph[source][i] != self._no_edge and i != source]
+            if self.graph[source][i] != self._no_edge]
 
     def getBackwardEdges(self, sink: int) -> List[int]:
         return [i for i in range(len(self.graph)) \
@@ -201,45 +201,44 @@ class BaseOnlineChecker(IPathChecker):
 
 #### Quick tests ####
 
-def testBuildCompactBkdGraph():
+def testBuildCompactBkdGraph(testDefinition: TestDefinition):
     checker = BaseOnlineChecker()
-    checker.graph = test_graph
+    checker.graph = testDefinition.test_graph
     checker.buildCompactBkdGraph()
-    assert(checker.graph == test_graph)
+    assertEqual(testDefinition.test_graph, checker.graph)
     if __debug__ and checker._debug:
         # Assumption: lists ordered from smallest to largest element.
         print("expected: {}".format(str(expected_bkd_graph)))
         print("got: {}".format(str(checker.compactBkdGraph)))
-    assert(str(checker.compactBkdGraph) == str(expected_bkd_graph))
+    assertEqual(testDefinition.expected_bkd_graph, checker.compactBkdGraph)
 
-def testBuildCompactFwdGraph():
+def testBuildCompactFwdGraph(testDefinition: TestDefinition):
     checker = BaseOnlineChecker()
-    checker.graph = test_graph
+    checker.graph = testDefinition.test_graph
     checker.buildCompactFwdGraph()
-    assert(checker.graph == test_graph)
-    # Assumption: lists ordered from smallest to largest element.
-    assert(str(checker.compactFwdGraph) == str(expected_fwd_graph))
+    assert(checker.graph == testDefinition.test_graph)
[…]

### 2020-05-25 currency-exchange/optimalSetPathChecker.py

@@ -2,9 +2,8 @@ from baseOnlineChecker import BaseOnlineChecker
 from colorama import Fore
 import copy
 from testUtilities import assertEqual, test_graph, test_s, test_t,\
-    expected_solution
+    expected_solution, TestDefinition, defaultTestSuite
 import time
-import traceback
 from typing import List, Set, Tuple
 
 class OptimalSetPathChecker(BaseOnlineChecker):
@@ -83,23 +82,24 @@ def testGetRootPairs():
     checker.setEdge(test_s, test_t)
     assertEqual({(2, 2), (3, 3), (6, 6), (7, 7)}, checker.getRootPairs())
     
-def testGetPathsToCheck():
+def testGetPathsToCheck(testDefinition: TestDefinition):
     checker = OptimalSetPathChecker()
-    checker.setGraph(test_graph)
-    checker.setEdge(test_s, test_t)
-    assertEqual(expected_solution, checker.getPathsToCheck())
+    checker.setGraph(testDefinition.test_graph)
+    checker.setEdge(testDefinition.test_s, testDefinition.test_t)
+    assertEqual(testDefinition.expected_solution, checker.getPathsToCheck())
     assert(checker.timeTaken > 0)
 
-def runAllTests():
+def runAllTests(testSuite: List[TestDefinition]):
     print('\033[0m' + "Running optimalSetPathChecker Tests")
     testGetSuccessors()
     testGetRootPairs()
-    testGetPathsToCheck()
-    print(Fore.GREEN + 'Run Completed')    
+    for testDefinition in testSuite:
+        testGetPathsToCheck(testDefinition)
+    print(Fore.GREEN + 'Run Completed')
 
 #### Execute ####
 def main():
-    runAllTests()
+    runAllTests(defaultTestSuite)
 
 if __name__=="__main__":
     main()

### 2020-05-25 currency-exchange/testUtilities.py

@@ -1,7 +1,9 @@
 from colorama import Fore
+import traceback
 
 NO_EDGE = -1
 
+# Constants for default test.
 test_graph = [
     [NO_EDGE,       1, NO_EDGE, NO_EDGE, NO_EDGE, NO_EDGE, NO_EDGE, NO_EDGE],
     [NO_EDGE, NO_EDGE,       1, NO_EDGE, NO_EDGE, NO_EDGE, NO_EDGE, NO_EDGE],
@@ -111,3 +113,90 @@ def assertActualIsSuperset(expected, actual):
         print("Expected: {}".format(expected))
         print("Actual: {}".format(actual))
         traceback.print_stack()
+
+class TestDefinition():
+    """Define all the constant necessary for a single test case."""
+
+    def __init__(
+        self, 
+        test_graph, 
+        test_s, 
+        test_t, 
+        expected_fwd_graph = [],
+        expected_bkd_graph = [],
+        expected_paths_to_s = [],
+        expected_paths_from_t = [],
+        expected_solution = [],
+        expected_solution_no_identity = []):
+        self.NO_EDGE = NO_EDGE
+        self.test_graph = test_graph
+        self.test_s = test_s
+        self.test_t = test_t
+        self.test_new_edge = (test_s, test_t)
+        self.expected_fwd_graph = expected_fwd_graph
+        self.expected_bkd_graph = expected_bkd_graph
+        self.expected_paths_to_s = expected_paths_to_s
+        self.expected_paths_from_t = expected_paths_from_t,
+        self.expected_solution = expected_solution
+
+# Test Definitions
+
+defaultTestDefinition = TestDefinition(
+    test_graph = test_graph,
+    test_s = test_s,
+    test_t = test_t,
+    expected_fwd_graph = expected_fwd_graph,
+    expected_bkd_graph = expected_bkd_graph,
+    expected_paths_to_s = expected_paths_to_s,
+    expected_paths_from_t = expected_paths_from_t,
+    expected_solution = expected_solution)
+
+testCaseOne = TestDefinition(
+    test_graph = [
+        [NO_EDGE, NO_EDGE,       1,       1],
+        [NO_EDGE, NO_EDGE,       1,       1],
+        [1,       NO_EDGE, NO_EDGE, NO_EDGE],
[…]

### 2020-05-25 currency-exchange/noIdentityOptimalSetPathChecker.py
some more pieces for testing

@@ -2,7 +2,7 @@ from colorama import Fore
 from nonIdentityUtilities import NonIdentityPathChecker
 from optimalSetPathChecker import OptimalSetPathChecker
 from testUtilities import assertActualIsSuperset, test_graph, test_s, test_t,\
-    expected_solution_no_identity
+    expected_solution_no_identity, TestDefinition, defaultTestSuite
 import time
 from typing import List, Tuple
 
@@ -26,21 +26,22 @@ class NoIdentityOptimalSetPathChecker(OptimalSetPathChecker, NonIdentityPathChec
         self.timeTaken = endTime - startTime
         return pathPairs
 
-def testGetPathsToCheck():
+def testGetPathsToCheck(testDefinition: TestDefinition):
     checker = NoIdentityOptimalSetPathChecker()
-    checker.setGraph(test_graph)
-    checker.setEdge(test_s, test_t)
-    assertActualIsSuperset(expected_solution_no_identity,\
+    checker.setGraph(testDefinition.test_graph)
+    checker.setEdge(testDefinition.test_s, testDefinition.test_t)
+    assertActualIsSuperset(testDefinition.expected_solution_no_identity,\
         checker.getPathsToCheck())
     assert(checker.timeTaken > 0)
 
-def runAllTests():
+def runAllTests(testSuite: List[TestDefinition]):
     print('\033[0m' + "Running NoIdentityOptimalSetPathChecker Tests")
-    testGetPathsToCheck()
+    for testDefinition in testSuite:
+        testGetPathsToCheck(testDefinition)
     print(Fore.GREEN + 'Run Completed')
 
 def main():
-    runAllTests()
+    runAllTests(defaultTestSuite)
 
 if __name__=="__main__":
     main()

### 2020-05-25 currency-exchange/testUtilities.py

@@ -137,7 +137,8 @@ class TestDefinition():
         self.expected_bkd_graph = expected_bkd_graph
         self.expected_paths_to_s = expected_paths_to_s
         self.expected_paths_from_t = expected_paths_from_t,
-        self.expected_solution = expected_solution
+        self.expected_solution = expected_solution,
+        self.expected_solution_no_identity = expected_solution_no_identity
 
 # Test Definitions
 
@@ -149,7 +150,8 @@ defaultTestDefinition = TestDefinition(
     expected_bkd_graph = expected_bkd_graph,
     expected_paths_to_s = expected_paths_to_s,
     expected_paths_from_t = expected_paths_from_t,
-    expected_solution = expected_solution)
+    expected_solution = expected_solution,
+    expected_solution_no_identity = expected_solution_no_identity)
 
 testCaseOne = TestDefinition(
     test_graph = [
@@ -191,6 +193,11 @@ testCaseOne = TestDefinition(
         # [0, 1, 2] = [0, 2] => [0, 1, 2, 0, 3] = [0, 2, 0, 3].
         # From existing graph [0, 2, 0, 3] = [0, 3] and [1, 2, 0, 3] = [1, 3].
         # Put together, [0, 1, 3] = [0, 3].
+    ],
+    expected_solution_no_identity = [
+        ([0, 1, 2], [0, 2]),
+        ([1, 2, 0, 1], [1, 2, 0, 1, 2, 0, 1])
+        # TODO add missing entries.
     ]
 )
 

### 2020-05-26 currency-exchange/PolynomialPathChecker.py
partial fix for non-identity case

@@ -11,6 +11,7 @@ class PolynomialPathChecker(BaseOnlineChecker):
 
     def __init__(self):
         super().__init__()
+        self.noIdentity = True
 
     def getPathsToCheck(self) ->  List[Tuple[List[int], List[int]]]:
         """Return the pairs of path whose equality implies path independence of 

### 2020-05-26 currency-exchange/nonIdentityUtilities.py

@@ -1,4 +1,7 @@
 from baseOnlineChecker import BaseOnlineChecker
+from colorama import Fore
+from testUtilities import test_graph, test_s, test_t,\
+    expected_solution_no_identity, TestDefinition, defaultTestSuite, assertEqual
 from typing import List, Tuple
 
 class NonIdentityPathChecker(BaseOnlineChecker):
@@ -8,21 +11,73 @@ class NonIdentityPathChecker(BaseOnlineChecker):
         cycle = self.pathsToNewEdgeSource[terminalNode] +\
             self.pathsFromNewEdgeSink[terminalNode] 
         pathsToCheck +=[(cycle, cycle + cycle[1:])]
-        for cycleExit in self.getForwardEdges(terminalNode):
-            pathsToCheck += [(cycle, cycle + [cycleExit])]
-        for cycleEntry in self.getBackwardEdges(terminalNode):
-            pathsToCheck += [(cycle, [cycleEntry] + cycle)]
+        # for cycleExit in self.getForwardEdges(terminalNode):
+        #     pathsToCheck += [(cycle, cycle + [cycleExit])]
+        # for cycleEntry in self.getBackwardEdges(terminalNode):
+        #     pathsToCheck += [(cycle, [cycleEntry] + cycle)]
         return pathsToCheck
     
     def getSinkCycles(self) -> List[Tuple[List[int], List[int]]]:
-        (cycleExists, cycle) = self.findPath(self.newEdgeSink, self.newEdgeSink)
+        (cycleExists, cycle) = self.findCycle(self.newEdgeSink)
         if cycleExists:
-            return [(cycle, [self.newEdgeSource] + cycle)]
+            return [([self.newEdgeSource, self.newEdgeSink], [self.newEdgeSource] + cycle)]
         return []
     
     def getSourceCycles(self) -> List[Tuple[List[int], List[int]]]:
-        (cycleExists, cycle) = self.findPath(
-            self.newEdgeSource, self.newEdgeSource)
+        (cycleExists, cycle) = self.findCycle(
+            self.newEdgeSource)
         if cycleExists:
-            return [(cycle, cycle+[self.newEdgeSink])]
+            return [([self.newEdgeSource, self.newEdgeSink], cycle+[self.newEdgeSink])]
         return []
+    
+    def findCycle(self, source: int) -> Tuple[bool, List[int]]:
+        """Find cycle."""
+        # DFS.
+        path = []
+        visited = set()
+        stack = [source]
+        firstRun = True
+        while (len(stack)>0):
[…]

### 2020-05-26 currency-exchange/optimalSetPathChecker.py

@@ -12,6 +12,7 @@ class OptimalSetPathChecker(BaseOnlineChecker):
 
     def __init__(self):
         super().__init__()
+        self.noIdentity = False
 
     def getSuccessors(self, source: int, sink: int) -> Set[int]:
         """Returns a set of pairs such that verifying a path between given
@@ -19,6 +20,9 @@ class OptimalSetPathChecker(BaseOnlineChecker):
         equal."""
         predecessors = self.getAllPredecessors(source)
         successors = self.getAllSuccessors(sink)
+        if (self.noIdentity and source == sink):
+            return {(src, snk) for src in predecessors for snk in successors 
+            if src == snk}
         return {(src, snk) for src in predecessors for snk in successors}
 
     def getRootPairs(self) -> List[List[int]]:

### 2020-05-26 currency-exchange/testUtilities.py

@@ -69,8 +69,8 @@ expected_solution_no_identity = [
     ([3, 6, 7, 2, 3], [3, 6, 7, 2, 3, 6, 7, 2, 3]),
     ([6, 7, 2, 3, 6], [6, 7, 2, 3, 6, 7, 2, 3, 6]),
     ([7, 2, 3, 6, 7], [7, 2, 3, 6, 7, 2, 3, 6, 7]),
-    ([2, 3, 6, 7, 2], [1, 2, 3, 6, 7, 2]),
-    ([3, 6, 7, 2, 3], [3, 6, 7, 2, 3, 4])
+    ([1, 2], [1, 2, 3, 6, 7, 2]),
+    ([3, 4], [3, 6, 7, 2, 3, 4])
 ]
 
 # This method is from stack overflow.
@@ -80,6 +80,8 @@ def compareLists(expected, actual):
         for elem in expected:
             actual.remove(elem)
     except ValueError:
+        print(Fore.RED + "Missing element")
+        traceback.print_stack()
         return False
     return not actual
 
@@ -89,6 +91,8 @@ def actualListIsSuperset(expected, actual):
         for elem in expected:
             actual.remove(elem)
     except ValueError:
+        print(Fore.RED + "Missing element")
+        traceback.print_stack()
         return False
     return True
 
@@ -108,6 +112,9 @@ def assertActualIsSuperset(expected, actual):
     try:
         if isinstance(expected, list):
             assert(actualListIsSuperset(expected, actual))
+        else:
+            raise Exception("no handling for checking superset for {}"
+            .format(expected))
     except:
         print(Fore.RED + "Error")
         print("Expected: {}".format(expected))
@@ -127,7 +134,9 @@ class TestDefinition():
         expected_paths_to_s = [],
         expected_paths_from_t = [],
         expected_solution = [],
-        expected_solution_no_identity = []):
+        expected_solution_no_identity = [],
+        source_cycles = [],
+        sink_cycles = []):
         self.NO_EDGE = NO_EDGE
         self.test_graph = test_graph
         self.test_s = test_s
@@ -136,9 +145,11 @@ class TestDefinition():
         self.expected_fwd_graph = expected_fwd_graph
         self.expected_bkd_graph = expected_bkd_graph
         self.expected_paths_to_s = expected_paths_to_s
[…]

### 2020-05-26 currency-exchange/PolynomialPathChecker.py
Yay i think no identity optimal case is fixed

@@ -11,7 +11,7 @@ class PolynomialPathChecker(BaseOnlineChecker):
 
     def __init__(self):
         super().__init__()
-        self.noIdentity = True
+        self.noIdentity = False
 
     def getPathsToCheck(self) ->  List[Tuple[List[int], List[int]]]:
         """Return the pairs of path whose equality implies path independence of 

### 2020-05-26 currency-exchange/baseOnlineChecker.py

@@ -174,6 +174,7 @@ class BaseOnlineChecker(IPathChecker):
         # sink node -> {source node -> path}
         self.pathsToNode = dict()
         self._no_edge = _default_no_edge
+        self.noIdentity = False
     
     def setGraph(
         self,

### 2020-05-26 currency-exchange/noIdentityOptimalSetPathChecker.py

@@ -2,11 +2,15 @@ from colorama import Fore
 from nonIdentityUtilities import NonIdentityPathChecker
 from optimalSetPathChecker import OptimalSetPathChecker
 from testUtilities import assertActualIsSuperset, test_graph, test_s, test_t,\
-    expected_solution_no_identity, TestDefinition, defaultTestSuite
+    expected_solution_no_identity, TestDefinition, defaultTestSuite, assertEqual
 import time
 from typing import List, Tuple
 
 class NoIdentityOptimalSetPathChecker(OptimalSetPathChecker, NonIdentityPathChecker):
+
+    def __init__(self):
+        super().__init__()
+        self.noIdentity = True
 
     def getPathsToCheck(self) ->  List[Tuple[List[int], List[int]]]:
         """Return the pairs of path whose equality implies path independence of 
@@ -30,7 +34,7 @@ def testGetPathsToCheck(testDefinition: TestDefinition):
     checker = NoIdentityOptimalSetPathChecker()
     checker.setGraph(testDefinition.test_graph)
     checker.setEdge(testDefinition.test_s, testDefinition.test_t)
-    assertActualIsSuperset(testDefinition.expected_solution_no_identity,\
+    assertEqual(testDefinition.expected_solution_no_identity,\
         checker.getPathsToCheck())
     assert(checker.timeTaken > 0)
 

### 2020-05-26 currency-exchange/nonIdentityUtilities.py

@@ -6,6 +6,10 @@ from typing import List, Tuple
 
 class NonIdentityPathChecker(BaseOnlineChecker):
 
+    def __init__(self):
+        super().__init__()
+        self.noIdentity = True
+
     def handleCycle(self, terminalNode: int) -> List[Tuple[List[int], List[int]]]:
         pathsToCheck = []
         cycle = self.pathsToNewEdgeSource[terminalNode] +\
@@ -72,7 +76,7 @@ def runAllTests(testSuite: List[TestDefinition]):
     print('\033[0m' + "Running NoIdentityPolynomialPathChecker Tests")
     for testDefinition in testSuite:
         testGetSourceCycles(testDefinition)
-        testGetSinkCycles
+        testGetSinkCycles(testDefinition)
     print(Fore.GREEN + 'Run Completed')    
 
 #### Execute ####

### 2020-05-26 currency-exchange/testUtilities.py

@@ -69,8 +69,11 @@ expected_solution_no_identity = [
     ([3, 6, 7, 2, 3], [3, 6, 7, 2, 3, 6, 7, 2, 3]),
     ([6, 7, 2, 3, 6], [6, 7, 2, 3, 6, 7, 2, 3, 6]),
     ([7, 2, 3, 6, 7], [7, 2, 3, 6, 7, 2, 3, 6, 7]),
-    ([1, 2], [1, 2, 3, 6, 7, 2]),
-    ([3, 4], [3, 6, 7, 2, 3, 4])
+    ([1, 2, 3, 6, 7, 2], [1, 2]),
+    ([3, 6, 7, 2, 3, 4], [3, 4]),
+    ([7, 2, 3, 6, 7, 2], [7, 2]),
+    ([3, 6, 7, 2, 3, 6], [3, 6]),
+    ([6, 7, 2, 3, 6, 7], [6, 7])
 ]
 
 # This method is from stack overflow.
@@ -83,15 +86,21 @@ def compareLists(expected, actual):
         print(Fore.RED + "Missing element")
         traceback.print_stack()
         return False
+    if actual:
+        print(Fore.RED + "Extra element(s) {}".format(actual))
+        traceback.print_stack()
+        return False 
     return not actual
 
 def actualListIsSuperset(expected, actual):
     actual = list(actual)   # make a mutable copy
+    currentElem = None
     try:
         for elem in expected:
+            currentElem = elem
             actual.remove(elem)
     except ValueError:
-        print(Fore.RED + "Missing element")
+        print(Fore.RED + "Missing element {}".format(currentElem))
         traceback.print_stack()
         return False
     return True

### 2020-05-26 currency-exchange/noIdentityPolynomialPathChecker.py
Added tests for noIdentityPolynomialPathChecker

@@ -1,7 +1,7 @@
 from nonIdentityUtilities import NonIdentityPathChecker
 from colorama import Fore
 from testUtilities import assertActualIsSuperset, test_graph, test_s, test_t,\
-    expected_solution_no_identity
+    expected_solution_no_identity, defaultTestSuite, TestDefinition
 import time
 from typing import List, Tuple
 
@@ -42,22 +42,23 @@ class NoIdentityPolynomialPathChecker(NonIdentityPathChecker):
 
 #### Quick Tests ####
 
-def testGetPathsToCheck():
+def testGetPathsToCheck(testDefinition: TestDefinition):
     checker = NoIdentityPolynomialPathChecker()
-    checker.setGraph(test_graph)
-    checker.setEdge(test_s, test_t)
-    assertActualIsSuperset(expected_solution_no_identity,\
+    checker.setGraph(testDefinition.test_graph)
+    checker.setEdge(testDefinition.test_s, testDefinition.test_t)
+    assertActualIsSuperset(testDefinition.expected_solution_no_identity,\
         checker.getPathsToCheck())
     assert(checker.timeTaken > 0)
 
-def runAllTests():
+def runAllTests(testSuite: List[TestDefinition]):
     print('\033[0m' + "Running NoIdentityPolynomialPathChecker Tests")
-    testGetPathsToCheck()
+    for testDefinition in testSuite:
+        testGetPathsToCheck(testDefinition)
     print(Fore.GREEN + 'Run Completed')    
 
 #### Execute ####
 def main():
-    runAllTests()
+    runAllTests(defaultTestSuite)
 
 if __name__=="__main__":
     main()

### 2020-05-26 currency-exchange/nonIdentityUtilities.py

@@ -15,10 +15,6 @@ class NonIdentityPathChecker(BaseOnlineChecker):
         cycle = self.pathsToNewEdgeSource[terminalNode] +\
             self.pathsFromNewEdgeSink[terminalNode] 
         pathsToCheck +=[(cycle, cycle + cycle[1:])]
-        # for cycleExit in self.getForwardEdges(terminalNode):
-        #     pathsToCheck += [(cycle, cycle + [cycleExit])]
-        # for cycleEntry in self.getBackwardEdges(terminalNode):
-        #     pathsToCheck += [(cycle, [cycleEntry] + cycle)]
         return pathsToCheck
     
     def getSinkCycles(self) -> List[Tuple[List[int], List[int]]]:

### 2020-05-26 currency-exchange/PolynomialPathChecker.py
Currency graph code

@@ -11,7 +11,6 @@ class PolynomialPathChecker(BaseOnlineChecker):
 
     def __init__(self):
         super().__init__()
-        self.noIdentity = False
 
     def getPathsToCheck(self) ->  List[Tuple[List[int], List[int]]]:
         """Return the pairs of path whose equality implies path independence of 

### 2020-05-26 currency-exchange/currencyGraph.py

@@ -1,7 +1,10 @@
+from colorama import Fore
 import requests 
 import numpy as np
 import copy
 from optimalSetPathChecker import OptimalSetPathChecker
+from testUtilities import TestDefinition, defaultTestSuite
+from typing import List
 
 class CurrencyGraph():
     """Class that holds, builds and manipulates the graph with currency 
@@ -15,10 +18,123 @@ class CurrencyGraph():
     def __init__(self):
         self.graph = []
         self.currencyList = []
+        self.base = 'USD' # Default but can change
         np.set_printoptions(formatter={
             'float': lambda x: "{0:0.4f}".format(x)}, threshold=10000)
+    
+    def setup(self):
+        """Setup graph with initial row of data from base currency"""
+        rates = self.makeRequest(self.base)
+        self.initialize(list(rates))
+        self.addRow(self.base, rates)
+        if self.DEBUG:
+            print("Initialized Graph:")
+            print(np.matrix(self.currencyList))
+            print(np.matrix(self.graph))
+
+    def initialize(self, currencies):
+        """Initialize graph and currency list."""
+        self.currencyList = currencies
+        self.constructGraph()
     
     def indexOf(self, currency):
         """Gets node in graph that a currency corresponds to."""
         return self.currencyList.index(currency)
     
+    def constructGraph(self):
+        """Initialize empty graph."""
+        for _ in range(len(self.currencyList)):
+            row = []
+            for _ in range(len(self.currencyList)):
+                row += [self.NO_EDGE]
+            self.graph += [row]
+    
+    def saveEntry(self, base, target, rate, checkIndependence=False):
+        isValid = True
+        if (checkIndependence):
+            isValid = self.checkIndependenceFunc(base, target, rate)
+        if (isValid):
+            self.graph[self.indexOf(base)][self.indexOf(target)] = rate    
+        return isValid
+    
+    # Does not call addRates for efficiency.
+    def addRow(self, base, rates, checkIndependence=False):
+        for currency, rate in rates.items():
+            isValid = True
+            if (checkIndependence):
+                isValid = self.checkIndependenceFunc(base, currency, rate)
+            if (isValid):
+                self.graph[self.indexOf(base)][self.indexOf(currency)] = rate
+            else:
+                return False
+        return True
+
+    def makeUrl(self, base, symbols):
+        URL = self.endPoint
[…]

### 2020-05-26 currency-exchange/main.py

@@ -1,18 +1,16 @@
-import requests 
-import numpy as np
 import copy
+from currencyGraph import CurrencyGraph
+import numpy as np
 from optimalSetPathChecker import OptimalSetPathChecker
+from iPathChecker import IPathChecker
+from polynomialPathChecker import PolynomialPathChecker
+import requests 
+from typing import List
 
 # Assume that a graph doesn't have more edges nodes than this
 INFTY = 1000000
 NO_EDGE = -1
 DEBUG = False
-DEV_DEBUG = True
-GRAPH = []
-# To get indices of currencies
-CURRENCY_LIST = []
-# api-endpoint 
-endPoint = "https://api.exchangeratesapi.io/latest?"
 
 # TODO add:
 # 1. Commandline arguments.
@@ -21,95 +19,47 @@ endPoint = "https://api.exchangeratesapi.io/latest?"
 
 ###############################################
 
-# Helpers
-def indexOf(currency):
-    global CURRENCY_LIST
-    return CURRENCY_LIST.index(currency)
+epsilon = 0.001 # allowed error percent
 
-def constructGraph(currencies):
-    global CURRENCY_LIST
-    global GRAPH
-    CURRENCY_LIST = currencies
-    # GRAPH = [[0]*len(CURRENCY_LIST)]*len(CURRENCY_LIST)
-    for _ in range(len(CURRENCY_LIST)):
-        row = []
-        for _ in range(len(CURRENCY_LIST)):
-            row += [NO_EDGE]
-        GRAPH += [row]
+checkers = [
+    OptimalSetPathChecker(),
+    PolynomialPathChecker()
+]
 
-def addRates(base, target, rate, checkIndependence=False):
-    global GRAPH
-    isValid = True
-    if (checkIndependence):
-        isValid = checkIndependenceFunc((base, target, rate))
-    if (isValid):
-        GRAPH[indexOf(base)][indexOf(target)] = rate    
-    return isValid
+def getPathValue(path: List[int], graph: CurrencyGraph):
+    """Composition oracle defintion."""
+    result = 1
+    for i in range(len(path)-1):
+        result *= graph.graph[path[i]][path[i+1]]
+    return result
 
-# Does not call addRates for efficiency.
-def addRow(base, rates, checkIndependence=False):
-    global GRAPH
-    for currency,rate in rates.iteritems():
-        isValid = True
-        if (checkIndependence):
-            isValid = checkIndependenceFunc((base, currency, rate))
[…]

### 2020-05-26 currency-exchange/optimalSetPathChecker.py

@@ -12,7 +12,6 @@ class OptimalSetPathChecker(BaseOnlineChecker):
 
     def __init__(self):
         super().__init__()
-        self.noIdentity = False
 
     def getSuccessors(self, source: int, sink: int) -> Set[int]:
         """Returns a set of pairs such that verifying a path between given

### 2020-06-07 currency-exchange/currencyGraph.py
improvements to path checker

@@ -51,16 +51,17 @@ class CurrencyGraph():
     
     def saveEntry(self, base, target, rate, checkIndependence=False):
         isValid = True
+        extraData = ''
         copyWithNewEdge = self.getCopyWithNewEdge(base, target, rate)
         if (checkIndependence):
-            isValid = self.checkIndependenceFunc(
+            (isValid, extraData) = self.checkIndependenceFunc(
                 base, 
                 target, 
                 rate, 
                 copyWithNewEdge)
         if (isValid):
             self.graph = copyWithNewEdge    
-        return isValid
+        return (isValid, extraData)
 
     def getCopyWithNewEdge(self, base, target, rate):
         newGraph = copy.deepcopy(self.graph)
@@ -70,9 +71,10 @@ class CurrencyGraph():
     def saveRow(self, base, rates, checkIndependence=False):
         for currency, rate in rates.items():
             isValid = True
-            if not self.saveEntry(base, currency, rate):
-                return False
-        return True
+            (isValid, extraInfo) = self.saveEntry(base, currency, rate)
+            if not isValid:
+                return (isValid, extraInfo)
+        return (True, 'Succeeded')
 
     def makeUrl(self, base, symbols):
         URL = self.endPoint
@@ -107,7 +109,7 @@ class CurrencyGraph():
             rates = self.makeRequest(base, [target])
         except:
             if retries > 0:
-                self.addEntry(base, target, check, retries-1)
+                return self.addEntry(base, target, check, retries-1)
             else:
                 raise Exception("api call failed")
         return self.saveEntry(base, target, rates[target], check)        
@@ -118,7 +120,9 @@ class CurrencyGraph():
         raise NotImplementedError
 
     def printGraph(self):
-        print(self.graph)
+        for i in range(len(self.graph)):
+            print(i)
+            print(self.graph[i])
 
 #### Quick Tests ####
 

### 2020-06-07 currency-exchange/main.py

@@ -1,3 +1,4 @@
+from colorama import Fore
 import copy
 from currencyGraph import CurrencyGraph
 import numpy as np
@@ -20,11 +21,11 @@ ID_FUNCTION = [-1]
 
 ###############################################
 
-epsilon = 0.0001 # allowed error percent
+epsilon = 0.000000001 # allowed error multiplier
 
 checkers = [
-    PolynomialPathChecker(),
-    OptimalSetPathChecker()
+    OptimalSetPathChecker(),
+    PolynomialPathChecker()
 ]
 
 def getPathValue(path: List[int], graph: List[List[int]]):
@@ -51,9 +52,18 @@ def getIndependenceFromChecker(checker: IPathChecker, graph: CurrencyGraph):
         for (path1, path2) in pairsToCheck:
             path1Value = getPathValue(path1, newGraph)
             path2Value = getPathValue(path2, newGraph)
-            if (path1Value - path2Value) > epsilon * path2Value:
-                return False
-        return True
+            if abs(path1Value - path2Value) > epsilon * path2Value:
+                return (False, "path 1: {}: {} \n path 2: {}: {} \n difference: {}, new edge: {} -> {}: {}"
+                .format(
+                path1Value,
+                path1,
+                path2Value,
+                path2,
+                path1Value - path2Value,
+                base,
+                target,
+                rate))
+        return (True, "Success for base {} target {}".format(base, target))
     return IndependenceFunction
             
 
@@ -62,13 +72,26 @@ for checker in checkers:
     graph = CurrencyGraph()
     graph.setup()
     graph.checkIndependenceFunc = getIndependenceFromChecker(checker, graph)
-    for baseCurrency in graph.currencyList:
-        if baseCurrency == graph.base:
-            continue
-        for targetCurrency in graph.currencyList:
-            print("Adding edge from {} to {}"
-                .format(baseCurrency, targetCurrency))
-            success = graph.addEntry(baseCurrency, targetCurrency, True)
-            print(success)
-    print("Completed graph for {}".format(checker.__class__.__name__))
+    try:    
+        for baseCurrency in graph.currencyList:
+            if baseCurrency == graph.base:
+                continue
+            for targetCurrency in graph.currencyList:
+                if (DEBUG):
+                    print("Adding edge from {} to {}"
+                        .format(baseCurrency, targetCurrency))
+                (success, info) = graph.addEntry(baseCurrency, targetCurrency, True)
+                if not success:
[…]

### 2020-06-07 acmPaper/paper.tex
Capitalize titles
A bit about currency graph

@@
-\documentclass[sigconf,authordraft]{acmart}
+\documentclass[sigplan,review,anonymous]{acmart}
+\setcopyright{none}
 %
@@
 \settopmatter{printacmref=false}
-\title{Online verification of Commutativity}
+\title{Online Verification of Commutativity}
 %
@@
 
+To demonstrate our algorithm applied to a real world situation, 
+we found inconsistencies in a diagram of the exchange rate between currencies.
+Consider a diagram with nodes as currencies and a directed edge being the
+conversion rate from its source node's currency to its sink node's currency.
+Since the transformation exchange rate of money from any given base currency to a target currency
+can be expected to be the same regardless of which intermediate currency transformations
+are used, this diagram should commute.
+
+Using an api, https://exchangeratesapi.io, we built a fully connected diagram for 32 currencies.
+To ensure that it indeed commuted, before the addition of each edge, we used the algorithms presented
+to ensure the addition of a new edge did not introduce inconsistencies in the existing diagram.
+If a new edge was, in fact, problematic, the algorithm returned an example inconsistent
+pair that would arise from the addition of the edge.
+The pair would consist of two currency transformation sequences with the 
+same source currency and ultimate destination currency, but with different effective 
+exchange rates values, as computed by taking the product of all the exchange rates
+encountered through the chain.
+
+We allowed an "error tolerance" for floating point error.
+Graphs were generally consistent up to an error tolerance to the order of magnitude $10^{-6}$.
+However, below this threshold we found inconsistencies.
+For both algorithms, computation completed in minutes.
+
 \section{Evaluation}

### 2020-06-07 currency-exchange/currencyGraph.py
A few printing adjustments

@@ -121,7 +121,7 @@ class CurrencyGraph():
 
     def printGraph(self):
         for i in range(len(self.graph)):
-            print(i)
+            print("row {}:".format(i))
             print(self.graph[i])
 
 #### Quick Tests ####

### 2020-06-07 currency-exchange/main.py

@@ -17,11 +17,10 @@ ID_FUNCTION = [-1]
 # TODO add:
 # 1. Commandline arguments.
 # 2. Historic rate.
-# 3. Connection to path checking algorithm.
 
 ###############################################
 
-epsilon = 0.000000001 # allowed error multiplier
+epsilon = 1E-7 # allowed error multiplier
 
 checkers = [
     OptimalSetPathChecker(),
@@ -53,7 +52,7 @@ def getIndependenceFromChecker(checker: IPathChecker, graph: CurrencyGraph):
             path1Value = getPathValue(path1, newGraph)
             path2Value = getPathValue(path2, newGraph)
             if abs(path1Value - path2Value) > epsilon * path2Value:
-                return (False, "path 1: {}: {} \n path 2: {}: {} \n difference: {}, new edge: {} -> {}: {}"
+                return (False, "path 1: {}: {} \npath 2: {}: {} \n difference: {}, new edge: {} -> {}: rate: {}"
                 .format(
                 path1Value,
                 path1,
@@ -89,9 +88,11 @@ for checker in checkers:
                             checker.__class__.__name__))
                     print(info)
                     raise StopIteration
-    except:
-        pass    
-    print(Fore.GREEN + "Completed graph for {}"
+        print(Fore.GREEN + "Completed graph for {}"
         .format(checker.__class__.__name__))
+    except:
+        print(Fore.RED + "Aborted building graph for {}"
+        .format(checker.__class__.__name__))  
     print('\033[0m')
+    print('Resultant graph:')
     graph.printGraph()

### 2020-06-07 currency-exchange/currencyGraph.py
Another small refactoring

@@ -52,18 +52,14 @@ class CurrencyGraph():
     def saveEntry(self, base, target, rate, checkIndependence=False):
         isValid = True
         extraData = ''
-        copyWithNewEdge = self.getCopyWithNewEdge(base, target, rate)
         if (checkIndependence):
-            (isValid, extraData) = self.checkIndependenceFunc(
-                base, 
-                target, 
-                rate, 
-                copyWithNewEdge)
+            (isValid, extraData) = self.checkIndependenceFunc(base, target, rate)
         if (isValid):
-            self.graph = copyWithNewEdge    
+            self.graph[self.indexOf(base)][self.indexOf(target)] = rate
         return (isValid, extraData)
 
     def getCopyWithNewEdge(self, base, target, rate):
+        """Get a copy of this graph with a new edge from base to target added"""
         newGraph = copy.deepcopy(self.graph)
         newGraph[self.indexOf(base)][self.indexOf(target)] = rate
         return newGraph
@@ -115,7 +111,7 @@ class CurrencyGraph():
         return self.saveEntry(base, target, rates[target], check)        
 
     # New edge is (source, target, rate)
-    def checkIndependenceFunc(self, source, target, newGraph):
+    def checkIndependenceFunc(self, source, target):
         """This should be set by caller, much like an observer."""
         raise NotImplementedError
 

### 2020-06-07 currency-exchange/main.py

@@ -41,13 +41,14 @@ def getPathValue(path: List[int], graph: List[List[int]]):
         result *= graph[path[i]][path[i+1]]
     return result
 
-def getIndependenceFromChecker(checker: IPathChecker, graph: CurrencyGraph):
+def independenceFuncFactory(checker: IPathChecker, graph: CurrencyGraph):
     """Equality check oracle definition."""
     checker.setGraph(graph.graph)
     checker.setIdFunction(ID_FUNCTION)
-    def IndependenceFunction(base, target, rate, newGraph):
+    def IndependenceFunction(base, target, rate):
         checker.setEdge(graph.indexOf(base), graph.indexOf(target))
         pairsToCheck = checker.getPathsToCheck()
+        newGraph = graph.getCopyWithNewEdge(base, target, rate)
         for (path1, path2) in pairsToCheck:
             path1Value = getPathValue(path1, newGraph)
             path2Value = getPathValue(path2, newGraph)
@@ -70,7 +71,7 @@ for checker in checkers:
     print("Working with {}".format(checker.__class__.__name__))
     graph = CurrencyGraph()
     graph.setup()
-    graph.checkIndependenceFunc = getIndependenceFromChecker(checker, graph)
+    graph.checkIndependenceFunc = independenceFuncFactory(checker, graph)
     try:    
         for baseCurrency in graph.currencyList:
             if baseCurrency == graph.base:

### 2020-06-08 acmPaper/paper.tex
Pass on sections 1/2

--- 2020-05-22 acmPaper/paper.tex
@@
+\newcount\draft\draft=1 % set to 0 for submission or publication
 \documentclass[sigconf,authordraft]{acmart}
@@
 \usepackage{algorithm}
+\usepackage{xxx}
 %
@@
 \section{Introduction}
-Many systems can be understood as a category, with a collection of sets, and some transformation functions between the sets- for example, in programming languages, type systems with coercion.
-If a directed graph is constructed with the sets as nodes and functions as edges between the nodes, then by construction, in the new graph, choice of path should not matter- any choice of path should yield the same transformation between any two nodes.
-Each object has a unique representation in a given set in the system, so the transformation mapping between any two nodes should not depend on choice of path used to obtain the transformation.
-We will use the term "diagram" to refer to such graphs with their associated sets (vertices) and transformations (edges).
+Many systems can be understood as graphs: a set of objects with directed functions between members.
+With programming, for example, type systems with coercions can be drawn as a graph with type nodes and coercion edges.
+A desirable trait of such a type system and many other similar systems is that traversing the graph from a given start and end node results in the same value independent of path chosen; a property equivalent to saying the graph diagram is commutative.
+With our type coercion example, it could be problematic if casting to a supposedly equivalent type as an intermediate step resulted in a different answer than a direct cast.
 
-Since such commutative diagrams arise in many systems, it is often useful to be able to perform verification and ensure the graphs in question are indeed path independent. 
-We formalize the problem as checking the commutativity of a diagram incrementally.
-A diagram is said to commute when on selecting any two vertices, any directed path leads to the same result by composition.
+Given the utility of system diagram commutativity, it would be helpful to check that the property holds for real systems.
+Since diagrams may change over time, and verifying the entire system from scratch may be computationally expensive, it would be additionally convenient if new edges could be checked online as they are added.
+
 Naively checking if all paths in a given graph are path independent could require a number of function equality checks as bad as factorial in the number of nodes, since a path consists of an ordering of nodes.
-Further, the presence of cycles implies a potentially infinite number of paths.
-Previous work ~\cite{commutative} has identified an $O(|E|^2|V|^4)$ algorithm to verify that an acyclic diagram commutes.
+Previous work ~\cite{commutative} has identified an $O(|E|^2|V|^4)$ algorithm to verify that a complete acyclic diagram commutes; however, this solution neither manages the batch case nor the existence of cycles.
+Unfortunately, the presence of cycles implies a potentially infinite number of paths.
 
-We present a polynomial time $O(|V|(|E|+|V|)$ algorithm to verify that commutativity of a graph is maintained over the course of online addition, assuming an oracle to check the abstract notion of equality of functions.
-The number of calls to the equality checking oracle required by the algorithm is asymptotically tight.
-With an additional $O(|V|^4)$ optimization step, is is possible to use instead the minimal possible number of equality checks, should these checks be expensive.
-We compare our solution to three baselines: a naive checking of all path pairs, with only special handling for cycles, a check for all path pairs that involve the new edge, and with an algorithm suggested by previous work to solve the batch version of the problem for acyclic diagrams.
+We present a polynomial time $O(|V|(|E|+|V|)$ algorithm to verify that commutativity of a graph (potentially with cycles) is maintained over the course of online addition, assuming an oracle to check the abstract notion of equality of functions.
+We show the number of calls to the equality checking oracle is asymptotically tight.
+Should the calls to the oracle be expensive, we also show, with an additional $O(|V|^4)$ optimization step, an algorithm to use instead the minimal possible number of equality checks.
 
 To evaluate our results we use two case studies.
-First, in the domain specific geometry type language \textit{Gator}, our algorithm is used to ensure that user defined transformations between spaces stay consistent, and the compiler, deterministic as it makes automatic conversion choices.
-As a second case study, the the algorithm is used to identify inefficiencies in a currency conversion graph.
+First, our algorithm is used for the domain specific geometry type language \textit{Gator} to ensure that user defined transformations between spaces stay consistent.
+Second, our algorithm is used to identify inefficiencies in a currency conversion graph.
+
+We compare our solution to three baseline implementations: a naive checking of all path pairs with only special handling for cycles, a check for all path pairs that involve the new edge, and an algorithm suggested by previous work to solve the batch version of the problem for acyclic diagrams.
 
@@
 
-First, we formalize the notion of a diagram, drawing inspiration from Murota ~\cite{commutative}.
+We start by formalizing the notion of a diagram, drawing inspiration from the previous acyclic work by Murota ~\cite{commutative}.
 
@@
 
-We start with a directed graph G=(V,E) with sets attached to vertices and maps, to arcs. Each edge (u, v) in E corresponds to a function that maps elements of u to elements in v.
-All the functions form a semigroup F, with multiplication denoted by $\circ$.
-A semigroup is an algebraic structure with a set and an associative binary operation. We use it to represent functions in order to capture the one allowed operation- function composition.
+We start with a directed graph G=(V,E), where V consists of sets of elements and edges (u, v) in E correspond to functions that maps elements of u to elements in v.
+All these functions form a semigroup F, where multiplication is function composition.
+A semigroup consists of a set and an associative binary operation, which we use to capture function composition.
 
-The correspondence between edges and functions is stored as a mapping f:$E\rightarrow F$. f maps each edge to the function it represents.
+The correspondence between edges and functions is stored as a mapping f:$E\rightarrow F$, where f maps each edge to the function it represents.
 
-A path is a sequence of edges. The edge-to-function mapping f can be naturally extended to paths. If path p=$e_1\circ e_2 ... e_n$ then $f(p)=f(e_1) \circ f(e_2) ... f(e_n)$.
+A path is a sequence of edges. The edge-to-function mapping f can be naturally extended to paths: if path p=$e_1\circ e_2 ... e_n$ then $f(p)=f(e_1) \circ f(e_2) ... f(e_n)$.
 
-Let $P_{all}$ is the set of all paths in G.
+Let $P_{all}$ be the set of all paths in G.\xxx[dg]{Do we need this notation?  I don't see it anywhere.  Delete stuff that doesn't show up in the rest of the paper}
 
-The notation $\partial^{+}(p)$ is used to indicate the start node of path p, and similarly $\partial^-(p)$ is the end node of path p. $\partial(p)$ maps to the tuple (start node of p, end node of p).
+The notation $\partial^{+}(p)$ is used to indicate the start node of path p, and similarly $\partial^-(p)$ is the end node of path p.
+\xxx[dg]{Same for this notation.
+I see $\partial$ being useful, but not $\partial^-$} $\partial(p)$ maps to the tuple (start node of p, end node of p).
 
 A pair of paths p and q is called \textit{parallel} iff their terminal nodes are the same, i.e., $\partial(p)=\partial(q)$.
-The problem of checking if diagrams commute amounts to checking if the two members of every parallel pair that can be constructed is equal.
-
 Let $R_{all}$ be the set of all parallel pairs of paths in the diagram.
 
-The diagram commutes iff f(p)=f(q) $\forall (p,q)\in R_{all}$, which is to say, the compositions of the maps along any path connecting u to v is independent of choice of path.
-Here, equality of functions is defined by an equality oracle whose exact specifics will vary with the nature of the functions and the domain to which the diagram belongs.
+The diagram commutes iff f(p)=f(q) $\forall (p,q)\in R_{all}$; that is, the composition of maps along any path connecting u to v is independent of path choice.
+Here, equality of functions is defined by some oracle whose specifics will vary by target functions and domain.
 
-\section{Baselines}
+\section{Baseline Algorithms}
 
-\subsection{Naive all-path check}
-The naive solution is to verify that in every parallel pair in the diagram, the two paths are equal.
-However since cycles potentially introduce an infinite number of pairs, they must be dealt with separately.
+To examine the efficacy of our proposed algorithm, we compare it to some potential alternatives.
+Specifically, we examine a na\"{i}ve factorial algorithm, which we identify to be a two-flip tolerant path search, and Murota's historical batch solution~\cite{commutative}.
 
-Two situation can be imagined- the first, where the existence of the identity function from every node to itself is assumed as would describe most real world situations, and the second, where no such identity needs to exist.
+\subsection{Na\"{i}ve algorithm for online addition}
 
-When the identity is assumed, every cycle forms a parallel pair with an identity function on the terminal node.
-Therefore, the cycle must be verified to be equal to the identity on every node in the cycle.
-If this check succeeds for every elementary cycle in the diagram, then paths with multiple instance of the cycle must be equal to the version of the path without the cycle.
-It is therefore then enough to perform verification for all pairs where none of the path involves cycles.
-This is a finite set, bounded by $^{|E|!}C_2$.
-All elementary cycles in a diagram can be found in $O((|V|+|E|)(C+1))$ using Johnson's algorithm where C is the number of cycles, that can be greater than $2^{|V|}$.
-Thus the baseline is a factorial time algorithm that is nevertheless guaranteed to terminate.
-
-Without the identity assumed, a cycle c forms a parallel pair with itself repeated twice- the pair (c, c;c).
-This verification needs to be performed for all elementary cycles.
-Further, if there is any "exit" or "entrance" into an elementary cycle- an edge that is not a part of the cycle whose one end is a node that belongs to the cycle- then it must be verified that the transformation represented by the edge is equal to the transformation represented by the edge and cycle composed (the cycle is pre-pended in the case of an edge directed away from the cycle and appended in the case of an edge directed into the cycle).
-Should this verification succeed, by logic similar to the previous case, it must be that all paths (including complex cycles) that include the elementary cycle must be equal to version of the paths without the cycle (since a segment of the path with the cycle has already been verified to be equal to a version of that segment without the cycle).
-Once again this is a factorial time, terminating algorithm.
-
-\subsection{Naive algorithm for online addition}
-
-Since we solve the question of online addition, where the original diagram is path independent and a new edge is added, the naive baseline can be improved to verify only parallel pairs that have been affected by the addition of the new edge.
+We solve the question of online addition, where the original diagram is path independent and a new edge is added.
+Any implementation can verify only parallel pairs that have been affected by the addition of the new edge.
 Pairs that do not involve the new edge must continue to be path independent because they already belonged to the old diagram.
@@
 
-a "two-flip tolerant" path search could be used to identify the pairs of paths where exactly one path includes the new edge.
+\subsubsection{Two flip tolerant path search}
+We use a "two-flip tolerant" path search to identify the pairs of paths where exactly one path includes the new edge.
 
-\subsubsection{Two flip tolerant path search}
 In a normal directed graph path, only forward edges, i.e. edges that go outward from the current node, are considered. 
-On the other hand, the \textit{two flip path} goes through up to three phases: in the first phase, only backward edges- pointing inward to the current node- are accepted.
+A \textit{two flip path} consists of up to three phases: in the first phase, only backward edges- pointing inward to the current node- are accepted.
 In the second phase, only forward edges are accepted, and in the third phase, again only backward edges are accepted.
-As a consequence, there is a node in the path for which both path edges are oriented outwards, that lies between the first and second phase.
-This will be referred to as the \textit{first flipping point}.
-Similarly, there is a \textit{second flipping point} between the second and third phase, for which both path edges point inwards.
+The node between the first two phases we refer to as the \textit{first flipping point}, which has both edges pointing outward; similarly, we refer to the node between the latter two phases as the \textit{second flipping point}, at which both edges point inwards.
 

### 2020-06-08 currency-exchange/tester.py

@@ -3,27 +3,33 @@ from noIdentityPolynomialPathChecker import NoIdentityPolynomialPathChecker
 from optimalSetPathChecker import OptimalSetPathChecker
 from PolynomialPathChecker import PolynomialPathChecker
 
-graph = [[-1, -1, -1], 
-         [1, -1, -1], 
-         [-1, -1, -1]]
+graph = [[-1, -1, 1, 1, -1], 
+         [-1, -1, 1, 1, -1], 
+         [1, -1, -1, -1, -1],
+         [-1, -1, -1, -1, -1],
+         [-1, -1, -1, -1, -1]]
+
+# graph = [[-1, -1, 1],
+#          [-1, -1, 1],
+#          [1, -1, -1]]
 
 poly = PolynomialPathChecker()
 
 poly.setGraph(graph)
 poly.setEdge(0, 1)
-print("polynomial:\n" + str(poly.getPathsToCheck()))
+print("polynomial:\n" + str(poly.getPathsToCheck()) + "\n\n")
 
 opti = OptimalSetPathChecker()
 
 opti.setGraph(graph)
 opti.setEdge(0, 1)
-print("optimal:\n" + str(opti.getPathsToCheck()))
+print("optimal:\n" + str(opti.getPathsToCheck()) + "\n\n")
 
 polyid = NoIdentityPolynomialPathChecker()
 
 polyid.setGraph(graph)
 polyid.setEdge(0, 1)
-print("polynomial noid:\n" + str(polyid.getPathsToCheck()))
+print("polynomial noid:\n" + str(polyid.getPathsToCheck()) + "\n\n")
 
 optiid = NoIdentityOptimalSetPathChecker()
 

### 2020-06-08 acmPaper/paper.tex

@@
 \newcount\draft\draft=1 % set to 0 for submission or publication
-\documentclass[sigconf,authordraft]{acmart}
+\documentclass[sigplan,review,anonymous]{acmart}
+\setcopyright{none}
 %

### 2020-06-09 acmPaper/paper.tex
Updates to section 3

@@
 
-% A lot of this exposition is from the cited paper, I feel like I need to make this clearer to give proper credit, but not sure how.
+\xxx[AK]{A lot of this exposition is from the cited paper, I feel like I need to make this clearer to give proper credit, but not sure how}
 
@@
 
-We solve the question of online addition, where the original diagram is path independent and a new edge is added.
-Any implementation can verify only parallel pairs that have been affected by the addition of the new edge.
+We start with a reference baseline solution for online addition, where the original diagram is path independent and a new edge is added.
 Pairs that do not involve the new edge must continue to be path independent because they already belonged to the old diagram.
+Any implementation can thus verify only parallel pairs that have been affected by the addition of the new edge.
+
+That the original diagram commutes can be used to narrow down the set of pairs to verify.
+Cases of pairs where either both paths or neither path includes the new edge must both remain equal with straightforward reasoning.
 
@@
 When the identity is not assumed, cycles that include one of the nodes of the new edge must be checked for the new exit or entry point that the new edge becomes (if the cycle includes the source of the new edge then the cycle followed by the new edge must be equal to the new edge alone and if the cycle include the sink then the new edge followed by the cycle must be equal to the new edge).
-Additionally, all cycles that include the new edge would need full verification like in the batch algorithm- the cycle would have to be equal to itself composed with itself, and would have to be verified for all its entry and exit points.
+Additionally, all cycles that include the new edge would need full verification like in the batch algorithm- the cycle would have to be equal to itself composed with itself, and would have to be verified for all its entry and exit points.\xxx[dg]{This paragraph needs to be rewritten to make cycles clear without the identity stuff.
+I'm not sure how to do it; I'll come back to it later}
 
-Now, the commutativity of the original diagram could be used to narrow down the set of pairs to verify.
-Pairs where both paths do not involve the new edge would remain equal, but also, pairs where both paths involved the new edge would have to be equal.
-To see why this is true, each path could be thought of as consisting of the composition of three segments, the first segment from the common source to the source of the new edge, the second, the new edge itself, and the third, from the sink of the new edge to the common sink of the parallel pair.
-The first segment of both pairs would have to be equal because they existed in as parallel pairs in the original diagram, and similarly the third segments would also have to be equal.
-The second segments, consisting of the same edge, would also have to be equal.
-Therefore the composition of these three equal components would be equal.
-Note that the new edge could only appear once because cycles have already been dealt with and none of the paths under consideration hit a node twice.
 Thus we are left only to verify the pairs where exactly one of the paths includes the new edge.
 
-\subsubsection{Two flip tolerant path search}
+\paragraph{Two flip tolerant path search}
 We use a "two-flip tolerant" path search to identify the pairs of paths where exactly one path includes the new edge.
@@
 The node between the first two phases we refer to as the \textit{first flipping point}, which has both edges pointing outward; similarly, we refer to the node between the latter two phases as the \textit{second flipping point}, at which both edges point inwards.
-
-Consider the situation where a new edge from source node S to sink T is being added to a diagram.
-A two flip tolerant path from S to T would correspond to a new pair of parallel paths- flipping point 1 would be the source of both paths, and flipping point 2, the sink.
-The first path would correspond to the second phase of the flip tolerant path search, and the second path, involving the new edge, would correspond to composing the first phase of the two flip path, the new edge, and the third phase.
 
@@
 Either the path is oriented from S to T, in which case the conflicting paths are simply the edge (S, T) and the entire flip-less path, or the path is oriented from T to S.
-In this case, we have found a cycle that has already been dealt with.
+In this case, we have found a cycle, which has already been handled.
 
-\paragraph{Theorem} The result of the two-flip tolerant path search from the source to the sink node of an edge that is to be newly added, followed with the path extraction algorithm has a one-to-one correspondence to the set of new parallel pairs with exactly one path passing through the new edge, with neither paths containing any cycles.
+\paragraph{Theorem} Consider the result of two-flip tolerant path search from the source to sink node of an edge that is to be newly added followed with the path extraction algorithm described.
+This result has a one-to-one correspondence to the set of new parallel pairs with exactly one path passing through the new edge and neither paths containing any cycles.
 
@@
 \paragraph{Implementation}
-*Copy in python code from implementation? Appendix?*
+\xxx[AK]{Copy in python code from implementation? Appendix?}
+\xxx[DG]{Not sure we need this paragraph, the implementation is pretty clear from the description of the algorithm}
 
@@
 
-\subsection{Optimal solution for batch problem}
+\subsection{Optimal batch solution}
 Murota's main result ~\cite{commutative} is a solution to the question of whether a given acyclic diagram commutes that returns the minimal number of equality checks.
@@
 
-The approach in the algorithm, at a high level, is to define a function that takes in a subset of pairs and returns the subset of pairs whose verification is implied by the verification of the input set.
+\xxx[dg]{Should be a figure}
+The approach in this algorithm, at a high level, is to define a function that takes in a subset of pairs and returns the subset of pairs whose verification is implied by verifying the input set.
 Then, greedy elimination of redundancies is performed until a minimal set is reached.
@@
 Define $r_1>r_2$ for bilinkings $r_1$ = $\{p_l,q_l\}$, $r_2$= $\{p_2, q_2)$ $\in$ $R_0$, if there exists a path p such that $\partial p=\partial r_1$ and p contains $p_2$.
-The $\langle\rangle$ function is defined as:
+Define $\langle\rangle$ as:
 $\langle r \rangle = \{ s\in R_0| r>s\}$.
 
-For bilinking s, let F(s) be the vector in $\mathbb{F}2^{|E|}$ representing the edges present in s (there is a dimension corresponding to each edge, the $n^{th}$ dimension of F(s) is 1 if the corresponding edge is in s, and 0 otherwise).
-Let this function be extended to sets, so that for some set of bilinkings S, F(S) = $\{ F(s) | s\in S \}$.
+For bilinking $s$, let $F(s)$ be the vector in $\mathbb{F}2^{|E|}$ representing the edges present in s (the $n^{th}$ dimension of $F(s)$ is 1 if the corresponding edge is in $s$, and 0 otherwise).
+Let this function be extended to sets, so that for some set of bilinkings $S, F(S) = \{ F(s) | s\in S \}$.
 
-for a set of bilinkings r, the closure function cl is defined as:
-cl(r) = $\{ s\in R_0|$ s is linearly dependent on $F(r) \}$.
+for a set of bilinkings $r$, the closure function $cl$ is defined as:
+\xxx[dg]{I think this definition is incomplete}
+$cl(r) = \{ s\in R_0| s$ is linearly dependent on $F(r) \}$.
 
-The closure function on r basically captures all the pairs that can be made by made by composing or "gluing together" the bilinkings in r. 
-The aforementioned $\langle\rangle$ function captures the orientation of paths, and can be used to constrain compositions to be in the right direction.
+The closure function on $r$ basically captures all the pairs that can be made by made by composing or "gluing together" the bilinkings in r.
 
@@
 
-$\sigma$ is used to iteratively check if a given pair is redundant. bilinkings are eliminated until a minimum "spanning" subset is reached.
+$\sigma$ is used to iteratively check if a given pair is redundant. Bilinkings are eliminated until a minimum "spanning" subset is reached.
 
-An implementation of the described algorithm is
+A complete implementation of the described algorithm is included in the appendix.
 
+\xxx[dg]{Probably put this in the appendix}
+\xxx[dg]{Give a summary of the algorithm though, breaking down important parts.  I think that's mostly been done above, but I'm not sure}
 \begin{verbatim}

### 2020-06-10 acmPaper/paper.tex
Added currency checker changes, data

@@
 \usepackage{amssymb}
-\usepackage{algorithm}
+\usepackage[norelsize, section]{algorithm2e}
 \usepackage{xxx}
@@
 \end{verbatim}
+\begin{algorithm}
+    \caption{Polynomial time algorithm}
+    \label{polynomial}
+    \KwData{placeholder}
+    \KwResult{placeholder}
+\end{algorithm}
 
@@
 
-Note that the proof is not affected if an of these paths is the identity, eg. if $f_1$ is the identity and S and $P_1$ are actually the same node.
+Note that the proof is not affected if any of these paths is the identity, eg. if $f_1$ is the identity and S and $P_1$ are actually the same node.
 
@@
 \end{verbatim}
+\begin{algorithm}
+    \caption{Optimal subset algorithm}
+    \label{optimal}
+    \KwData{placeholder}
+    \KwResult{placeholder}
+\end{algorithm}
 
@@
 
-To demonstrate our algorithm applied to a real world situation, 
+To demonstrate our algorithms applied to a real world situation, 
 we found inconsistencies in a diagram of the exchange rate between currencies.
@@
 
-Using an api, https://exchangeratesapi.io, we built a fully connected diagram for 32 currencies.
-To ensure that it indeed commuted, before the addition of each edge, we used the algorithms presented
-to ensure the addition of a new edge did not introduce inconsistencies in the existing diagram.
-If a new edge was, in fact, problematic, the algorithm returned an example inconsistent
+Using an API ~\footnote{https://exchangeratesapi.io}, we built the fully connected diagram of exchange rates between 32 currencies on a given day.
+To ensure that it indeed commuted, we started with an empty diagram, and added in edges one by one.
+Before the addition of each edge, we used the algorithms (~\ref{polynomial} and ~\ref{optimal}) to ensure the addition of a new edge did not introduce inconsistencies in the existing diagram.
+If a new edge was, in fact, problematic, the algorithms returned an example inconsistent
 pair that would arise from the addition of the edge.
@@
 We allowed an "error tolerance" for floating point error.
-Graphs were generally consistent up to an error tolerance to the order of magnitude $10^{-6}$.
+Graphs were generally found to be consistent up to an error tolerance to the order of magnitude $10^{-6}$.
 However, below this threshold we found inconsistencies.
-For both algorithms, computation completed in minutes.
+Averaging over evaluation for the first 9 days of 2020, verifying and building a graph to completion (inclusive of the time required by network calls) took 223$\pm$34 seconds using algorithm ~\ref{optimal}, and in 128$\pm$7 seconds with algorithm ~\ref{polynomial}.
+Computation was performed on a Macbook Pro 2015, 2.9 GHz Dual-Core Intel Core i5.
+More information about implementation can be found in the evaluation section.
 

### 2020-06-10 currency-exchange/currencyGraph.py

@@ -2,8 +2,11 @@ from colorama import Fore
 import requests 
 import numpy as np
 import copy
+from datetime import datetime
 from optimalSetPathChecker import OptimalSetPathChecker
 from testUtilities import TestDefinition, defaultTestSuite
+import time
+import traceback
 from typing import List
 
 class CurrencyGraph():
@@ -13,14 +16,21 @@ class CurrencyGraph():
     INFTY = 1000000
     NO_EDGE = -1
     DEBUG = False
-    endPoint = "https://api.exchangeratesapi.io/latest?"
+    baseEndPoint = "https://api.exchangeratesapi.io/"
+    endPoint = baseEndPoint + "latest?"
     
     def __init__(self):
         self.graph = []
         self.currencyList = []
         self.base = 'USD' # Default but can change
+        self.sleepTime = 0
         np.set_printoptions(formatter={
             'float': lambda x: "{0:0.4f}".format(x)}, threshold=10000)
+    
+    def setupWithDay(self, year, month, day):
+        date = datetime(year, month, day)
+        self.endPoint = self.baseEndPoint + date.strftime("%Y-%m-%d") + "?"
+        self.setup()
     
     def setup(self):
         """Setup graph with initial row of data from base currency"""
@@ -87,7 +97,7 @@ class CurrencyGraph():
         if base == 'EUR' and symbols == ['EUR']:
             return { 'EUR' : 1 }
         # sending get request and saving the response as response object 
-        r = requests.get(url = self.makeUrl(base, symbols))   
+        r = requests.get(url = self.makeUrl(base, symbols), timeout = 1)
         # extracting data in json format 
         data = r.json() 
         if (self.DEBUG): 
@@ -100,11 +110,15 @@ class CurrencyGraph():
                 print(rate)
         return rates
     
-    def addEntry(self, base, target, check = False, retries = 3):
+    def addEntry(self, base, target, check = False, retries = 10):
         try:
             rates = self.makeRequest(base, [target])
-        except:
+        except Exception as e:
+            if (self.DEBUG):
+                print(e)
             if retries > 0:
+                self.sleepTime += 2 # 1 second because of the request timeout.
+                time.sleep(1)
                 return self.addEntry(base, target, check, retries-1)
             else:
                 raise Exception("api call failed")

### 2020-06-10 currency-exchange/main.py

@@ -5,7 +5,9 @@ import numpy as np
 from optimalSetPathChecker import OptimalSetPathChecker
 from iPathChecker import IPathChecker
 from polynomialPathChecker import PolynomialPathChecker
-import requests 
+import requests
+import time
+import traceback
 from typing import List
 
 # Assume that a graph doesn't have more edges nodes than this
@@ -20,7 +22,7 @@ ID_FUNCTION = [-1]
 
 ###############################################
 
-epsilon = 1E-7 # allowed error multiplier
+epsilon = 1E-2 # allowed error multiplier
 
 checkers = [
     OptimalSetPathChecker(),
@@ -52,7 +54,7 @@ def independenceFuncFactory(checker: IPathChecker, graph: CurrencyGraph):
         for (path1, path2) in pairsToCheck:
             path1Value = getPathValue(path1, newGraph)
             path2Value = getPathValue(path2, newGraph)
-            if abs(path1Value - path2Value) > epsilon * path2Value:
+            if abs(path1Value - path2Value) > epsilon:
                 return (False, "path 1: {}: {} \npath 2: {}: {} \n difference: {}, new edge: {} -> {}: rate: {}"
                 .format(
                 path1Value,
@@ -66,34 +68,62 @@ def independenceFuncFactory(checker: IPathChecker, graph: CurrencyGraph):
         return (True, "Success for base {} target {}".format(base, target))
     return IndependenceFunction
             
+totalCheckerTimes = dict() # checker -> (total seconds, number of runs)
 
-for checker in checkers:
-    print("Working with {}".format(checker.__class__.__name__))
-    graph = CurrencyGraph()
-    graph.setup()
-    graph.checkIndependenceFunc = independenceFuncFactory(checker, graph)
-    try:    
-        for baseCurrency in graph.currencyList:
-            if baseCurrency == graph.base:
-                continue
-            for targetCurrency in graph.currencyList:
-                if (DEBUG):
-                    print("Adding edge from {} to {}"
-                        .format(baseCurrency, targetCurrency))
-                (success, info) = graph.addEntry(baseCurrency, targetCurrency, True)
-                if not success:
-                    print(Fore.RED + "Adding edge from {} to {} failed for checker {}."
-                        .format(
-                            baseCurrency,
-                            targetCurrency,
-                            checker.__class__.__name__))
-                    print(info)
-                    raise StopIteration
-        print(Fore.GREEN + "Completed graph for {}"
-        .format(checker.__class__.__name__))
-    except:
[…]

### 2020-06-11 acmPaper/paper.tex
better evaluation for currency graph.

@@
 However, below this threshold we found inconsistencies.
-Averaging over evaluation for the first 9 days of 2020, verifying and building a graph to completion (inclusive of the time required by network calls) took 223$\pm$34 seconds using algorithm ~\ref{optimal}, and in 128$\pm$7 seconds with algorithm ~\ref{polynomial}.
+Averaging over evaluation for the first 30 days of 2020, building and verifying a diagram to completion (inclusive of the time required by network calls) took 243$\pm$19 seconds using algorithm ~\ref{optimal}, and in 133$\pm$13 seconds with algorithm ~\ref{polynomial}.
 Computation was performed on a Macbook Pro 2015, 2.9 GHz Dual-Core Intel Core i5.

### 2020-06-11 currency-exchange/main.py

@@ -70,7 +70,7 @@ def independenceFuncFactory(checker: IPathChecker, graph: CurrencyGraph):
             
 totalCheckerTimes = dict() # checker -> (total seconds, number of runs)
 
-for day in range(1, 10):
+for day in range(1, 31):
     month = 1
     year = 2020
     if (DEBUG):

### 2020-06-17 acmPaper/paper.tex
edits

@@
 \section{Introduction}
-Many systems can be understood as graphs: a set of objects with directed functions between members.
-With programming, for example, type systems with coercions can be drawn as a graph with type nodes and coercion edges.
-A desirable trait of such a type system and many other similar systems is that traversing the graph from a given start and end node results in the same value independent of path chosen; a property equivalent to saying the graph diagram is commutative.
+Many systems can be understood as diagrams: graphs where nodes represent objects, and edges, directed transformation functions between nodes.
+In programming, for example, type systems with coercions can be drawn as a graph with type nodes and coercion edges.
+A desirable trait of such a type system and many other similar systems is that traversing the graph from a given start and end node results in the same value independent of path chosen.
+A diagram with this property is said to commute.
 With our type coercion example, it could be problematic if casting to a supposedly equivalent type as an intermediate step resulted in a different answer than a direct cast.
 
-Given the utility of system diagram commutativity, it would be helpful to check that the property holds for real systems.
+Given the utility of system diagram commutativity, it would be useful to check that the property holds for real systems.
 Since diagrams may change over time, and verifying the entire system from scratch may be computationally expensive, it would be additionally convenient if new edges could be checked online as they are added.
 
-Naively checking if all paths in a given graph are path independent could require a number of function equality checks as bad as factorial in the number of nodes, since a path consists of an ordering of nodes.
-Previous work ~\cite{commutative} has identified an $O(|E|^2|V|^4)$ algorithm to verify that a complete acyclic diagram commutes; however, this solution neither manages the batch case nor the existence of cycles.
-Unfortunately, the presence of cycles implies a potentially infinite number of paths.
+Naively checking if all paths in a given diagram are path independent could require a number of function equality checks as bad as factorial in the number of nodes, since a path consists of an ordering of nodes.
+Further, the presence of cycles implies a potentially infinite number of paths.
 
-We present a polynomial time $O(|V|(|E|+|V|)$ algorithm to verify that commutativity of a graph (potentially with cycles) is maintained over the course of online addition, assuming an oracle to check the abstract notion of equality of functions.
-We show the number of calls to the equality checking oracle is asymptotically tight.
-Should the calls to the oracle be expensive, we also show, with an additional $O(|V|^4)$ optimization step, an algorithm to use instead the minimal possible number of equality checks.
+In order to verify commutativity efficiently over online addition, we identified two key insights.
+The first was that in a commutative diagram, when a new edge is to be added, at most one path needs to be checked against for a given source and sink pair.
+Because the diagram commutes, all the paths between a given source and sink are equal and a representative to check against can arbitrarily be chosen.
+It lead to an $O(|V|^2(|E|+|V|)$ algorithm to verify a diagram remains commutative over the course of online addition, assuming an oracle to check the equality of functions.
+The algorithm makes an asymptotically optimal number of calls to the oracle.
+
+The second insight was that there is a single rule that places a partial, transitive ordering on the informativeness of paths.
+It allowed for the creation of a greedy, $O(|V|^4)$ optimization step, that results in the number of calls to the oracle being minimal.
+The optimization is very useful when the equality checking oracle is expensive.
+
+Previous work ~\cite{commutative} has identified an $O(|E|^2|V|^4)$ algorithm to verify that a complete acyclic diagram commutes; however, it addresses neither online addition nor cyclic diagrams.
 

### 2020-06-17 acmPaper/paper.tex

@@
 
-In order to verify commutativity efficiently over online addition, we identified two key insights.
+In the course of efficiently verifying commutativity over online addition, we identified two key insights.
 The first was that in a commutative diagram, when a new edge is to be added, at most one path needs to be checked against for a given source and sink pair.
@@
 
-We allowed an "error tolerance" for floating point error.
-Graphs were generally found to be consistent up to an error tolerance to the order of magnitude $10^{-6}$.
-However, below this threshold we found inconsistencies.
+We allowed an "error tolerance", so that differences reported would not be the trivial consequences of a floating point error.
+However, this relaxation of the equality oracle into imprecision meant that the mathematical reasoning that allow the algorithms to remove redundant path checks no longer applied.
+For instance, composing a new function with two approximately equal functions does not lead to equal results, so theorem ~\ref{reductionRule} fails with this approximate equality.
+When the algorithms reported no inconsistencies, it was still possible that the graph possessed inconsistencies above the given threshold and did not commute.
+Nonetheless, both algorithms were effective in catching inconsistencies.
+Algorithm ~\ref{polynomial} started finding inconsistencies at error tolerances to the order of $10^{-3}$, and ~\ref{optimal}, which makes more invalid redundant path check removals, at error tolerances to the order of $10^{-7}$.
+
 Averaging over evaluation for the first 30 days of 2020, building and verifying a diagram to completion (inclusive of the time required by network calls) took 243$\pm$19 seconds using algorithm ~\ref{optimal}, and in 133$\pm$13 seconds with algorithm ~\ref{polynomial}.

### 2020-06-21 acmPaper/paper.tex
Small updates

--- 2020-06-11 acmPaper/paper.tex
@@
 \usepackage{amssymb}
-\usepackage[norelsize, section]{algorithm2e}
+\usepackage[norelsize]{algorithm2e}
 \usepackage{xxx}
@@
 \xxx[dg]{Probably put this in the appendix}
-\xxx[dg]{Give a summary of the algorithm though, breaking down important parts.  I think that's mostly been done above, but I'm not sure}
+\xxx[dg]{Give a summary of the algorithm though, breaking down important parts.  I think that's mostly been done above, but some actual code snippets might help}
 \begin{verbatim}
@@
 
+\xxx[dg]{Insert an introduction paragraph here about what we're solving before jumping into the details}
+
+\xxx[dg]{As we've discussed, I think the existence of an identity function doesn't have to be assumed, since every node "exists" as itself.
+I guess another way to put it is that we assume the 0-edge path exists as the "identity" function}
 The solution presented here assumes that there is an implicit identity function on every node, as would be the case in most applications.
@@
 Like with the online baseline, we do not concern ourselves with parallel pairs where neither or both paths pass through the new edge.
-The key observation that allows us to improve on the online baseline is that for a given (source, sink) pair only a single parallel pair needs to be verified.
+The key observation that allows us to improve on the online baseline is that, for a given source and sink pair, only a single parallel pair needs to be verified. 
 This is an implication of Theorem \ref{reductionRule}, expanded on later.

### 2020-06-30 acmPaper/paper.tex
Use BibTeX!
3.1

@@
 
-Let $P_{all}$ be the set of all paths in G.\xxx[dg]{Do we need this notation?  I don't see it anywhere.  Delete stuff that doesn't show up in the rest of the paper}
-
-The notation $\partial^{+}(p)$ is used to indicate the start node of path p, and similarly $\partial^-(p)$ is the end node of path p.
-\xxx[dg]{Same for this notation.
-I see $\partial$ being useful, but not $\partial^-$} $\partial(p)$ maps to the tuple (start node of p, end node of p).
+The notation $\partial(p)$ maps to the tuple (start node of p, end node of p).
 
@@
 The diagram commutes iff f(p)=f(q) $\forall (p,q)\in R_{all}$; that is, the composition of maps along any path connecting u to v is independent of path choice.
-Here, equality of functions is defined by some oracle whose specifics will vary by target functions and domain.
+
+The \textsc{Online addition problem}, given a commuting diagram and a new edge, returns whether the diagram commutes.
+We make an abstraction - equality of functions is defined by some oracle whose specifics will vary by target functions and domain - and reduce the \textsc{Online addition problem} to the \textsc{Verification set problem}, which given a diagram and a new edge, returns the set of parallel pairs of paths, such that if and only if the members in each pair are equal, then the new graph must commute.
+The output to the \textsc{Online addition problem} can then be obtained as whether equality checking for all pairs succeeds.
 
@@
 
-To examine the efficacy of our proposed algorithm, we compare it to some potential alternatives.
-Specifically, we examine a na\"{i}ve factorial algorithm, which we identify to be a two-flip tolerant path search, and Murota's historical batch solution~\cite{commutative}.
+To examine the efficacy of our proposed solution to the \textsc{Verification set problem}, we compare it to some potential alternatives.
+Specifically, we examine a na\"{i}ve factorial algorithm, a slightly less na\"{i}ve factorial algorithm which we identify to be a two-flip tolerant path search, and Murota's historical batch solution~\cite{commutative}.
 
-\subsection{Na\"{i}ve algorithm for online addition}
+\subsection{Very Na\"{i}ve baseline}
 
-We start with a reference baseline solution for online addition, where the original diagram is path independent and a new edge is added.
-Pairs that do not involve the new edge must continue to be path independent because they already belonged to the old diagram.
-Any implementation can thus verify only parallel pairs that have been affected by the addition of the new edge.
+We start with the set of all parallel pairs in the diagram (with the new edge added in), and pare it down to be finite by handling cycles.
+We find all simple cycles in the diagram, using a procedure like Johnson's algorithm. 
+%TODO cite the paper.
+We verify for each cycle that a single traversal starting on any node is equal to the identity function by adding pairs of (cycle starting at node, identity function) for each node in the function to the output verification set.
+\xxx[AK]{Need cleaner way to put this.}
+As cycles must now be the identity, for any pair in the set of all parallel paths, any instance of a cycle can be removed to obtain an equivalent pair with shorter, cycle free paths.
+If the shorter pair has equal paths then the paths in the original pair must also be equal to each other.
+It is therefore safe to remove all pairs of paths with cycles, leaving only parallel pairs where neither path has a cycle.
+Since there are only a finite number of paths in the graph with no cycles, (to the order of n!
+as a path is an ordering on nodes), and the parallel pairs are a subset of the possibilities generated by choosing a first and second member from this finite set, the resultant set of all parallel pairs must be finite.
+%TODO Check if O(n!n) = O(n). What about O((n!)^2)?
+The output of the algorithm, then, is the union of the set of cycle verification pairs, and the parallel pairs found by doing an all-acyclic-path search from every node in the diagram with the new edge added.
 
-That the original diagram commutes can be used to narrow down the set of pairs to verify.
-Cases of pairs where either both paths or neither path includes the new edge must both remain equal with straightforward reasoning.
-
-The cycle checking steps can also be improved upon.
-In the case where the identity is assumed, only elementary cycles passing through the new edge need to be verified to be equal to the identity.
-When the identity is not assumed, cycles that include one of the nodes of the new edge must be checked for the new exit or entry point that the new edge becomes (if the cycle includes the source of the new edge then the cycle followed by the new edge must be equal to the new edge alone and if the cycle include the sink then the new edge followed by the cycle must be equal to the new edge).
-Additionally, all cycles that include the new edge would need full verification like in the batch algorithm- the cycle would have to be equal to itself composed with itself, and would have to be verified for all its entry and exit points.\xxx[dg]{This paragraph needs to be rewritten to make cycles clear without the identity stuff.
-I'm not sure how to do it; I'll come back to it later}
-
+\subsection{Less Na\"{i}ve baseline}
+We further refine the output of the very na\"{i}ve baseline.
+The commutativity of the original diagram could be used to narrow down the set of pairs to verify.
+Pairs where both paths do not involve the new edge would remain equal (this would apply to cycles too).
+Also, pairs where both paths involved the new edge would have to be equal.
+To see why this is true, each path could be thought of as consisting of the composition of three segments, the first segment from the common source to the source of the new edge, the second, the new edge itself, and the third, from the sink of the new edge to the common sink of the parallel pair.
+The first segment of both pairs would have to be equal because they existed in as parallel pairs in the original diagram, and similarly the third segments would also have to be equal.
+The second segments, consisting of the same edge, would also have to be equal.
+Therefore the composition of these three equal components would be equal.
+Note that the new edge could only appear once because cycles have already been dealt with and none of the paths under consideration hit a node twice.
 Thus we are left only to verify the pairs where exactly one of the paths includes the new edge.
+Further, only pairs where the path includes the new edge once need be checked (lemma ~\ref{one_occurence_lemma}).
 
@@
 
-\xxx[dg]{Insert an introduction paragraph here about what we're solving before jumping into the details}
-
-\xxx[dg]{As we've discussed, I think the existence of an identity function doesn't have to be assumed, since every node "exists" as itself.
-I guess another way to put it is that we assume the 0-edge path exists as the "identity" function}
-The solution presented here assumes that there is an implicit identity function on every node, as would be the case in most applications.
-However, a modified version of the algorithm for diagrams without this assumption is presented in the appendix.
-The modification is similar in spirit to what was described in the description of the naive baselines.
-
-Like with the online baseline, we do not concern ourselves with parallel pairs where neither or both paths pass through the new edge.
+We present a polynomial time solution to the \textsc{verification set} problem.
+Like with the baseline, we do not concern ourselves with parallel pairs where neither or both paths pass through the new edge.
 The key observation that allows us to improve on the online baseline is that, for a given source and sink pair, only a single parallel pair needs to be verified. 
@@
 
-An efficient implementation of the algorithm can run in $O(|V|(|V|+|E|))$ time at a $O(|V|^3)$ space cost.
-In such an implementation, path finding from a given source node to all potential sink nodes would be done in a single $O(|V|+|E|)$ breadth first search and the results memoized in $O(|V|^2)$ space.
+An efficient implementation of the algorithm can run in $O(|V|^2(|V|+|E|))$ time at a $O(|V|^3)$ space cost.
+In such an implementation, path finding from a given source node to all potential sink nodes could be done in a single $O(|V|+|E|)$ breadth first search and the results memoized in $O(|V|^2)$ space.
 
@@
 
-%
-% ---- Bibliography ----
-%
-% BibTeX users should specify bibliography style 'splncs04'.
-% References will then be sorted and formatted in the correct style.
-%
-% \bibliographystyle{splncs04}
-% \bibliography{mybibliography}
-%
-\begin{thebibliography}{8}
+\bibliography{refs}
+\bibliographystyle{plain}
 
-\bibitem{commutative} 
-Kazuo Murota: {HOMOTOPY BASE OF ACYCLIC GRAPHS - A COMBINATORIAL ANALYSIS OF COMMUTATIVE BY MEANS OF PREORDERED MATROID}. Discrete Applied Mathematics, vol. 17, Issues 1--2, 135--155 (1987).
-
-\end{thebibliography}
 \end{document}

### 2020-06-30 acmPaper/paper.tex
cleanup i think

@@
 We find all simple cycles in the diagram, using a procedure like Johnson's algorithm. 
-%TODO cite the paper.
+\xxx[AK]{TODO cite the paper.}
 We verify for each cycle that a single traversal starting on any node is equal to the identity function by adding pairs of (cycle starting at node, identity function) for each node in the function to the output verification set.
@@
 as a path is an ordering on nodes), and the parallel pairs are a subset of the possibilities generated by choosing a first and second member from this finite set, the resultant set of all parallel pairs must be finite.
-%TODO Check if O(n!n) = O(n). What about O((n!)^2)?
+\xxx[AK]{TODO Check if O(n!n) = O(n). What about $O((n!)^2)$?}
 The output of the algorithm, then, is the union of the set of cycle verification pairs, and the parallel pairs found by doing an all-acyclic-path search from every node in the diagram with the new edge added.
@@
 
-\paragraph{Implementation}
-\xxx[AK]{Copy in python code from implementation? Appendix?}
-\xxx[DG]{Not sure we need this paragraph, the implementation is pretty clear from the description of the algorithm}
-
+\paragraph{Analysis}
 This algorithm is also in the worst case as bad as factorial in the number of nodes.
@@
 \subsection{Optimal batch solution}
-Murota's main result ~\cite{commutative} is a solution to the question of whether a given acyclic diagram commutes that returns the minimal number of equality checks.
+Murota's main result ~\cite{commutative} is a solution to a batch version of \textsc{verification set} - given an acyclic diagram, it returns the minimal set of equality checks that would succeed if and only if the diagram commutes.
 The paper describes the following algorithm to find the ($V^2E$ bounded) minimal set of pairs that needs to be checked.
@@
 For bilinking $s$, let $F(s)$ be the vector in $\mathbb{F}2^{|E|}$ representing the edges present in s (the $n^{th}$ dimension of $F(s)$ is 1 if the corresponding edge is in $s$, and 0 otherwise).
-Let this function be extended to sets, so that for some set of bilinkings $S, F(S) = \{ F(s) | s\in S \}$.
+Let this function be extended to sets, so that for some set of bilinkings $S, F(S) = \{ F(s) | s\in S \}$. A notion of linear independence in this vector field exists.
 
-for a set of bilinkings $r$, the closure function $cl$ is defined as:
-\xxx[dg]{I think this definition is incomplete}
+For a set of bilinkings $r$, the closure function $cl$ is defined as:
+\xxx[dg]{I think this definition is incomplete}\xxx[ak]{sorry, what is it missing?}
 $cl(r) = \{ s\in R_0| s$ is linearly dependent on $F(r) \}$.

### 2020-06-30 implementation/evaluator.py

@@ -10,9 +10,9 @@ from typing import List
 
 NUM_TRIES = 1
 
-densityStep = 0.1
+densityStep = 0.2
 sizeStep = 10
-maxSize = 1000
+maxSize = 100
 
 densities = [i*densityStep for i in range(1, int(1.0/densityStep))] #[0.1, 0.5, 0.7, 0.8]
 sizes = [i*sizeStep for i in range(1, int(maxSize/sizeStep))]
@@ -68,7 +68,7 @@ def plot(checkerName, results):
 
     fig = plt.figure()
     ax = plt.axes(projection='3d')
-    ax.contour3D(X, Y, Z, 50, cmap='binary')
+    ax.contour3D(X, Y, Z, 50, zdir='z')
     ax.set_xlabel('Density')
     ax.set_ylabel('Size')
     ax.set_zlabel('time')

### 2020-07-01 implementation/evaluator.py
Added execution time vs size graphs

@@ -6,16 +6,19 @@ from pathCheckers.iPathChecker import IPathChecker
 from pathCheckers.optimalSetPathChecker import OptimalSetPathChecker
 from pathCheckers.polynomialPathChecker import PolynomialPathChecker
 from randomGraphGenerator import generateGraph
+import traceback
 from typing import List
 
-NUM_TRIES = 1
+NUM_TRIES = 10
 
 densityStep = 0.2
 sizeStep = 10
 maxSize = 100
 
-densities = [i*densityStep for i in range(1, int(1.0/densityStep))] #[0.1, 0.5, 0.7, 0.8]
+# densities = [i*densityStep for i in range(1, int(1.0/densityStep))] 
+densities = [0.1, 0.5, 0.9]
 sizes = [i*sizeStep for i in range(1, int(maxSize/sizeStep))]
+# sizes = [10, 50, 100]
 
 evaluationList = [(density, size) for size in sizes for density in densities]
 
@@ -60,9 +63,8 @@ def getEvaluateFunction(checker: IPathChecker, numTries: int):
         return info
     return evaluate
 
-
-
-def plot(checkerName, results):
+def plot3d(checkerName, results):
+    plt.clf()
     X, Y = np.meshgrid(densities, sizes)
     Z = [[results[(x, y)].getAvgTime() for x in densities] for y in sizes]
 
@@ -73,8 +75,32 @@ def plot(checkerName, results):
     ax.set_ylabel('Size')
     ax.set_zlabel('time')
 
-    plt.show()
-    fig.savefig("results/result{}{}.png".format(checkerName, datetime.utcnow()))
+    # plt.show()
+    fig.savefig("results/result3d{}{}.png".format(checkerName, datetime.utcnow()))
+    plt.clf()
+
+def plotTimeVsSize(checkerName, results, densities):
+    plt.clf()
+    handles = []
+
+    X = sizes
+    for density in densities:
+        try:
+            Y = [results[(density, size)].getAvgTime() for size in sizes]
+            handles += plt.plot(X, Y, label="density {}".format(density))
+        except:
+            print("No results for density {}".format(density))
+            print(traceback.print_stack())
+    plt.legend(handles=handles)
+    plt.xlabel('Number of Nodes')
+    plt.ylabel('Execution Time (seconds)')
+    plt.savefig("results/timeVsSize_{}_{}{}.png".format(NUM_TRIES, checkerName, datetime.utcnow()))
+    plt.clf()
+
[…]

### 2020-07-14 acmPaper/paper.tex
Transcribe algorithms

@@
     \normalfont B\kern-0.5em{\scshape i\kern-0.25em b}\kern-0.8em\TeX}}}
+
+\labelformat{algocf}{algorithm\,(#1)}
 
@@
 \end{abstract}
-%
-%
-%
 
@@
 as a path is an ordering on nodes), and the parallel pairs are a subset of the possibilities generated by choosing a first and second member from this finite set, the resultant set of all parallel pairs must be finite.
-\xxx[AK]{TODO Check if O(n!n) = O(n). What about $O((n!)^2)$?}
+\xxx[AK]{TODO Check if O(n!n) = O(n!). What about $O((n!)^2)$?}
 The output of the algorithm, then, is the union of the set of cycle verification pairs, and the parallel pairs found by doing an all-acyclic-path search from every node in the diagram with the new edge added.
@@
 
-To present this idea diagrammatically, we represent path phases with arrows (note that these are the composition of many edges, not a single edge). The new edge is represented with a dashed arrow.
+We present the idea diagrammatically in figure \ref{figure_two_flip}.
+Arrows represent path phases (note that these are the composition of many edges, not a single edge).
+The new edge is represented with a dashed arrow.
 
+\begin{figure}
 \begin{center}
@@
 \end{center}
+\caption{Two flip tolerant path.}
+\label{figure_two_flip}
+\end{figure}
+
 Here, (S, $F_1$, $F_2$, T) is a two flip tolerant path. ($F_1$, S, T, $F_2$) is a new path, created because of the addition of (S,T), that conflicts with ($F_1$, $F_2$).
@@
 A diagram's structure can result in some redundancy, in the sense that verifying a subset of pairs can imply the verification of path pairs that aren't in the subset.
-For example, in the figure below, verifying $p_1$ and $p_2$ implies that the paths in $p_3$ are equal too.
+For example, in figure ~\ref{figure_redundancies}, verifying $p_1$ and $p_2$ implies that the paths in $p_3$ are equal too.
+
+\begin{figure}
 
@@
 \end{center}
+\caption{Path checking redundancies: verifying $p_1$ and $p_2$ implies that the paths in $p_3$ are equal.}
+\label{figure_redundancies}
+\end{figure}
 
-\xxx[dg]{Should be a figure}
 The approach in this algorithm, at a high level, is to define a function that takes in a subset of pairs and returns the subset of pairs whose verification is implied by verifying the input set.
@@
 
-A complete implementation of the described algorithm is included in the appendix.
+% A complete implementation of the described algorithm is included in the appendix.
+
+Roughly, it proceeds by first efficiently finding a "spanning" set of bilinkings (a subset whose verification implies the verification of all bilinkings in the graph).
+It does this by, starting at every node, finding the reachable subsection of the graph, and a spanning tree for the subsection.
+From each edges in the reachable section that is not a part of the tree, it generates a bilinking using the edge and a path in the tree that is parallel to the edge (algorithm ~\ref{algo_spanning_set}).
 
 \xxx[dg]{Probably put this in the appendix}
-\xxx[dg]{Give a summary of the algorithm though, breaking down important parts.  I think that's mostly been done above, but some actual code snippets might help}
-\begin{verbatim}
-Graph existingGraph;
-R_s = {}
-for each node v in V:
-    # Let the subsection of the graph reachable from v be S.
-    S = existingGraph.extractReachableSection(v)
-    # Create a minimum spanning tree of S
-    T = createMinimumSpanningTree(S)
-    # Find the excluded edges
-    excludedEdges = S.edges - T.edges
-    # Use these edges to create bilinkings
-    for each edge e in excludedEdges:
-        firstPath = T.findPath(source: e.source, sink: e.sink)
-        R_s.addElement(new Bilinking(firstPath, e)
-return R_s
-\end{verbatim}
+\begin{algorithm}
+\label{alg_spanning_set}
+\DontPrintSemicolon
+\KwResult{Find a spanning set Rs = [$r_1$, ... ,$r_k$].}
 
-\begin{verbatim}
-Find a spanning set Rs, = [r_1, ... ,r_k].
+Graph existingGraph\;
+$R_s$ $\gets$ \{\}\;
+\ForEach{node v in V}{
+    S $\gets$ existingGraph.extractReachableSection(v)\;
+    T $\gets$ createMinimumSpanningTree(S)\;
+    excludedEdges = S.edges - T.edges\;
+    \ForEach{edge e $\in$ excludedEdges}{
+        firstPath = T.findPath(source: e.source, sink: e.sink)\;
+        $R_s$.addElement(new Bilinking(firstPath, e)\;
+    } 
+}
+\KwRet{$R_s$}\;
 
-R := Rs,
-    for i:=l to K do
-        if r_i in sigma(R - r_i) then R:=R - r_i;
-return R.
+\caption{finding spanning set}\label{algo_spanning_set}
+\end{algorithm}
 
-sigma(inputSet S, codomain = Rs):
-    output = {}
-    for bilinking in Rs:
-        smallerPairs = {}
-        for candidateBilinking in Rs:
-            if exists path(source: bilinking.source, sink:candidateBilinking.source) and
-            exists path(source: candidateBilinking.sink, sink: bilinking.sink):
-                smallerPairs.add(candidateBilinking)
-        # Check if there's a linear dependence between the bilinking and smallerPairs
-        # vectorize(pair) returns a vector with a dimension for each edge in the graph,
-        # with a 1 if pair includes that edge and 0 if it doesn't.
-        matrix = [vectorize(pair) for pair in smallerPair]
-        matrix = matrix+[vectorize(bilinking)]
-        # if there is a linear dependence
-        if determinant(matrix)%2==0:
+With the spanning set thus initialized, it greedily tries to remove each pair from the spanning set if the set remains spanning even after removing the edge (algorithm ~\ref{algo_minimal_spanning_set}).
+
+\begin{algorithm}
+\DontPrintSemicolon
+\KwResult{Find a minimal spanning set R.}
+
+\SetKwFunction{FSigma}{$\sigma$}
+\SetKwProg{Fn}{Function}{:}{}
+\Fn{\FSigma{inputSet S, codomain = Rs}}{
+    output $\gets$ \{\}\;
+    \For{bilinking $\in$ Rs}{
+        smallerPairs $\gets$ \{\}\;
+        \For{candidateBilinking $\in$ Rs}{
+            \If{$\exists$ path from bilinking.source to candidateBilinking.source
+            and
+            $\exists$ path from candidateBilinking.sink to bilinking.sink}{
+                \tcc{Check if there's a linear dependence between the bilinking and smallerPairs.
+                    Vectorize(pair) returns a vector with a dimension for each edge in the graph,
+                    with a 1 if pair includes that edge and 0 if it doesn't.}    
+                matrix $\gets$ [vectorize(pair) for pair in smallerPair]\;
+                matrix $\gets$ matrix+[vectorize(bilinking)]\;
+                \If{determinant(matrix)\%2==0}{
                     output.add(bilinking)
-    return output
+                }
+            }
+        }
+    }
+    \KwRet{output}
+}
 
-\end{verbatim}
+R $\gets$ Rs\;
+\For{i=1 to K}{
+    \If{$r_i \in \sigma$(R-$r_i$)}{
+       R $\gets$ R-$r_i$\;
+    }
+}
+\KwRet{R}\;
+
+\caption{Minimal spanning set}\label{algo_minimal_spanning_set}
+\end{algorithm}
 
@@
 Theorem \ref{verifyingSet} shows that should our selected set of pairs along with cycles passing through the new edge be verified commutative, the entire diagram must commute. 
-The approach is to identify a parallel pair with exactly one path through the edge for each (source, sink) pair.
+The approach is to identify a parallel pair with exactly one path through the edge for each (source, sink) pair (algorithm ~\ref{algo_online_polynomial}).
 
-\begin{verbatim}
+\begin{algorithm}
+\DontPrintSemicolon
+\KwData{existing graph, new edge.}
+\KwResult{Set of parallel pairs to verify.}
+\SetKwProg{try}{try}{:}{}
+\SetKwProg{catch}{catch}{:}{end}
 Graph existingGraph;
@@
 
-Set parallelPairs = new Set();
-for S in existingGraph.Nodes():
-    for T in existingGraph.Nodes():
-        try:
-            Path pathWithNewEdge = 
-                FindPath(
+parallelPairs $\gets$ \{\}\;
+\For{S in existingGraph.Nodes}{
+    \For{T in existingGraph}{
+        \try{}{
+            Path pathWithNewEdge $\gets$ FindPath(
                 graph: existingGraph, 
                 sourceNode: S,
-                    sinkNode: newEdge.Source()) +
-                newEdge as Path +
+                sinkNode: newEdge.Source()) +\;
+            newEdge as Path +\;
             FindPath(
@@
                 sourceNode: newEdge.Sink(), 
-                    sinkNode: T)
-            if (S==T):
-                // Cycles are a speacial case.
-                pathInOldGraph = identity(S)
-            else:
-                Path pathInOldGraph = FindPath(
+                sinkNode: T)\;
+            \If{S == T}{
+                pathInOldGraph $\gets$ identity(S)\;
+            }
+            \Else(){
+                Path pathInOldGraph $\gets$ FindPath(
                     graph: existingGraph, 
                     sourceNode: S, 
-                    sinkNode: T);
-            parallelPairs.add((pathInOldGraph, pathWithNewEdge));
-        catch (pathFindingFailed Exception):
-            // No comparable pairs from node {S} to node {T} 
-            // that need to be checked
-            continue
-
-output parallelPairs;
-\end{verbatim}
-\begin{algorithm}
-    \caption{Polynomial time algorithm}
-    \label{polynomial}
-    \KwData{placeholder}
-    \KwResult{placeholder}
+                    sinkNode: T)\;
+            }
+            parallelPairs.add((pathInOldGraph, pathWithNewEdge))\;
+        }
+        \catch{PathFindingFailedException}{
+           \tcc{No comparable pairs from node S to node T that need to be checked}
+           continue\;
+        }
+    }
+}
+\KwRet{parallelPairs}\;
+\caption{Online polynomial time algorithm to find parallel pair set}\label{algo_online_polynomial}
 \end{algorithm}
@@
 
-Consider the following situation:
+Consider the situation in figure ~\ref{figure_reduction_rule}.
 
+\begin{figure}
 \begin{center}
@@
 \end{center}
+\caption{Reduction rule.}
+\label{figure_reduction_rule}
+\end{figure}
+
 \begin{theorem}
@@
 
-Note that the proof is not affected if any of these paths is the identity, eg. if $f_1$ is the identity and S and $P_1$ are actually the same node.
+The proof is not affected if any of these paths is the identity, eg. if $f_1$ is the identity and S and $P_1$ are actually the same node.
 
@@
 
-Using this information it is possible to choose a minimal subset of path pairs to verify.
+Using this information it is possible to choose a minimal subset of path pairs to verify, as in algorithm ~\ref{algo_online_minimal}.
 
@@
 
-\begin{verbatim}
-# Existng graph
-Graph existingGraph;
-# New edge
-Edge (S, T);
+\begin{algorithm}
+\DontPrintSemicolon
+\KwData{Existing graph, new edge.}
+\KwResult{Set of parallel pairs to verify.}
+Graph existingGraph\;
+Edge (S, T)\;
 
-predecessors = S.predecessors(existingGraph)
-successors = T.successors(existingGraph)
+predecessors $\gets$ S.predecessors(existingGraph)\;
+successors $\gets$ T.successors(existingGraph)\;
 
-# Construct the graph
-Graph terminalPairGraph;
-for q in successors:
-    for p in predecessors:
-        terminalPairGraph.addNode(q, p)
-        # addEdge function should handle addition of edges between nodes that 
-        # haven't been defined yet
-        for q_pred in q.predecessors(existingGraph):
-            for p_succ in p.successors(existingGraph):
-                terminalPairGraph.addEdge((q_pred, p_succ))
+Graph terminalPairGraph\;
+\For{q $\in$ successors}{
+    \For{p $\in$ predecessors}{
+        terminalPairGraph.addNode(q, p)\;
+        \For{qPred $\in$ q.predecessors(existingGraph)}{
+            \For{pSucc $\in$ p.successors(existingGraph)}{
+                terminalPairGraph.addEdge((qPred, pSucc))\;
+            }
+        }
+    }
+}
 
-# Now we reduce the graph.
-verificationSet = {}
-while (len(terminalPairGraph.nodes) > 0):
-    # Choose a random node to start reduction from.
-    current_node = terminalPairGraph.nodes[0]
-    visited_nodes = set()
-    # Climb up the graph till a "root" node is found
-    while (len(currentNode.parents()) > 0):
-        visited_nodes.add(currentNode)
-        # Choose a parent at random.
-        currentNode = currentNode.parents()[0]
-        # Deal with cycles by reducing them
-        if currentNode in visited_nodes:
-            edges = getAllEdges(visited_nodes, terminalPairGraph)
-            terminalPairGraph.removeNodes(visitedNodes)
-            # All the incoming and outgoing edges from the entire visited set are now 
-            # transferred to a single node that represents the class
-            terminalPairGraph.addNode(currentNode, edges)
-    # No predecessor! We should have reached a root!
-    verificationSet.add(currentNode)
-    terminalPairGraph.removeNodes(currentNode.successors(terminalPairGraph)
-
-# This contains the minimal set of (source, sink) pairs whose verification 
-# implies verification of the entire graph. 
-return verificationSet
-
-\end{verbatim}
-\begin{algorithm}
-    \caption{Optimal subset algorithm}
-    \label{optimal}
-    \KwData{placeholder}
-    \KwResult{placeholder}
+verificationSet $\gets$ \{\}\;
+\While{len(terminalPairGraph.nodes) $>$ 0}{
+    currentNode $\gets$ terminalPairGraph.nodes[0]\;
+    visitedNodes $\gets$ \{\}\;
+    \While{len(currentNode.parents()) $>$ 0}{
+        visitedNodes.add(currentNode)\;
+        currentNode $\gets$ currentNode.parents()[0]\;
+        \If{currentNode $\in$ visitedNodes}{
+            edges $\gets$ getAllEdges(visitedNodes, terminalPairGraph)\;
+            terminalPairGraph.removeNodes(visitedNodes)\;
+            terminalPairGraph.addNode(currentNode, edges)\; 
+        }
+    }
+    verificationSet.add(currentNode)\;
+    terminalPairGraph.removeNodes( currentNode.successors(terminalPairGraph))\;
+}
+\KwRet{verificationSet}\;
+\caption{Minimal set finding algorithm}\label{algo_online_minimal}
 \end{algorithm}
@@
 The result of the algorithm cannot be further reduced.
-Also, the verification of the parallel pairs returned in the algorithm implies that the output of the previous algorithm must commute, and transitively with Theorem \ref{verifyingSet} that the entire diagram must commute.
+Also, the verification of the parallel pairs returned in the algorithm implies that the output of the previous algorithm must commute, and transitively with Theorem ~\ref{verifyingSet} that the entire diagram must commute.
 
@@
 \subsection{Verification}
-Ultimately verification is performed by calling an equality oracle (that is beyond the scope of this work) for every comparable pair returned by the previous stage.
+Ultimately verification is performed by calling an equality oracle (that is beyond the scope of this work) for every comparable pair returned by the previous stage, as in algorithm ~\ref{algo_verification}.
 
-\begin{verbatim}
-for (path_1, path_2) in parallelPairs:
-    if (path_1 != path_2):
-        print($"Edge addition does not maintain path independence. 
-        Counterexample: {path_1}, {path_2}")
-        return False
-print("Edge addition maintains path independence.")
-return True
-\end{verbatim}
+\begin{algorithm}
+\DontPrintSemicolon
+\For{(path1, path2) $\in$ parallelPairs}{
+    \If{path1 $!$= path2}{
+        \KwRet{False}\;
+    }
+}
+\KwRet{True}\;
+\caption{Verification algorithm}\label{algo_verification}
+\end{algorithm}
 
@@
 TODO.
+
 \subsection{Gator}
+
 \subsection{Currency Graph}
@@
 To ensure that it indeed commuted, we started with an empty diagram, and added in edges one by one.
-Before the addition of each edge, we used the algorithms (~\ref{polynomial} and ~\ref{optimal}) to ensure the addition of a new edge did not introduce inconsistencies in the existing diagram.
+Before the addition of each edge, we used the algorithms (~\ref{algo_online_polynomial} and ~\ref{algo_online_minimal}) to ensure the addition of a new edge did not introduce inconsistencies in the existing diagram.
 If a new edge was, in fact, problematic, the algorithms returned an example inconsistent
@@
 Nonetheless, both algorithms were effective in catching inconsistencies.
-Algorithm ~\ref{polynomial} started finding inconsistencies at error tolerances to the order of $10^{-3}$, and ~\ref{optimal}, which makes more invalid redundant path check removals, at error tolerances to the order of $10^{-7}$.
+Algorithm ~\ref{algo_online_polynomial} started finding inconsistencies at error tolerances to the order of $10^{-3}$, and ~\ref{algo_online_minimal}, which makes more invalid redundant path check removals, at error tolerances to the order of $10^{-7}$.
 
-Averaging over evaluation for the first 30 days of 2020, building and verifying a diagram to completion (inclusive of the time required by network calls) took 243$\pm$19 seconds using algorithm ~\ref{optimal}, and in 133$\pm$13 seconds with algorithm ~\ref{polynomial}.
+Averaging over evaluation for the first 30 days of 2020, building and verifying a diagram to completion (inclusive of the time required by network calls) took 243$\pm$19 seconds using algorithm ~\ref{algo_online_minimal}, and in 133$\pm$13 seconds with algorithm ~\ref{algo_online_polynomial}.
 Computation was performed on a Macbook Pro 2015, 2.9 GHz Dual-Core Intel Core i5.
@@
 
-Modified algorithm for diagrams without the identity:
-
-\begin{verbatim}
-Graph existingGraph;
-Edge newEdge;
-
-Set parallelPairs = new Set();
-for S in existingGraph.Nodes():
-    for T in existingGraph.Nodes():
-        try:
-            Path pathWithNewEdge = 
-                FindPath(
-                    graph: existingGraph, 
-                    sourceNode: S,
-                    sinkNode: newEdge.Source()) +
-                newEdge as Path +
-                FindPath(
-                    graph: existingGraph, 
-                    sourceNode: newEdge.Sink(), 
-                    sinkNode: T)
-            if (S==T):
-                // Cycles are a special case.
-                parallelPairs.add(pathWithNewEdge, pathWithNewEdge+pathWithNewEdge)
-                for edge in (S.forwardEdges - pathWithNewEdge[0]):
-                    parallelPairs.add(edge, edge+pathWithNewEdge)
-                for edge in (S.backwardEdges - pathWithNewEdge[-1]):
-                    parallelPairs.add(edge, pathWithNewEdge+edge)
-            else:
-                Path pathInOldGraph = FindPath(
-                    graph: existingGraph, 
-                    sourceNode: S, 
-                    sinkNode: T);
-                parallelPairs.add((pathInOldGraph, pathWithNewEdge));
-        catch (pathFindingFailed Exception):
-            // No comparable pairs from node {S} to node {T} 
-            // that need to be checked
-            continue
-    cycle = findCycleWithTerminalNode(newEdge.Source()):
-    if cycle:    
-        parallelPairs.add(newEdge, cycle+newEdge)
-    cycle = findCycleWithTerminalNode(newEdge.Source()):
-    if cycle:    
-        parallelPairs.add(newEdge, newEdge+cycle)
-output parallelPairs;
-\end{verbatim}
-
-The optional minimization algorithm is modified to exclude nodes representing cycles where the source and sink are equal, since the reduction case no longer will apply in the absence of an implicit identity on the terminal node of the cycle.
-
 \section{Case Studies}

### 2020-07-21 currency-exchange/PolynomialPathChecker.py
Murota implementation:

@@ -1,58 +1,0 @@
-from baseOnlineChecker import BaseOnlineChecker
-from colorama import Fore
-from testUtilities import assertActualIsSuperset, test_graph, test_s, test_t,\
-    expected_solution
-import time
-from typing import List, Tuple
-
-class PolynomialPathChecker(BaseOnlineChecker):
-    """Path checker that returns a set for verification for online problem
-    in time O(|V|.(|V|+|E|)))."""
-
-    def __init__(self):
-        super().__init__()
-
-    def getPathsToCheck(self) ->  List[Tuple[List[int], List[int]]]:
-        """Return the pairs of path whose equality implies path independence of 
-        the new graph."""
-        # First find set of all predecessors and successors.
-        startTime = time.time()
-        pathsToCheck = []
-        predecessors = self.getAllPredecessors(self.newEdgeSource)
-        successors = self.getAllSuccessors(self.newEdgeSink)
-        predecessorSet = set(predecessors)
-        # For each successor, find paths to all valid predecessors.
-        for sink in successors:
-            predecessorsForNode = self.getAllPredecessorsUnordered(
-                sink, 
-                memoize = True).intersection(predecessorSet)
-            while len(predecessorsForNode) > 0:
-                src = predecessorsForNode.pop()
-                pathsToCheck += [(self.pathsToNewEdgeSource[src] +
-                    self.pathsFromNewEdgeSink[sink],
-                    self.pathsToNode[sink][src] if sink != src 
-                        else self.identityFunction)]
-        endTime = time.time()
-        self.timeTaken = endTime - startTime
-        return pathsToCheck
-
-#### Quick Tests ####
-
-def testGetPathsToCheck():
-    checker = PolynomialPathChecker()
-    checker.setGraph(test_graph)
-    checker.setEdge(test_s, test_t)
-    assertActualIsSuperset(expected_solution, checker.getPathsToCheck())
-    assert(checker.timeTaken > 0)
-
-def runAllTests():
-    print('\033[0m' + "Running optimalSetPathChecker Tests")
-    testGetPathsToCheck()
-    print(Fore.GREEN + 'Run Completed')    
-
-#### Execute ####
-def main():
-    runAllTests()
-
-if __name__=="__main__":
-    main()

### 2020-07-21 currency-exchange/baseOnlineChecker.py

@@ -1,311 +1,0 @@
-from colorama import Fore
-import copy
-from iPathChecker import IPathChecker
-from testUtilities import assertEqual, test_graph, test_s, test_t,\
-    expected_paths_to_s, expected_paths_from_t, expected_fwd_graph,\
-        expected_bkd_graph, NO_EDGE, TestDefinition, defaultTestSuite
-from typing import List, Set, Tuple
-
-_default_no_edge = -1
-
-class BaseOnlineChecker(IPathChecker):
-    """Implements base functionality common to all PathCheckers."""
-
-    _invalid_node = -1
-    _debug = False
-
-    #### Utility functions ####
-
-    def getForwardEdges(self, source: int) -> List[int]:
-        return [i for i in range(len(self.graph)) \
-            if self.graph[source][i] != self._no_edge]
-
-    def getBackwardEdges(self, sink: int) -> List[int]:
-        return [i for i in range(len(self.graph)) \
-            if self.graph[i][sink] != self._no_edge and i != sink]
-
-    def getEmptyPathList(self):
-        paths = []
-        for _ in range(len(self.graph)):
-            paths += [[]]
-        return paths
-
-    def buildCompactFwdGraph(self):
-        self.compactFwdGraph = self.getEmptyPathList()
-        for node in range(len(self.graph)):
-            self.compactFwdGraph[node] = self.getForwardEdges(node)
-    
-    def buildCompactBkdGraph(self):
-        self.compactBkdGraph = self.getEmptyPathList()
-        for node in range(len(self.graph)):
-            self.compactBkdGraph[node] = self.getBackwardEdges(node)
-    
-    def getAllPredecessors(self, node, memoize = False) -> List[int]:
-        """Returns a list of nodes that are predecessors of this one.
-        Memoize if source of new edge."""
-        memoize = memoize or node == self.newEdgeSource
-        if memoize:
-            pathDict = dict()
-        visited = set()
-        # To store visitedNodes ordered closest to new edge first.
-        visitedList = []
-        currentNode = node
-        stack = [currentNode]
-        path = []
-        while (len(stack) > 0):
-            currentNode = stack.pop()
-            if currentNode == self._invalid_node:
-                path = path[1:]
-            elif currentNode not in visited:
-                visited.add(currentNode)
-                visitedList += [currentNode]
-                path = [currentNode] + path
-                if (memoize):
[…]

### 2020-07-21 currency-exchange/currencyGraph.py

@@ -1,169 +1,0 @@
-from colorama import Fore
-import requests 
-import numpy as np
-import copy
-from datetime import datetime
-from optimalSetPathChecker import OptimalSetPathChecker
-from testUtilities import TestDefinition, defaultTestSuite
-import time
-import traceback
-from typing import List
-
-class CurrencyGraph():
-    """Class that holds, builds and manipulates the graph with currency 
-    conversion rates"""
-
-    INFTY = 1000000
-    NO_EDGE = -1
-    DEBUG = False
-    baseEndPoint = "https://api.exchangeratesapi.io/"
-    endPoint = baseEndPoint + "latest?"
-    
-    def __init__(self):
-        self.graph = []
-        self.currencyList = []
-        self.base = 'USD' # Default but can change
-        self.sleepTime = 0
-        np.set_printoptions(formatter={
-            'float': lambda x: "{0:0.4f}".format(x)}, threshold=10000)
-    
-    def setupWithDay(self, year, month, day):
-        date = datetime(year, month, day)
-        self.endPoint = self.baseEndPoint + date.strftime("%Y-%m-%d") + "?"
-        self.setup()
-    
-    def setup(self):
-        """Setup graph with initial row of data from base currency"""
-        rates = self.makeRequest(self.base)
-        self.initialize(list(rates))
-        self.saveRow(self.base, rates)
-        if self.DEBUG:
-            print("Initialized Graph:")
-            print(np.matrix(self.currencyList))
-            print(np.matrix(self.graph))
-
-    def initialize(self, currencies):
-        """Initialize graph and currency list."""
-        self.currencyList = currencies
-        self.constructGraph()
-    
-    def indexOf(self, currency):
-        """Gets node in graph that a currency corresponds to."""
-        return self.currencyList.index(currency)
-    
-    def constructGraph(self):
-        """Initialize empty graph."""
-        for _ in range(len(self.currencyList)):
-            row = []
-            for _ in range(len(self.currencyList)):
-                row += [self.NO_EDGE]
-            self.graph += [row]
-    
-    def saveEntry(self, base, target, rate, checkIndependence=False):
-        isValid = True
-        extraData = ''
-        if (checkIndependence):
-            (isValid, extraData) = self.checkIndependenceFunc(base, target, rate)
-        if (isValid):
-            self.graph[self.indexOf(base)][self.indexOf(target)] = rate
-        return (isValid, extraData)
-
[…]

### 2020-07-21 currency-exchange/iPathChecker.py

@@ -1,26 +1,0 @@
-from typing import List
-from typing import Tuple
-
-class IPathChecker:
-    """Interface for path checkers"""
-
-    def setGraph(self, graph: List[List[int]]) -> None:
-       """Set the graph to which an edge is to be added"""
-       pass
-    
-    def setEdge(self, source: int, sink: int) -> None:
-        """Set the edge that is to be added"""
-        pass
-
-    def getPathsToCheck(self) -> List[Tuple[List[int], List[int]]]:
-        """Return the pairs of path whose equality implies path independence
-        of the new graph"""
-        pass
-
-    def getComputeTime(self) -> int:
-        """Return time required to compute latest path check"""
-        pass
-
-    def setIdFunction(self, idFunction: List[int]):
-        """Store the special representation of the identity function/path."""
-        pass

### 2020-07-21 currency-exchange/main.py

@@ -1,129 +1,0 @@
-from colorama import Fore
-import copy
-from currencyGraph import CurrencyGraph
-import numpy as np
-from optimalSetPathChecker import OptimalSetPathChecker
-from iPathChecker import IPathChecker
-from polynomialPathChecker import PolynomialPathChecker
-import requests
-import time
-import traceback
-from typing import List
-
-# Assume that a graph doesn't have more edges nodes than this
-INFTY = 1000000
-NO_EDGE = -1
-DEBUG = False
-ID_FUNCTION = [-1]
-
-# TODO add:
-# 1. Commandline arguments.
-# 2. Historic rate.
-
-###############################################
-
-epsilon = 1E-2 # allowed error multiplier
-
-checkers = [
-    OptimalSetPathChecker(),
-    PolynomialPathChecker()
-]
-
-def getPathValue(path: List[int], graph: List[List[int]]):
-    """Composition oracle defintion."""
-    result = 1
-    # Special case for identity function.
-    if path == ID_FUNCTION:
-        return result
-    # Special case for just new edge.
-    if len(path) == 1:
-        return graph[path[0]][path[0]]
-    # Regular case.
-    for i in range(len(path)-1):
-        result *= graph[path[i]][path[i+1]]
-    return result
-
-def independenceFuncFactory(checker: IPathChecker, graph: CurrencyGraph):
-    """Equality check oracle definition."""
-    checker.setGraph(graph.graph)
-    checker.setIdFunction(ID_FUNCTION)
-    def IndependenceFunction(base, target, rate):
-        checker.setEdge(graph.indexOf(base), graph.indexOf(target))
-        pairsToCheck = checker.getPathsToCheck()
-        newGraph = graph.getCopyWithNewEdge(base, target, rate)
-        for (path1, path2) in pairsToCheck:
-            path1Value = getPathValue(path1, newGraph)
-            path2Value = getPathValue(path2, newGraph)
-            if abs(path1Value - path2Value) > epsilon:
-                return (False, "path 1: {}: {} \npath 2: {}: {} \n difference: {}, new edge: {} -> {}: rate: {}"
-                .format(
-                path1Value,
-                path1,
-                path2Value,
-                path2,
-                path1Value - path2Value,
-                base,
-                target,
-                rate))
-        return (True, "Success for base {} target {}".format(base, target))
-    return IndependenceFunction
-            
[…]

### 2020-07-21 currency-exchange/noIdentityOptimalSetPathChecker.py

@@ -1,51 +1,0 @@
-from colorama import Fore
-from nonIdentityUtilities import NonIdentityPathChecker
-from optimalSetPathChecker import OptimalSetPathChecker
-from testUtilities import assertActualIsSuperset, test_graph, test_s, test_t,\
-    expected_solution_no_identity, TestDefinition, defaultTestSuite, assertEqual
-import time
-from typing import List, Tuple
-
-class NoIdentityOptimalSetPathChecker(OptimalSetPathChecker, NonIdentityPathChecker):
-
-    def __init__(self):
-        super().__init__()
-        self.noIdentity = True
-
-    def getPathsToCheck(self) ->  List[Tuple[List[int], List[int]]]:
-        """Return the pairs of path whose equality implies path independence of 
-        the new graph."""
-        startTime = time.time()
-        subset = self.getRootPairs()
-        # Find paths for each pair in the subset
-        pathPairs = []
-        for node in subset:
-            (source, sink) = node
-            if (source == sink):
-                pathPairs += self.handleCycle(source)
-            else:
-                pathPairs += [self.findPair(source, sink)]
-        pathPairs += self.getSourceCycles() + self.getSinkCycles()
-        endTime = time.time()
-        self.timeTaken = endTime - startTime
-        return pathPairs
-
-def testGetPathsToCheck(testDefinition: TestDefinition):
-    checker = NoIdentityOptimalSetPathChecker()
-    checker.setGraph(testDefinition.test_graph)
-    checker.setEdge(testDefinition.test_s, testDefinition.test_t)
-    assertEqual(testDefinition.expected_solution_no_identity,\
-        checker.getPathsToCheck())
-    assert(checker.timeTaken > 0)
-
-def runAllTests(testSuite: List[TestDefinition]):
-    print('\033[0m' + "Running NoIdentityOptimalSetPathChecker Tests")
-    for testDefinition in testSuite:
-        testGetPathsToCheck(testDefinition)
-    print(Fore.GREEN + 'Run Completed')
-
-def main():
-    runAllTests(defaultTestSuite)
-
-if __name__=="__main__":
-    main()

### 2020-07-21 currency-exchange/noIdentityPolynomialPathChecker.py

@@ -1,64 +1,0 @@
-from nonIdentityUtilities import NonIdentityPathChecker
-from colorama import Fore
-from testUtilities import assertActualIsSuperset, test_graph, test_s, test_t,\
-    expected_solution_no_identity, defaultTestSuite, TestDefinition
-import time
-from typing import List, Tuple
-
-class NoIdentityPolynomialPathChecker(NonIdentityPathChecker):
-    """Path checker that returns a set for verification for online problem
-    in time O(|V|.(|V|+|E|)))."""
-
-    def __init__(self):
-        super().__init__()
-
-    def getPathsToCheck(self) ->  List[Tuple[List[int], List[int]]]:
-        """Return the pairs of path whose equality implies path independence of 
-        the new graph."""
-        # First find set of all predecessors and successors.
-        startTime = time.time()
-        pathsToCheck = []
-        predecessors = self.getAllPredecessors(self.newEdgeSource)
-        successors = self.getAllSuccessors(self.newEdgeSink)
-        predecessorSet = set(predecessors)
-        # For each successor, find paths to all valid predecessors.
-        for sink in successors:
-            predecessorsForNode = self.getAllPredecessorsUnordered(
-                sink, 
-                memoize = True).intersection(predecessorSet)
-            # Handle cycles.
-            if sink in predecessorsForNode:
-                predecessorsForNode.remove(sink)
-                pathsToCheck += self.handleCycle(sink)
-            while len(predecessorsForNode) > 0:
-                src = predecessorsForNode.pop()
-                pathsToCheck += [(self.pathsToNewEdgeSource[src] +
-                    self.pathsFromNewEdgeSink[sink],
-                    self.pathsToNode[sink][src])]
-        pathsToCheck += self.getSinkCycles() + self.getSourceCycles()
-        endTime = time.time()
-        self.timeTaken = endTime - startTime
-        return pathsToCheck
-
-#### Quick Tests ####
-
-def testGetPathsToCheck(testDefinition: TestDefinition):
-    checker = NoIdentityPolynomialPathChecker()
-    checker.setGraph(testDefinition.test_graph)
-    checker.setEdge(testDefinition.test_s, testDefinition.test_t)
-    assertActualIsSuperset(testDefinition.expected_solution_no_identity,\
-        checker.getPathsToCheck())
-    assert(checker.timeTaken > 0)
-
[…]

### 2020-07-21 currency-exchange/nonIdentityUtilities.py

@@ -1,83 +1,0 @@
-from baseOnlineChecker import BaseOnlineChecker
-from colorama import Fore
-from testUtilities import test_graph, test_s, test_t,\
-    expected_solution_no_identity, TestDefinition, defaultTestSuite, assertEqual
-from typing import List, Tuple
-
-class NonIdentityPathChecker(BaseOnlineChecker):
-
-    def __init__(self):
-        super().__init__()
-        self.noIdentity = True
-
-    def handleCycle(self, terminalNode: int) -> List[Tuple[List[int], List[int]]]:
-        pathsToCheck = []
-        cycle = self.pathsToNewEdgeSource[terminalNode] +\
-            self.pathsFromNewEdgeSink[terminalNode] 
-        pathsToCheck +=[(cycle, cycle + cycle[1:])]
-        return pathsToCheck
-    
-    def getSinkCycles(self) -> List[Tuple[List[int], List[int]]]:
-        (cycleExists, cycle) = self.findCycle(self.newEdgeSink)
-        if cycleExists:
-            return [([self.newEdgeSource, self.newEdgeSink], [self.newEdgeSource] + cycle)]
-        return []
-    
-    def getSourceCycles(self) -> List[Tuple[List[int], List[int]]]:
-        (cycleExists, cycle) = self.findCycle(
-            self.newEdgeSource)
-        if cycleExists:
-            return [([self.newEdgeSource, self.newEdgeSink], cycle+[self.newEdgeSink])]
-        return []
-    
-    def findCycle(self, source: int) -> Tuple[bool, List[int]]:
-        """Find cycle."""
-        # DFS.
-        path = []
-        visited = set()
-        stack = [source]
-        firstRun = True
-        while (len(stack)>0):
-            currentNode = stack.pop()
-            if currentNode == self._invalid_node:
-                path.pop()
-            elif currentNode not in visited:
-                path = path+[currentNode]
-                if not firstRun:
-                    if currentNode == source:
-                        return (True, path)
-                    visited.add(currentNode)
-                else:
-                    firstRun = False
-                stack += [self._invalid_node] + self.compactFwdGraph[currentNode]
-        return (False, [])
-
-#### Quick Tests ####
-
-def testGetSourceCycles(testDefintion: TestDefinition):
-    checker = NonIdentityPathChecker()
-    checker.setGraph(testDefintion.test_graph)
[…]

### 2020-07-21 currency-exchange/optimalSetPathChecker.py

@@ -1,108 +1,0 @@
-from baseOnlineChecker import BaseOnlineChecker
-from colorama import Fore
-import copy
-from testUtilities import assertEqual, test_graph, test_s, test_t,\
-    expected_solution, TestDefinition, defaultTestSuite
-import time
-from typing import List, Set, Tuple
-
-class OptimalSetPathChecker(BaseOnlineChecker):
-    """Path checker that returns minimal set for verification 
-    for online problem."""
-
-    def __init__(self):
-        super().__init__()
-
-    def getSuccessors(self, source: int, sink: int) -> Set[int]:
-        """Returns a set of pairs such that verifying a path between given
-        source and sink implies also that all the pairs in the returned set are
-        equal."""
-        predecessors = self.getAllPredecessors(source)
-        successors = self.getAllSuccessors(sink)
-        if (self.noIdentity and source == sink):
-            return {(src, snk) for src in predecessors for snk in successors 
-            if src == snk}
-        return {(src, snk) for src in predecessors for snk in successors}
-
-    def getRootPairs(self) -> List[List[int]]:
-        """Returns a graph where each node represents a (source, sink) pair in
-        self.graph. 
-        Map from (source, sink) to node number in self.dependencyNodeMap.
-        Node A succeeds node B if checking a pair of B implies that 
-        the paths pairs of A must also be equal."""
-        acceptedPairs = set()
-        # Get set of all predecessors, successors.
-        predecessors = self.getAllPredecessors(self.newEdgeSource)
-        successors = self.getAllSuccessors(self.newEdgeSink)
-        # Go through each element in predecessors X successors.
-        # For efficiency, closest pairs to new edge should appear first.
-        potentialPairs = {(source, sink) for source in predecessors \
-            for sink in successors}
-        orderedPotentialPairs = [(source, sink) for source in predecessors \
-            for sink in successors]
-        while (len(potentialPairs) != 0):
-            # using ordering is a heuristic to (hopefully) improve performance,
-            # so parent is usually checked before child.
-            (source, sink) = orderedPotentialPairs.pop(0)
-            if (source, sink) in potentialPairs:
-                potentialPairs.remove((source, sink))
-                (pathExists, _) = self.findPath(source, sink)
-                if pathExists:
[…]

### 2020-07-21 currency-exchange/slowPolynomialPathChecker.py

@@ -1,56 +1,0 @@
-from baseOnlineChecker import BaseOnlineChecker
-from colorama import Fore
-from testUtilities import assertActualIsSuperset, test_graph, test_s, test_t,\
-    expected_solution
-import time
-from typing import List, Tuple
-
-class PolynomialPathChecker(BaseOnlineChecker):
-    """Path checker that returns a set for verification for online problem
-    without memoization in time O(|V|^2.(|V|+|E|)))."""
-
-    def __init__(self):
-        super().__init__()
-
-    def getPathsToCheck(self) ->  List[Tuple[List[int], List[int]]]:
-        """Return the pairs of path whose equality implies path independence of 
-        the new graph."""
-        # First find set of all predecessors and successors.
-        startTime = time.time()
-        pathsToCheck = []
-        predecessors = self.getAllPredecessors(self.newEdgeSource)
-        successors = self.getAllSuccessors(self.newEdgeSink)
-        # predecessorSet = set(predecessors)
-        # For each successor, find paths to all valid predecessors.
-        for sink in successors:
-            for source in predecessors:
-                (isPath, path) = self.findPath(source, sink)
-                if isPath:
-                    pathsToCheck += [(self.pathsToNewEdgeSource[source] +
-                        self.pathsFromNewEdgeSink[sink],
-                        path)]
-        endTime = time.time()
-        self.timeTaken = endTime - startTime
-        return pathsToCheck
-
-#### Quick Tests ####
-
-def testGetPathsToCheck():
-    checker = PolynomialPathChecker()
-    checker.setGraph(test_graph)
-    checker.setEdge(test_s, test_t)
-    assertActualIsSuperset(expected_solution, checker.getPathsToCheck())
-    assert(checker.timeTaken > 0)
-
-def runAllTests():
-    print('\033[0m' + "Running optimalSetPathChecker Tests")
-    testGetPathsToCheck()
-    print(Fore.GREEN + 'Run Completed')    
-
-#### Execute ####
-
-def main():
-    runAllTests()
-
-if __name__=="__main__":
-    main()

### 2020-07-21 currency-exchange/testUtilities.py

@@ -1,236 +1,0 @@
-from colorama import Fore
-import traceback
-
-NO_EDGE = -1
-
-# Constants for default test.
-test_graph = [
-    [NO_EDGE,       1, NO_EDGE, NO_EDGE, NO_EDGE, NO_EDGE, NO_EDGE, NO_EDGE],
-    [NO_EDGE, NO_EDGE,       1, NO_EDGE, NO_EDGE, NO_EDGE, NO_EDGE, NO_EDGE],
-    [NO_EDGE, NO_EDGE, NO_EDGE, NO_EDGE, NO_EDGE, NO_EDGE, NO_EDGE, NO_EDGE],
-    [NO_EDGE, NO_EDGE, NO_EDGE, NO_EDGE,       1, NO_EDGE,       1, NO_EDGE],
-    [NO_EDGE, NO_EDGE, NO_EDGE, NO_EDGE, NO_EDGE,       1, NO_EDGE, NO_EDGE],
-    [NO_EDGE, NO_EDGE, NO_EDGE, NO_EDGE, NO_EDGE, NO_EDGE, NO_EDGE, NO_EDGE],
-    [NO_EDGE, NO_EDGE, NO_EDGE, NO_EDGE, NO_EDGE, NO_EDGE, NO_EDGE,       1],
-    [NO_EDGE, NO_EDGE,       1, NO_EDGE, NO_EDGE, NO_EDGE, NO_EDGE, NO_EDGE]]
-test_s = 2
-test_t = 3
-test_new_edge = (2, 3)
-
-expected_fwd_graph = [
-    [1],
-    [2],
-    [],
-    [4, 6],
-    [5],
-    [],
-    [7],
-    [2]
-]
-
-expected_bkd_graph = [
-    [],
-    [0],
-    [1, 7],
-    [],
-    [3],
-    [4],
-    [3],
-    [6]
-]
-
-expected_paths_to_s = {
-    0 : [0, 1, 2],
-    1 : [1, 2],
-    2 : [2],
-    3 : [3, 6, 7, 2],
-    6 : [6, 7, 2],
-    7 : [7, 2]
-}
-
-expected_paths_from_t = {
-    2 : [3, 6, 7, 2],
-    3 : [3],
-    4 : [3, 4],
-    5 : [3, 4, 5],
-    6 : [3, 6],
-    7 : [3, 6, 7]
-}
-
[…]

### 2020-07-21 currency-exchange/tester.py

@@ -1,38 +1,0 @@
-from noIdentityOptimalSetPathChecker import NoIdentityOptimalSetPathChecker
-from noIdentityPolynomialPathChecker import NoIdentityPolynomialPathChecker
-from optimalSetPathChecker import OptimalSetPathChecker
-from PolynomialPathChecker import PolynomialPathChecker
-
-graph = [[-1, -1, 1, 1, -1], 
-         [-1, -1, 1, 1, -1], 
-         [1, -1, -1, -1, -1],
-         [-1, -1, -1, -1, -1],
-         [-1, -1, -1, -1, -1]]
-
-# graph = [[-1, -1, 1],
-#          [-1, -1, 1],
-#          [1, -1, -1]]
-
-poly = PolynomialPathChecker()
-
-poly.setGraph(graph)
-poly.setEdge(0, 1)
-print("polynomial:\n" + str(poly.getPathsToCheck()) + "\n\n")
-
-opti = OptimalSetPathChecker()
-
-opti.setGraph(graph)
-opti.setEdge(0, 1)
-print("optimal:\n" + str(opti.getPathsToCheck()) + "\n\n")
-
-polyid = NoIdentityPolynomialPathChecker()
-
-polyid.setGraph(graph)
-polyid.setEdge(0, 1)
-print("polynomial noid:\n" + str(polyid.getPathsToCheck()) + "\n\n")
-
-optiid = NoIdentityOptimalSetPathChecker()
-
-optiid.setGraph(graph)
-optiid.setEdge(0, 1)
-print("optimal noid:\n" + str(optiid.getPathsToCheck()))

### 2020-07-21 implementation/pathCheckers/baseOnlineChecker.py

@@ -44,6 +44,8 @@ class BaseOnlineChecker(IPathChecker):
     def getAllPredecessors(self, node, memoize = False) -> List[int]:
         """Returns a list of nodes that are predecessors of this one.
         Memoize if source of new edge."""
+        if not self.compactBkdGraph:
+            self.buildCompactBkdGraph()
         memoize = memoize or node == self.newEdgeSource
         if memoize:
             pathDict = dict()
@@ -105,6 +107,8 @@ class BaseOnlineChecker(IPathChecker):
     def getAllSuccessors(self, node) -> List[int]:
         """Returns a list of nodes that are successors of this one.
         Memoize if sink of new edge."""
+        if not self.compactFwdGraph:
+            self.buildCompactFwdGraph()
         visited = set()
         # maintain a separate list to get ordering where closest node is first.
         visitedList = []
@@ -127,6 +131,39 @@ class BaseOnlineChecker(IPathChecker):
             print("getAllSuccessors: node = {} and visited list: {}".format(
                 node, visitedList))
         return visitedList
+    
+    def getTree(self, node: int):
+        """Similar to get all successors, but returns a spanning tree in 
+        the form of a set of edges."""
+        if not self.compactFwdGraph:
+            self.buildCompactFwdGraph()
+        visited = set()
+        # maintain a separate list to get ordering where closest node is first.
+        visitedList = []
+        treeEdges = set()
+        currentNode = node
+        previousNode = _invalid_node
+        stack = [node]
+        path = []
+        # More or less DFS.
+        while (len(stack) > 0):
+            currentNode = stack.pop()
+            if currentNode == self._invalid_node:
+                path.pop()
+            elif currentNode not in visited:
+                visited.add(currentNode)
+                visitedList += [currentNode]
+                path += [currentNode]
+                if (node == self.newEdgeSink):
+                    self.pathsFromNewEdgeSink[currentNode] = copy.deepcopy(path)
+                stack += [self._invalid_node] + self.compactFwdGraph[currentNode]
+                if (previousNode != _invalid_node):
+                    treeEdges.add((previousNode, currentNode))
+                    previousNode = currentNode
+        if __debug__ and self._debug:
+            print("getAllSuccessors: node = {} and visited list: {}".format(
+                node, visitedList))
+        return treeEdges
 
     def findPath(self, source: int, sink: int) -> Tuple[bool, List[int]]:
         """Find path from source ot sink in old graph,

### 2020-07-21 implementation/pathCheckers/batchChecker.py

@@ -1,0 +1,200 @@
+import setup
+from colorama import Fore
+import copy
+from pathCheckers.naiveChecker import NaiveChecker
+from pathCheckers.baseOnlineChecker import BaseOnlineChecker
+from pathCheckers.pathFinding import findEdgeConflicts
+from testing.testUtilities import assertEqual, assertActualIsSuperset,\
+    test_graph, test_s, test_t, expected_paths_to_s, expected_paths_from_t,\
+    expected_fwd_graph, expected_bkd_graph, NO_EDGE, TestDefinition,\
+    defaultTestSuite
+import time
+from typing import List, Set, Tuple
+
+_default_no_edge = -1
+_source = 0
+_sink = 1
+
+class BatchChecker(NaiveChecker, BaseOnlineChecker):
+    """Set of paths to check using two flip tolerant search."""
+
+    _invalid_node = -1
+    _debug = False
+
+    def __init__(self):
+        super().__init__()
+
+    #### Utility functions ####
+
+    def vectorize(self, path):
+        vector = [0 for i in range(len(self.graph)**2)]
+        for i in range(len(path)-1):
+            vector[path[i]][path[i+1]] = (vector[path[i]][path[i+1]] + 1) % 2
+        return vector
+
+    def findCoTreePairs(self, node):
+        """Find co tree starting at node."""
+        allNodes = self.getSuccessors(node)
+        tree = self.getTree(node)
+        coTree = {}
+        for src in allNodes:
+            for snk in allNodes:
+                if self.graph[src][snk] != _invalid_node and (src, snk) not in tree:
+                    pair = (
+                        self.getTreePath(node, snk, tree),
+                        self.getTreePath(node, src, tree) + [snk]
+                    )
+                    coTree.add(pair)
+        return coTree
+    
+    def getTreePath(self, src, snk, tree):
+        """Reconstruct path in tree from source to sink given tree edge set."""
+        currentNode = snk
+        path = [snk]
+        while currentNode != snk:
+            edge = self.getBackEdge(snk, tree)
+            currentNode = edge[0]
+            path += [currentNode]
+        return path
+    
+    def getBackEdge(self, snk, st):
+        """Get some edge (*, snk) from set st."""
+        for edge in st:
+            if edge[1] == snk:
+                return edge
+
+    def findPath(self, source: int, sink: int) -> Tuple[bool, List[int]]:
+        """Find path from source ot sink in old graph,
+        without involving new edge."""
[…]

### 2020-07-28 implementation/pathCheckers/baseOnlineChecker.py
A little more fixing

@@ -135,14 +135,15 @@ class BaseOnlineChecker(IPathChecker):
     def getTree(self, node: int):
         """Similar to get all successors, but returns a spanning tree in 
         the form of a set of edges."""
-        if not self.compactFwdGraph:
-            self.buildCompactFwdGraph()
+        # if not self.compactFwdGraph:
+        #     self.buildCompactFwdGraph()
+        self.buildCompactFwdGraph()
         visited = set()
         # maintain a separate list to get ordering where closest node is first.
         visitedList = []
         treeEdges = set()
         currentNode = node
-        previousNode = _invalid_node
+        previousNode = self._invalid_node
         stack = [node]
         path = []
         # More or less DFS.
@@ -154,12 +155,10 @@ class BaseOnlineChecker(IPathChecker):
                 visited.add(currentNode)
                 visitedList += [currentNode]
                 path += [currentNode]
-                if (node == self.newEdgeSink):
-                    self.pathsFromNewEdgeSink[currentNode] = copy.deepcopy(path)
                 stack += [self._invalid_node] + self.compactFwdGraph[currentNode]
-                if (previousNode != _invalid_node):
+                if (previousNode != self._invalid_node):
                     treeEdges.add((previousNode, currentNode))
-                    previousNode = currentNode
+                previousNode = currentNode
         if __debug__ and self._debug:
             print("getAllSuccessors: node = {} and visited list: {}".format(
                 node, visitedList))
@@ -217,6 +216,8 @@ class BaseOnlineChecker(IPathChecker):
         self.pathsToNode = dict()
         self._no_edge = _default_no_edge
         self.noIdentity = False
+        # Hack to make sub classes work
+        self._invalid_node = -1
     
     def setGraph(
         self,

### 2020-07-28 implementation/pathCheckers/batchChecker.py

@@ -24,6 +24,7 @@ class BatchChecker(NaiveChecker, BaseOnlineChecker):
 
     def __init__(self):
         super().__init__()
+        BaseOnlineChecker.__init__(self)
 
     #### Utility functions ####
 
@@ -44,17 +45,17 @@ class BatchChecker(NaiveChecker, BaseOnlineChecker):
         vector1, vector2 = (self.vectorize(path) for path in pair)
         return addVectors(vector1, vector2)
 
-    def findCoTreePairs(self, node):
-        """Find co tree starting at node."""
+    def findCoTreePairs(self, node) -> Set[Tuple[Tuple[int], Tuple[int]]]:
+        """Find co tree starting at node. Paths are tuple because hashing."""
         allNodes = self.getSuccessors(node)
         tree = self.getTree(node)
-        coTree = {}
+        coTree = set()
         for src in allNodes:
             for snk in allNodes:
-                if self.graph[src][snk] != _invalid_node and (src, snk) not in tree:
+                if self.graph[src][snk] != self._invalid_node and (src, snk) not in tree:
                     pair = (
-                        self.getTreePath(node, snk, tree),
-                        self.getTreePath(node, src, tree) + [snk]
+                        tuple(self.getTreePath(src, snk, tree)),
+                        (src, snk)
                     )
                     coTree.add(pair)
         return coTree
@@ -63,11 +64,11 @@ class BatchChecker(NaiveChecker, BaseOnlineChecker):
         """Reconstruct path in tree from source to sink given tree edge set."""
         currentNode = snk
         path = [snk]
-        while currentNode != snk:
-            edge = self.getBackEdge(snk, tree)
+        while currentNode != src:
+            edge = self.getBackEdge(currentNode, tree)
             currentNode = edge[0]
             path += [currentNode]
-        return path
+        return path[::-1]
     
     def getBackEdge(self, snk, st):
         """Get some edge (*, snk) from set st."""
@@ -118,19 +119,22 @@ class BatchChecker(NaiveChecker, BaseOnlineChecker):
                 return False
         return True
 
-    def sigma(self, subst: Set, allPairs: Set):
-        return {
+    def sigma(self, subst: List[Tuple[List, List]], allPairs: List[Tuple[List, List]]) -> List:
+        """Implementation of Sigma closure function. Using lists not sets 
[…]

### 2020-07-28 implementation/testing/testUtilities.py

@@ -179,7 +179,12 @@ class BatchTestDefinition():
         test_graph,
         paths = [],
         expected_vectors = [],
-        expected_matrix = [[]]):
+        expected_matrix = [[]],
+        co_tree_root = 0,
+        expected_co_tree = {},
+        sigma_test_subset = {},
+        all_pairs = {},
+        expected_sigma = {}):
         self.test_graph = test_graph
         self.expected_vectors = expected_vectors
         self.paths = paths
@@ -187,7 +192,11 @@ class BatchTestDefinition():
         self.pieces = self.pairs[:-1]
         self.finalPair = self.pairs[-1]
         self.expected_matrix = expected_matrix
-
+        self.co_tree_root = co_tree_root
+        self.expected_co_tree = expected_co_tree
+        self.sigma_test_subset = sigma_test_subset
+        self.all_pair = all_pairs
+        self.expected_sigma = expected_sigma
 # Test Definitions
 
 defaultTestDefinition = TestDefinition(
@@ -215,12 +224,13 @@ defaultTestDefinition = TestDefinition(
     ])
 
 # An example to test Batch algorithm.
+allPairs = [([0, 1, 3, 2],[0, 1, 2]), ([1, 3, 2],[1, 2])],
 acyclicTest = BatchTestDefinition(
     test_graph = [
-    [0, 1, 0, 0],
-    [0, 0, 1, 1],
-    [0, 0, 0, 0],
-    [0, 0, 1, 0]],
+    [-1,  1, -1, -1],
+    [-1, -1,  1,  1],
+    [-1, -1, -1, -1],
+    [-1, -1,  1, -1]],
     paths = [[0, 1, 2], [0, 1, 3, 2], [1, 3, 2], [1, 2]],
     expected_vectors = [
     [0, 1, 0, 0,
@@ -257,8 +267,12 @@ acyclicTest = BatchTestDefinition(
         [0, 0],
         [1, 1],
         [0, 0]
-    ]
-    
+    ],
+    co_tree_root = 0,
+    expected_co_tree = {((1, 3, 2), (1, 2))},
+    sigma_test_subset = [([1, 3, 2],[1, 2])],
+    all_pairs = allPairs,
+    expected_sigma = allPairs
 )
 
 testCaseOne = TestDefinition(

### 2020-07-28 implementation/pathCheckers/baseOnlineChecker.py
batch checker tested and working

@@ -143,7 +143,6 @@ class BaseOnlineChecker(IPathChecker):
         visitedList = []
         treeEdges = set()
         currentNode = node
-        previousNode = self._invalid_node
         stack = [node]
         path = []
         # More or less DFS.
@@ -156,9 +155,8 @@ class BaseOnlineChecker(IPathChecker):
                 visitedList += [currentNode]
                 path += [currentNode]
                 stack += [self._invalid_node] + self.compactFwdGraph[currentNode]
-                if (previousNode != self._invalid_node):
-                    treeEdges.add((previousNode, currentNode))
-                previousNode = currentNode
+                if (len(path) > 1):
+                    treeEdges.add((path[-2], currentNode))
         if __debug__ and self._debug:
             print("getAllSuccessors: node = {} and visited list: {}".format(
                 node, visitedList))

### 2020-07-28 implementation/pathCheckers/batchChecker.py

@@ -14,7 +14,9 @@ from typing import List, Set, Tuple, Union
 
 _default_no_edge = -1
 _source = 0
-_sink = 1
+_sink = -1
+_first = 0
+_second = 1
 
 class BatchChecker(NaiveChecker, BaseOnlineChecker):
     """Set of paths to check using two flip tolerant search."""
@@ -54,8 +56,8 @@ class BatchChecker(NaiveChecker, BaseOnlineChecker):
             for snk in allNodes:
                 if self.graph[src][snk] != self._invalid_node and (src, snk) not in tree:
                     pair = (
-                        tuple(self.getTreePath(src, snk, tree)),
-                        (src, snk)
+                        tuple(self.getTreePath(node, snk, tree)),
+                        tuple(self.getTreePath(node, src, tree) + [snk])
                     )
                     coTree.add(pair)
         return coTree
@@ -122,19 +124,25 @@ class BatchChecker(NaiveChecker, BaseOnlineChecker):
     def sigma(self, subst: List[Tuple[List, List]], allPairs: List[Tuple[List, List]]) -> List:
         """Implementation of Sigma closure function. Using lists not sets 
         because paths are unhashable lists."""
-        return [
-            pair for pair in allPairs if self.gaussianReduce(self.matrixify(
-                self.smallerPiecesIntersection(subst,
-                pair), pair)) ]
+        # return [
+        #     pair for pair in allPairs if self.gaussianReduce(self.matrixify(
+        #         self.smallerPiecesIntersection(subst,
+        #         pair), pair)) ]
+        result = []
+        for pair in allPairs:
+            if self.gaussianReduce(self.matrixify(self.smallerPiecesIntersection(
+                subst, pair), pair)):
+                result.append(pair)
+        return result
     
     def smallerPiecesIntersection(self, subst: List[Tuple[List, List]],
         pair: Tuple[List, List]):
         """Get subset of subst such that the members are smaller than pair. 
         (<> function intersected with subset)."""
-        sources = self.getSuccessors(pair[_source])
-        sinks = self.getPredecessors(pair[_sink])
-        return [bilinking for bilinking in subst if bilinking[_source] in sources 
-            and bilinking[_sink] in sinks]
+        sources = self.getSuccessors(pair[_first][_source])
+        sinks = self.getPredecessors(pair[_second][_sink])
[…]

### 2020-07-28 implementation/testing/testUtilities.py

@@ -183,8 +183,9 @@ class BatchTestDefinition():
         co_tree_root = 0,
         expected_co_tree = {},
         sigma_test_subset = {},
-        all_pairs = {},
-        expected_sigma = {}):
+        all_pairs = [],
+        expected_sigma = {},
+        expected_batch_result = []):
         self.test_graph = test_graph
         self.expected_vectors = expected_vectors
         self.paths = paths
@@ -195,8 +196,9 @@ class BatchTestDefinition():
         self.co_tree_root = co_tree_root
         self.expected_co_tree = expected_co_tree
         self.sigma_test_subset = sigma_test_subset
-        self.all_pair = all_pairs
+        self.all_pairs = all_pairs
         self.expected_sigma = expected_sigma
+        self.expected_batch_result = expected_batch_result
 # Test Definitions
 
 defaultTestDefinition = TestDefinition(
@@ -224,10 +226,10 @@ defaultTestDefinition = TestDefinition(
     ])
 
 # An example to test Batch algorithm.
-allPairs = [([0, 1, 3, 2],[0, 1, 2]), ([1, 3, 2],[1, 2])],
+allPairs = [([0, 3, 2],[0, 1, 2]), ([1, 3, 2],[1, 2]), ([0, 3], [0, 1, 3])]
 acyclicTest = BatchTestDefinition(
     test_graph = [
-    [-1,  1, -1, -1],
+    [-1,  1, -1,  1],
     [-1, -1,  1,  1],
     [-1, -1, -1, -1],
     [-1, -1,  1, -1]],
@@ -269,10 +271,12 @@ acyclicTest = BatchTestDefinition(
         [0, 0]
     ],
     co_tree_root = 0,
-    expected_co_tree = {((1, 3, 2), (1, 2))},
-    sigma_test_subset = [([1, 3, 2],[1, 2])],
+    expected_co_tree = {((0, 3), (0, 1, 3)), ((0, 3, 2), (0, 1, 2))},
+    sigma_test_subset = [([1, 3, 2],[1, 2]), ([0, 1, 3], [0, 3])],
+    # IDK why but there's an extra tuple sign.
     all_pairs = allPairs,
-    expected_sigma = allPairs
+    expected_sigma = allPairs,
+    expected_batch_result = [([1, 3, 2],[1, 2]), ([0, 3], [0, 1, 3])] 
 )
 
 testCaseOne = TestDefinition(

### 2020-07-28 implementation/evaluator.py
Added some evaluation code

@@ -2,12 +2,13 @@ from datetime import datetime
 import functools
 import matplotlib.pyplot as plt
 import numpy as np
+from pathCheckers.batchChecker import BatchChecker
 from pathCheckers.iPathChecker import IPathChecker
 from pathCheckers.optimalSetPathChecker import OptimalSetPathChecker
 from pathCheckers.polynomialPathChecker import PolynomialPathChecker
 from pathCheckers.naiveChecker import NaiveChecker
 from pathCheckers.twoFlipChecker import TwoFlipPathChecker
-from randomGraphGenerator import generateGraph
+from randomGraphGenerator import generateGraph, generateAcyclicGraph
 import traceback
 from typing import List
 
@@ -18,7 +19,7 @@ sizeStep = 2
 maxSize = 10
 
 # densities = [i*densityStep for i in range(1, int(1.0/densityStep))] 
-densities = [0.1, 0.5, 0.9]
+densities = [0.1, 0.4]
 sizes = [i*sizeStep for i in range(1, int(maxSize/sizeStep))]
 # sizes = [10, 50, 100]
 
@@ -26,9 +27,10 @@ evaluationList = [(density, size) for size in sizes for density in densities]
 
 checkers : List[IPathChecker] = [
     #PolynomialPathChecker(),
-    #OptimalSetPathChecker(),
-    NaiveChecker(),
-    #TwoFlipPathChecker()
+    OptimalSetPathChecker(),
+    # NaiveChecker(),
+    #TwoFlipPathChecker(),
+    BatchChecker()
 ]
 
 class attemptInfo:
@@ -54,11 +56,25 @@ class attemptInfo:
         return self.avgList(self.sizes)
 
 
-def getEvaluateFunction(checker: IPathChecker, numTries: int):
+def getEvaluateFunction(checker: IPathChecker, numTries: int, acyclic=False):
+    if isinstance(checker, BatchChecker):
+        def evaluate(density, size):
+            info = attemptInfo()
+            for attempt in range(numTries):
+                inp = generateAcyclicGraph(density, size)
+                checker.setGraph(inp.graph)
+                paths = checker.getPathsToCheck()
+                info.addSize(len(paths))
+                info.addTime(checker.getComputeTime())
+            return info
+        return evaluate
     def evaluate(density, size):
         info = attemptInfo()
         for attempt in range(numTries):
-            inp = generateGraph(density, size)
+            if (acyclic):
+                inp = generateAcyclicGraph(density, size)
[…]

### 2020-07-28 implementation/randomGraphGenerator.py

@@ -1,7 +1,9 @@
 """Define a script that generates random graphs."""
 import random
+from typing import List, Tuple
 
 NO_EDGE = -1
+_EDGE = 1
 
 class evaluationGraph:
     """Type for input to evaluation script."""
@@ -11,6 +13,8 @@ class evaluationGraph:
         self.newEdge = newEdge
         self.newEdgeSource = newEdge[0]
         self.newEdgeSink = newEdge[1]
+        self.graphWithEdge = graph
+        self.graphWithEdge[self.newEdgeSource][self.newEdgeSink] = _EDGE
 
 def generateGraph(density: float, nodes: int) -> evaluationGraph:
     """generate a random graph of the given density, with a given number of 
@@ -23,5 +27,21 @@ def generateGraph(density: float, nodes: int) -> evaluationGraph:
     randomList = random.sample(allPaths, max(int(density * nodes**2), 1))
     for edge in randomList[1:]:
         (src, snk) = edge
-        graph[src][snk] = 1
+        graph[src][snk] = _EDGE
+    return evaluationGraph(graph, randomList[0])
+
+def generateAcyclicGraph(density: float, nodes: int) -> evaluationGraph:
+    """Generate a random acyclic graph of the given density, with a given number 
+    of nodes along with a new edge to add to the graph.
+    Density = number of edges / total possible number of edges, i.e. n^2."""
+    # Generate a graph that is a lower triangular matrix.
+    # https://mathematica.stackexchange.com/questions/608/how-to-generate-random-directed-acyclic-graphs.
+    allPaths = [(src, snk) for src in range(nodes) for snk in range(src)]
+    graph = [[NO_EDGE for col in range(nodes)] for row in range(nodes)]
+    #TODO handle density>0.5
+    randomList = random.sample(allPaths, min(max(int(density * nodes**2), 1), len(allPaths)))
+    #TODO factor out addition to graph.
+    for edge in randomList[1:]:
+        (src, snk) = edge
+        graph[src][snk] = _EDGE
     return evaluationGraph(graph, randomList[0])

### 2020-07-28 acmPaper/paper.tex
Add fncylab to use labelformat
Murota implementation:
Added a stub for evaluation

@@
 \usepackage{xxx}
+\usepackage{fncylab}
 %
@@
     output $\gets$ \{\}\;
-    \For{bilinking $\in$ Rs}{
+    \For{bilinking $\in$ S}{
         smallerPairs $\gets$ \{\}\;
@@
 Averaging over evaluation for the first 30 days of 2020, building and verifying a diagram to completion (inclusive of the time required by network calls) took 243$\pm$19 seconds using algorithm ~\ref{algo_online_minimal}, and in 133$\pm$13 seconds with algorithm ~\ref{algo_online_polynomial}.
-Computation was performed on a Macbook Pro 2015, 2.9 GHz Dual-Core Intel Core i5.
-More information about implementation can be found in the evaluation section.
 
 \section{Evaluation}
+We compare performance of the following path checking algorithms: (1) naive baseline (2) less naive, two-flip baseline (3) batch baseline (4) algorithm ~\ref{algo_online_polynomial}, and (5) algorithm ~\ref{algo_online_minimal}.
+The two aspects we look at are time for response and size of response set (smaller sets -- tighter output results -- would mean less calls to the oracle).
+We use randomly generated graphs for benchmarking, and vary their size.
+All computation was performed on a Macbook Pro 2015, 2.9 GHz Dual-Core Intel Core i5.
+
+The naive baseline performs poorly, taking well over a thousand seconds for even small graphs of 10 nodes.
+The batch algorithm brings timing down to seconds. Algorithm ~\ref{algo_online_polynomial} performs only slightly better.
+Surprisingly, the optimal set algorithm cuts time cost by several orders of magnitude.
+All implementations are sensitive to density, performing better when density is very low or very high.
+
+%TODO add figures
+
+\begin{tabular}{|c|c|}
+    \hline
+    Algorithm & average seconds taken for ten nodes, density close to 0.5 \\
+    \hline
+    Naive baseline & 1,500 \\
+    Batch algorithm & 1.4 \\
+    algorithm ~\ref{algo_online_polynomial} & 1.2 \\
+    algorithm ~\ref{algo_online_minimal} & 0.0006 \\
+    \hline
+\end{tabular}
 

### 2020-07-31 acmPaper/paper.tex
sketch of evaluation

@@
 \usepackage{fncylab}
+\usepackage{subcaption}
 %
@@
 We use randomly generated graphs for benchmarking, and vary their size.
+Given a graph and a new edge, we time how long it takes for an algorithm to return 
 All computation was performed on a Macbook Pro 2015, 2.9 GHz Dual-Core Intel Core i5.
@@
 The batch algorithm brings timing down to seconds. Algorithm ~\ref{algo_online_polynomial} performs only slightly better.
-Surprisingly, the optimal set algorithm cuts time cost by several orders of magnitude.
+Surprisingly, the optimal set algorithm cuts time cost by several orders of magnitude, and runs in milliseconds for small graphs.
 All implementations are sensitive to density, performing better when density is very low or very high.
 
-%TODO add figures
+\section{Comparison of algorithm time cost}
+The average time taken by each algorithm over the course of 10 runs over randomly generated graphs with 8 nodes and 32 edges is listed in table ~\ref{times}.
 
+%TODO get this data.
 \begin{tabular}{|c|c|}
     \hline
-    Algorithm & average seconds taken for ten nodes, density close to 0.5 \\
+    Algorithm & average seconds of computation \\
     \hline
     Naive baseline & 1,500 \\
+    Two Flip tolerant & 1 \\
     Batch algorithm & 1.4 \\
     algorithm ~\ref{algo_online_polynomial} & 1.2 \\
-    algorithm ~\ref{algo_online_minimal} & 0.0006 \\
+    algorithm ~\ref{algo_online_minimal} & 0.0006 
     \hline
+    \label{times}
+\end{tabular}
+
+\section{Scaling of time with input size}
+
+Algorithms scale as expected with time, as seen in figure ~\ref{fig:timeVsSize}. 
+Both algorithm ~\ref{algo_online_minimal} and ~\ref{algo_minimal_spanning_set} exhibit graphs that are polynomial in appearance, as does the batch baseline.
+\xxx[AK]{Do I need to justify this by doing some kind of fit?}
+The naive baseline as well as the two flip tolerant baseline display catastrophic growth that quickly makes computation untenable.
+
+Further, run time is found to relate to the density of edges in the input graph.
+We define density to be the ratio of the number of edges in the graph to the total possible number of edges ($|V|^2$, where $|V|$ is the number of nodes).
+Generally, time cost is found to be higher at middling densities and lower as density approaches 0 or 1.
+
+\begin{figure}
+    % TODO format nicely.
+    \begin{subfigure}{.2\textwidth}
+      \includegraphics[width=.8\linewidth]{timeVsSize_10_OptimalSet.png}
+      \caption{~\ref{algo_online_minimal}}
+      \label{fig:sfig1}
+    \end{subfigure}
+    \begin{subfigure}{.2\textwidth}
+      \includegraphics[width=.8\linewidth]{timeVsSize_10_Polynomial.png}
+      \caption{~\ref{algo_online_polynomial}}
+      \label{fig:sfig2}
+    \end{subfigure}
+    \caption{Plots of scaling of time with input size and density}
+    \label{fig:timeVsSize}
+    %TODO add plots for two flip, naive, batch checker
+\end{figure}
+
+\section{Variance}
+For all algorithms, there is considerable spread in the time taken for a given graph size, varying with graph structure.
+The plots in figure ~\ref{fig:timeVsSize} %TODO replace
+depict the situation.
+After computation on 10 randomly generated graphs of density 0.5, each outcome is plotted as a point and error bars indicate standard deviation.
+
+%TODO add a figure for each of the checkers.
+
+\section{Size of Output}
+Output size is found to be along similar lines to time cost.
+Table ~\ref{sizes}, the size analog of table ~\ref{times} summarizes the number of output pairs that the algorithms returned on average over 10 runs, for graphs with 8 nodes and 32 edges.
+% TODO get this data.
+\begin{tabular}{|c|c|}
+    \hline
+    Algorithm & average number of output pairs \\
+    \hline
+    Naive baseline & 1,500 \\
+    Two Flip tolerant & 1 \\
+    Batch algorithm & 1.4 \\
+    algorithm ~\ref{algo_online_polynomial} & 1.2 \\
+    algorithm ~\ref{algo_online_minimal} & 0.0006
+    \hline
+    \label{sizes}
 \end{tabular}

### 2020-08-02 acmPaper/paper.tex
Evaluation draft

--- 2020-07-28 acmPaper/paper.tex
@@
 The batch algorithm brings timing down to seconds. Algorithm ~\ref{algo_online_polynomial} performs only slightly better.
-Surprisingly, the optimal set algorithm cuts time cost by several orders of magnitude.
+Surprisingly, the optimal set algorithm cuts time cost by several orders of magnitude, and runs in milliseconds for small graphs.
 All implementations are sensitive to density, performing better when density is very low or very high.
 
-%TODO add figures
+\subsection{Comparison of algorithm time cost}
+The average time taken by each algorithm over the course of 10 runs over randomly generated graphs with 9 nodes and 32 edges is listed in table ~\ref{times}.
 
+\begin{figure}
 \begin{tabular}{|c|c|}
     \hline
-    Algorithm & average seconds taken for ten nodes, density close to 0.5 \\
+    Algorithm & average seconds of computation \\
     \hline
-    Naive baseline & 1,500 \\
-    Batch algorithm & 1.4 \\
-    algorithm ~\ref{algo_online_polynomial} & 1.2 \\
-    algorithm ~\ref{algo_online_minimal} & 0.0006 \\
+    Naive baseline & 0.77 \\
+    Two Flip tolerant & 0.075 \\
+    Batch algorithm & 7.55 \\
+    ~\ref{algo_online_polynomial} & 0.0038 \\
+    ~\ref{algo_online_minimal} & 0.00086 \\
     \hline
 \end{tabular}
+\label{times} %TODO debug and fix label not working.
+\caption{Computation time for 9 node graph of density 0.4, averaged over ten runs}
+\end{figure}
+
+\subsection{Scaling of time with input size}
+
+Algorithms scale as expected with time, as seen in figure ~\ref{fig:timeVsSize}. 
+Both algorithm ~\ref{algo_online_minimal} and ~\ref{algo_minimal_spanning_set} exhibit graphs that are polynomial in appearance, as does the batch baseline.
+\xxx[AK]{Do I need to justify this by doing some kind of fit?}
+The naive baseline as well as the two flip tolerant baseline display growth that quickly makes computation untenable.
+The batch algorithm also grows quickly.
+
+We define density to be the ratio of the number of edges in the graph to the total possible number of edges ($|V|^2$, where $|V|$ is the number of nodes).
+Run time relates to the density of edges in the input graph.
+The degree of the effect differs with the algorithms, as figure ~\ref{fig:timeVsSize} shows.
+Generally, denser graphs entail longer computation time.
+% Our proposed algorithms, ~\ref{algo_minimal_spanning_set} and ~\ref{algo_online_polynomial} are less affected because they aggressively remove redundancies.
+
+\begin{figure}
+    % TODO format nicely.
+    \begin{subfigure}{.4\textwidth}
+      \includegraphics[width=.8\linewidth]{timeVsSize_10_OptimalSet.png}
+      \caption{~\ref{algo_online_minimal}}
+      \label{fig:sfigOptimal}
+    \end{subfigure}
+
+    \begin{subfigure}{.4\textwidth}
+      \includegraphics[width=.8\linewidth]{timeVsSize_10_Polynomial.png}
+      \caption{~\ref{algo_online_polynomial}}
+      \label{fig:sfigPolynomial}
+    \end{subfigure}
+
+    \begin{subfigure}{.4\textwidth}
+      \includegraphics[width=.8\linewidth]{timeVsSize_10_NaiveChecker.png}
+      \caption{Naive baseline}
+      \label{fig:sfigNaive}
+    \end{subfigure}
+
+    \begin{subfigure}{.4\textwidth}
+      \includegraphics[width=.8\linewidth]{timeVsSize_10_TwoFlipPathChecker.png}
+      \caption{Two flip tolerant baseline}
+      \label{fig:sfigTwoFlip}
+    \end{subfigure}
+
+    \begin{subfigure}{.4\textwidth}
+      \includegraphics[width=.8\linewidth]{timeVsSize_10_BatchChecker.png}
+      \caption{Batch algorithm baseline}
+      \label{fig:sfigBatch}
+    \end{subfigure}
+
+    \label{fig:timeVsSize}
+    \caption{Plots of scaling of time with input size and density}
+\end{figure}
+
+\subsection{Variance}
+The periodic spikes in figure ~\ref{fig:sfigOptimal} are striking. 
+We plot the spread of results to understand what is happening.
+We find ~\ref{algo_online_minimal} has outliers about two standard deviation above the mean responsible for the spikes in the average.
+The outliers themselves follow a polynomial curve.
+The plots in figure ~\ref{fig:variance} depict the situation.
+Grey points are the results of evaluation on individual points, and error bars show standard deviation. 
+The black curve traces the mean.
+No such effect is observed in ~\ref{algo_online_polynomial}.
+\begin{figure}
+    % TODO format nicely.
+    \begin{subfigure}{.4\textwidth}
+      \includegraphics[width=.8\linewidth]{variance_10_OptimalSet.png}
+      \caption{~\ref{algo_online_minimal}}
+      \label{fig:sfigOptimal}
+    \end{subfigure}
+
+    \begin{subfigure}{.4\textwidth}
+      \includegraphics[width=.8\linewidth]{variance_10_Polynomial.png}
+      \caption{~\ref{algo_online_polynomial}}
+      \label{fig:sfigPolynomial}
+    \end{subfigure}
+
+    \begin{subfigure}{.4\textwidth}
+      \includegraphics[width=.8\linewidth]{variance_10_TwoFlip.png}
+      \caption{Two flip tolerant baseline}
+      \label{fig:sfigTwoFlip}
+    \end{subfigure}
+
+    \label{fig:variance}
+    \caption{Plots of spread of results}
+\end{figure}
+
+\subsection{Size of Output}
+Output size is a metric of interest, should the equality checking oracle be expensive.
+Table ~\ref{sizes}, summarizes the number of output pairs that the algorithms returned on average over 10 runs, for graphs with 9 nodes and 32 edges.
+\begin{figure}
+\begin{tabular}{|c|c|}
+    \hline
+    Algorithm & average number of output pairs \\
+    \hline
+    Naive baseline & 39754.9 \\
+    Two Flip tolerant & 748.9 \\
+    Batch algorithm & 23 \\
+    algorithm ~\ref{algo_online_polynomial} & 78.3 \\
+    algorithm ~\ref{algo_online_minimal} & 1 \\
+    \hline
+\end{tabular}
+\label{sizes} % Now working. TODO fix label.
+\caption{Output size for 9 node graph of density 0.4, averaged over ten runs}
+\end{figure}
 

### 2020-08-02 implementation/evaluator.py

@@ -1,5 +1,6 @@
 from datetime import datetime
 import functools
+import json
 import matplotlib.pyplot as plt
 import numpy as np
 from pathCheckers.batchChecker import BatchChecker
@@ -12,24 +13,29 @@ from randomGraphGenerator import generateGraph, generateAcyclicGraph
 import traceback
 from typing import List
 
-NUM_TRIES = 1
+NUM_TRIES = 10
+DO_ACYCLIC = True
+OUTFILE = ''
 
 densityStep = 0.5
-sizeStep = 2
-maxSize = 10
+sizeStep = 1
+maxSize = 12
 
 # densities = [i*densityStep for i in range(1, int(1.0/densityStep))] 
-densities = [0.1, 0.4]
-sizes = [i*sizeStep for i in range(1, int(maxSize/sizeStep))]
-# sizes = [10, 50, 100]
+densities = [0.4]
+# sizes = [i*sizeStep for i in range(1, int(maxSize/sizeStep))]
+sizes = [9]
 
 evaluationList = [(density, size) for size in sizes for density in densities]
 
+# First three colors look good in black and white.
+colors = ["black", "purple", "orange", "blue", "green", "red"]
+
 checkers : List[IPathChecker] = [
-    #PolynomialPathChecker(),
-    OptimalSetPathChecker(),
+    # PolynomialPathChecker(),
+    # OptimalSetPathChecker(),
     # NaiveChecker(),
-    #TwoFlipPathChecker(),
+    # TwoFlipPathChecker(),
     BatchChecker()
 ]
 
@@ -47,34 +53,75 @@ class attemptInfo:
         self.sizes += [sze]
     
     def avgList(self, lst):
-        return functools.reduce(lambda  acc, n: acc + n, lst, 0)/len(lst)
+        self.avg = functools.reduce(lambda  acc, n: acc + n, lst, 0)/len(lst)
+        return self.avg
     
-    def getAvgTime(self):
+    def getAvgTime(self, median=False): 
+        if median:
+            return self.getMedianTime()
         return self.avgList(self.times)
 
     def getAvgSizes(self):
         return self.avgList(self.sizes)
+    
+    def getTimes(self):
+        return self.times
 
+    def getMedianTime(self):
+        copy = self.times
+        copy.sort()
+        if len(copy)%2 == 1:
+            self.median = copy[int(len(copy)/2)+1]
+        else:
+            floor = int(len(copy)/2)
+            self.median = (copy[floor] + copy[floor+1])/2
[…]

### 2020-08-02 implementation/pathCheckers/batchChecker.py

@@ -164,6 +164,16 @@ class BatchChecker(NaiveChecker, BaseOnlineChecker):
 
     #### Public interface ####
 
+    def setGraph(
+        self,
+        graph: List[List[int]],
+        noEdge: int = _default_no_edge) -> None:
+        """Set the graph to which an edge is to be added."""
+        self.graph = graph
+        self._no_edge = noEdge
+        self.buildCompactBkdGraph()
+        self.buildCompactFwdGraph()
+
     def getPathsToCheck(self) ->  Set[Tuple[List[int], List[int]]]:
         """Return the pairs of path whose equality implies path independence of 
         the new graph."""

### 2020-08-02 implementation/randomGraphGenerator.py

@@ -16,6 +16,12 @@ class evaluationGraph:
         self.graphWithEdge = graph
         self.graphWithEdge[self.newEdgeSource][self.newEdgeSink] = _EDGE
 
+class batchEvaluationGraph:
+    """Type for input to evaluation script for batchChecker (no new edge)."""
+
+    def __init__(self, graph):
+        self.graph = graph
+
 def generateGraph(density: float, nodes: int) -> evaluationGraph:
     """generate a random graph of the given density, with a given number of 
     nodes along with a new edge to add to the graph.
@@ -36,7 +42,9 @@ def generateAcyclicGraph(density: float, nodes: int) -> evaluationGraph:
     Density = number of edges / total possible number of edges, i.e. n^2."""
     # Generate a graph that is a lower triangular matrix.
     # https://mathematica.stackexchange.com/questions/608/how-to-generate-random-directed-acyclic-graphs.
-    allPaths = [(src, snk) for src in range(nodes) for snk in range(src)]
+    allPaths = []
+    for src in range(1, nodes):
+        allPaths += [(src, snk) for snk in range(src)]
     graph = [[NO_EDGE for col in range(nodes)] for row in range(nodes)]
     #TODO handle density>0.5
     randomList = random.sample(allPaths, min(max(int(density * nodes**2), 1), len(allPaths)))
@@ -44,4 +52,4 @@ def generateAcyclicGraph(density: float, nodes: int) -> evaluationGraph:
     for edge in randomList[1:]:
         (src, snk) = edge
         graph[src][snk] = _EDGE
-    return evaluationGraph(graph, randomList[0])
+    return batchEvaluationGraph(graph)

### 2020-08-02 acmPaper/paper.tex
half a sentence

@@
 We use randomly generated graphs for benchmarking, and vary their size.
-Given a graph and a new edge, we time how long it takes for an algorithm to return 
+Given a graph and a new edge, we time how long it takes for an algorithm to return  the set of pairs that need to be verified.
 All computation was performed on a Macbook Pro 2015, 2.9 GHz Dual-Core Intel Core i5.

### 2020-08-02 acmPaper/paper.tex
fix labeling/refs

@@
 \subsection{Comparison of algorithm time cost}
-The average time taken by each algorithm over the course of 10 runs over randomly generated graphs with 9 nodes and 32 edges is listed in table ~\ref{times}.
+The average time taken by each algorithm over the course of 10 runs over randomly generated graphs with 9 nodes and 32 edges is listed in table ~\ref{tab:times}.
 
-\begin{figure}
+\begin{table}
 \begin{tabular}{|c|c|}
@@
 \end{tabular}
-\label{times} %TODO debug and fix label not working.
 \caption{Computation time for 9 node graph of density 0.4, averaged over ten runs}
-\end{figure}
+\label{tab:times} %TODO debug and fix label not working.
+\end{table}
 
@@
 Algorithms scale as expected with time, as seen in figure ~\ref{fig:timeVsSize}. 
-Both algorithm ~\ref{algo_online_minimal} and ~\ref{algo_minimal_spanning_set} exhibit graphs that are polynomial in appearance, as does the batch baseline.
+Both ~\ref{algo_online_minimal} and ~\ref{algo_minimal_spanning_set} exhibit graphs that are polynomial in appearance, as does the batch baseline.
 \xxx[AK]{Do I need to justify this by doing some kind of fit?}
 The naive baseline as well as the two flip tolerant baseline display growth that quickly makes computation untenable.
-The batch algorithm also grows quickly.
+The batch algorithm also grows fast, though not as much as the online checking baselines.
 
@@
 Generally, denser graphs entail longer computation time.
+For the batch algorithm we use lower densities since the input graph must be acyclic. This puts an upper bound on density that approaches 0.5 in large graphs.
 % Our proposed algorithms, ~\ref{algo_minimal_spanning_set} and ~\ref{algo_online_polynomial} are less affected because they aggressively remove redundancies.
@@
       \caption{~\ref{algo_online_minimal}}
-      \label{fig:sfigOptimal}
+      \label{fig:sfigOptimalTvsS}
     \end{subfigure}
@@
       \caption{~\ref{algo_online_polynomial}}
-      \label{fig:sfigPolynomial}
+      \label{fig:sfigPolynomialTvsS}
     \end{subfigure}
@@
       \caption{Naive baseline}
-      \label{fig:sfigNaive}
+      \label{fig:sfigNaiveTvsS}
     \end{subfigure}
@@
       \caption{Two flip tolerant baseline}
-      \label{fig:sfigTwoFlip}
+      \label{fig:sfigTwoFlipTvsS}
     \end{subfigure}
@@
       \caption{Batch algorithm baseline}
-      \label{fig:sfigBatch}
+      \label{fig:sfigBatchTvsS}
     \end{subfigure}
-
+    \caption{Plots of scaling of time with input size and density}
     \label{fig:timeVsSize}
-    \caption{Plots of scaling of time with input size and density}
 \end{figure}
@@
 \subsection{Variance}
-The periodic spikes in figure ~\ref{fig:sfigOptimal} are striking. 
+The periodic spikes in figure ~\ref{fig:sfigOptimalTvsS} are striking. 
 We plot the spread of results to understand what is happening.
@@
       \caption{~\ref{algo_online_minimal}}
-      \label{fig:sfigOptimal}
+      \label{fig:sfigOptimalSpread}
     \end{subfigure}
@@
       \caption{~\ref{algo_online_polynomial}}
-      \label{fig:sfigPolynomial}
+      \label{fig:sfigPolynomialSpread}
     \end{subfigure}
@@
       \caption{Two flip tolerant baseline}
-      \label{fig:sfigTwoFlip}
+      \label{fig:sfigTwoFlipSpread}
     \end{subfigure}
 
+    \caption{Plots of spread of results}
     \label{fig:variance}
-    \caption{Plots of spread of results}
 \end{figure}
@@
 Output size is a metric of interest, should the equality checking oracle be expensive.
-Table ~\ref{sizes}, summarizes the number of output pairs that the algorithms returned on average over 10 runs, for graphs with 9 nodes and 32 edges.
-\begin{figure}
+Table ~\ref{tab:sizes}, summarizes the number of output pairs that the algorithms returned on average over 10 runs, for graphs with 9 nodes and 32 edges.
+\begin{table}
 \begin{tabular}{|c|c|}
@@
     Batch algorithm & 23 \\
-    algorithm ~\ref{algo_online_polynomial} & 78.3 \\
-    algorithm ~\ref{algo_online_minimal} & 1 \\
+    ~\ref{algo_online_polynomial} & 78.3 \\
+    ~\ref{algo_online_minimal} & 1 \\
     \hline
 \end{tabular}
-\label{sizes} % Now working. TODO fix label.
 \caption{Output size for 9 node graph of density 0.4, averaged over ten runs}
-\end{figure}
+\label{tab:sizes}
+\end{table}
 

### 2020-08-04 acmPaper/paper.tex
section 2

@@
 Previous work provides an algorithm for the special case of acyclic diagrams that verifies that a given diagram commutes in $O(|V|^4|E|^2)$ time, where V is the set of vertices in the diagram, and E, the set of edges.
-We present an algorithm that given a diagram and an edge to add to it, verifies that the resultant diagram commutes in $O(|V|(|E| + |V|)$ time.
-For the case when checking the equality of paths is expensive, we present an optimization that runs in $O(V^4)$ time but reduces to the minimum possible number of equality checks.
+We present an algorithm that given a diagram and an edge to add to it, and an oracle to check the equality of a given pair of paths, verifies that the resultant diagram commutes in $O(|V|^2(|E| + |V|))$ time.
+For the case when checking the equality of paths is expensive, we present an optimization that runs in $O(|V|^4)$ time but reduces to the minimum possible number of equality checks.
 We evaluate the algorithms against baselines, and see its practical application in the compiler of a domain specific language for geometry types.
@@
 \section{Introduction}
-Many systems can be understood as diagrams: graphs where nodes represent objects, and edges, directed transformation functions between nodes.
-In programming, for example, type systems with coercions can be drawn as a graph with type nodes and coercion edges.
+Many systems can be understood as diagrams: graphs where nodes represent domains, and edges, directed transformation functions between nodes.
+In programming, for example, type systems with coercions can be drawn as a graph with type nodes and coercion edges, as demonstrated in figure ~\ref{fig:typeExample}.
 A desirable trait of such a type system and many other similar systems is that traversing the graph from a given start and end node results in the same value independent of path chosen.
-A diagram with this property is said to commute.
+A diagram with this property is said to \textit{commute}.
 With our type coercion example, it could be problematic if casting to a supposedly equivalent type as an intermediate step resulted in a different answer than a direct cast.
+
+\begin{figure}[]
+    \centering
+    \begin{subfigure}{.4\textwidth}
+        \begin{verbatim}
+            var a : meters = 1;
+            var b : miles = (miles) a;
+            var c : feet = (feet) a;
+            define wugs:
+                1 miles = 10 wugs;
+                1 foot = 10000 wugs;
+            var d : wugs = (wugs) a;
+        \end{verbatim}
+        \caption{A sample program with user defined type conversion.}
+    \end{subfigure}
+
+    \begin{subfigure}{.4\textwidth}
+        \begin{tikzpicture}[scale=0.2]
+            \tikzstyle{every node}+=[inner sep=0pt]
+            \draw [black] (32.8,-8.8) circle (3);
+            \draw (32.8,-8.8) node {$meters$};
+            \draw [black] (21.5,-21.4) circle (3);
+            \draw (21.5,-21.4) node {$feet$};
+            \draw [black] (43.7,-21.4) circle (3);
+            \draw (43.7,-21.4) node {$miles$};
+            \draw [black] (32.4,-33.6) circle (3);
+            \draw (32.4,-33.6) node {$wugs$};
+            \draw [black] (30.8,-11.03) -- (23.5,-19.17);
+            \fill [black] (23.5,-19.17) -- (24.41,-18.9) -- (23.66,-18.24);
+            \draw (26.61,-13.64) node [left] {$\times 3.28$};
+            \draw [black] (34.76,-11.07) -- (41.74,-19.13);
+            \fill [black] (41.74,-19.13) -- (41.59,-18.2) -- (40.84,-18.85);
+            \draw (43.7,-13) node [left] {$\times 6.21 \times 10^{-4}$};
+            \draw [black] (23.5,-23.64) -- (30.4,-31.36);
+            \fill [black] (30.4,-31.36) -- (30.24,-30.43) -- (29.5,-31.1);
+            \draw (26.41,-28.96) node [left] {$\times 10$};
+            \draw [black] (41.66,-23.6) -- (34.44,-31.4);
+            \fill [black] (34.44,-31.4) -- (35.35,-31.15) -- (34.62,-30.47);
+            \draw (37.52,-26.04) node [left] {$\times 10^4$};
+            \draw [dashed] (29.441,-34.049) arc (-88.02206:-273.82603:12.915);
+            \fill [black] (29.44,-34.05) -- (28.62,-33.58) -- (28.66,-34.58);
+            \draw (15.56,-20.93) node [left] {$\times 32.8$};
+            \draw [dashed] (35.77,-8.425) arc (90.49177:-92.33986:12.828);
+            \fill [black] (35.36,-34.07) -- (36.13,-34.6) -- (36.18,-33.6);
+            \draw (49.23,-21.46) node [right] {$\times 6.21$};
+            \end{tikzpicture}
+        \caption{Diagram for the type conversions in the program}
+    \end{subfigure}
+
+    \caption{In this sample program, the user implicitly defines two ways to cast variable a from meters to the new unit wugs.
+    The definitions are different, and a compiler performing implicit conversion would not know which to choose.}
+    \label{fig:typeExample}
+\end{figure}
 
@@
 Since diagrams may change over time, and verifying the entire system from scratch may be computationally expensive, it would be additionally convenient if new edges could be checked online as they are added.
+In figure ~\ref{fig:typeExample}, for example, an interpreter that caught the first instance of a user making a bad definition would need to build and verify the conversion graph at run time.
 
-Naively checking if all paths in a given diagram are path independent could require a number of function equality checks as bad as factorial in the number of nodes, since a path consists of an ordering of nodes.
+Naively checking if all path pairs that begin and end at the same node in a given diagram commute could require a number of function equality checks that grow as factorial in the number of nodes, since a path consists of an ordering of nodes.
 Further, the presence of cycles implies a potentially infinite number of paths. 
+Previous work ~\cite{commutative} has identified an $O(|E|^2|V|^4)$ algorithm to verify that a complete acyclic diagram commutes; however, it addresses neither online addition nor cyclic diagrams.
 
@@
 Because the diagram commutes, all the paths between a given source and sink are equal and a representative to check against can arbitrarily be chosen.
-It lead to an $O(|V|^2(|E|+|V|)$ algorithm to verify a diagram remains commutative over the course of online addition, assuming an oracle to check the equality of functions.
+It lead to an $O(|V|^2(|E|+|V|))$ algorithm to verify a diagram remains commutative over the course of online addition, assuming an oracle to check the equality of functions.
 The algorithm makes an asymptotically optimal number of calls to the oracle.
@@
 
-Previous work ~\cite{commutative} has identified an $O(|E|^2|V|^4)$ algorithm to verify that a complete acyclic diagram commutes; however, it addresses neither online addition nor cyclic diagrams.
-
 To evaluate our results we use two case studies.
-First, our algorithm is used for the domain specific geometry type language \textit{Gator} to ensure that user defined transformations between spaces stay consistent.
+First, our algorithm is used for the domain specific geometry type language \textit{Gator} ~\cite{gator} to ensure that user defined transformations between spaces stay consistent.
 Second, our algorithm is used to identify inefficiencies in a currency conversion graph.
@@
 We compare our solution to three baseline implementations: a naive checking of all path pairs with only special handling for cycles, a check for all path pairs that involve the new edge, and an algorithm suggested by previous work to solve the batch version of the problem for acyclic diagrams.
+Our proposed algorithms run orders of magnitude faster than the baselines.
 
@@
 
-We start by formalizing the notion of a diagram, drawing inspiration from the previous acyclic work by Murota ~\cite{commutative}.
+We start by formalizing the notion of a diagram, drawing terminology from the previous acyclic work by Murota ~\cite{commutative}.
 
-\xxx[AK]{A lot of this exposition is from the cited paper, I feel like I need to make this clearer to give proper credit, but not sure how}
-
-We start with a directed graph G=(V,E), where V consists of sets of elements and edges (u, v) in E correspond to functions that maps elements of u to elements in v.
-All these functions form a semigroup F, where multiplication is function composition.
+We start with a directed graph $G=(V,E)$, where $V$ consists of sets of elements and edges $(u, v)$ in $E$ correspond to functions that maps elements of $u$ to elements in $v$.
+All these functions form a semigroup $F$, where multiplication is function composition.
 A semigroup consists of a set and an associative binary operation, which we use to capture function composition.
 
-The correspondence between edges and functions is stored as a mapping f:$E\rightarrow F$, where f maps each edge to the function it represents.
+The correspondence between edges and functions is stored as a mapping $f:E\rightarrow F$, where $f$ maps each edge to the function it represents.
 
-A path is a sequence of edges. The edge-to-function mapping f can be naturally extended to paths: if path p=$e_1\circ e_2 ... e_n$ then $f(p)=f(e_1) \circ f(e_2) ... f(e_n)$.
+A path is a sequence of edges. The edge-to-function mapping $f$ can be naturally extended to paths: if path $p=e_1\circ e_2 ... e_n$ then $f(p)=f(e_1) \circ f(e_2) ... f(e_n)$.
 
-The notation $\partial(p)$ maps to the tuple (start node of p, end node of p).
+The notation $\partial(p)$ maps to the tuple (start node of $p$, end node of $p$).
 
-A pair of paths p and q is called \textit{parallel} iff their terminal nodes are the same, i.e., $\partial(p)=\partial(q)$.
+A pair of paths $p$ and $q$ is called \textit{parallel} iff their terminal nodes are the same, i.e., $\partial(p)=\partial(q)$.
 Let $R_{all}$ be the set of all parallel pairs of paths in the diagram.
 
-The diagram commutes iff f(p)=f(q) $\forall (p,q)\in R_{all}$; that is, the composition of maps along any path connecting u to v is independent of path choice.
+The diagram commutes iff $f(p)=f(q)$ $\forall (p,q)\in R_{all}$; that is, the composition of maps along any path connecting $u$ to $v$ is independent of path choice.
 
 The \textsc{Online addition problem}, given a commuting diagram and a new edge, returns whether the diagram commutes.
+Checking function equality is a domain specific, potentially hard problem, dependent on the nature of the graph.
+For graphics programs in Gator, for example, where edges are matrices, function equality checking would be comparing matrix elements.
 We make an abstraction - equality of functions is defined by some oracle whose specifics will vary by target functions and domain - and reduce the \textsc{Online addition problem} to the \textsc{Verification set problem}, which given a diagram and a new edge, returns the set of parallel pairs of paths, such that if and only if the members in each pair are equal, then the new graph must commute.
@@
 
-\subsection{Very Na\"{i}ve baseline}
+\subsection{Very Na\"{i}ve Baseline}
 
@@
 \KwRet{parallelPairs}\;
-\caption{Online polynomial time algorithm to find parallel pair set}\label{algo_online_polynomial}
+\caption{Online polynomial time algorithm to find parallel pair set}
+\label{algo_online_polynomial}
 \end{algorithm}
@@
 \KwRet{verificationSet}\;
-\caption{Minimal set finding algorithm}\label{algo_online_minimal}
+\caption{Minimal set finding algorithm}
+\label{algo_online_minimal}
 \end{algorithm}

### 2020-08-05 acmPaper/paper.tex
Remove amssymb import
Was causing this error:

    ! LaTeX Error: Command `\Bbbk' already defined.

And it's not necessary AFAICT.
Abstract pass

@@
 \usepackage{tikz}
-\usepackage{amssymb}
 \usepackage[norelsize]{algorithm2e}
@@
 \begin{abstract}
-Commutative diagrams arise in many systems, such as an implicit type conversion graph in some programming languages.
-It can be useful in program analysis to verify that a diagram representing conversions remains commutative over the course of its construction.
-A naive approach to solving the problem would take factorial time, and the presence of cycles could potentially lead to an infinite number of paths to check.
-Previous work provides an algorithm for the special case of acyclic diagrams that verifies that a given diagram commutes in $O(|V|^4|E|^2)$ time, where V is the set of vertices in the diagram, and E, the set of edges.
-We present an algorithm that given a diagram and an edge to add to it, and an oracle to check the equality of a given pair of paths, verifies that the resultant diagram commutes in $O(|V|^2(|E| + |V|))$ time.
-For the case when checking the equality of paths is expensive, we present an optimization that runs in $O(|V|^4)$ time but reduces to the minimum possible number of equality checks.
-We evaluate the algorithms against baselines, and see its practical application in the compiler of a domain specific language for geometry types.
-As a demonstration of the more general applicability of the algorithm, we use it to identify discrepancies in a  currency conversion graph.
+Systems of transformations arise in many programming systems, such as in type graphs of implicit type conversion functions.
+It is important to ensure that these diagrams commute: that any composing path of transformations from the same source to the same destination yields the same result.
+However, a straightforward approach to verifying commutativity must contend with cycles, and even so it runs in factorial time.
+While previous work has shown how to verify commutativity in the special case of acyclic diagrams in $O(|V|^4|E|^2)$ time,
+% I think this is clear enough (at least for the abstract). --AS
+% where V is the set of vertices in the diagram, and E, the set of edges.
+this is a \emph{batch} algorithm: the entire system must be known ahead of time.
+We present an \emph{online} algorithm that efficiently verifies that a commutative diagram remains commutative when adding a new edge.
+The new incremental algorithm runs in $O(|V|^2(|E| + |V|))$ time.
+For the case when checking the equality of paths is expensive, we also present an optimization that runs in $O(|V|^4)$ time but reduces to the minimum possible number of equality checks.
+We implement the algorithms and compare them to batch baselines, and we demonstrate their practical application in the compiler of a domain-specific language for geometry types.
+To study the algorithms' scalability to large diagrams, we apply them to discover discrepancies in currency conversion graphs.
 \end{abstract}

### 2020-08-08 transpilingSource/checker/__target__/hello.js
javascript for gator

@@ -1,0 +1,26 @@
+'use strict';var time={};import{AssertionError,AttributeError,BaseException,DeprecationWarning,Exception,IndexError,IterableError,KeyError,NotImplementedError,RuntimeWarning,StopIteration,UserWarning,ValueError,Warning,__JsIterator__,__PyIterator__,__Terminal__,__add__,__and__,__call__,__class__,__envir__,__eq__,__floordiv__,__ge__,__get__,__getcm__,__getitem__,__getslice__,__getsm__,__gt__,__i__,__iadd__,__iand__,__idiv__,__ijsmod__,__ilshift__,__imatmul__,__imod__,__imul__,__in__,__init__,__ior__,
+__ipow__,__irshift__,__isub__,__ixor__,__jsUsePyNext__,__jsmod__,__k__,__kwargtrans__,__le__,__lshift__,__lt__,__matmul__,__mergefields__,__mergekwargtrans__,__mod__,__mul__,__ne__,__neg__,__nest__,__or__,__pow__,__pragma__,__proxy__,__pyUseJsNext__,__rshift__,__setitem__,__setproperty__,__setslice__,__sort__,__specialattrib__,__sub__,__super__,__t__,__terminal__,__truediv__,__withblock__,__xor__,abs,all,any,assert,bool,bytearray,bytes,callable,chr,copy,deepcopy,delattr,dict,dir,divmod,enumerate,
+filter,float,getattr,hasattr,input,int,isinstance,issubclass,len,list,map,max,min,object,ord,pow,print,property,py_TypeError,py_iter,py_metatype,py_next,py_reversed,py_typeof,range,repr,round,set,setattr,sorted,str,sum,tuple,zip}from"./org.transcrypt.__runtime__.js";import*as __module_time__ from"./time.js";__nest__(time,"",__module_time__);var __name__="__main__";export var _default_no_edge=-1;export var BaseOnlineChecker=__class__("BaseOnlineChecker",[object],{__module__:__name__,_invalid_node:-1,
[…]

### 2020-08-08 transpilingSource/checker/checker.py

@@ -1,0 +1,315 @@
+# from copy import copy
+import time
+
+_default_no_edge = -1
+
+class OptimalSetPathChecker(BaseOnlineChecker):
+    """Path checker that returns minimal set for verification 
+    for online problem."""
+
+    def __init__(self):
+        super().__init__()
+
+    def getSuccessors(self, source: int, sink: int) -> Set[int]:
+        """Returns a set of pairs such that verifying a path between given
+        source and sink implies also that all the pairs in the returned set are
+        equal."""
+        predecessors = self.getAllPredecessors(source)
+        successors = self.getAllSuccessors(sink)
+        if (self.noIdentity and source == sink):
+            return {(src, snk) for src in predecessors for snk in successors 
+            if src == snk}
+        return {(src, snk) for src in predecessors for snk in successors}
+
+    def getRootPairs(self) -> List[List[int]]:
+        """Returns a graph where each node represents a (source, sink) pair in
+        self.graph. 
+        Map from (source, sink) to node number in self.dependencyNodeMap.
+        Node A succeeds node B if checking a pair of B implies that 
+        the paths pairs of A must also be equal."""
+        acceptedPairs = set()
+        # Get set of all predecessors, successors.
+        predecessors = self.getAllPredecessors(self.newEdgeSource)
+        successors = self.getAllSuccessors(self.newEdgeSink)
+        # Go through each element in predecessors X successors.
+        # For efficiency, closest pairs to new edge should appear first.
+        potentialPairs = {(source, sink) for source in predecessors \
+            for sink in successors}
+        orderedPotentialPairs = [(source, sink) for source in predecessors \
+            for sink in successors]
+        while (len(potentialPairs) != 0):
+            # using ordering is a heuristic to (hopefully) improve performance,
+            # so parent is usually checked before child.
+            (source, sink) = orderedPotentialPairs.pop(0)
+            if (source, sink) in potentialPairs:
+                potentialPairs.remove((source, sink))
+                (pathExists, _) = self.findPath(source, sink)
+                if pathExists:
+                    currentPairSuccessors = self.getSuccessors(source, sink)
+                    for redundantPair in acceptedPairs.intersection(
+                        currentPairSuccessors):
+                        acceptedPairs.remove(redundantPair)
+                    for redundantPair in potentialPairs.intersection(
+                        currentPairSuccessors):
[…]

### 2020-08-08 transpilingSource/checker/hello.py

@@ -1,0 +1,394 @@
+import time
+
+_default_no_edge = -1
+
+
+class BaseOnlineChecker():
+    """Implements base functionality common to all PathCheckers."""
+
+    _invalid_node = -1
+    _debug = False
+
+    #### Utility functions ####
+
+    def getForwardEdges(self, source: int) -> List[int]:
+        return [i for i in range(len(self.graph)) \
+            if self.graph[source][i] != self._no_edge]
+
+    def getBackwardEdges(self, sink: int) -> List[int]:
+        return [i for i in range(len(self.graph)) \
+            if self.graph[i][sink] != self._no_edge and i != sink]
+
+    def getEmptyPathList(self):
+        paths = []
+        for _ in range(len(self.graph)):
+            paths += [[]]
+        return paths
+
+    def buildCompactFwdGraph(self):
+        self.compactFwdGraph = self.getEmptyPathList()
+        for node in range(len(self.graph)):
+            self.compactFwdGraph[node] = self.getForwardEdges(node)
+    
+    def buildCompactBkdGraph(self):
+        self.compactBkdGraph = self.getEmptyPathList()
+        for node in range(len(self.graph)):
+            self.compactBkdGraph[node] = self.getBackwardEdges(node)
+    
+    def getAllPredecessors(self, node, memoize = False) -> List[int]:
+        """Returns a list of nodes that are predecessors of this one.
+        Memoize if source of new edge."""
+        if not self.compactBkdGraph:
+            self.buildCompactBkdGraph()
+        memoize = memoize or node == self.newEdgeSource
+        if memoize:
+            pathDict = dict()
+        visited = set()
+        # To store visitedNodes ordered closest to new edge first.
+        visitedList = []
+        currentNode = node
+        stack = [currentNode]
+        path = []
+        while (len(stack) > 0):
+            currentNode = stack.pop()
+            if currentNode == self._invalid_node:
+                path = path[1:]
+            elif currentNode not in visited:
+                visited.add(currentNode)
+                visitedList += [currentNode]
+                path = [currentNode] + path
+                if (memoize):
+                    pathDict[currentNode] = self.deepcopyList(path)
+                stack += [self._invalid_node] + self.compactBkdGraph[currentNode]
+        if self._debug:
+            print("getAllPredecessors: node: {} and visitedList: {}".format(
+                node, visitedList))
+        if node == self.newEdgeSource:
+            self.pathsToNewEdgeSource = pathDict
[…]

### 2020-08-08 transpilingSource/evaluator.py

@@ -1,0 +1,247 @@
+from datetime import datetime
+import functools
+import json
+import matplotlib.pyplot as plt
+import numpy as np
+from pathCheckers.batchChecker import BatchChecker
+from pathCheckers.iPathChecker import IPathChecker
+from pathCheckers.optimalSetPathChecker import OptimalSetPathChecker
+from pathCheckers.polynomialPathChecker import PolynomialPathChecker
+from pathCheckers.naiveChecker import NaiveChecker
+from pathCheckers.twoFlipChecker import TwoFlipPathChecker
+from randomGraphGenerator import generateGraph, generateAcyclicGraph
+import traceback
+from typing import List
+
+NUM_TRIES = 10
+DO_ACYCLIC = True
+OUTFILE = ''
+
+densityStep = 0.5
+sizeStep = 1
+maxSize = 12
+
+# densities = [i*densityStep for i in range(1, int(1.0/densityStep))] 
+densities = [0.4]
+# sizes = [i*sizeStep for i in range(1, int(maxSize/sizeStep))]
+sizes = [9]
+
+evaluationList = [(density, size) for size in sizes for density in densities]
+
+# First three colors look good in black and white.
+colors = ["black", "purple", "orange", "blue", "green", "red"]
+
+checkers : List[IPathChecker] = [
+    # PolynomialPathChecker(),
+    # OptimalSetPathChecker(),
+    # NaiveChecker(),
+    # TwoFlipPathChecker(),
+    BatchChecker()
+]
+
+class attemptInfo:
+    """Stores information for a given density, size"""
+
+    def __init__(self):
+        self.times = []
+        self.sizes = []
+    
+    def addTime(self,time):
+        self.times += [time]
+    
+    def addSize(self, sze):
+        self.sizes += [sze]
+    
+    def avgList(self, lst):
+        self.avg = functools.reduce(lambda  acc, n: acc + n, lst, 0)/len(lst)
+        return self.avg
+    
+    def getAvgTime(self, median=False): 
+        if median:
+            return self.getMedianTime()
+        return self.avgList(self.times)
+
+    def getAvgSizes(self):
+        return self.avgList(self.sizes)
+    
+    def getTimes(self):
+        return self.times
+
+    def getMedianTime(self):
+        copy = self.times
+        copy.sort()
+        if len(copy)%2 == 1:
+            self.median = copy[int(len(copy)/2)+1]
+        else:
+            floor = int(len(copy)/2)
[…]

### 2020-08-08 transpilingSource/example/__target__/hello.js

@@ -1,0 +1,2 @@
+
+//# sourceMappingURL=hello.map

### 2020-08-08 transpilingSource/example/__target__/itertools.js

@@ -1,0 +1,2 @@
+
+//# sourceMappingURL=itertools.map

### 2020-08-08 transpilingSource/main.py

@@ -1,0 +1,137 @@
+"""Currency graph evaluation."""
+
+import copy
+import time
+import traceback
+from typing import List
+
+import numpy as np
+import requests
+from colorama import Fore
+
+from currencyGraph.currencyGraph import CurrencyGraph
+from pathCheckers.iPathChecker import IPathChecker
+from pathCheckers.optimalSetPathChecker import OptimalSetPathChecker
+from pathCheckers.polynomialPathChecker import PolynomialPathChecker
+
+# Assume that a graph doesn't have more edges nodes than this
+INFTY = 1000000
+NO_EDGE = -1
+DEBUG = False
+
+ID = [-1]
+ID_PATH = [ID]
+ID_FUNCTION = lambda source : ID_PATH
+
+# TODO add:
+# 1. Commandline arguments.
+# 2. Historic rate.
+
+###############################################
+
+epsilon = 1E-4 # allowed error multiplier
+
+checkers = [
+    OptimalSetPathChecker(),
+    PolynomialPathChecker()
+]
+
+def getPathValue(path: List[int], graph: List[List[int]]):
+    """Composition oracle defintion."""
+    result = 1
+    # Special case for identity function.
+    if path == ID_PATH:
+        return result
+    # Special case for just new edge.
+    if len(path) == 1:
+        return graph[path[0]][path[0]]
+    # Regular case.
+    for i in range(len(path)-1):
+        result *= graph[path[i]][path[i+1]]
+    return result
+
+def independenceFuncFactory(checker: IPathChecker, graph: CurrencyGraph):
+    """Equality check oracle definition."""
+    checker.setGraph(graph.graph)
+    checker.setIdFunction(ID_FUNCTION)
+    def IndependenceFunction(base, target, rate):
+        checker.setEdge(graph.indexOf(base), graph.indexOf(target))
+        pairsToCheck = checker.getPathsToCheck()
+        newGraph = graph.getCopyWithNewEdge(base, target, rate)
+        for (path1, path2) in pairsToCheck:
+            path1Value = getPathValue(path1, newGraph)
+            path2Value = getPathValue(path2, newGraph)
+            if abs(path1Value - path2Value) > epsilon:
+                return (False, "path 1: {}: {} \npath 2: {}: {} \n difference: {}, new edge: {} -> {}: rate: {}"
+                .format(
+                path1Value,
+                path1,
+                path2Value,
+                path2,
+                path1Value - path2Value,
+                base,
+                target,
+                rate))
[…]

### 2020-08-08 transpilingSource/checker/__target__/hello.js
some fixes

@@ -1,26 +1,27 @@
 'use strict';var time={};import{AssertionError,AttributeError,BaseException,DeprecationWarning,Exception,IndexError,IterableError,KeyError,NotImplementedError,RuntimeWarning,StopIteration,UserWarning,ValueError,Warning,__JsIterator__,__PyIterator__,__Terminal__,__add__,__and__,__call__,__class__,__envir__,__eq__,__floordiv__,__ge__,__get__,__getcm__,__getitem__,__getslice__,__getsm__,__gt__,__i__,__iadd__,__iand__,__idiv__,__ijsmod__,__ilshift__,__imatmul__,__imod__,__imul__,__in__,__init__,__ior__,
 __ipow__,__irshift__,__isub__,__ixor__,__jsUsePyNext__,__jsmod__,__k__,__kwargtrans__,__le__,__lshift__,__lt__,__matmul__,__mergefields__,__mergekwargtrans__,__mod__,__mul__,__ne__,__neg__,__nest__,__or__,__pow__,__pragma__,__proxy__,__pyUseJsNext__,__rshift__,__setitem__,__setproperty__,__setslice__,__sort__,__specialattrib__,__sub__,__super__,__t__,__terminal__,__truediv__,__withblock__,__xor__,abs,all,any,assert,bool,bytearray,bytes,callable,chr,copy,deepcopy,delattr,dict,dir,divmod,enumerate,
 filter,float,getattr,hasattr,input,int,isinstance,issubclass,len,list,map,max,min,object,ord,pow,print,property,py_TypeError,py_iter,py_metatype,py_next,py_reversed,py_typeof,range,repr,round,set,setattr,sorted,str,sum,tuple,zip}from"./org.transcrypt.__runtime__.js";import*as __module_time__ from"./time.js";__nest__(time,"",__module_time__);var __name__="__main__";export var _default_no_edge=-1;export var BaseOnlineChecker=__class__("BaseOnlineChecker",[object],{__module__:__name__,_invalid_node:-1,
[…]

### 2020-08-09 transpilingSource/checker/__target__/hello.js
path checker and frontend working

@@ -4,24 +4,24 @@ filter,float,getattr,hasattr,input,int,isinstance,issubclass,len,list,map,max,mi
 _debug:false,get extend(){return __get__(this,function(self,list1,list2){for(var elem of list2)list1.append(elem);return list1})},get getForwardEdges(){return __get__(this,function(self,source){return function(){var __accu0__=[];for(var i=0;i<len(self.graph);i++)if(self.graph[source][i]!=self._no_edge)__accu0__.append(i);return __accu0__}()})},get getBackwardEdges(){return __get__(this,function(self,sink){return function(){var __accu0__=[];for(var i=0;i<len(self.graph);i++)if(self.graph[i][sink]!=
 self._no_edge&&i!=sink)__accu0__.append(i);return __accu0__}()})},get getEmptyPathList(){return __get__(this,function(self){var paths=[];for(var _=0;_<len(self.graph);_++)paths.append([]);return paths})},get buildCompactFwdGraph(){return __get__(this,function(self){self.compactFwdGraph=self.getEmptyPathList();for(var node=0;node<len(self.graph);node++)self.compactFwdGraph[node]=self.getForwardEdges(node)})},get buildCompactBkdGraph(){return __get__(this,function(self){self.compactBkdGraph=self.getEmptyPathList();
 for(var node=0;node<len(self.graph);node++)self.compactBkdGraph[node]=self.getBackwardEdges(node)})},get getAllPredecessors(){return __get__(this,function(self,node,memoize){if(typeof memoize=="undefined"||memoize!=null&&memoize.hasOwnProperty("__kwargtrans__"))var memoize=false;if(!self.compactBkdGraph)self.buildCompactBkdGraph();var memoize=memoize||node==self.newEdgeSource;if(memoize)var pathDict=dict();var visited=set();var visitedList=[];var currentNode=node;var stack=[currentNode];var path=
-[];while(len(stack)>0){var currentNode=stack.py_pop();if(currentNode==self._invalid_node)var path=path.__getslice__(1,null,1);else if(!__in__(currentNode,visited)){visited.add(currentNode);visitedList.append(currentNode);var path=[currentNode]+path;if(memoize)pathDict[currentNode]=self.deepcopyList(path);stack.append(self._invalid_node);var stack=self.extend(stack,self.compactBkdGraph[currentNode])}}if(self._debug)print("getAllPredecessors: node: {} and visitedList: {}".format(node,visitedList));
[…]

### 2020-08-09 transpilingSource/checker/hello.py

@@ -11,6 +11,11 @@ class BaseOnlineChecker():
 
     #### Utility functions ####
 
+    def extend(self, list1, list2):
+        for elem in list2:
+            list1.append(elem)
+        return list1
+
     def getForwardEdges(self, source: int) -> List[int]:
         return [i for i in range(len(self.graph)) \
             if self.graph[source][i] != self._no_edge]
@@ -22,7 +27,7 @@ class BaseOnlineChecker():
     def getEmptyPathList(self):
         paths = []
         for _ in range(len(self.graph)):
-            paths += [[]]
+            paths.append([])
         return paths
 
     def buildCompactFwdGraph(self):
@@ -55,11 +60,12 @@ class BaseOnlineChecker():
                 path = path[1:]
             elif currentNode not in visited:
                 visited.add(currentNode)
-                visitedList += [currentNode]
-                path = [currentNode] + path
+                visitedList.append(currentNode)
+                path = self.extend([currentNode], path)
                 if (memoize):
                     pathDict[currentNode] = self.deepcopyList(path)
-                stack += [self._invalid_node] + self.compactBkdGraph[currentNode]
+                stack.append(self._invalid_node)
+                stack = self.extend(stack, self.compactBkdGraph[currentNode])
         if self._debug:
             print("getAllPredecessors: node: {} and visitedList: {}".format(
                 node, visitedList))
@@ -88,7 +94,8 @@ class BaseOnlineChecker():
                 path = [currentNode] + path
                 if (memoize):
                     pathDict[currentNode] = self.deepcopyList(path)
-                stack += [self._invalid_node] + self.compactBkdGraph[currentNode]
+                stack.append(self._invalid_node)
+                stack = self.extend(stack, self.compactBkdGraph[currentNode])
         if self._debug:
             print("getAllPredecessors: node: {} and visitedList: {}".format(
                 node, visited))
@@ -116,11 +123,12 @@ class BaseOnlineChecker():
                 path.pop()
             elif currentNode not in visited:
                 visited.add(currentNode)
-                visitedList += [currentNode]
-                path += [currentNode]
+                visitedList.append(currentNode)
+                path.append(currentNode)
                 if (node == self.newEdgeSink):
                     self.pathsFromNewEdgeSink[currentNode] = self.deepcopyList(path)
[…]

### 2020-08-09 transpilingSource/checker/__target__/hello.js
sanity check for gator hook

@@ -22,6 +22,7 @@ getPathListString(paths)}else document.getElementById("paths").innerHTML="Graph 
 edgeString.py_split();var source=__left0__[0];var sink=__left0__[1];checker.setEdge(int(source),int(sink))};export var scalarMultiply=function(scalar,matrix){var ans=[];for(var row of matrix){var ansRow=[];for(var elem of row)ansRow.append(scalar*elem);ans.append(ansRow)}return ans};export var metrixMultiple=function(mat1,mat2){if(isinstance(mat1,list))if(isinstance(mat2,list)){var ans=[];for(var row of len(mat1)){var ansRow=[];var vector1=mat1[row];for(var col of len(mat2[0])){var vector2=function(){var __accu0__=
 [];for(var i=0;i<len(mat2);i++)__accu0__.append(mat2[i][row]);return __accu0__}();var sumOfEntries=0;for(var i=0;i<len(vector2);i++)sumOfEntries+=vector1[i]*vector2[i];ansRow.append(sumOfEntries)}ans.append(ansRow)}return ans}else return scalarMultiply(__kwargtrans__({scalar:mat2,matrix:mat1}));else if(isinstance(mat2,list))return scalarMultiply(__kwargtrans__({scalar:mat1,matrix:mat2}));else return mat1*mat2};export var getMatrixOfPaths=function(path,graphValues){var productMatrix=1;if(len(path)<
 2)return productMatrix;for(var i=1;i<len(path);i++){var edgeMatrix=graphValues();matrixMultiply(productMatrix,edgeMatrix)}};export var checkPaths=function(pathsToCheck,graphValues,oracle){for(var pathPair of pathsToCheck){var __left0__=pathPair;var path1=__left0__[0];var path2=__left0__[1];var matrix1=getMatrixOfPaths(path1,graphValues);var matrix2=getMatrixOfPaths(path2,graphValues);if(!oracle(matrix1,matrix2))return false}return true};export var verify=function(graph,graphValues,newEdge,oracle){checker.setGraph(graph);
-checker.setEdge(newEdge[0],newEdge[1]);var pathsToCheck=checker.getPathsToCheck();return checkPaths(pathsToCheck,graphValues,newEdge,oracle)};
[…]

### 2020-08-09 transpilingSource/checker/hello.py

@@ -421,3 +421,40 @@ def verify(graph, graphValues, newEdge, oracle):
     checker.setEdge(newEdge[0], newEdge[1])
     pathsToCheck = checker.getPathsToCheck()
     return checkPaths(pathsToCheck, graphValues, newEdge, oracle)
+
+def gatorTest():
+    """Check for functionality of gator hook code.
+    Please add to test case list instead of replacing."""
+    NO_EDGE = _default_no_edge
+    graph1 = [
+        [-1,  1, -1],
+        [-1, -1,  1],
+        [-1, -1, -1]]
+    newEdge1 = (0, 2)
+    graphValue1 = [
+        [1, 1, 1],
+        [1, 1, 1],
+        [1, 1, 1]]
+    oracle1 = lambda a, b: a==b
+    expectedOutcome1 = True
+
+    # If expanding on this later, put in "testCase" class
+    testGraphs = [graph1]
+    newEdge = [newEdge1]
+    oracles = [oracle1]
+    graphValues = [graphValue1]
+    expectedOutcomes = [expectedOutcome1]
+    
+    testsSucceeded  = True
+    
+    for i in range(testGraphs):
+        outcome = verify(
+            testGraphs[i],
+            graphValues[i],
+            newEdge[i],
+            oracles[i])
+        testsSucceeded = testsSucceeded and (outcome == expectedOutcomes[i])
+    
+    document.getElementById('gator').innerHTML = (
+            'succeeded: '+testsSucceeded
+        )

### 2020-08-09 transpilingSource/checker/checker.py
remove unwanted file

@@ -1,315 +1,0 @@
-# from copy import copy
-import time
-
-_default_no_edge = -1
-
-class OptimalSetPathChecker(BaseOnlineChecker):
-    """Path checker that returns minimal set for verification 
-    for online problem."""
-
-    def __init__(self):
-        super().__init__()
-
-    def getSuccessors(self, source: int, sink: int) -> Set[int]:
-        """Returns a set of pairs such that verifying a path between given
-        source and sink implies also that all the pairs in the returned set are
-        equal."""
-        predecessors = self.getAllPredecessors(source)
-        successors = self.getAllSuccessors(sink)
-        if (self.noIdentity and source == sink):
-            return {(src, snk) for src in predecessors for snk in successors 
-            if src == snk}
-        return {(src, snk) for src in predecessors for snk in successors}
-
-    def getRootPairs(self) -> List[List[int]]:
-        """Returns a graph where each node represents a (source, sink) pair in
-        self.graph. 
-        Map from (source, sink) to node number in self.dependencyNodeMap.
-        Node A succeeds node B if checking a pair of B implies that 
-        the paths pairs of A must also be equal."""
-        acceptedPairs = set()
-        # Get set of all predecessors, successors.
-        predecessors = self.getAllPredecessors(self.newEdgeSource)
-        successors = self.getAllSuccessors(self.newEdgeSink)
-        # Go through each element in predecessors X successors.
-        # For efficiency, closest pairs to new edge should appear first.
-        potentialPairs = {(source, sink) for source in predecessors \
-            for sink in successors}
-        orderedPotentialPairs = [(source, sink) for source in predecessors \
-            for sink in successors]
-        while (len(potentialPairs) != 0):
-            # using ordering is a heuristic to (hopefully) improve performance,
-            # so parent is usually checked before child.
-            (source, sink) = orderedPotentialPairs.pop(0)
-            if (source, sink) in potentialPairs:
-                potentialPairs.remove((source, sink))
-                (pathExists, _) = self.findPath(source, sink)
-                if pathExists:
-                    currentPairSuccessors = self.getSuccessors(source, sink)
-                    for redundantPair in acceptedPairs.intersection(
-                        currentPairSuccessors):
-                        acceptedPairs.remove(redundantPair)
-                    for redundantPair in potentialPairs.intersection(
-                        currentPairSuccessors):
[…]

### 2020-08-09 transpilingSource/evaluator.py
remove two more unwanted files

@@ -1,247 +1,0 @@
-from datetime import datetime
-import functools
-import json
-import matplotlib.pyplot as plt
-import numpy as np
-from pathCheckers.batchChecker import BatchChecker
-from pathCheckers.iPathChecker import IPathChecker
-from pathCheckers.optimalSetPathChecker import OptimalSetPathChecker
-from pathCheckers.polynomialPathChecker import PolynomialPathChecker
-from pathCheckers.naiveChecker import NaiveChecker
-from pathCheckers.twoFlipChecker import TwoFlipPathChecker
-from randomGraphGenerator import generateGraph, generateAcyclicGraph
-import traceback
-from typing import List
-
-NUM_TRIES = 10
-DO_ACYCLIC = True
-OUTFILE = ''
-
-densityStep = 0.5
-sizeStep = 1
-maxSize = 12
-
-# densities = [i*densityStep for i in range(1, int(1.0/densityStep))] 
-densities = [0.4]
-# sizes = [i*sizeStep for i in range(1, int(maxSize/sizeStep))]
-sizes = [9]
-
-evaluationList = [(density, size) for size in sizes for density in densities]
-
-# First three colors look good in black and white.
-colors = ["black", "purple", "orange", "blue", "green", "red"]
-
-checkers : List[IPathChecker] = [
-    # PolynomialPathChecker(),
-    # OptimalSetPathChecker(),
-    # NaiveChecker(),
-    # TwoFlipPathChecker(),
-    BatchChecker()
-]
-
-class attemptInfo:
-    """Stores information for a given density, size"""
-
-    def __init__(self):
-        self.times = []
-        self.sizes = []
-    
-    def addTime(self,time):
-        self.times += [time]
-    
-    def addSize(self, sze):
-        self.sizes += [sze]
-    
-    def avgList(self, lst):
-        self.avg = functools.reduce(lambda  acc, n: acc + n, lst, 0)/len(lst)
-        return self.avg
-    
-    def getAvgTime(self, median=False): 
-        if median:
-            return self.getMedianTime()
-        return self.avgList(self.times)
-
-    def getAvgSizes(self):
-        return self.avgList(self.sizes)
-    
-    def getTimes(self):
-        return self.times
-
-    def getMedianTime(self):
-        copy = self.times
-        copy.sort()
-        if len(copy)%2 == 1:
-            self.median = copy[int(len(copy)/2)+1]
-        else:
-            floor = int(len(copy)/2)
[…]

### 2020-08-09 transpilingSource/main.py

@@ -1,137 +1,0 @@
-"""Currency graph evaluation."""
-
-import copy
-import time
-import traceback
-from typing import List
-
-import numpy as np
-import requests
-from colorama import Fore
-
-from currencyGraph.currencyGraph import CurrencyGraph
-from pathCheckers.iPathChecker import IPathChecker
-from pathCheckers.optimalSetPathChecker import OptimalSetPathChecker
-from pathCheckers.polynomialPathChecker import PolynomialPathChecker
-
-# Assume that a graph doesn't have more edges nodes than this
-INFTY = 1000000
-NO_EDGE = -1
-DEBUG = False
-
-ID = [-1]
-ID_PATH = [ID]
-ID_FUNCTION = lambda source : ID_PATH
-
-# TODO add:
-# 1. Commandline arguments.
-# 2. Historic rate.
-
-###############################################
-
-epsilon = 1E-4 # allowed error multiplier
-
-checkers = [
-    OptimalSetPathChecker(),
-    PolynomialPathChecker()
-]
-
-def getPathValue(path: List[int], graph: List[List[int]]):
-    """Composition oracle defintion."""
-    result = 1
-    # Special case for identity function.
-    if path == ID_PATH:
-        return result
-    # Special case for just new edge.
-    if len(path) == 1:
-        return graph[path[0]][path[0]]
-    # Regular case.
-    for i in range(len(path)-1):
-        result *= graph[path[i]][path[i+1]]
-    return result
-
-def independenceFuncFactory(checker: IPathChecker, graph: CurrencyGraph):
-    """Equality check oracle definition."""
-    checker.setGraph(graph.graph)
-    checker.setIdFunction(ID_FUNCTION)
-    def IndependenceFunction(base, target, rate):
-        checker.setEdge(graph.indexOf(base), graph.indexOf(target))
-        pairsToCheck = checker.getPathsToCheck()
-        newGraph = graph.getCopyWithNewEdge(base, target, rate)
-        for (path1, path2) in pairsToCheck:
-            path1Value = getPathValue(path1, newGraph)
-            path2Value = getPathValue(path2, newGraph)
-            if abs(path1Value - path2Value) > epsilon:
-                return (False, "path 1: {}: {} \npath 2: {}: {} \n difference: {}, new edge: {} -> {}: rate: {}"
-                .format(
-                path1Value,
-                path1,
-                path2Value,
-                path2,
-                path1Value - path2Value,
-                base,
-                target,
-                rate))
[…]

### 2020-08-09 transpilingSource/checker/__target__/hello.js
more bugfixes

@@ -5,24 +5,23 @@ _debug:false,get extend(){return __get__(this,function(self,list1,list2){for(var
 self._no_edge&&i!=sink)__accu0__.append(i);return __accu0__}()})},get getEmptyPathList(){return __get__(this,function(self){var paths=[];for(var _=0;_<len(self.graph);_++)paths.append([]);return paths})},get buildCompactFwdGraph(){return __get__(this,function(self){self.compactFwdGraph=self.getEmptyPathList();for(var node=0;node<len(self.graph);node++)self.compactFwdGraph[node]=self.getForwardEdges(node)})},get buildCompactBkdGraph(){return __get__(this,function(self){self.compactBkdGraph=self.getEmptyPathList();
 for(var node=0;node<len(self.graph);node++)self.compactBkdGraph[node]=self.getBackwardEdges(node)})},get getAllPredecessors(){return __get__(this,function(self,node,memoize){if(typeof memoize=="undefined"||memoize!=null&&memoize.hasOwnProperty("__kwargtrans__"))var memoize=false;if(!self.compactBkdGraph)self.buildCompactBkdGraph();var memoize=memoize||node==self.newEdgeSource;if(memoize)var pathDict=dict();var visited=set();var visitedList=[];var currentNode=node;var stack=[currentNode];var path=
 [];while(len(stack)>0){var currentNode=stack.py_pop();if(currentNode==self._invalid_node)var path=path.__getslice__(1,null,1);else if(!__in__(currentNode,visited)){visited.add(currentNode);visitedList.append(currentNode);var path=self.extend([currentNode],path);if(memoize)pathDict[currentNode]=self.deepcopyList(path);stack.append(self._invalid_node);var stack=self.extend(stack,self.compactBkdGraph[currentNode])}}if(self._debug)print("getAllPredecessors: node: {} and visitedList: {}".format(node,visitedList));
-if(node==self.newEdgeSource)self.pathsToNewEdgeSource=pathDict;if(memoize)self.pathsToNode[node]=pathDict;return visitedList})},get getAllPredecessorsUnordered(){return __get__(this,function(self,node,memoize){if(typeof memoize=="undefined"||memoize!=null&&memoize.hasOwnProperty("__kwargtrans__"))var memoize=true;var memoize=memoize||node==self.newEdgeSource;if(memoize)var pathDict=dict();var visited=set();var currentNode=node;var stack=[currentNode];var path=[];while(len(stack)>0){var currentNode=
[…]

### 2020-08-09 transpilingSource/checker/hello.py

@@ -71,40 +71,8 @@ class BaseOnlineChecker():
                 node, visitedList))
         if node == self.newEdgeSource:
             self.pathsToNewEdgeSource = pathDict
-        if memoize:
-            self.pathsToNode[node] = pathDict
         return visitedList
     
-    def getAllPredecessorsUnordered(self, node, memoize = True) -> Set[int]:
-        """Returns a list of nodes that are predecessors of this one.
-        Memoize if source of new edge."""
-        memoize = memoize or node == self.newEdgeSource
-        if memoize:
-            pathDict = dict()
-        visited = set()
-        currentNode = node
-        stack = [currentNode]
-        path = []
-        while (len(stack) > 0):
-            currentNode = stack.pop()
-            if currentNode == self._invalid_node:
-                path = path[1:]
-            elif currentNode not in visited:
-                visited.add(currentNode)
-                path = [currentNode] + path
-                if (memoize):
-                    pathDict[currentNode] = self.deepcopyList(path)
-                stack.append(self._invalid_node)
-                stack = self.extend(stack, self.compactBkdGraph[currentNode])
-        if self._debug:
-            print("getAllPredecessors: node: {} and visitedList: {}".format(
-                node, visited))
-        if node == self.newEdgeSource:
-            self.pathsToNewEdgeSource = pathDict
-        if memoize:
-            self.pathsToNode[node] = pathDict
-        return visited
-
     def getAllSuccessors(self, node) -> List[int]:
         """Returns a list of nodes that are successors of this one.
         Memoize if sink of new edge."""
@@ -169,7 +137,7 @@ class BaseOnlineChecker():
         if not bool(self.pathsFromNewEdgeSink):
             self.getAllSuccessors(self.newEdgeSink)
         if not bool(self.pathsToNewEdgeSource):
-            self.getAllPredecessors(self.newEdgeSource)
+            self.getAllPredecessors(self.newEdgeSource, memoize=False)
         firstSegment = self.pathsToNewEdgeSource[source]
         lastSegment = self.pathsFromNewEdgeSink[sink]
         (_, secondPath) = self.findPath(source, sink)
@@ -191,8 +159,6 @@ class BaseOnlineChecker():
         self.compactBkdGraph = []
         self.pathsToNewEdgeSource = dict()
         self.pathsFromNewEdgeSink = dict()
-        # sink node -> {source node -> path}
-        self.pathsToNode = dict()
         self._no_edge = _default_no_edge
         self.noIdentity = False
         # Hack to make sub classes works
[…]

### 2020-08-10 acmPaper/paper.tex
section 3 revision

@@
 \usepackage{tikz}
+\usetikzlibrary{snakes}
 \usepackage[norelsize]{algorithm2e}
@@
 % where V is the set of vertices in the diagram, and E, the set of edges.
-this is a \emph{batch} algorithm: the entire system must be known ahead of time.
+but this is a \emph{batch} algorithm: the entire diagram must be known ahead of time.
 We present an \emph{online} algorithm that efficiently verifies that a commutative diagram remains commutative when adding a new edge.
@@
 Many systems can be understood as diagrams: graphs where nodes represent domains, and edges, directed transformation functions between nodes.
-In programming, for example, type systems with coercions can be drawn as a graph with type nodes and coercion edges, as demonstrated in figure ~\ref{fig:typeExample}.
+In programming, for example, type systems with coercions can be drawn as a graph with type nodes and coercion edges, as demonstrated in Figure ~\ref{fig:typeExample}.
 A desirable trait of such a type system and many other similar systems is that traversing the graph from a given start and end node results in the same value independent of path chosen.
@@
     \centering
-    \begin{subfigure}{.4\textwidth}
+    \begin{subfigure}{0.4\textwidth}
         \begin{verbatim}
@@
 
-    \begin{subfigure}{.4\textwidth}
+    \begin{subfigure}{0.4\textwidth}
         \begin{tikzpicture}[scale=0.2]
@@
 Since diagrams may change over time, and verifying the entire system from scratch may be computationally expensive, it would be additionally convenient if new edges could be checked online as they are added.
-In figure ~\ref{fig:typeExample}, for example, an interpreter that caught the first instance of a user making a bad definition would need to build and verify the conversion graph at run time.
+In Figure ~\ref{fig:typeExample}, for example, an interpreter that caught the first instance of a user making a bad definition would need to build and verify the conversion graph at run time.
 
-Naively checking if all path pairs that begin and end at the same node in a given diagram commute could require a number of function equality checks that grow as factorial in the number of nodes, since a path consists of an ordering of nodes.
+Na\"{i}vely checking if all path pairs that begin and end at the same node in a given diagram commute could require a number of function equality checks that grow as factorial in the number of nodes, since a path consists of an ordering of nodes.
 Further, the presence of cycles implies a potentially infinite number of paths. 
@@
 
-We compare our solution to three baseline implementations: a naive checking of all path pairs with only special handling for cycles, a check for all path pairs that involve the new edge, and an algorithm suggested by previous work to solve the batch version of the problem for acyclic diagrams.
+We compare our solution to three baseline implementations: a Na\"{i}ve checking of all path pairs with only special handling for cycles, a check for all path pairs that involve the new edge, and an algorithm suggested by previous work to solve the batch version of the problem for acyclic diagrams.
 Our proposed algorithms run orders of magnitude faster than the baselines.
@@
 The notation $\partial(p)$ maps to the tuple (start node of $p$, end node of $p$).
+$\partial(p)^{+}$ is the start node, and $\partial(p)^{-}$ is end node).
 
 A pair of paths $p$ and $q$ is called \textit{parallel} iff their terminal nodes are the same, i.e., $\partial(p)=\partial(q)$.
+$\partial$, $\partial^{+}$ and $\partial^{-}$ are extended to apply to parallel pairs.
+For parallel pair $p$ = ($p_1$, $p_2$), $\partial(p)=\partial(p_1)=\partial(p_2)=(\partial(p)^{+}, \partial(p)^{-})$.
 Let $R_{all}$ be the set of all parallel pairs of paths in the diagram.
@@
 For graphics programs in Gator, for example, where edges are matrices, function equality checking would be comparing matrix elements.
-We make an abstraction - equality of functions is defined by some oracle whose specifics will vary by target functions and domain - and reduce the \textsc{Online addition problem} to the \textsc{Verification set problem}, which given a diagram and a new edge, returns the set of parallel pairs of paths, such that if and only if the members in each pair are equal, then the new graph must commute.
+We make an abstraction --- equality of functions is defined by some oracle whose specifics will vary by target functions and domain --- and reduce the \textsc{Online addition problem} to the \textsc{Verification set problem}, which given a diagram and a new edge, returns the set of parallel pairs of paths, such that if and only if the members in each pair are equal, then the new graph must commute.
 The output to the \textsc{Online addition problem} can then be obtained as whether equality checking for all pairs succeeds.
+
+The oracle would needs to preserve that equality is reflexive, symmetric and transitive in order to benefit from the guarantees of the algorithms below.
+For example, the oracle for currency graph (Section ~\ref{sec:currency}) only approximately preserves transitivity because of floating point error magnification, so while the algorithm catches mistakes, it does not guarantee that the diagram commutes.
 
@@
 
-\subsection{Very Na\"{i}ve Baseline}
+\subsection{Na\"{i}ve Baseline}
 
 We start with the set of all parallel pairs in the diagram (with the new edge added in), and pare it down to be finite by handling cycles.
-We find all simple cycles in the diagram, using a procedure like Johnson's algorithm. 
-\xxx[AK]{TODO cite the paper.}
-We verify for each cycle that a single traversal starting on any node is equal to the identity function by adding pairs of (cycle starting at node, identity function) for each node in the function to the output verification set.
-\xxx[AK]{Need cleaner way to put this.}
-As cycles must now be the identity, for any pair in the set of all parallel paths, any instance of a cycle can be removed to obtain an equivalent pair with shorter, cycle free paths.
+Then, using a procedure like Johnson's algorithm ~\cite{johnson}, we find all simple cycles in the diagram. 
+We create a cycle verification set, $C$, and verify for each cycle that a single traversal is equal to the identity function by adding $(v \rightarrow v, 1)$ for each node v in the cycle to $C$.
+Here, $v \rightarrow v$ is a simple cycle starting and ending at v, 1 is the identity function, and these must be verified to be equal to each other.
+We then create a set $P$ of all the paths in the diagram with no cycles, and filter the set $P \times P$, excluding pairs where the paths begin or end on different nodes, or are identical, to get the set of all cycle-free parallel pairs $Q$.
+The output of the algorithm is $C \cup Q$, the cycle verification pairs and non-cyclic parallel pairs.
+After verifying $C$, it is sufficient to verify only $Q$ (lemma ~\ref{one_occurence_lemma}) because cycles must now be the identity, so for any pair in the set of all parallel paths, any instance of a cycle can be removed to obtain an equivalent pair with shorter, cycle free paths.
 If the shorter pair has equal paths then the paths in the original pair must also be equal to each other.
 It is therefore safe to remove all pairs of paths with cycles, leaving only parallel pairs where neither path has a cycle.
-Since there are only a finite number of paths in the graph with no cycles, (to the order of n!
-as a path is an ordering on nodes), and the parallel pairs are a subset of the possibilities generated by choosing a first and second member from this finite set, the resultant set of all parallel pairs must be finite.
-\xxx[AK]{TODO Check if O(n!n) = O(n!). What about $O((n!)^2)$?}
-The output of the algorithm, then, is the union of the set of cycle verification pairs, and the parallel pairs found by doing an all-acyclic-path search from every node in the diagram with the new edge added.
+$P$ is finite, bounded by $2^{|V|}$, as a path without cycles is an ordering on nodes, each node occurring at most once.
+$|P \times P|$, and consequently, $|Q|$, are also finite, bounded by $2^{2|V|}$.
+Thus the algorithm terminates and returns a finite (if large) set.
 
-\subsection{Less Na\"{i}ve baseline}
-We further refine the output of the very na\"{i}ve baseline.
-The commutativity of the original diagram could be used to narrow down the set of pairs to verify.
-Pairs where both paths do not involve the new edge would remain equal (this would apply to cycles too).
+\subsection{Baseline Incremental}
+This baseline refines the output of the na\"{i}ve baseline.
+Like before, it start by creating a cycle verification set $C'$, but includes only the simple cycles that pass through the new edge.
+Then, instead of $Q$, the set of all non-cyclic parallel pairs, the algorithm obtains its subset $Q'$ consisting of all non-cyclic parallel pairs such that exactly one path in each pair passes through the new edge.
+To this end, the algorithm performs a \textit{two-flip tolerant path search} whose output is passed into a \textit{path extraction algorithm}.
+The result of the path extraction algorithm to get the final output $Q' \cup C'$.
+
+This narrowing can be done because the original diagram commutes.
+Pairs where both paths do not involve the new edge would remain equal (this would apply to cycles too; cycles that do not pass through the new edge must be the identity).
 Also, pairs where both paths involved the new edge would have to be equal.
-To see why this is true, each path could be thought of as consisting of the composition of three segments, the first segment from the common source to the source of the new edge, the second, the new edge itself, and the third, from the sink of the new edge to the common sink of the parallel pair.
+To see why this is true, each path could be thought of as consisting of the composition of three segments.
+For path pair $p$, and new edge from node $S$ to node $T$, the first segment extends from $\partial(p)^+$ to the $S$, the second, the new edge $(S, T)$ itself, and the third, from $T$ to $\partial(p)^-$.
+The new edge could only appear once because cycles have already been dealt with so only pairs where the path includes the new edge once need be checked (lemma ~\ref{one_occurence_lemma}).
 The first segment of both pairs would have to be equal because they existed in as parallel pairs in the original diagram, and similarly the third segments would also have to be equal.
-The second segments, consisting of the same edge, would also have to be equal.
-Therefore the composition of these three equal components would be equal.
-Note that the new edge could only appear once because cycles have already been dealt with and none of the paths under consideration hit a node twice.
-Thus we are left only to verify the pairs where exactly one of the paths includes the new edge.
-Further, only pairs where the path includes the new edge once need be checked (lemma ~\ref{one_occurence_lemma}).
+The second segments, consisting of the same edge, would also have to be equal because the oracle preserves the reflexivity of equality.
+Therefore the two paths of the pair, a composition of these three equal components, would be equal, since the oracle would preserve transitivity of equality.
+We are left only with parallel pairs where exactly one of the path passes through the new edge.
 
 \paragraph{Two flip tolerant path search}
-We use a "two-flip tolerant" path search to identify the pairs of paths where exactly one path includes the new edge.
+We use a ``two-flip tolerant'' path search from the source ($S$) to the sink ($T$) of the new edge to identify the pairs of paths where exactly one path includes the new edge.
+The idea is that every such pair corresponds to a two flip tolerant path from $S$ to $T$.
 
-In a normal directed graph path, only forward edges, i.e. edges that go outward from the current node, are considered. 
-A \textit{two flip path} consists of up to three phases: in the first phase, only backward edges- pointing inward to the current node- are accepted.
+In a normal directed graph path search, only forward edges, i.e. edges that go outward from the current node while executing the search are considered. 
+A \textit{two flip path} consists of up to three phases: in the first phase, only backward edges --- pointing inward to the source of the search --- are accepted.
 In the second phase, only forward edges are accepted, and in the third phase, again only backward edges are accepted.
+For a two flip tolerant path $p$, let $t_1(p)$ map to the first phase, $t_2(p)$, to the second, and $t_3(p)$, to the third.
 The node between the first two phases we refer to as the \textit{first flipping point}, which has both edges pointing outward; similarly, we refer to the node between the latter two phases as the \textit{second flipping point}, at which both edges point inwards.
@@
 We present the idea diagrammatically in figure \ref{figure_two_flip}.
-Arrows represent path phases (note that these are the composition of many edges, not a single edge).
+Squiggly arrows represent path phases (these are the composition of 0 or more edges, not a single edge).
 The new edge is represented with a dashed arrow.
+Here, $f_1 \circ f_2 \circ f_3$ is a two flip path, and $f_1 \circ (S, T) \circ f_2$ is a new path created because of the addition of ($S$,$T$) that forms a parallel pair with ($f_2$).
 
@@
 \draw (34.15,-35.6) node [above] {new edge};
-\draw [black] (24.46,-25.47) -- (19.64,-33.53);
+\draw [decoration=snake,decorate] (24.46,-25.47) -- (19.64,-33.53);
+\draw (20.5,-29.5) node {$f_1$};
 \fill [black] (19.64,-33.53) -- (20.48,-33.1) -- (19.62,-32.58);
-\draw [black] (29,-22.9) -- (38.3,-22.9);
+\draw [decoration=snake,decorate] (29,-22.9) -- (38.3,-22.9);
+\draw (34,-21) node {$f_2$};
 \fill [black] (38.3,-22.9) -- (37.5,-22.4) -- (37.5,-23.4);
-\draw [black] (48.52,-33.61) -- (42.98,-25.39);
+\draw [decoration=snake,decorate] (48.52,-33.61) -- (42.98,-25.39);
+\draw (47,-29.5) node {$f_3$};
 \fill [black] (42.98,-25.39) -- (43.01,-26.33) -- (43.84,-25.77);
@@
 
-Here, (S, $F_1$, $F_2$, T) is a two flip tolerant path. ($F_1$, S, T, $F_2$) is a new path, created because of the addition of (S,T), that conflicts with ($F_1$, $F_2$).
-
 The two flip tolerant path search returns the set of all paths between a given source and sink that have up to two flips (paths that omit one or more of the three phases are also accepted).
 
-The \textit{path extraction algorithm} then transforms the output of the two flip path search to the set of new pairs of parallel paths that arise in the diagram due to its addition, that need to be verified.
+The \textit{path extraction algorithm} then transforms the output of the two flip path search to the verification set, $Q' \cup C'$.
 
 \paragraph{Path extraction algorithm}
-The purpose of this algorithm is to derive the conflicting pairs that a given two-flip tolerant path corresponds to.
+Given a set of two flip tolerant paths from the new edge source to sink, the algorithm outputs a set of pairs to verify.
 
-Let the new edge added to the diagram be (S, T).
-It has already been shown how each path corresponds to a conflicting pair in the case where a path has two flips.
-When a path has only the first flip (which is to say, the third phase of the path is missing), but not the second flip, sink node T can be treated as the sink of the two conflicting path, while the first flipping point remains the source.
-Similarly when only the second flipping point is present then it is the sink, and S is the source.
+Let the new edge added to the diagram be $(S, T)$ and the input set of paths, $P$.
+The algorithm processes every two flip tolerant path $p$ in $P$ case-wise to obtain pairs to add to the output set.
+In the case where p has two flips, $t_1(p) \circ (S, T) \circ t_3(p)$ and $t_2(p)$ form a parallel pair.
+When p has only the first flip (which is to say, the third phase of the path is missing), the parallel paths are $t_1(p) \circ (S, T)$ and $t_2(p)$.
+Similarly when only the second flipping point is present (so that there is no first phase), then the parallel pair is  $(S, T) \circ t_3(p)$ and $t_2(p)$.
 Finally when no flipping points are present, there are two possibilities.
-Either the path is oriented from S to T, in which case the conflicting paths are simply the edge (S, T) and the entire flip-less path, or the path is oriented from T to S.
-In this case, we have found a cycle, which has already been handled.
+Either $p$ is a path from $S$ to $T$, in which case the parallel paths are simply the edge $(S, T)$ and $p$, or $p$ is a path from $T$ to $S$.
+In this case, we have found a cycle, $p \circ (S,T)$, to be paired with the identity function.
+Like with the na\"{i}ve algorithm, for every node $v$ in the cycle, we add the pair $(v \rightarrow v, 1)$, where $v \rightarrow v$ is the cycle $p \circ (S,T)$ written to start and end at $v$.
 
-\paragraph{Theorem} Consider the result of two-flip tolerant path search from the source to sink node of an edge that is to be newly added followed with the path extraction algorithm described.
-This result has a one-to-one correspondence to the set of new parallel pairs with exactly one path passing through the new edge and neither paths containing any cycles.
+\begin{theorem}
+    Perform the two-flip tolerant path search from the source to sink node of the edge that is to be added followed, and on the output, apply the path extraction algorithm. 
+    The result is the set $O = Q'\cup C'$ of new parallel pairs with exactly one path passing through the new edge and neither paths containing any cycles, and the set of simple cycles passing through the new edge.
+\end{theorem}
 
-\paragraph{Proof}
+\begin{proof}
+    Every element in the output of the path extraction algorithm was by construction an element of $O$.
 
-Every element in the output of the path extraction algorithm was by construction a conflicting pair.
+    It is also clear that every cycle in $C'$ can be expressed as $(S, T) \circ p$, and corresponds to the input two flip tolerant path p.
 
-It remains to show that every new conflicting pair corresponds to a two flip tolerant path. Let the common source be $F_1$ and common sink be $F_2$.
-
-Only one path passes through (S, T), which we will name path 1.
-We construct a two flip tolerant path from S to T: phase 1 is the segment of path 1 from $F_1$ to S, phase 2 is path 2, and phase 3 is the segment of path 1 from T to $F_2$.
-It is possible that some of $F_1, F_2, S$ and $T$ coincide, in which case the corresponding segments between them disappear, and the resultant path has fewer than two flips.
-
-We have shown that all paths that need verification are caught by the two flip tolerant search followed with path extraction.
+    It remains to show that every new parallel pair $p$ in $Q'$ corresponds to a two flip tolerant path.
+    Let $\partial(p)^+ = F_1$ and $\partial(p)^- = F_2$. 
+    Only one path passes through $(S, T)$. Let it be called $p_1$, and the other path, $p_2$.
+    The two flip tolerant path from $S$ to $T$ can be constructed as follows: phase 1 is the segment of $p_1$ from $F_1$ to $S$, phase 2 is $p_2$, and phase 3 is the segment of $p_1$ from T to $F_2$.
+    Effectively, $F_1$ corresponds to the first flipping point, and $F_2$, to the second.
+    It is possible that some of $F_1, F_2, S$ and $T$ coincide (e.g., p starts at S --- $F_1 = S$), in which case the corresponding segments between the coinciding nodes can be considered the identity; the resultant path simply has fewer than two flips.
+\end{proof}
 
 \paragraph{Analysis}
-This algorithm is also in the worst case as bad as factorial in the number of nodes.
-However, in the average case, it significantly outperforms the naive batch baseline.
-For sparse diagrams it is even competitive with our ultimate polynomial solution.
-Average case heuristic analysis will be presented in the evaluation section.
+An upper bound on the number of pairs that this algorithm return is $O(|V|^{2} 2^{|V|})$, since two flip tolerant paths are an ordering on nodes, each node appearing at most once, followed by a selection of the flip points.
+In practice, the algorithm significantly outperforms the Na\"{i}ve batch baseline because it looks only at parallel pairs that involve the new edge, usually a small subset of all parallel pairs.
+Heuristic results are presented in the Section ~\ref{sec:evaluation}.
 
 \subsection{Optimal batch solution}
-Murota's main result ~\cite{commutative} is a solution to a batch version of \textsc{verification set} - given an acyclic diagram, it returns the minimal set of equality checks that would succeed if and only if the diagram commutes.
-The paper describes the following algorithm to find the ($V^2E$ bounded) minimal set of pairs that needs to be checked.
+Murota's main result ~\cite{commutative} is a solution to a batch version of \textsc{verification set}; given an acyclic diagram, it returns the minimal set of equality checks that would succeed if and only if the diagram commutes.
+The paper describes an algorithm to find the ($|V|^2|E|$ bounded) minimal set of pairs that needs to be checked.
 
-A diagram's structure can result in some redundancy, in the sense that verifying a subset of pairs can imply the verification of path pairs that aren't in the subset.
-For example, in figure ~\ref{figure_redundancies}, verifying $p_1$ and $p_2$ implies that the paths in $p_3$ are equal too.
+% A diagram's structure can result in some redundancy, in the sense that equivalence of a subset of pairs can imply the equivalence of path pairs that aren't in the subset.
+% For example, in Figure ~\ref{figure_redundancies}, verifying $p_1$ and $p_2$ implies that the paths in $p_3$ are equal too.
 
-\begin{figure}
+% \begin{figure}
 
-\begin{center}
-\begin{tikzpicture}[scale=0.2]
-\tikzstyle{every node}+=[inner sep=0pt]
-\draw [black] (12,-15.8) circle (3);
-\draw [black] (30.5,-30.5) circle (3);
-\draw [black] (46.4,-15.8) circle (3);
-\draw [black] (29.9,-15.8) circle (3);
-\draw [black] (15,-15.8) -- (26.9,-15.8);
-\fill [black] (26.9,-15.8) -- (26.1,-15.3) -- (26.1,-16.3);
-\draw (20.95,-16.3) node [below] {$f_1$};
-\draw [black] (14.35,-17.67) -- (28.15,-28.63);
-\fill [black] (28.15,-28.63) -- (27.84,-27.74) -- (27.21,-28.53);
-\draw (19.94,-23.64) node [below] {$g_1$};
-\draw [black] (29.063,-27.873) arc (-157.62406:-197.70132:13.638);
-\fill [black] (29.06,-27.87) -- (29.22,-26.94) -- (28.3,-27.32);
-\draw (27.49,-23.25) node [left] {$f_2$};
-\draw [black] (31.53,-18.309) arc (25.84176:-21.16715:11.992);
-\fill [black] (31.53,-18.31) -- (31.43,-19.25) -- (32.33,-18.81);
-\draw (33.28,-23.03) node [right] {$j_1$};
-\draw [black] (32.9,-15.8) -- (43.4,-15.8);
-\fill [black] (43.4,-15.8) -- (42.6,-15.3) -- (42.6,-16.3);
-\draw (38.15,-16.3) node [below] {$j_2$};
-\draw [black] (32.7,-28.46) -- (44.2,-17.84);
-\fill [black] (44.2,-17.84) -- (43.27,-18.01) -- (43.95,-18.75);
-\draw (39.77,-23.64) node [below] {$h_1$};
-\end{tikzpicture}
+% \begin{center}
+% \begin{tikzpicture}[scale=0.2]
+% \tikzstyle{every node}+=[inner sep=0pt]
+% \draw [black] (12,-15.8) circle (3);
+% \draw [black] (30.5,-30.5) circle (3);
+% \draw [black] (46.4,-15.8) circle (3);
+% \draw [black] (29.9,-15.8) circle (3);
+% \draw [black] (15,-15.8) -- (26.9,-15.8);
+% \fill [black] (26.9,-15.8) -- (26.1,-15.3) -- (26.1,-16.3);
+% \draw (20.95,-16.3) node [below] {$f_1$};
+% \draw [black] (14.35,-17.67) -- (28.15,-28.63);
+% \fill [black] (28.15,-28.63) -- (27.84,-27.74) -- (27.21,-28.53);
+% \draw (19.94,-23.64) node [below] {$g_1$};
+% \draw [black] (29.063,-27.873) arc (-157.62406:-197.70132:13.638);
+% \fill [black] (29.06,-27.87) -- (29.22,-26.94) -- (28.3,-27.32);
+% \draw (27.49,-23.25) node [left] {$f_2$};
+% \draw [black] (31.53,-18.309) arc (25.84176:-21.16715:11.992);
+% \fill [black] (31.53,-18.31) -- (31.43,-19.25) -- (32.33,-18.81);
+% \draw (33.28,-23.03) node [right] {$j_1$};
+% \draw [black] (32.9,-15.8) -- (43.4,-15.8);
+% \fill [black] (43.4,-15.8) -- (42.6,-15.3) -- (42.6,-16.3);
+% \draw (38.15,-16.3) node [below] {$j_2$};
+% \draw [black] (32.7,-28.46) -- (44.2,-17.84);
+% \fill [black] (44.2,-17.84) -- (43.27,-18.01) -- (43.95,-18.75);
+% \draw (39.77,-23.64) node [below] {$h_1$};
+% \end{tikzpicture}
 
-$p_1$=($f_1;f_2$, $g_1$)\\
-$p_2$=($j_1;j_2$, $h_1$)\\
-$p_3$=($f_1;h_1$, $f_1;j_2$)
-\end{center}
-\caption{Path checking redundancies: verifying $p_1$ and $p_2$ implies that the paths in $p_3$ are equal.}
-\label{figure_redundancies}
-\end{figure}
+% $p_1$=($f_1;f_2$, $g_1$)\\
+% $p_2$=($j_1;j_2$, $h_1$)\\
+% $p_3$=($g_1;h_1$, $f_1;j_2$)
+% \end{center}
+% \caption{Path checking redundancies: verifying $p_1$ and $p_2$ implies that the paths in $p_3$ are equal.}
+% \label{figure_redundancies}
+% \end{figure}
 
-The approach in this algorithm, at a high level, is to define a function that takes in a subset of pairs and returns the subset of pairs whose verification is implied by verifying the input set.
-Then, greedy elimination of redundancies is performed until a minimal set is reached.
+The approach in this algorithm, at a high level, is to define a function that takes in a subset of pairs and returns the subset of pairs whose equivalence is implied by the equivalence of the pairs in the input set.
+Then the algorithm greedily eliminates redundancies until a minimal set is reached.
 
@@
 
-Define $r_1>r_2$ for bilinkings $r_1$ = $\{p_l,q_l\}$, $r_2$= $\{p_2, q_2)$ $\in$ $R_0$, if there exists a path p such that $\partial p=\partial r_1$ and p contains $p_2$.
+Define $r_1>r_2$ for bilinkings $r_1$ = $\{p_1,q_1\}$, $r_2$= $\{p_2, q_2)$ $\in$ $R_0$, if there exists a path $p$ such that $\partial(p) = \partial(r_1)$ and $p$ contains $p_2$.
 Define $\langle\rangle$ as:
@@
 
-For bilinking $s$, let $F(s)$ be the vector in $\mathbb{F}2^{|E|}$ representing the edges present in s (the $n^{th}$ dimension of $F(s)$ is 1 if the corresponding edge is in $s$, and 0 otherwise).
+For bilinking $s$, let $F(s)$ be the vector in $\text{GF}(2)^{|E|}$ representing the edges present in s (the $n^{\text{th}}$ dimension of $F(s)$ is 1 if the corresponding edge is in $s$, and 0 otherwise).
+\xxx[AK]{GF(2) or $\mathbb{F}_2$ is standard notation for the Galois field of two elements. Do I need to write more about it here?}
 Let this function be extended to sets, so that for some set of bilinkings $S, F(S) = \{ F(s) | s\in S \}$. A notion of linear independence in this vector field exists.
 
-For a set of bilinkings $r$, the closure function $cl$ is defined as:
-\xxx[dg]{I think this definition is incomplete}\xxx[ak]{sorry, what is it missing?}
-$cl(r) = \{ s\in R_0| s$ is linearly dependent on $F(r) \}$.
+For a set of bilinkings $R$, the closure function $cl$ is defined as:
+$cl(R) = \{ s\in R_0| s$ is linearly {dependent on $F(R) \}$.
 
-The closure function on $r$ basically captures all the pairs that can be made by made by composing or "gluing together" the bilinkings in r.
+The closure function on $R$ basically captures all the pairs that can be made by made by composing or ``gluing together'' the bilinkings in $R$.
 
-Using these two function we finally define the function $\sigma$ on a set of bilinkings r as
-$\sigma(r) = \{s \in R_0 | s\in cl(R\cap \langle s \rangle) \}$.
-This is the function used to capture all the pairs whose verification is implied by the verification of pairs in r.
+Using these two function we finally define the function $\sigma$ on a set of bilinkings R as
+$\sigma(R) = \{s \in R_0 | s\in cl(R\cap \langle s \rangle) \}$.
+This is the function used to capture all the pairs whose equivalence is implied by the equivalence of pairs in $R$.
 
-$\sigma$ is used to iteratively check if a given pair is redundant. Bilinkings are eliminated until a minimum "spanning" subset is reached.
+$\sigma$ is used to iteratively check if a given pair is redundant.
+Bilinkings are eliminated until a minimum ``spanning'' subset is reached.
 
-% A complete implementation of the described algorithm is included in the appendix.
-
-Roughly, it proceeds by first efficiently finding a "spanning" set of bilinkings (a subset whose verification implies the verification of all bilinkings in the graph).
+Roughly, the algorithm proceeds by first efficiently finding a \textit{spanning} set of bilinkings (a subset whose verification implies the verification of all bilinkings in the graph).
 It does this by, starting at every node, finding the reachable subsection of the graph, and a spanning tree for the subsection.
-From each edges in the reachable section that is not a part of the tree, it generates a bilinking using the edge and a path in the tree that is parallel to the edge (algorithm ~\ref{algo_spanning_set}).
+From each edges in the reachable section that is not a part of the tree, it generates a bilinking using the edge and a path in the tree that is parallel to the edge (~\ref{algo_spanning_set}).
 
@@
 
-With the spanning set thus initialized, it greedily tries to remove each pair from the spanning set if the set remains spanning even after removing the edge (algorithm ~\ref{algo_minimal_spanning_set}).
+With the spanning set thus initialized, it greedily tries to remove each pair from the spanning set if the set remains spanning even after removing the edge (~\ref{algo_minimal_spanning_set}).
 
@@
 \SetKwProg{Fn}{Function}{:}{}
-\Fn{\FSigma{inputSet S, codomain = Rs}}{
+\Fn{\FSigma{inputSet S, codomain Rs}}{
     output $\gets$ \{\}\;
-    \For{bilinking $\in$ S}{
-        smallerPairs $\gets$ \{\}\;
-        \For{candidateBilinking $\in$ Rs}{
-            \If{$\exists$ path from bilinking.source to candidateBilinking.source
-            and
-            $\exists$ path from candidateBilinking.sink to bilinking.sink}{
-                \tcc{Check if there's a linear dependence between the bilinking and smallerPairs.
-                    Vectorize(pair) returns a vector with a dimension for each edge in the graph,
-                    with a 1 if pair includes that edge and 0 if it doesn't.}    
-                matrix $\gets$ [vectorize(pair) for pair in smallerPair]\;
-                matrix $\gets$ matrix+[vectorize(bilinking)]\;
-                \If{determinant(matrix)\%2==0}{
-                    output.add(bilinking)
-                }
-            }
+    \For{bilinking $\in$ Rs}{
+        \tcc{Get fragments that could build up to the bilinking.}
+        smallerPairs $\gets$ getSmallerPairs(bilinking)\;
+        consideredPieces $\gets$ smallerPairs $\cap$ S \;
+        \tcc{See if bilinking can be built from these pieces.}
+        \If{linearlyDependent(consideredPieces, bilinking)}{
+            output.add(bilinking)\;
         }
@@
 
-The proof of correctness can be found in Murota\cite{commutative}, 
+The proof of correctness can be found in Murota\cite{commutative}.
 The number of checks returned by the algorithm is at worst $O(|V|^2|E|)$. The overall run time of an optimized implementation is $O(|V|^4|E|^2)$.
@@
 
-Consider the situation in figure ~\ref{figure_reduction_rule}.
+Consider the situation in Figure ~\ref{figure_reduction_rule}.
 
@@
 \label{reductionRule}
-If conflicting paths $g_2 = f_2; (S,T); h_2$ then it must be that $g_1 = f_1; (S,T); h_1$.
+If parallel paths $g_2 = f_2; (S,T); h_2$ then it must be that $g_1 = f_1; (S,T); h_1$.
 \end{theorem} 
@@
 \subsection{Currency Graph}
+\label{sec:currency}
 
@@
 \section{Evaluation}
-We compare performance of the following path checking algorithms: (1) naive baseline (2) less naive, two-flip baseline (3) batch baseline (4) algorithm ~\ref{algo_online_polynomial}, and (5) algorithm ~\ref{algo_online_minimal}.
+\label{sec:evaluation}
+We compare performance of the following path checking algorithms: (1) Na\"{i}ve baseline (2) less Na\"{i}ve, two-flip baseline (3) batch baseline (4) algorithm ~\ref{algo_online_polynomial}, and (5) algorithm ~\ref{algo_online_minimal}.
 The two aspects we look at are time for response and size of response set (smaller sets -- tighter output results -- would mean less calls to the oracle).
@@
 
-The naive baseline performs poorly, taking well over a thousand seconds for even small graphs of 10 nodes.
+The Na\"{i}ve baseline performs poorly, taking well over a thousand seconds for even small graphs of 10 nodes.
 The batch algorithm brings timing down to seconds. Algorithm ~\ref{algo_online_polynomial} performs only slightly better.
@@
 \subsection{Comparison of algorithm time cost}
-The average time taken by each algorithm over the course of 10 runs over randomly generated graphs with 9 nodes and 32 edges is listed in table ~\ref{tab:times}.
+The average time taken by each algorithm over the course of 10 runs over randomly generated graphs with 9 nodes and 32 edges is listed in Table ~\ref{tab:times}.
 
@@
     \hline
-    Naive baseline & 0.77 \\
+    Na\"{i}ve baseline & 0.77 \\
     Two Flip tolerant & 0.075 \\
@@
 
-Algorithms scale as expected with time, as seen in figure ~\ref{fig:timeVsSize}. 
-Both ~\ref{algo_online_minimal} and ~\ref{algo_minimal_spanning_set} exhibit graphs that are polynomial in appearance, as does the batch baseline.
-\xxx[AK]{Do I need to justify this by doing some kind of fit?}
-The naive baseline as well as the two flip tolerant baseline display growth that quickly makes computation untenable.
+Algorithms scale as expected with time, as seen in Figure ~\ref{fig:timeVsSize}. 
+Both ~\ref{algo_online_minimal} and ~\ref{algo_minimal_spanning_set} exhibit graphs that are polynomial in appearance.
+The Na\"{i}ve baseline as well as the two flip tolerant baseline display quick growth.
 The batch algorithm also grows fast, though not as much as the online checking baselines.
@@
 Run time relates to the density of edges in the input graph.
-The degree of the effect differs with the algorithms, as figure ~\ref{fig:timeVsSize} shows.
+The degree of the effect differs with the algorithms, as Figure ~\ref{fig:timeVsSize} shows.
 Generally, denser graphs entail longer computation time.
@@
     % TODO format nicely.
-    \begin{subfigure}{.4\textwidth}
+    \begin{subfigure}{0.5\linewidth}
       \includegraphics[width=.8\linewidth]{timeVsSize_10_OptimalSet.png}
@@
 
-    \begin{subfigure}{.4\textwidth}
+    \begin{subfigure}{0.5\linewidth}
       \includegraphics[width=.8\linewidth]{timeVsSize_10_Polynomial.png}
@@
 
-    \begin{subfigure}{.4\textwidth}
+    \begin{subfigure}{0.5\linewidth}
       \includegraphics[width=.8\linewidth]{timeVsSize_10_NaiveChecker.png}
       \caption{Naive baseline}
-      \label{fig:sfigNaiveTvsS}
+      \label{fig:sfigNa\"{i}veTvsS}
     \end{subfigure}
 
-    \begin{subfigure}{.4\textwidth}
+    \begin{subfigure}{0.5\linewidth}
       \includegraphics[width=.8\linewidth]{timeVsSize_10_TwoFlipPathChecker.png}
@@
 
-    \begin{subfigure}{.4\textwidth}
+    \begin{subfigure}{0.5\linewidth}
       \includegraphics[width=.8\linewidth]{timeVsSize_10_BatchChecker.png}
@@
 \subsection{Variance}
-The periodic spikes in figure ~\ref{fig:sfigOptimalTvsS} are striking. 
+The periodic spikes in Figure ~\ref{fig:sfigOptimalTvsS} are striking. 
 We plot the spread of results to understand what is happening.
@@
 The outliers themselves follow a polynomial curve.
-The plots in figure ~\ref{fig:variance} depict the situation.
+The plots in Figure ~\ref{fig:variance} depict the situation.
 Grey points are the results of evaluation on individual points, and error bars show standard deviation. 
@@
     % TODO format nicely.
-    \begin{subfigure}{.4\textwidth}
+    \begin{subfigure}{0.5\linewidth}
       \includegraphics[width=.8\linewidth]{variance_10_OptimalSet.png}
@@
 
-    \begin{subfigure}{.4\textwidth}
+    \begin{subfigure}{0.5\linewidth}
       \includegraphics[width=.8\linewidth]{variance_10_Polynomial.png}
@@
 
-    \begin{subfigure}{.4\textwidth}
+    \begin{subfigure}{0.5\linewidth}
       \includegraphics[width=.8\linewidth]{variance_10_TwoFlip.png}
@@
     \hline
-    Naive baseline & 39754.9 \\
+    Na\"{i}ve baseline & 39754.9 \\
     Two Flip tolerant & 748.9 \\
@@
 
-Let V, the set of pairs to verify, be $C \cup R_0$.
+Let $V$, the set of pairs to verify, be $C \cup R_0$.
 

### 2020-08-10 acmPaper/paper.tex
typo

@@
             define wugs:
-                1 miles = 10 wugs;
-                1 foot = 10000 wugs;
+                1 miles = 10000 wugs;
+                1 foot = 10 wugs;
             var d : wugs = (wugs) a;

### 2020-08-10 acmPaper/paper.tex
idk
Added space complexity

@@
 The notation $\partial(p)$ maps to the tuple (start node of $p$, end node of $p$).
-$\partial(p)^{+}$ is the start node, and $\partial(p)^{-}$ is end node).
+Additionally, $\partial(p)^{+}$ is the start node, and $\partial(p)^{-}$ is end node of p.
 
@@
 
-An efficient implementation of the algorithm can run in $O(|V|^2(|V|+|E|))$ time at a $O(|V|^3)$ space cost.
-In such an implementation, path finding from a given source node to all potential sink nodes could be done in a single $O(|V|+|E|)$ breadth first search and the results memoized in $O(|V|^2)$ space.
+An efficient implementation of the algorithm can run in $O(|V|^2(|V|+|E|))$ time, space complexity not exceeding the asymptotic $O(|V|^2)$ bound on the output.
+In such an implementation, path finding from a given source node to all potential sink nodes could be done in a single $O(|V|+|E|)$ breadth first search.
 
@@
 The run time of the first step is $O(|V|^4)$, and that of the second step is $O(|V|)$, so that the overall bound is $O(|V|^4)$.
+Space complexity remains $O(|V|^2)$.
 

### 2020-08-10 acmPaper/paper.tex
Small changes in abstract

@@
 It is important to ensure that these diagrams commute: that any composing path of transformations from the same source to the same destination yields the same result.
-However, a straightforward approach to verifying commutativity must contend with cycles, and even so it runs in factorial time.
-While previous work has shown how to verify commutativity in the special case of acyclic diagrams in $O(|V|^4|E|^2)$ time,
+However, a straightforward approach to verifying commutativity must contend with cycles, and even so it runs in exponential time.
+Previous work has shown how to verify commutativity in the special case of acyclic diagrams in $O(|V|^4|E|^2)$ time,
 % I think this is clear enough (at least for the abstract). --AS

### 2020-08-10 acmPaper/paper.tex
small changes to introduction

@@
 
-To evaluate our results we use two case studies.
-First, our algorithm is used for the domain specific geometry type language \textit{Gator} ~\cite{gator} to ensure that user defined transformations between spaces stay consistent.
+We evaluate our algorithms against random graphs, and use them in two case studies.
+First, our algorithm is used in the domain specific geometry type language \textit{Gator} ~\cite{gator} to ensure that user defined transformations between spaces stay consistent.
 Second, our algorithm is used to identify inefficiencies in a currency conversion graph.
@@
 Our proposed algorithms run orders of magnitude faster than the baselines.
+
+\section{Related Work}
+
+Absolutely nothing related to this was ever done.
 

### 2020-08-11 acmPaper/paper.tex
Conclusion placeholder
Intro pass

--- 2020-08-10 acmPaper/paper.tex
@@
 \section{Introduction}
-Many systems can be understood as diagrams: graphs where nodes represent domains, and edges, directed transformation functions between nodes.
-In programming, for example, type systems with coercions can be drawn as a graph with type nodes and coercion edges, as demonstrated in Figure ~\ref{fig:typeExample}.
-A desirable trait of such a type system and many other similar systems is that traversing the graph from a given start and end node results in the same value independent of path chosen.
-A diagram with this property is said to \textit{commute}.
-With our type coercion example, it could be problematic if casting to a supposedly equivalent type as an intermediate step resulted in a different answer than a direct cast.
+Many systems can be understood as diagrams: graphs where nodes represent domains and edges represent directed transformation functions between nodes.
+A type system with coercions, for example, corresponds to a graph whose nodes are types and whose edges are coercions.
+Figure~\ref{fig:typeExample} illustrates an example in a simple language with units-of-measure types.
+\xxx[as]{Refer here to some citations of real type systems for units of measure.}
+In such a system, an important correctness criterion is that the diagram \emph{commutes:} when traversing the graph from any start node to any end node, applying every transformation along the path to any input value, the result is the same output value \emph{independent of the path chosen between the two nodes}.
+With our type coercion example, it is a problem if casting to a supposedly equivalent type as an intermediate step resulted in a different answer than a direct cast.
+\xxx[as]{Next, let's make the example specific: say that the expression ``(wugs) a'', for example, can produce value X or value Y, depending on whether the compiler uses path A or path B to do the conversion.}
 
@@
 
-Given the utility of system diagram commutativity, it would be useful to check that the property holds for real systems.
-Since diagrams may change over time, and verifying the entire system from scratch may be computationally expensive, it would be additionally convenient if new edges could be checked online as they are added.
-In Figure ~\ref{fig:typeExample}, for example, an interpreter that caught the first instance of a user making a bad definition would need to build and verify the conversion graph at run time.
+This paper is about efficiently checking diagrams that arise in real systems for commutativity.
+We assume a simple equivalence checker for individual transformation functions: in our type system example, for instance, it is possible to check transformation equivalence by comparing the conversion factors.
+Our aim is to analyze the graph of transformations and minimize the number of times we need to perform an equivalence check.
+Since diagrams may change over time in real systems, as new conversions are added, and verifying the entire system from scratch may be computationally expensive, we want an online method that only checks the impact of new edges.
+In Figure~\ref{fig:typeExample}, for example, a run-time system can catch the point where the programmer adds a bad conversion definition by verifying each new conversion edge as it is created.
 
-Na\"{i}vely checking if all path pairs that begin and end at the same node in a given diagram commute could require a number of function equality checks that grow as factorial in the number of nodes, since a path consists of an ordering of nodes.
+Efficient commutativity checking is not trivial.
+\xxx[as]{Let's flip these next two thoughts, because the first thing you have to deal with is resolving cycles---then you can worry about an efficient algorithm for acyclic graphs.}
+Na\"{i}vely checking if all path pairs that begin and end at the same node in a given diagram commute could require a number of function equality checks that grow as factorial in the number of nodes, because a path consists of an ordering of nodes.
 Further, the presence of cycles implies a potentially infinite number of paths.
@@
 
+\xxx[as]{Let's stick to present tense, even if it sounds a little weird.
+And we can state up front what the paper accomplishes.
+Something like this: ``This paper presents an algorithm for online checking of diagram commutativity and applies it to verifying systems of value conversions.
+We identify two key insights:''}
 In the course of efficiently verifying commutativity over online addition, we identified two key insights.
-The first was that in a commutative diagram, when a new edge is to be added, at most one path needs to be checked against for a given source and sink pair.
+The first was that in a commutative diagram, when a new edge is to be added, at most one path needs to be checked against \xxx[as]{can we rephrase this passive voice, which makes the sentence somewhat tricky to process?} for a given source and sink pair.
 Because the diagram commutes, all the paths between a given source and sink are equal and a representative to check against can arbitrarily be chosen.
@@
 
+\xxx[as]{This paragraph is a little confusing to me.
+First, what is ``informativeness''?
+Maybe we can be a little more specific.
+Second, I thought the previous paragraph said we already had an optimal number of calls to the oracle, but the second half of this paragraphs says that we need an even more expensive algorithm to obtain that minimum.}
 The second insight was that there is a single rule that places a partial, transitive ordering on the informativeness of paths.
-It allowed for the creation of a greedy, $O(|V|^4)$ optimization step, that results in the number of calls to the oracle being minimal.
+It allowed for the creation of a greedy, $O(|V|^4)$ optimization step that results in the number of calls to the oracle being minimal.
 The optimization is very useful when the equality checking oracle is expensive.
 
-To evaluate our results we use two case studies.
-First, our algorithm is used for the domain specific geometry type language \textit{Gator} ~\cite{gator} to ensure that user defined transformations between spaces stay consistent.
+We evaluate our algorithms against random graphs and use them in two case studies.
+\xxx[as]{Let's avoid passive voice in these next two sentences by saying ``we use the algorithm to do this'' instead of ``the algorithm is used to do this.''}
+First, our algorithm is used in the domain specific geometry type language \textit{Gator} ~\cite{gator} to ensure that user defined transformations between spaces stay consistent.
 Second, our algorithm is used to identify inefficiencies in a currency conversion graph.
 
-We compare our solution to three baseline implementations: a Na\"{i}ve checking of all path pairs with only special handling for cycles, a check for all path pairs that involve the new edge, and an algorithm suggested by previous work to solve the batch version of the problem for acyclic diagrams.
+We empirically compare our solution to three baseline implementations: a na\"{i}ve checking of all path pairs with only special handling for cycles, a check for all path pairs that involve the new edge, and an algorithm suggested by previous work to solve the batch version of the problem for acyclic diagrams.
 Our proposed algorithms run orders of magnitude faster than the baselines.
+
+\section{Related Work}
+
+\xxx{Absolutely nothing related to this was ever done.}
 
@@
 
-\section{Case Studies}
+\section{Conclusion}
+
+\xxx[as]{Write a conclusion.}
 

### 2020-08-11 acmPaper/paper.tex
Move RW to the end
Notation pass

@@
 
-\section{Related Work}
-
-\xxx{Absolutely nothing related to this was ever done.}
-
 \section{Formal Problem Setup and Terminology}
 
-We start by formalizing the notion of a diagram, drawing terminology from the previous acyclic work by Murota ~\cite{commutative}.
+We start by formalizing the notion of a diagram, drawing terminology from the previous acyclic work by Murota~\cite{commutative}.
 
-We start with a directed graph $G=(V,E)$, where $V$ consists of sets of elements and edges $(u, v)$ in $E$ correspond to functions that maps elements of $u$ to elements in $v$.
+\paragraph{Notation.}
+
+We start with a directed graph $G=(V,E)$, where $V$ corresponds to sets of elements and edges $(u, v)$ in $E$ correspond to functions that maps elements of $u$ to elements in $v$.
 All these functions form a semigroup $F$, where multiplication is function composition.
 A semigroup consists of a set and an associative binary operation, which we use to capture function composition.
-
+%
 The correspondence between edges and functions is stored as a mapping $f:E\rightarrow F$, where $f$ maps each edge to the function it represents.
 
-A path is a sequence of edges. The edge-to-function mapping $f$ can be naturally extended to paths: if path $p=e_1\circ e_2 ... e_n$ then $f(p)=f(e_1) \circ f(e_2) ... f(e_n)$.
-
-The notation $\partial(p)$ maps to the tuple (start node of $p$, end node of $p$).
-Additionally, $\partial(p)^{+}$ is the start node, and $\partial(p)^{-}$ is end node of p.
+A path is a sequence of edges.
+The edge-to-function mapping $f$ can be naturally extended to paths: if path $p=e_1\circ e_2 \circ \cdots \circ e_n$ then $f(p)=f(e_1) \circ f(e_2) \circ \cdots \circ f(e_n)$.
+%
+We write $\partial(p)^{+}$ for $p$'s start node, $\partial(p)^{-}$ for its end node, and $\partial(p)$ to denote the pair
+$(\partial(p)^{+}, \partial(p)^{-})$.
+\xxx[as]{Simplified this explanation; I hope it's OK.}
 
@@
 $\partial$, $\partial^{+}$ and $\partial^{-}$ are extended to apply to parallel pairs.
+\xxx[as]{Let's use a name other than $p$, which we have already used for paths, for pairs.
+(Or use something else for paths above.) Also, since we just used $p$ and $q$ for two paths, maybe that would be better than $p_1$ and $p_2$.}
 For parallel pair $p$ = ($p_1$, $p_2$), $\partial(p)=\partial(p_1)=\partial(p_2)=(\partial(p)^{+}, \partial(p)^{-})$.
-Let $R_{all}$ be the set of all parallel pairs of paths in the diagram.
 
-The diagram commutes iff $f(p)=f(q)$ $\forall (p,q)\in R_{all}$; that is, the composition of maps along any path connecting $u$ to $v$ is independent of path choice.
+Let $R_{all}$ be the set of all parallel pairs of paths in a given diagram.
+The diagram commutes iff
+$\forall (p,q)\in R_{all}$,
+$f(p)=f(q)$; that is, the composition of maps along any path connecting any pair $u$ to $v$ is independent of path choice.
+
+\paragraph{Problems.}
 
@@
 
+\section{Related Work}
+
+\xxx{Absolutely nothing related to this was ever done.}
+
 \section{Conclusion}

### 2020-08-11 acmPaper/paper.tex
Start problem description pass

@@
 Checking function equality is a domain specific, potentially hard problem, dependent on the nature of the graph.
-For graphics programs in Gator, for example, where edges are matrices, function equality checking would be comparing matrix elements.
-We make an abstraction --- equality of functions is defined by some oracle whose specifics will vary by target functions and domain --- and reduce the \textsc{Online addition problem} to the \textsc{Verification set problem}, which given a diagram and a new edge, returns the set of parallel pairs of paths, such that if and only if the members in each pair are equal, then the new graph must commute.
+For our case study in graphics programming (see Section~\ref{sec:gator}), edges are matrices and nodes are vector spaces, so function composition uses matrix multiplication and equivalence checking simply compares matrix values.
+We therefore assume some oracle for checking transformation function equivalence that will vary by domain.
+%
+We therefore reduce the \textsc{Online addition problem} to the \textsc{Verification set problem}.
+The latter, when given a diagram and a new edge, returns the set of parallel pairs of paths, such that if and only if the members in each pair are equal, then the new graph must commute.
 The output to the \textsc{Online addition problem} can then be obtained as whether equality checking for all pairs succeeds.
@@
 \section{Applications}
-TODO.
+
+\xxx[as]{Just a brief overview here to explain to the reader why these applications are important and appropriate (and what we hope to learn from the case studies).}
 
 \subsection{Gator}
+\label{sec:gator}
+
+\xxx{Write the Gator section!}
 

### 2020-08-11 acmPaper/paper.tex
Finish problem pass

@@
 We therefore reduce the \textsc{Online addition problem} to the \textsc{Verification set problem}.
+\xxx[as]{Although it's not technically a ``reduction,'' is it? Maybe better to say that the latter is what we solve, and we assume an oracle that uses those results to solve the former.}
 The latter, when given a diagram and a new edge, returns the set of parallel pairs of paths, such that if and only if the members in each pair are equal, then the new graph must commute.
+\xxx[as]{Let's be precise around the ``equal'' terminology in the previous and next sentences.
+It's not that the pairs are equal---it's that we check function equivalence for the pairs' corresponding functions.}
 The output to the \textsc{Online addition problem} can then be obtained as whether equality checking for all pairs succeeds.
 
-The oracle would needs to preserve that equality is reflexive, symmetric and transitive in order to benefit from the guarantees of the algorithms below.
-For example, the oracle for currency graph (Section ~\ref{sec:currency}) only approximately preserves transitivity because of floating point error magnification, so while the algorithm catches mistakes, it does not guarantee that the diagram commutes.
+The algorithms in the rest of the paper assume that the function equivalence oracle is reflexive, symmetric, and transitive.
+\xxx[as]{The following is a little tricky to understand here... maybe this is best left to the currency section?}
+For example, the oracle for currency graph (Section~\ref{sec:currency}) only approximately preserves transitivity because of floating point error magnification, so while the algorithm catches mistakes, it does not guarantee that the diagram commutes.
 

### 2020-08-11 acmPaper/paper.tex
3.1 pass

@@
 
-We start with the set of all parallel pairs in the diagram (with the new edge added in), and pare it down to be finite by handling cycles.
-Then, using a procedure like Johnson's algorithm ~\cite{johnson}, we find all simple cycles in the diagram. 
+\xxx[as]{Start with an overview of the steps.
+Something like: deal with cycles, then deal with the remaining paths in an acyclic graph.
+Maybe also give overall context to say that this algorithm is going to be exponential but at least it will be finite, unlike trying to check ``all paths.'' Then break up the discussion so there is one paragraph per step instead of one long paragraph.}
+
+We start with the set of all parallel pairs in the diagram (with the new edge added in) \xxx[as]{clarify: is this a batch or incremental algorithm?
+If it's a batch algorithm, then there is no need to discuss a ``new edge''}, and pare it down to be finite by handling cycles.
+Then, using a procedure like Johnson's algorithm~\cite{johnson}, we find all simple cycles in the diagram.
+\xxx[as]{I'm a tad confused here because this sentence says ``then...'' as if we had moved on from the previous step, but the previous step says it handled cycles already?}
 We create a cycle verification set, $C$, and verify for each cycle that a single traversal is equal to the identity function by adding $(v \rightarrow v, 1)$ for each node v in the cycle to $C$.
-Here, $v \rightarrow v$ is a simple cycle starting and ending at v, 1 is the identity function, and these must be verified to be equal to each other.
+\xxx[as]{It seems important to have already explained the ``type'' of $C$ (i.e., the output format of the algorithm?}
+Here, $v \rightarrow v$ is a simple cycle starting and ending at $v$, 1 is the identity function, and these must be verified to be equal to each other.
 We then create a set $P$ of all the paths in the diagram with no cycles, and filter the set $P \times P$, excluding pairs where the paths begin or end on different nodes, or are identical, to get the set of all cycle-free parallel pairs $Q$.
-The output of the algorithm is $C \cup Q$, the cycle verification pairs and non-cyclic parallel pairs.
-After verifying $C$, it is sufficient to verify only $Q$ (lemma ~\ref{one_occurence_lemma}) because cycles must now be the identity, so for any pair in the set of all parallel paths, any instance of a cycle can be removed to obtain an equivalent pair with shorter, cycle free paths.
+The output of the algorithm is $C \cup Q$, the cycle verification pairs and acyclic parallel pairs.
+After verifying $C$, it is sufficient to verify only $Q$ (Lemma~\ref{one_occurence_lemma}) because cycles must now be the identity, so for any pair in the set of all parallel paths, any instance of a cycle can be removed to obtain an equivalent pair with shorter, cycle-free paths.
+\xxx[as]{What does it mean to ``verify only $Q$''? This could be a bit more precise...}
 If the shorter pair has equal paths then the paths in the original pair must also be equal to each other.

### 2020-08-11 acmPaper/paper.tex
Pass through 3.2

@@
 
-\subsection{Na\"{i}ve Baseline}
+\subsection{Na\"{i}ve Baseline Algorithm}
 
@@
 
-\subsection{Baseline Incremental}
-This baseline refines the output of the na\"{i}ve baseline.
+\subsection{Baseline Incremental Algorithm}
+
+Our second baseline refines the output of the na\"{i}ve baseline.
+\xxx[as]{But even before saying that, let's give some overall context: remind the reader what the incremental setup looks like.
+Then we can explain that we are going to use something like the previous section.
+Perhaps we can also be a little more precise than ``refining'' the output, which could mean many things.}
 Like before, it start by creating a cycle verification set $C'$, but includes only the simple cycles that pass through the new edge.
@@
 To this end, the algorithm performs a \textit{two-flip tolerant path search} whose output is passed into a \textit{path extraction algorithm}.
+\xxx[as]{Maybe we can give a one-phrase explanation of what TFTPS does: it finds all the parallel pairs of which only one path includes the new edge, right?}
 The result of the path extraction algorithm to get the final output $Q' \cup C'$.
@@
 For path pair $p$, and new edge from node $S$ to node $T$, the first segment extends from $\partial(p)^+$ to the $S$, the second, the new edge $(S, T)$ itself, and the third, from $T$ to $\partial(p)^-$.
-The new edge could only appear once because cycles have already been dealt with so only pairs where the path includes the new edge once need be checked (lemma ~\ref{one_occurence_lemma}).
-The first segment of both pairs would have to be equal because they existed in as parallel pairs in the original diagram, and similarly the third segments would also have to be equal.
-The second segments, consisting of the same edge, would also have to be equal because the oracle preserves the reflexivity of equality.
-Therefore the two paths of the pair, a composition of these three equal components, would be equal, since the oracle would preserve transitivity of equality.
-We are left only with parallel pairs where exactly one of the path passes through the new edge.
+The new edge could only appear once because cycles have already been dealt with so only pairs where the path includes the new edge once need be checked (Lemma~\ref{one_occurence_lemma}).
+The first segment of both pairs would have to be equal because they existed as parallel pairs in the original diagram, and similarly the third segments would also have to be equal.
+The second segments, consisting of the same edge, would also have to be equal because the equivalence oracle is reflexive.
+Therefore, the two paths of the pair, a composition of these three equal components, would be equal, since the oracle would preserve transitivity of equality.
+We are left only with parallel pairs where exactly one of the paths passes through the new edge.
+
+\xxx[as]{Maybe a quick paragraph here to remind the reader what is still coming in the rest of this section?}
 
@@
 The idea is that every such pair corresponds to a two flip tolerant path from $S$ to $T$.
+\xxx[as]{However, ``two flip tolerant path'' has not yet been defined, so this remark sounds a little strange.}
 
-In a normal directed graph path search, only forward edges, i.e. edges that go outward from the current node while executing the search are considered. 
-A \textit{two flip path} consists of up to three phases: in the first phase, only backward edges --- pointing inward to the source of the search --- are accepted.
+In a normal directed graph path search, only forward edges, i.e., edges that go outward from the current node while executing the search are considered.
+A \textit{two flip path} consists of up to three phases: in the first phase, only backward edges---pointing inward to the source of the search---are accepted.
 In the second phase, only forward edges are accepted, and in the third phase, again only backward edges are accepted.
@@
 
-We present the idea diagrammatically in figure \ref{figure_two_flip}.
-Squiggly arrows represent path phases (these are the composition of 0 or more edges, not a single edge).
+We present the idea diagrammatically in Figure~\ref{figure_two_flip}.
+Squiggly arrows represent path phases (these are the composition of zero or more edges, not a single edge).
 The new edge is represented with a dashed arrow.
-Here, $f_1 \circ f_2 \circ f_3$ is a two flip path, and $f_1 \circ (S, T) \circ f_2$ is a new path created because of the addition of ($S$,$T$) that forms a parallel pair with ($f_2$).
+Here, $f_1 \circ f_2 \circ f_3$ is a two flip path, and $f_1 \circ (S, T) \circ f_2$ \xxx[as]{should that be $f_2$ at the end?} is a new path created because of the addition of ($S$,$T$) that forms a parallel pair with ($f_2$).
 
@@
 
-The \textit{path extraction algorithm} then transforms the output of the two flip path search to the verification set, $Q' \cup C'$.
-
 \paragraph{Path extraction algorithm}
+Next, the \textit{path extraction algorithm} then transforms the output of the two flip path search to the verification set, $Q' \cup C'$.
 Given a set of two flip tolerant paths from the new edge source to sink, the algorithm outputs a set of pairs to verify.
@@
 The algorithm processes every two flip tolerant path $p$ in $P$ case-wise to obtain pairs to add to the output set.
+\xxx[as]{Bullet points (itemize) are often a good way to lay out cases?}
 In the case where p has two flips, $t_1(p) \circ (S, T) \circ t_3(p)$ and $t_2(p)$ form a parallel pair.
@@
 Like with the na\"{i}ve algorithm, for every node $v$ in the cycle, we add the pair $(v \rightarrow v, 1)$, where $v \rightarrow v$ is the cycle $p \circ (S,T)$ written to start and end at $v$.
+
+\xxx[as]{A little signposting here might be helpful... maybe a new {\textbackslash}paragraph and a brief reminder to the reader about why we need a theorem to tie things up?}
 
@@
     Only one path passes through $(S, T)$. Let it be called $p_1$, and the other path, $p_2$.
-    The two flip tolerant path from $S$ to $T$ can be constructed as follows: phase 1 is the segment of $p_1$ from $F_1$ to $S$, phase 2 is $p_2$, and phase 3 is the segment of $p_1$ from T to $F_2$.
+    The two flip tolerant path from $S$ to $T$ can be constructed as follows: phase 1 is the segment of $p_1$ from $F_1$ to $S$, phase 2 is $p_2$, and phase 3 is the segment of $p_1$ from $T$ to $F_2$.
     Effectively, $F_1$ corresponds to the first flipping point, and $F_2$, to the second.
-    It is possible that some of $F_1, F_2, S$ and $T$ coincide (e.g., p starts at S --- $F_1 = S$), in which case the corresponding segments between the coinciding nodes can be considered the identity; the resultant path simply has fewer than two flips.
+    It is possible that some of $F_1, F_2, S$ and $T$ coincide (e.g., $p$ starts at $S$, i.e., $F_1 = S$), in which case the corresponding segments between the coinciding nodes can be considered the identity; the resultant path simply has fewer than two flips.
 \end{proof}
@@
 \paragraph{Analysis}
-An upper bound on the number of pairs that this algorithm return is $O(|V|^{2} 2^{|V|})$, since two flip tolerant paths are an ordering on nodes, each node appearing at most once, followed by a selection of the flip points.
-In practice, the algorithm significantly outperforms the Na\"{i}ve batch baseline because it looks only at parallel pairs that involve the new edge, usually a small subset of all parallel pairs.
-Heuristic results are presented in the Section ~\ref{sec:evaluation}.
+An upper bound on the number of pairs that this algorithm returns is $O(|V|^{2} 2^{|V|})$, since two flip tolerant paths are an ordering on nodes, each node appearing at most once, followed by a selection of the flip points.
+In practice, the algorithm significantly outperforms the na\"{i}ve batch baseline because it looks only at parallel pairs that involve the new edge, which is usually a small subset of all parallel pairs.
+Empirical results are presented in Section~\ref{sec:evaluation}.
 
-\subsection{Optimal batch solution}
-Murota's main result ~\cite{commutative} is a solution to a batch version of \textsc{verification set}; given an acyclic diagram, it returns the minimal set of equality checks that would succeed if and only if the diagram commutes.
+\subsection{Optimal Batch Solution}
+
+Murota's main result~\cite{commutative} is a solution to a batch version of \textsc{verification set}; given an acyclic diagram, it returns the minimal set of equality checks that would succeed if and only if the diagram commutes.
 The paper describes an algorithm to find the ($|V|^2|E|$ bounded) minimal set of pairs that needs to be checked.

### 2020-08-11 acmPaper/paper.tex
Pass through 3.3

@@
 \xxx[AK]{GF(2) or $\mathbb{F}_2$ is standard notation for the Galois field of two elements. Do I need to write more about it here?}
+\xxx[as]{Probably just to briefly tell the reader what it is, not to fully explain it.}
 Let this function be extended to sets, so that for some set of bilinkings $S, F(S) = \{ F(s) | s\in S \}$. A notion of linear independence in this vector field exists.
@@
 $cl(R) = \{ s\in R_0| s$ is linearly {dependent on $F(R) \}$.
-
+%
 The closure function on $R$ basically captures all the pairs that can be made by made by composing or ``gluing together'' the bilinkings in $R$. 
-
-Using these two function we finally define the function $\sigma$ on a set of bilinkings R as
+%
+Using these two functions, we finally define the function $\sigma$ on a set of bilinkings $R$ as
 $\sigma(R) = \{s \in R_0 | s\in cl(R\cap \langle s \rangle) \}$.
 This is the function used to capture all the pairs whose equivalence is implied by the equivalence of pairs in $R$.
-
-$\sigma$ is used to iteratively check if a given pair is redundant.
+%
+$\sigma$ is used \xxx[as]{passive voice, and a somewhat awkward start to the sentence} to iteratively check if a given pair is redundant.
 Bilinkings are eliminated until a minimum ``spanning'' subset is reached.
@@
 It does this by, starting at every node, finding the reachable subsection of the graph, and a spanning tree for the subsection.
-From each edges in the reachable section that is not a part of the tree, it generates a bilinking using the edge and a path in the tree that is parallel to the edge (~\ref{algo_spanning_set}).
+From each edges in the reachable section that is not a part of the tree, it generates a bilinking using the edge and a path in the tree that is parallel to the edge (\ref{algo_spanning_set}).
 

### 2020-08-11 acmPaper/paper.tex
Notation fixes

@@
 
-\section{Solving the online addition problem}
+\section{Solving the Online Addition Problem}
 
 We present a polynomial time solution to the \textsc{verification set} problem.
-Like with the baseline, we do not concern ourselves with parallel pairs where neither or both paths pass through the new edge.
+\xxx[as]{Elsewhere, ``problem'' is also in small-caps. Either way is fine, but consistency is important.}
+Like in the online baseline algorithm, we do not concern ourselves with parallel pairs where neither or both paths pass through the new edge.
 The key observation that allows us to improve on the online baseline is that, for a given source and sink pair, only a single parallel pair needs to be verified.
-This is an implication of Theorem \ref{reductionRule}, expanded on later.
-Theorem \ref{verifyingSet} shows that should our selected set of pairs along with cycles passing through the new edge be verified commutative, the entire diagram must commute. 
-The approach is to identify a parallel pair with exactly one path through the edge for each (source, sink) pair (algorithm ~\ref{algo_online_polynomial}).
+This is an implication of Theorem~\ref{reductionRule}, expanded on later.
+Theorem~\ref{verifyingSet} shows that should our selected set of pairs along with cycles passing through the new edge be verified commutative, the entire diagram must commute.
+The approach is to identify a parallel pair with exactly one path through the edge for each (source, sink) pair (\ref{algo_online_polynomial}).
 
@@
 The bound is asymptotically tight.
-This can be seen in the case where the graph contains 2N nodes besides S and T.
-We consider N of the nodes to be in group 1, and the other N to be in group 2.
-Every node in group 1 has a forward node to every node in group 2, as well as to S.
-T has a forward edge to every node in group 2.
-In this diagram, on the addition of edge (S, T), $N^2$ paths need to be verified which is polynomial in 2N+2.
+This can be seen in the case where the graph contains $2N$ nodes besides $S$ and $T$.
+We consider $N$ of the nodes to be in group 1, and the other $N$ to be in group 2.
+Every node in group 1 has a forward node to every node in group 2, as well as to $S$.
+$T$ has a forward edge to every node in group 2.
+In this diagram, on the addition of edge $(S, T)$, $N^2$ paths need to be verified which is polynomial in $2N+2$.
 
-Also notice that if trying to optimize for path length (say, if composing functions is expensive) then "find any path" can be replaced with "find shortest path".
+Also notice that if trying to optimize for path length (say, if composing functions is expensive) then ``find any path'' can be replaced with ``find the shortest path.''
 

### 2020-08-11 acmPaper/paper.tex
Fix references to algorithms
Sec 4 comments

@@
 
-\labelformat{algocf}{algorithm\,(#1)}
+\labelformat{algocf}{Algorithm~#1}
 
@@
 The approach is to identify a parallel pair with exactly one path through the edge for each (source, sink) pair (\ref{algo_online_polynomial}).
+\xxx[as]{Instead of putting the reference in parens, say ``Algorithm X is the strategy for doing Y. Some more details include Z.''}
 
@@
 The try block is executed at most $O(|V|^2)$ times, so that this is the bound on the number of pairs verified.
-
-The bound is asymptotically tight.
-This can be seen in the case where the graph contains $2N$ nodes besides $S$ and $T$.
-We consider $N$ of the nodes to be in group 1, and the other $N$ to be in group 2.
-Every node in group 1 has a forward node to every node in group 2, as well as to $S$.
+%
+The bound is asymptotically tight. This can be seen in the case where the graph contains $2N$ nodes besides $S$ and $T$.
+\xxx[as]{Introduce this concept by saying that ``we imagine dividing the nodes into two groups...''}
+We consider $N$ of the nodes to be in group 1 and the other $N$ to be in group 2.
+Every node in group 1 has a forward node to every node in group 2 as well as to $S$.
 $T$ has a forward edge to every node in group 2.
-In this diagram, on the addition of edge $(S, T)$, $N^2$ paths need to be verified which is polynomial in $2N+2$.
+In this diagram, on the addition of edge $(S, T)$, $N^2$ paths need to be verified which is polynomial in the total number of nodes, $2N+2$.
 
@@
 
-An efficient implementation of the algorithm can run in $O(|V|^2(|V|+|E|))$ time, space complexity not exceeding the asymptotic $O(|V|^2)$ bound on the output.
+An efficient implementation of the algorithm can run in $O(|V|^2(|V|+|E|))$ time, with space complexity not exceeding the asymptotic $O(|V|^2)$ bound on the output.
 In such an implementation, path finding from a given source node to all potential sink nodes could be done in a single $O(|V|+|E|)$ breadth first search.
@@
 We observe that there are some redundancies in the diagram.
-
-Consider the situation in Figure ~\ref{figure_reduction_rule}.
+%
+Consider the situation in Figure~\ref{figure_reduction_rule}.
+\xxx[as]{This needs a bit more explanation. What is the reader looking for? Also, the caption of this figure could probably be a little clearer.}
 
@@
 \end{tikzpicture}
-Each arrow represents a path, and (S,T) is the new edge being added.
+Each arrow represents a path, and $(S,T)$ is the new edge being added.
 \end{center}
@@
 
-The proof is not affected if any of these paths is the identity, eg. if $f_1$ is the identity and S and $P_1$ are actually the same node.
+The proof \xxx[as]{which proof?} is not affected if any of these \xxx[as]{which?} paths is the identity, e.g., if $f_1$ is the identity and $S$ and $P_1$ are actually the same node.
+\xxx[as]{Maybe this observation belongs somewhere else, where we are talking about the correctness argument/proof?}
 
@@
 
-Under the assumption that the only path operations allowed are composition and replacement of one path by a different, equal path, as would be true when edges are generic functions, and no other information is available, so that F is a semi-group, this "reduction" rule is also the only rule to reduce the set of path pairs to check by finding implications.
+Under the assumption that the only path operations allowed are composition and replacement of one path by a different, equal path, as would be true when edges are generic functions, and no other information is available, so that $F$ is a semi-group, this ``reduction'' rule is also the only rule to reduce the set of path pairs to check by finding implications.
 
@@
 
-Using this information it is possible to choose a minimal subset of path pairs to verify, as in algorithm ~\ref{algo_online_minimal}.
+Using this information it is possible to choose a minimal subset of path pairs to verify, as in \ref{algo_online_minimal}.
 
 We construct a graph with a node for each possible (source, sink) pair in the graph- each node then represents a possible choice for parallel pair endpoint pairs, and greedily search for the smallest set of nodes from which the entire graph would be reachable.
-The idea is to look for "roots" in the graph that have to be included in the ultimate verification set because they have no predecessor an cannot be verified "through" the verification of some other pair.
+The idea is to look for ``roots'' in the graph that have to be included in the ultimate verification set because they have no predecessor an cannot be verified ``through'' the verification of some other pair.
 Then all the successors whose verification is implied by the roots are eliminated.

### 2020-08-11 acmPaper/paper.tex
Sec 4 pass

@@
 
-The proof \xxx[as]{which proof?} is not affected if any of these \xxx[as]{which?} paths is the identity, e.g., if $f_1$ is the identity and $S$ and $P_1$ are actually the same node.
-\xxx[as]{Maybe this observation belongs somewhere else, where we are talking about the correctness argument/proof?}
+The proof is not affected if any of these \xxx[as]{which?} paths is the identity, e.g., if $f_1$ is the identity and $S$ and $P_1$ are actually the same node.
+\xxx[as]{Maybe this observation belongs inside the proof block, where it's most relevant?}
 
 We conclude that verifying a comparable pair of paths with end points ($P_1$, $P_2$) implies the verification of all path pairs ($Q_1$, $Q_2$) such that $Q_1$ is a successor of $P_1$ and $P_2$ is a successor of $Q_2$.
-A successor S to node N is any node such that there exists a path from N to S.
+A successor $S$ to node $N$ is any node such that there exists a path from $N$ to $S$.
 Nodes are also their own successors and predecessors.
+\xxx[as]{At some point, we have switched to capital letters for both paths and node variables... maybe make this consistent?}
 
+\xxx[as]{I couldn't quite follow the following, perhaps because the figure is not quite enough for me to understand what the ``rule'' is here.
+Which makes it hard to see why ``operations'' are relevant in this context.}
 Under the assumption that the only path operations allowed are composition and replacement of one path by a different, equal path, as would be true when edges are generic functions, and no other information is available, so that $F$ is a semi-group, this ``reduction'' rule is also the only rule to reduce the set of path pairs to check by finding implications.
@@
 Using this information it is possible to choose a minimal subset of path pairs to verify, as in \ref{algo_online_minimal}.
-
-We construct a graph with a node for each possible (source, sink) pair in the graph- each node then represents a possible choice for parallel pair endpoint pairs, and greedily search for the smallest set of nodes from which the entire graph would be reachable.
+%
+We construct a graph with a node for each possible (source, sink) pair in the graph: each node then represents a possible choice for parallel pair endpoint pairs, and we greedily search for the smallest set of nodes from which the entire graph would be reachable.
 The idea is to look for ``roots'' in the graph that have to be included in the ultimate verification set because they have no predecessor an cannot be verified ``through'' the verification of some other pair.
@@
 The result of the algorithm cannot be further reduced.
-Also, the verification of the parallel pairs returned in the algorithm implies that the output of the previous algorithm must commute, and transitively with Theorem ~\ref{verifyingSet} that the entire diagram must commute.
+\xxx[as]{Do we need an argument for why?}
+Also, the verification of the parallel pairs returned in the algorithm implies that the output of the previous algorithm must commute, and transitively with Theorem~\ref{verifyingSet} that the entire diagram must commute.
 
@@
 \subsection{Verification}
-Ultimately verification is performed by calling an equality oracle (that is beyond the scope of this work) for every comparable pair returned by the previous stage, as in algorithm ~\ref{algo_verification}.
+Ultimately verification is performed by calling an equality oracle (that is beyond the scope of this work) for every comparable pair returned by the previous stage, as in \ref{algo_verification}.
 

### 2020-08-11 acmPaper/paper.tex
Comments on 6

@@
 
+\xxx[as]{To perhaps excuse why this section doesn't have a proper programming language attached, maybe we should briefly say that we are imagining a units-of-measure type system---but to make it more interesting and scale to large graphs easily, we are using currencies as units.}
+
 To demonstrate our algorithms applied to a real world situation,
@@
 
-Using an API ~\footnote{https://exchangeratesapi.io}, we built the fully connected diagram of exchange rates between 32 currencies on a given day.
+Using a web API\footnote{https://exchangeratesapi.io} for currency data, we built the fully connected diagram of exchange rates between 32 currencies on a given day.
 To ensure that it indeed commuted, we started with an empty diagram, and added in edges one by one.
-Before the addition of each edge, we used the algorithms (~\ref{algo_online_polynomial} and ~\ref{algo_online_minimal}) to ensure the addition of a new edge did not introduce inconsistencies in the existing diagram.
+Before the addition of each edge, we used the algorithms (\ref{algo_online_polynomial} and ~\ref{algo_online_minimal}) \xxx[as]{remind the reader what these are: our algorithms?
+baselines?
+batch?
+incremental?
+etc.} to ensure the addition of a new edge did not introduce inconsistencies in the existing diagram.
 If a new edge was, in fact, problematic, the algorithms returned an example inconsistent
@@
 
-We allowed an "error tolerance", so that differences reported would not be the trivial consequences of a floating point error.
+We allowed an ``error tolerance'' so that differences reported would not be the trivial consequences of a floating point error.
 However, this relaxation of the equality oracle into imprecision meant that the mathematical reasoning that allow the algorithms to remove redundant path checks no longer applied.
-For instance, composing a new function with two approximately equal functions does not lead to equal results, so theorem ~\ref{reductionRule} fails with this approximate equality.
+For instance, composing a new function with two approximately equal functions does not lead to equal results, so Theorem~\ref{reductionRule} fails with this approximate equality.
 When the algorithms reported no inconsistencies, it was still possible that the graph possessed inconsistencies above the given threshold and did not commute.
 Nonetheless, both algorithms were effective in catching inconsistencies.
-Algorithm ~\ref{algo_online_polynomial} started finding inconsistencies at error tolerances to the order of $10^{-3}$, and ~\ref{algo_online_minimal}, which makes more invalid redundant path check removals, at error tolerances to the order of $10^{-7}$.
+\ref{algo_online_polynomial} started finding inconsistencies at error tolerances to the order of $10^{-3}$, and \ref{algo_online_minimal}, which makes more invalid redundant path check removals, at error tolerances to the order of $10^{-7}$.
 
-Averaging over evaluation for the first 30 days of 2020, building and verifying a diagram to completion (inclusive of the time required by network calls) took 243$\pm$19 seconds using algorithm ~\ref{algo_online_minimal}, and in 133$\pm$13 seconds with algorithm ~\ref{algo_online_polynomial}.
+Averaging over evaluation for the first 30 days of 2020, building and verifying a diagram to completion (inclusive of the time required by network calls) took 243$\pm$19 seconds using \ref{algo_online_minimal}, and in 133$\pm$13 seconds with \ref{algo_online_polynomial}.
+\xxx[as]{What is the conclusion the reader should take away?}
 
@@
 \label{sec:evaluation}
-We compare performance of the following path checking algorithms: (1) Na\"{i}ve baseline (2) less Na\"{i}ve, two-flip baseline (3) batch baseline (4) algorithm ~\ref{algo_online_polynomial}, and (5) algorithm ~\ref{algo_online_minimal}.
-The two aspects we look at are time for response and size of response set (smaller sets -- tighter output results -- would mean less calls to the oracle).
+We compare performance of the following path checking algorithms:
+(1) the na\"{i}ve baseline,
+(2) the less na\"{i}ve two-flip baseline,
+(3) the batch baseline,
+(4) \ref{algo_online_polynomial}, and
+(5) \ref{algo_online_minimal}.
+\xxx[as]{Again, the reader probably needs a reminder about what the latter ones are.}
+The two aspects we look at are time for response and size of response set (smaller sets---tighter output results---would mean less calls to the oracle).
 We use randomly generated graphs for benchmarking, and vary their size.
@@
 
-The Na\"{i}ve baseline performs poorly, taking well over a thousand seconds for even small graphs of 10 nodes.
-The batch algorithm brings timing down to seconds. Algorithm ~\ref{algo_online_polynomial} performs only slightly better.
+The na\"{i}ve baseline performs poorly, taking well over a thousand seconds for even small graphs of 10 nodes.
+The batch algorithm brings timing down to seconds.
+\ref{algo_online_polynomial} performs only slightly better.
 Surprisingly, the optimal set algorithm cuts time cost by several orders of magnitude, and runs in milliseconds for small graphs.
@@
 
-\subsection{Comparison of algorithm time cost}
-The average time taken by each algorithm over the course of 10 runs over randomly generated graphs with 9 nodes and 32 edges is listed in Table ~\ref{tab:times}.
+\subsection{Comparison of Algorithm Time Cost}
+The average time taken by each algorithm over the course of 10 runs over randomly generated graphs with 9 nodes and 32 edges is listed in Table~\ref{tab:times}.
+\xxx[as]{Perhaps the previous paragraph summarizing the findings from this table should be moved here?}
 
@@
 
-\subsection{Scaling of time with input size}
+\subsection{Scaling of Time with Input Size}
 
-Algorithms scale as expected with time, as seen in Figure ~\ref{fig:timeVsSize}. 
-Both ~\ref{algo_online_minimal} and ~\ref{algo_minimal_spanning_set} exhibit graphs that are polynomial in appearance.
-The Na\"{i}ve baseline as well as the two flip tolerant baseline display quick growth.
+\xxx[as]{In general, I recommend the syntax ``Figure X shows the time results.
+They are interesting.'' over ``The time results indicate something interesting, as Figure X shows.'' Putting the reference up front helps the reader immediately see what figure they should be cross-referencing.
+(Occurs in both paragraphs below.)}
+
+Algorithms scale as expected with time, as seen in Figure~\ref{fig:timeVsSize}.
+\xxx[as]{The plots are weirdly tiny on my machine... any chance your plotting tool can output PDF images?}
+Both \ref{algo_online_minimal} and \ref{algo_minimal_spanning_set} exhibit graphs that are polynomial in appearance.
+\xxx[as]{\ref{algo_minimal_spanning_set} does not appear in the figure.}
+The na\"{i}ve baseline as well as the two flip tolerant baseline display quick growth.
 The batch algorithm also grows fast, though not as much as the online checking baselines.
@@
 Run time relates to the density of edges in the input graph.
-The degree of the effect differs with the algorithms, as Figure ~\ref{fig:timeVsSize} shows.
+The degree of the effect differs with the algorithms, as Figure~\ref{fig:timeVsSize} shows.
 Generally, denser graphs entail longer computation time.
@@
 Table ~\ref{tab:sizes}, summarizes the number of output pairs that the algorithms returned on average over 10 runs, for graphs with 9 nodes and 32 edges.
+\xxx[as]{Conclusions. Are they as expected? If so, it's OK to just say why they were expected.}
+
 \begin{table}

### 2020-08-11 acmPaper/paper.tex
Move appendix to appendix
fix figures:

--- 2020-08-11 acmPaper/paper.tex
@@
 
-\labelformat{algocf}{Algorithm~#1}
+\labelformat{algocf}{algorithm\,(#1)}
 
@@
 
-\section{Solving the Online Addition Problem}
+\section{Solving the online addition problem}
 
 We present a polynomial time solution to the \textsc{verification set} problem.
-\xxx[as]{Elsewhere, ``problem'' is also in small-caps. Either way is fine, but consistency is important.}
-Like in the online baseline algorithm, we do not concern ourselves with parallel pairs where neither or both paths pass through the new edge.
+Like with the baseline, we do not concern ourselves with parallel pairs where neither or both paths pass through the new edge.
 The key observation that allows us to improve on the online baseline is that, for a given source and sink pair, only a single parallel pair needs to be verified. 
-This is an implication of Theorem~\ref{reductionRule}, expanded on later.
-Theorem~\ref{verifyingSet} shows that should our selected set of pairs along with cycles passing through the new edge be verified commutative, the entire diagram must commute.
-The approach is to identify a parallel pair with exactly one path through the edge for each (source, sink) pair (\ref{algo_online_polynomial}).
-\xxx[as]{Instead of putting the reference in parens, say ``Algorithm X is the strategy for doing Y. Some more details include Z.''}
+This is an implication of Theorem \ref{reductionRule}, expanded on later.
+Theorem \ref{verifyingSet} shows that should our selected set of pairs along with cycles passing through the new edge be verified commutative, the entire diagram must commute. 
+The approach is to identify a parallel pair with exactly one path through the edge for each (source, sink) pair (algorithm ~\ref{algo_online_polynomial}).
 
@@
 The try block is executed at most $O(|V|^2)$ times, so that this is the bound on the number of pairs verified.
-%
-The bound is asymptotically tight. This can be seen in the case where the graph contains $2N$ nodes besides $S$ and $T$.
-\xxx[as]{Introduce this concept by saying that ``we imagine dividing the nodes into two groups...''}
-We consider $N$ of the nodes to be in group 1 and the other $N$ to be in group 2.
-Every node in group 1 has a forward node to every node in group 2 as well as to $S$.
-$T$ has a forward edge to every node in group 2.
-In this diagram, on the addition of edge $(S, T)$, $N^2$ paths need to be verified which is polynomial in the total number of nodes, $2N+2$.
 
-Also notice that if trying to optimize for path length (say, if composing functions is expensive) then ``find any path'' can be replaced with ``find the shortest path.''
+The bound is asymptotically tight.
+This can be seen in the case where the graph contains 2N nodes besides S and T.
+We consider N of the nodes to be in group 1, and the other N to be in group 2.
+Every node in group 1 has a forward node to every node in group 2, as well as to S.
+T has a forward edge to every node in group 2.
+In this diagram, on the addition of edge (S, T), $N^2$ paths need to be verified which is polynomial in 2N+2.
 
-An efficient implementation of the algorithm can run in $O(|V|^2(|V|+|E|))$ time, with space complexity not exceeding the asymptotic $O(|V|^2)$ bound on the output.
+Also notice that if trying to optimize for path length (say, if composing functions is expensive) then "find any path" can be replaced with "find shortest path".
+
+An efficient implementation of the algorithm can run in $O(|V|^2(|V|+|E|))$ time, space complexity not exceeding the asymptotic $O(|V|^2)$ bound on the output.
 In such an implementation, path finding from a given source node to all potential sink nodes could be done in a single $O(|V|+|E|)$ breadth first search.
@@
 We observe that there are some redundancies in the diagram.
-%
-Consider the situation in Figure~\ref{figure_reduction_rule}.
-\xxx[as]{This needs a bit more explanation. What is the reader looking for? Also, the caption of this figure could probably be a little clearer.}
+
+Consider the situation in Figure ~\ref{figure_reduction_rule}.
 
@@
 \end{tikzpicture}
-Each arrow represents a path, and $(S,T)$ is the new edge being added.
+Each arrow represents a path, and (S,T) is the new edge being added.
 \end{center}
@@
 
-The proof \xxx[as]{which proof?} is not affected if any of these \xxx[as]{which?} paths is the identity, e.g., if $f_1$ is the identity and $S$ and $P_1$ are actually the same node.
-\xxx[as]{Maybe this observation belongs somewhere else, where we are talking about the correctness argument/proof?}
+The proof is not affected if any of these paths is the identity, eg. if $f_1$ is the identity and S and $P_1$ are actually the same node.
 
@@
 
-Under the assumption that the only path operations allowed are composition and replacement of one path by a different, equal path, as would be true when edges are generic functions, and no other information is available, so that $F$ is a semi-group, this ``reduction'' rule is also the only rule to reduce the set of path pairs to check by finding implications.
+Under the assumption that the only path operations allowed are composition and replacement of one path by a different, equal path, as would be true when edges are generic functions, and no other information is available, so that F is a semi-group, this "reduction" rule is also the only rule to reduce the set of path pairs to check by finding implications.
 
@@
 
-Using this information it is possible to choose a minimal subset of path pairs to verify, as in \ref{algo_online_minimal}.
+Using this information it is possible to choose a minimal subset of path pairs to verify, as in algorithm ~\ref{algo_online_minimal}.
 
 We construct a graph with a node for each possible (source, sink) pair in the graph- each node then represents a possible choice for parallel pair endpoint pairs, and greedily search for the smallest set of nodes from which the entire graph would be reachable.
-The idea is to look for ``roots'' in the graph that have to be included in the ultimate verification set because they have no predecessor an cannot be verified ``through'' the verification of some other pair.
+The idea is to look for "roots" in the graph that have to be included in the ultimate verification set because they have no predecessor an cannot be verified "through" the verification of some other pair.
 Then all the successors whose verification is implied by the roots are eliminated.
@@
     % TODO format nicely.
-    \begin{subfigure}{0.5\linewidth}
-      \includegraphics[width=.8\linewidth]{timeVsSize_10_OptimalSet.png}
+    \begin{subfigure}{\linewidth}
+      \includegraphics[width=\linewidth]{timeVsSize_10_OptimalSet.png}
       \caption{~\ref{algo_online_minimal}}
@@
 
-    \begin{subfigure}{0.5\linewidth}
-      \includegraphics[width=.8\linewidth]{timeVsSize_10_Polynomial.png}
+    \begin{subfigure}{\linewidth}
+      \includegraphics[width=\linewidth]{timeVsSize_10_Polynomial.png}
       \caption{~\ref{algo_online_polynomial}}
@@
 
-    \begin{subfigure}{0.5\linewidth}
-      \includegraphics[width=.8\linewidth]{timeVsSize_10_NaiveChecker.png}
+    \begin{subfigure}{\linewidth}
+      \includegraphics[width=\linewidth]{timeVsSize_10_NaiveChecker.png}
       \caption{Naive baseline}
@@
 
-    \begin{subfigure}{0.5\linewidth}
-      \includegraphics[width=.8\linewidth]{timeVsSize_10_TwoFlipPathChecker.png}
+\end{figure}
+\begin{figure}\ContinuedFloat
+
+    \begin{subfigure}{\linewidth}
+      \includegraphics[width=\linewidth]{timeVsSize_10_TwoFlipPathChecker.png}
       \caption{Two flip tolerant baseline}
@@
 
-    \begin{subfigure}{0.5\linewidth}
-      \includegraphics[width=.8\linewidth]{timeVsSize_10_BatchChecker.png}
+    \begin{subfigure}{\linewidth}
+      \includegraphics[width=\linewidth]{timeVsSize_10_BatchChecker.png}
       \caption{Batch algorithm baseline}
@@
 The periodic spikes in Figure ~\ref{fig:sfigOptimalTvsS} are striking. 
-We plot the spread of results to understand what is happening.
-We find ~\ref{algo_online_minimal} has outliers about two standard deviation above the mean responsible for the spikes in the average.
-The outliers themselves follow a polynomial curve.
-The plots in Figure ~\ref{fig:variance} depict the situation.
+We plot the spread of results in Figure~\ref{fig:sfigOptimalSpread} to understand what is happening.
 Grey points are the results of evaluation on individual points, and error bars show standard deviation. 
 The black curve traces the mean.
-No such effect is observed in ~\ref{algo_online_polynomial}.
+We find ~\ref{algo_online_minimal} has outliers about two standard deviation above the mean responsible for the spikes in the average.
+The outliers themselves follow a polynomial curve, appearing almost periodically.
+We have not yet identified the cause of the behavior.
+Figure ~\ref{fig:variance} depicts the situation for ~\ref{algo_online_polynomial}, where no such effect is observed.
+
 \begin{figure}
     % TODO format nicely.
-    \begin{subfigure}{0.5\linewidth}
-      \includegraphics[width=.8\linewidth]{variance_10_OptimalSet.png}
+    \begin{subfigure}{\linewidth}
+      \includegraphics[width=\linewidth]{variance_10_OptimalSet.png}
       \caption{~\ref{algo_online_minimal}}
@@
 
-    \begin{subfigure}{0.5\linewidth}
-      \includegraphics[width=.8\linewidth]{variance_10_Polynomial.png}
+    \begin{subfigure}{\linewidth}
+      \includegraphics[width=\linewidth]{variance_10_Polynomial.png}
       \caption{~\ref{algo_online_polynomial}}
@@
 
-    \begin{subfigure}{0.5\linewidth}
-      \includegraphics[width=.8\linewidth]{variance_10_TwoFlip.png}
-      \caption{Two flip tolerant baseline}
-      \label{fig:sfigTwoFlipSpread}
-    \end{subfigure}
+    % \begin{subfigure}{\linewidth}
+    %   \includegraphics[width=\linewidth]{variance_10_TwoFlip.png}
+    %   \caption{Two flip tolerant baseline}
+    %   \label{fig:sfigTwoFlipSpread}
+    % \end{subfigure}
 
@@
 
-\xxx{Absolutely nothing related to this was ever done.}
+As described in section ~\ref{sec:}
 

### 2020-08-11 acmPaper/paper.tex

--- 2020-08-11 acmPaper/paper.tex
@@
 
-\section{Appendix: Theorems and additional cases}
+\section{Related Work}
+
+\xxx{Absolutely nothing related to this was ever done.}
+
+\section{Conclusion}
+
+\xxx[as]{Write a conclusion.}
+
+\bibliography{refs}
+\bibliographystyle{plain}
+
+\appendix
+
+\section{Theorems and Additional Cases}
 
@@
 
-\section{Related Work}
-
-\xxx{Absolutely nothing related to this was ever done.}
-
 \section{Conclusion}

### 2020-08-11 acmPaper/paper.tex
Stub of related work

@@
 
+\label{sec:batchBaseline}
+
 Murota's main result~\cite{commutative} is a solution to a batch version of \textsc{verification set}; given an acyclic diagram, it returns the minimal set of equality checks that would succeed if and only if the diagram commutes.
 The paper describes an algorithm to find the ($|V|^2|E|$ bounded) minimal set of pairs that needs to be checked.
-
-% A diagram's structure can result in some redundancy, in the sense that equivalence of a subset of pairs can imply the equivalence of path pairs that aren't in the subset.
-% For example, in Figure ~\ref{figure_redundancies}, verifying $p_1$ and $p_2$ implies that the paths in $p_3$ are equal too.
-
-% \begin{figure}
-
-% \begin{center}
-% \begin{tikzpicture}[scale=0.2]
-% \tikzstyle{every node}+=[inner sep=0pt]
-% \draw [black] (12,-15.8) circle (3);
-% \draw [black] (30.5,-30.5) circle (3);
-% \draw [black] (46.4,-15.8) circle (3);
-% \draw [black] (29.9,-15.8) circle (3);
-% \draw [black] (15,-15.8) -- (26.9,-15.8);
-% \fill [black] (26.9,-15.8) -- (26.1,-15.3) -- (26.1,-16.3);
-% \draw (20.95,-16.3) node [below] {$f_1$};
-% \draw [black] (14.35,-17.67) -- (28.15,-28.63);
-% \fill [black] (28.15,-28.63) -- (27.84,-27.74) -- (27.21,-28.53);
-% \draw (19.94,-23.64) node [below] {$g_1$};
-% \draw [black] (29.063,-27.873) arc (-157.62406:-197.70132:13.638);
-% \fill [black] (29.06,-27.87) -- (29.22,-26.94) -- (28.3,-27.32);
-% \draw (27.49,-23.25) node [left] {$f_2$};
-% \draw [black] (31.53,-18.309) arc (25.84176:-21.16715:11.992);
-% \fill [black] (31.53,-18.31) -- (31.43,-19.25) -- (32.33,-18.81);
-% \draw (33.28,-23.03) node [right] {$j_1$};
-% \draw [black] (32.9,-15.8) -- (43.4,-15.8);
-% \fill [black] (43.4,-15.8) -- (42.6,-15.3) -- (42.6,-16.3);
-% \draw (38.15,-16.3) node [below] {$j_2$};
-% \draw [black] (32.7,-28.46) -- (44.2,-17.84);
-% \fill [black] (44.2,-17.84) -- (43.27,-18.01) -- (43.95,-18.75);
-% \draw (39.77,-23.64) node [below] {$h_1$};
-% \end{tikzpicture}
-
-% $p_1$=($f_1;f_2$, $g_1$)\\
-% $p_2$=($j_1;j_2$, $h_1$)\\
-% $p_3$=($g_1;h_1$, $f_1;j_2$)
-% \end{center}
-% \caption{Path checking redundancies: verifying $p_1$ and $p_2$ implies that the paths in $p_3$ are equal.}
-% \label{figure_redundancies}
-% \end{figure}
 
@@
 
-\xxx{Absolutely nothing related to this was ever done.}
+Section ~\ref{sec:batchBaseline} describes Murota's solution to efficiently finding the minimal set of path pairs that need to be compared to check if a given acyclic graph commutes ~\cite{commutative}.
+We did not find any other work that solves the question of verifying that diagrams commute.
 
+However, the question of commuting does come up in programming languages with implicit type conversion.
+Gator ~\cite{gator}, as described in Section ~\ref{sec:gator}, supports automatic type conversion between geometry types.
+The language implements some restrictions to eliminate obvious cases of non-commuting graphs, but does not verify that defined graphs commute, allowing scope for non-commuting graph definitions.
+Frink ~\cite{frink} is a language that supports automatic conversion between units and infinite precision floating point numbers.
+It does not appear to support the implicit definition of conversion between unites \xxx[AK]{Need to spend some time confirming this} but if extended to do so, would need to contend with the problem of commuting graphs.
+The same is true for F\# which has support for units of measure ~\cite{unitsOfMeasure} and Ada's GNAT compiler~\cite{gnat}.
+
+% Units of measure work resource: https://gmpreussner.com/research/dimensional-analysis-in-programming-languages
 \section{Conclusion}

### 2020-08-13 acmPaper/paper.tex
Bib fixes
Added section on Gator

@@
 \usepackage{subcaption}
+\usepackage{listings}
 %
@@
 \labelformat{algocf}{Algorithm~#1}
+
+\newcommand{\code}{\lstinline[keywordstyle=]}
 
@@
 
+To demonstrate our algorithms applied to a real world situation,
+we search for inconsistencies in diagrams of geometry transformations, and in a diagram of the exchange rate between currencies.
+
 \subsection{Gator}
@@
 
-\xxx{Write the Gator section!}
+Gator is a domain specific language designed around Geometry types, which are used to describe properties and transformations of geometric objects.
+A key feature of Gator is \code{in} expressions, which insert code to automatically transform between two geometry types.
+For example, given a point \code{p} represented in 2-dimensional Cartesian coordinates (which has type \code{cart2}), we can transform this point into polar coordinates using the expression \code{p in polar}.
+
+A key property of these \code{in} expressions is that the path chosen between two representations is arbitrary; thus we would like to ensure that every path is equivalent.
+This problem defines a natural application space for working with our commutative diagrams, so long as we can define a transformation graph and oracle.
+
+For simplicity, and to stay in scope for this paper, we focus on geometry transformations between \textit{reference frames}, which are the geometry equivalent of transforming between linear algebra basis vectors.
+Each edge on our transformation graph is a thus a matrix, with composition of edges as matrix multiplication and an oracle checking matrix equality (up to a rounding error $\epsilon$).
+
+These sorts of transformations are very common in graphics programming, and there are several examples of reasonably complicated transformation graphs that we can pick from.
+Gator includes several graphics examples as part of its examples package; for this evaluation, we looked at the \textit{phong}, \textit{reflection}, and \textit{shadow map} examples, sample images of which can be seen in Appendix~\ref{sec:gator-appendix}.
+
+We implemented a system for interfacing with the optimal set incremental checker in Gator \xxx[dg]{Which algorithm was it?}.
+Gator uses TypeScript for CPU computations, so it was reasonable to build a system to manage graphs for interoperation with the path independence algorithm.
+
+Using this algorithm, we found a bug in the reflection model, where a matrix was not being divided correctly by its determinant \xxx[dg]{List other things I find, if anything}.
+We also inserted previous bugs to check the algorithm was working correctly, and it identified all inserted bugs.
 
@@
 
-To demonstrate our algorithms applied to a real world situation,
-we found inconsistencies in a diagram of the exchange rate between currencies.
 Consider a diagram with nodes as currencies and a directed edge being the
@@
 
-\section{Conclusion}
+\section{Example Gator Images}
+\label{sec:gator-appendix}
+\begin{figure}
+	\begin{subfigure}{\linewidth}
+		\includegraphics[width=\linewidth]{phong.png}
+		\caption{Phong Lighting Model}
+	\end{subfigure}
 
-\xxx[as]{Write a conclusion.}
+	\begin{subfigure}{\linewidth}
+		\includegraphics[width=\linewidth]{reflection.png}
+		\caption{Reflection Model}
+	\end{subfigure}
 
-\bibliography{refs}
-\bibliographystyle{plain}
+	\begin{subfigure}{\linewidth}
+		\includegraphics[width=\linewidth]{shadowmap.png}
+		\caption{Shadow Map Model}
+	\end{subfigure}
+
+	% \begin{subfigure}{\linewidth}
+	%   \includegraphics[width=\linewidth]{variance_10_TwoFlip.png}
+	%   \caption{Two flip tolerant baseline}
+	%   \label{fig:sfigTwoFlipSpread}
+	% \end{subfigure}
+
+	\caption{Example Images from Gator Programs}
+	\label{fig:gator-images}
+\end{figure}
 

### 2020-08-13 acmPaper/paper.tex
Added conclusion

@@
             define wugs:
-                1 miles = 10000 wugs;
+                1 mile = 10000 wugs;
                 1 foot = 10 wugs;
@@
 
-\xxx[as]{Write a conclusion.}
+Being able to verify if diagrams commute can be a very useful program analysis tool.
+It allows a compiler to make deterministic automatic type conversions and can catch inconsistencies of definition in a program with user defined conversions.
+In this paper, we have presented verification algorithms that given a commuting diagram, a new edge, efficiently compute the set of paths that would equal to each other if and only if the diagram would still commute after addition of the edge.
+
+Integrating conversion consistency checks into widely used languages such as scala could provide a lot of value to the program.
+Beyond implementation work, this requires an answer to the question of how to write function equality oracles.
+
+In the evaluation section, we found that the graph for the optimal set path checker displayed a peculiar regularity in outliers.
+Future work might include investigating this behavior.
 

### 2020-08-14 transpilingSource/checker/hello.py
Updates to the transpiling source

@@ -71,8 +71,40 @@ class BaseOnlineChecker():
                 node, visitedList))
         if node == self.newEdgeSource:
             self.pathsToNewEdgeSource = pathDict
+        if memoize:
+            self.pathsToNode[node] = pathDict
         return visitedList
     
+    def getAllPredecessorsUnordered(self, node, memoize = True) -> Set[int]:
+        """Returns a list of nodes that are predecessors of this one.
+        Memoize if source of new edge."""
+        memoize = memoize or node == self.newEdgeSource
+        if memoize:
+            pathDict = dict()
+        visited = set()
+        currentNode = node
+        stack = [currentNode]
+        path = []
+        while (len(stack) > 0):
+            currentNode = stack.pop()
+            if currentNode == self._invalid_node:
+                path = path[1:]
+            elif currentNode not in visited:
+                visited.add(currentNode)
+                path = [currentNode] + path
+                if (memoize):
+                    pathDict[currentNode] = self.deepcopyList(path)
+                stack.append(self._invalid_node)
+                stack = self.extend(stack, self.compactBkdGraph[currentNode])
+        if self._debug:
+            print("getAllPredecessors: node: {} and visitedList: {}".format(
+                node, visited))
+        if node == self.newEdgeSource:
+            self.pathsToNewEdgeSource = pathDict
+        if memoize:
+            self.pathsToNode[node] = pathDict
+        return visited
+
     def getAllSuccessors(self, node) -> List[int]:
         """Returns a list of nodes that are successors of this one.
         Memoize if sink of new edge."""
@@ -137,7 +169,7 @@ class BaseOnlineChecker():
         if not bool(self.pathsFromNewEdgeSink):
             self.getAllSuccessors(self.newEdgeSink)
         if not bool(self.pathsToNewEdgeSource):
-            self.getAllPredecessors(self.newEdgeSource, memoize=False)
+            self.getAllPredecessors(self.newEdgeSource)
         firstSegment = self.pathsToNewEdgeSource[source]
         lastSegment = self.pathsFromNewEdgeSink[sink]
         (_, secondPath) = self.findPath(source, sink)
@@ -159,6 +191,8 @@ class BaseOnlineChecker():
         self.compactBkdGraph = []
         self.pathsToNewEdgeSource = dict()
         self.pathsFromNewEdgeSink = dict()
+        # sink node -> {source node -> path}
+        self.pathsToNode = dict()
         self._no_edge = _default_no_edge
[…]

### 2020-08-14 acmPaper/paper.tex

@@
 
-We implemented a system for interfacing with the optimal set incremental checker in Gator \xxx[dg]{Which algorithm was it?}.
+We implemented a system for interfacing with the optimal set incremental checker in Gator \xxx[dg]{Which algorithm was it?}\xxx[AK]{\ref{algo_online_minimal}}.
 Gator uses TypeScript for CPU computations, so it was reasonable to build a system to manage graphs for interoperation with the path independence algorithm.
 
-Using this algorithm, we found a bug in the reflection model, where a matrix was not being divided correctly by its determinant \xxx[dg]{List other things I find, if anything}.
+Using this algorithm, we found a bug in the reflection model, where a matrix was not being divided correctly by its determinant \xxx[dg]{List other things I find, if anything}\xxx[ak]{I think it may be nice to add more detail for at least one bug (code, what was the program doing, what the graph/conflict was, time taken, etc.)}.
 We also inserted previous bugs to check the algorithm was working correctly, and it identified all inserted bugs.

### 2020-09-04 implementation/evaluator.py
changes for profiling

@@ -1,4 +1,5 @@
 from datetime import datetime
+import cProfile
 import functools
 import json
 import matplotlib.pyplot as plt
@@ -22,9 +23,9 @@ sizeStep = 1
 maxSize = 5
 
 # densities = [i*densityStep for i in range(1, int(1.0/densityStep))] 
-densities = [0.1, 0.5, 0.9]
+densities = [0.3]
 # sizes = [i*sizeStep for i in range(1, int(maxSize/sizeStep))]
-sizes = [1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11]
+sizes = [9]
 
 evaluationList = [(density, size) for size in sizes for density in densities]
 
@@ -232,16 +233,34 @@ def dumpResult(results: dict, runDetails: runDetails):
 def getFileEndString():
     return "_{}_{}".format(NUM_TRIES, datetime.utcnow())
 
-checkerToResults = dict()
-graphs = { (density, size) : generateGraphs(size, density, acyclic=DO_ACYCLIC) for
-    density in densities for size in sizes }
-for checker in checkers:
+def main():
+    graphs = { (density, size) : generateGraphs(size, density, acyclic=DO_ACYCLIC) for
+        density in densities for size in sizes }
+    for checker in checkers:
+        evaluateForChecker(checker, graphs)
+
+def evaluateForChecker(checker, graphs):
     evaluate = getEvaluateFunction(checker, NUM_TRIES)
     # dictionary from (density, size) to attemptInfo
     results = { (density, size) : evaluate(density, size, graphs=graphs[(density, size)]) for 
         density in densities for size in sizes }
-    checkerToResults[checker.__class__.__name__] = results
     details = runDetails()
     details.checkerName = checker.__class__.__name__
     plot(checker.__class__.__name__, results)
     dumpResult(results, details)
+
+def profileBatchChecker():
+    """Profile the batch checker!"""
+    checker = BatchChecker()
+    graphs = { (density, size) : generateGraphs(size, density, acyclic=True) for
+        density in densities for size in sizes }
+    evaluateForChecker(checker, graphs) 
+
+if __name__=="__main__":
+    # main()
+    # cProfile.run('profileBatchChecker()')
+    pr = cProfile.Profile()
+    pr.enable()
+    profileBatchChecker()
+    pr.disable()
+    pr.print_stats(sort='cumtime')

### 2020-09-04 implementation/pathCheckers/batchChecker.py

@@ -1,6 +1,7 @@
 import setup
 from colorama import Fore
 import copy
+import cProfile
 from pathCheckers.naiveChecker import NaiveChecker
 from pathCheckers.baseOnlineChecker import BaseOnlineChecker
 from pathCheckers.batchTestCommonUtilities import addVectors

### 2020-09-04 acmPaper/paper.tex
added statement about profiling results

@@
 The batch algorithm brings timing down to seconds.
+However time cost still does not scale very well, with computation for a graph with 14 nodes and 0.4 density taking hours.
+Our implementation does not memoize the construction of the vector and matrix representation of paths in GF(2).
+Profiling reveals that this vector construction step is called many times and is chiefly responsible for the high time cost.
 \ref{algo_online_polynomial} performs only slightly better.

### 2020-09-04 acmPaper/paper.tex
Updates to section 4

--- 2020-08-14 acmPaper/paper.tex
@@
 \section{Introduction}
+\label{introduction}
 Many systems can be understood as diagrams: graphs where nodes represent domains and edges represent directed transformation functions between nodes.
@@
 
-We present a polynomial time solution to the \textsc{verification set} problem.
-\xxx[as]{Elsewhere, ``problem'' is also in small-caps. Either way is fine, but consistency is important.}
-Like in the online baseline algorithm, we do not concern ourselves with parallel pairs where neither or both paths pass through the new edge.
-The key observation that allows us to improve on the online baseline is that, for a given source and sink pair, only a single parallel pair needs to be verified.
-This is an implication of Theorem~\ref{reductionRule}, expanded on later.
-Theorem~\ref{verifyingSet} shows that should our selected set of pairs along with cycles passing through the new edge be verified commutative, the entire diagram must commute.
-The approach is to identify a parallel pair with exactly one path through the edge for each (source, sink) pair (\ref{algo_online_polynomial}).
-\xxx[as]{Instead of putting the reference in parens, say ``Algorithm X is the strategy for doing Y. Some more details include Z.''}
+We present a polynomial time solution to the \textsc{verification set problem} .
+As in the online baseline algorithm, we do not concern ourselves with parallel pairs where neither or both paths pass through the new edge.
+The key observation allowing us to improve on the online baseline is a result of Theorem~\ref{reductionRule} (which will be expanded on later): for a given source and sink pair, only a single parallel pair needs to be verified.
+Theorem~\ref{verifyingSet} shows that, should our selected set of pairs and cycles passing through the new edge be verified commutative, the entire diagram must commute.
+Algorithm \ref{algo_online_polynomial} uses this strategy of identifying a parallel pair with exactly one path through the edge for each (source, sink) pair.
 
@@
 
-The try block is executed at most $O(|V|^2)$ times, so that this is the bound on the number of pairs verified.
+The try block is executed at most $O(|V|^2)$ times, which is also the bound on the number of pairs verified.
 %
-The bound is asymptotically tight. This can be seen in the case where the graph contains $2N$ nodes besides $S$ and $T$.
-\xxx[as]{Introduce this concept by saying that ``we imagine dividing the nodes into two groups...''}
-We consider $N$ of the nodes to be in group 1 and the other $N$ to be in group 2.
+This bound is asymptotically tight, as can be seen in the case where the graph contains $2N$ nodes along $S$ and $T$.
+We imagine dividing the nodes into two groups of $N$ nodes each.
 Every node in group 1 has a forward node to every node in group 2 as well as to $S$.
@@
 
-Also notice that if trying to optimize for path length (say, if composing functions is expensive) then ``find any path'' can be replaced with ``find the shortest path.''
+If trying to optimize for path length (say, if composing functions is expensive) then ``find any path'' can be replaced with ``find the shortest path.''
 
@@
 
-We observe that there are some redundancies in the diagram.
-%
-Consider the situation in Figure~\ref{figure_reduction_rule}.
+We observe that there are some redundancies in the diagram shown in Figure~\ref{figure_reduction_rule}.  
 \xxx[as]{This needs a bit more explanation. What is the reader looking for? Also, the caption of this figure could probably be a little clearer.}
@@
 \end{tikzpicture}
-Each arrow represents a path, and $(S,T)$ is the new edge being added.
+Each arrow represents a path, where $(S,T)$ is the new edge being added.
 \end{center}
@@
 \[\Rightarrow l ; g_2 ; m = l ; f_2 ; (S,T) ; h_2 ; m \Rightarrow g_1 = f_1 ; (S,T) ; h_1\]
+
+The proof is not affected if any of the paths used are the identity, e.g., if $f_1$ is the identity and $S$ and $P_1$ are the same node.
 \end{proof}
-
-The proof is not affected if any of these \xxx[as]{which?} paths is the identity, e.g., if $f_1$ is the identity and $S$ and $P_1$ are actually the same node.
-\xxx[as]{Maybe this observation belongs inside the proof block, where it's most relevant?}
 
@@
 Nodes are also their own successors and predecessors.
-\xxx[as]{At some point, we have switched to capital letters for both paths and node variables... maybe make this consistent?}
+\xxx[as]{At some point, we have switched to capital letters for both paths and node variables... maybe make this consistent?} \xxx[dg]{I think these are all nodes, so it should be good}
 
@@
 %
-We construct a graph with a node for each possible (source, sink) pair in the graph: each node then represents a possible choice for parallel pair endpoint pairs, and we greedily search for the smallest set of nodes from which the entire graph would be reachable.
+To summarize this algorithm, we start by constructing a graph with a node for each possible (source, sink) pair in the graph: each node then represents a possible choice for parallel pair endpoint pairs, and we greedily search for the smallest set of nodes from which the entire graph would be reachable.
 The idea is to look for ``roots'' in the graph that have to be included in the ultimate verification set because they have no predecessor an cannot be verified ``through'' the verification of some other pair.
@@
 The result of the algorithm cannot be further reduced.
-\xxx[as]{Do we need an argument for why?}
+\xxx[as]{Do we need an argument for why?} \xxx[dg]{Yeah, we probably do}
 Also, the verification of the parallel pairs returned in the algorithm implies that the output of the previous algorithm must commute, and transitively with Theorem~\ref{verifyingSet} that the entire diagram must commute.
@@
 \subsection{Verification}
-Ultimately verification is performed by calling an equality oracle (that is beyond the scope of this work) for every comparable pair returned by the previous stage, as in \ref{algo_verification}.
+Ultimately verification is performed by calling an equality oracle (that is beyond the scope of this work) for every comparable pair returned by the previous stage, as in \ref{algo_verification}\xxx[as]{This algorithm can probably be summarized in one sentence and doesn't seem that helpful} .
 
@@
 
-\section{Applications}
+\section{Case Studies}
 
@@
 
-Gator is a domain specific language designed around Geometry types, which are used to describe properties and transformations of geometric objects.
+Gator is a domain specific language designed around Geometry types, which are used to describe properties and transformations of geometric objects~\cite{gator}.
 A key feature of Gator is \code{in} expressions, which insert code to automatically transform between two geometry types.
-For example, given a point \code{p} represented in 2-dimensional Cartesian coordinates (which has type \code{cart2}), we can transform this point into polar coordinates using the expression \code{p in polar}.
+For example, given a point \code{p} represented in 2-dimensional Cartesian coordinates (which has type \code{cart2}), we can transform this point into polar coordinates using the expression \underline{\code{p in polar}}.
+These \code{in} expressions create a structure of commutative diagrams, allowing use as introduced in Section~\ref{introduction}.
 
-A key property of these \code{in} expressions is that the path chosen between two representations is arbitrary; thus we would like to ensure that every path is equivalent.
-This problem defines a natural application space for working with our commutative diagrams, so long as we can define a transformation graph and oracle.
-
-For simplicity, and to stay in scope for this paper, we focus on geometry transformations between \textit{reference frames}, which are the geometry equivalent of transforming between linear algebra basis vectors.
+Specifically, Gator introduces transformations between \textit{reference frames}, which are the geometry equivalent of transforming between linear algebra basis vectors.
 Each edge on our transformation graph is a thus a matrix, with composition of edges as matrix multiplication and an oracle checking matrix equality (up to a rounding error $\epsilon$).
 
-These sorts of transformations are very common in graphics programming, and there are several examples of reasonably complicated transformation graphs that we can pick from.
-Gator includes several graphics examples as part of its examples package; for this evaluation, we looked at the \textit{phong}, \textit{reflection}, and \textit{shadow map} examples, sample images of which can be seen in Appendix~\ref{sec:gator-appendix}.
+There are several examples of reasonably complicated transformation graphs that we can pick from.
+Gator includes several graphics examples as part of its examples package; for this evaluation, we looked at the \textit{phong}, \textit{reflection}, and \textit{shadow map} examples, sample images of which can be seen in Appendix~\ref{sec:gator-appendix}\xxx[as]{Either add text to appendix B or just reference a figure}.
 
-We implemented a system for interfacing with the optimal set incremental checker in Gator \xxx[dg]{Which algorithm was it?}\xxx[AK]{\ref{algo_online_minimal}}.
-Gator uses TypeScript for CPU computations, so it was reasonable to build a system to manage graphs for interoperation with the path independence algorithm.
-
+We implemented a system for interfacing between the optimal set path checker (Algorithm 4) in the open-source implementation of Gator.
 Using this algorithm, we found a bug in the reflection model, where a matrix was not being divided correctly by its determinant \xxx[dg]{List other things I find, if anything}\xxx[ak]{I think it may be nice to add more detail for at least one bug (code, what was the program doing, what the graph/conflict was, time taken, etc.)}.
-We also inserted previous bugs to check the algorithm was working correctly, and it identified all inserted bugs.
+We also inserted previous bugs to check the algorithm was working correctly, and it identified all inserted bugs\xxx[as]{This general observation is less interesting than specific bugs, and should be omitted to just use the specifics}.
 
@@
 
-Being able to verify if diagrams commute can be a very useful program analysis tool.
-It allows a compiler to make deterministic automatic type conversions and can catch inconsistencies of definition in a program with user defined conversions.
+Being able to verify if diagrams commute allows a compiler to make
+deterministic automatic type conversions and can catch inconsistencies of definition in a program with user defined conversions.
 In this paper, we have presented verification algorithms that given a commuting diagram, a new edge, efficiently compute the set of paths that would equal to each other if and only if the diagram would still commute after addition of the edge.
 
-Integrating conversion consistency checks into widely used languages such as scala could provide a lot of value to the program.
-Beyond implementation work, this requires an answer to the question of how to write function equality oracles.
-
-In the evaluation section, we found that the graph for the optimal set path checker displayed a peculiar regularity in outliers.
-Future work might include investigating this behavior.
+Integrating conversion consistency checks into widely used languages such as Scala could provide a lot of value to the program.
+Beyond implementation work, this requires an answer to the question of how to write function equality oracles.\xxx[as]{Could we say a sentence or two about why this would be useful?
+Also maybe use Scala as an example of a broader languages with this property}
 

### 2020-09-07 acmPaper/paper.tex
Progress on paper

@@
 Using this algorithm, we found a bug in the reflection model, where a matrix was not being divided correctly by its determinant \xxx[dg]{List other things I find, if anything}\xxx[ak]{I think it may be nice to add more detail for at least one bug (code, what was the program doing, what the graph/conflict was, time taken, etc.)}.
-We also inserted previous bugs to check the algorithm was working correctly, and it identified all inserted bugs\xxx[as]{This general observation is less interesting than specific bugs, and should be omitted to just use the specifics}.
 
@@
 
-\xxx[as]{To perhaps excuse why this section doesn't have a proper programming language attached, maybe we should briefly say that we are imagining a units-of-measure type system---but to make it more interesting and scale to large graphs easily, we are using currencies as units.}
-
-Consider a diagram with nodes as currencies and a directed edge being the
-conversion rate from its source node's currency to its sink node's currency.
-Since the transformation exchange rate of money from any given base currency to a target currency
-can be expected to be the same regardless of which intermediate currency transformations
-are used, this diagram should commute.
+We imagine a units-of-measure type system as being an interesting application of concurrency graphs; however, to make this more interesting and scale nicely to large graphs with existing data, we focus on the specific unit of currencies.
+Consider a diagram with nodes as currencies and a directed edge being the conversion rate from its source node's currency to its sink node's currency.
+Since the transformation exchange rate of money from any given base currency to a target currency can be expected to be the same regardless of which intermediate currency transformations are used, this diagram should commute.
 
@@
 etc.} to ensure the addition of a new edge did not introduce inconsistencies in the existing diagram.
-If a new edge was, in fact, problematic, the algorithms returned an example inconsistent
-pair that would arise from the addition of the edge.
-The pair would consist of two currency transformation sequences with the 
-same source currency and ultimate destination currency, but with different effective 
-exchange rates values, as computed by taking the product of all the exchange rates
-encountered through the chain.
+If a new edge was problematic, the algorithms returned an example inconsistent pair that would arise from the addition of the edge.
+The pair would consist of two currency transformation sequences with the same source currency and ultimate destination currency, but with different effective exchange rates values, as computed by taking the product of all the exchange rates encountered through the chain.
 

### 2020-09-08 acmPaper/paper.tex
Finished first pass

@@
 There are several examples of reasonably complicated transformation graphs that we can pick from.
-Gator includes several graphics examples as part of its examples package; for this evaluation, we looked at the \textit{phong}, \textit{reflection}, and \textit{shadow map} examples, sample images of which can be seen in Appendix~\ref{sec:gator-appendix}\xxx[as]{Either add text to appendix B or just reference a figure}.
+Gator includes graphics examples as part of its examples package; for this evaluation, we looked at the \textit{phong}, \textit{reflection}, and \textit{shadow map} examples, sample images of which can be seen in Appendix~\ref{sec:gator-appendix}\xxx[as]{Either add text to appendix B or just reference a figure}.
 
 We implemented a system for interfacing between the optimal set path checker (Algorithm 4) in the open-source implementation of Gator.
-Using this algorithm, we found a bug in the reflection model, where a matrix was not being divided correctly by its determinant \xxx[dg]{List other things I find, if anything}\xxx[ak]{I think it may be nice to add more detail for at least one bug (code, what was the program doing, what the graph/conflict was, time taken, etc.)}.
+The system was tested with intentional bugs, of which it found them all, although no "real" bugs were found.
+The graphs used were of size 5 or less; for graphs of this size, the checker was able to run in real time with no noticeable loss of frames.
+Since the program is running at 60 frames per second, the checker was running at a rate faster than .01 seconds.
 
@@
 (3) the batch baseline,
-(4) \ref{algo_online_polynomial}, and
-(5) \ref{algo_online_minimal}.
+(4) \ref{algo_online_polynomial}, the non-minimal polynomial-time algorithm, and
+(5) \ref{algo_online_minimal}, the minimal set finding algorithm.
 \xxx[as]{Again, the reader probably needs a reminder about what the latter ones are.}
-The two aspects we look at are time for response and size of response set (smaller sets---tighter output results---would mean less calls to the oracle).
-We use randomly generated graphs for benchmarking, and vary their size.
-Given a graph and a new edge, we time how long it takes for an algorithm to return  the set of pairs that need to be verified.
-All computation was performed on a Macbook Pro 2015, 2.9 GHz Dual-Core Intel Core i5.
-
-The na\"{i}ve baseline performs poorly, taking well over a thousand seconds for even small graphs of 10 nodes.
-The batch algorithm brings timing down to seconds.
-However time cost still does not scale very well, with computation for a graph with 14 nodes and 0.4 density taking hours.
-Our implementation does not memoize the construction of the vector and matrix representation of paths in GF(2).
-Profiling reveals that this vector construction step is called many times and is chiefly responsible for the high time cost.
-\ref{algo_online_polynomial} performs only slightly better.
-Surprisingly, the optimal set algorithm cuts time cost by several orders of magnitude, and runs in milliseconds for small graphs.
-All implementations are sensitive to density, performing better when density is very low or very high.
+The two metrics we evaluate are time for response and size of response set (smaller sets---tighter output results---would mean less calls to the oracle).
+We use randomly generated graphs of varying size: given a graph and a new edge, we time how long it takes for an algorithm to return the set of pairs that need to be verified.
+All computations were performed on a Macbook Pro 2015, 2.9 GHz Dual-Core Intel Core i5.
 
@@
 
+The na\"{i}ve baseline performs poorly, taking well over a thousand seconds for even small graphs of 10 nodes.
+While the batch algorithm improves on this, it still does not scale very well, with computation for a graph with 14 nodes and 0.4 density taking hours.
+Our implementation does not memoize the construction of the vector and matrix representation of paths in GF(2); profiling indicates that this construction is a major factor in the high time cost for these algorithms.
+\ref{algo_online_polynomial} performs only slightly better than the batch algorithm.
+Surprisingly, the optimal set algorithm cuts time cost by several orders of magnitude, and runs in milliseconds for small graphs.
+All implementations are sensitive to density, performing better when density is low.
+
 \subsection{Scaling of Time with Input Size}
@@
 
-Algorithms scale as expected with time, as seen in Figure~\ref{fig:timeVsSize}.
+Figure~\ref{fig:timeVsSize} shows that the algorithms time scales with size, as expected.
 \xxx[as]{The plots are weirdly tiny on my machine... any chance your plotting tool can output PDF images?}
@@
 
-We define density to be the ratio of the number of edges in the graph to the total possible number of edges ($|V|^2$, where $|V|$ is the number of nodes).
+We define density to be the ratio of the number of edges in the graph to the total possible number of edges (which is $|V|^2$, where $|V|$ is the number of nodes).
 Run time relates to the density of edges in the input graph.

### 2020-09-08 acmPaper/paper.tex
section 4 pass

--- 2020-09-04 acmPaper/paper.tex
@@
 
-We observe that there are some redundancies in the diagram shown in Figure~\ref{figure_reduction_rule}.  
+If \ref{algo_online_polynomial} were applied to the diagram shown in Figure~\ref{figure_reduction_rule} there would be redundancies in the output.
+It turns out that verifying $g_2=f_2 \circ n \circ h_2$ is sufficient to ensure the diagram still commutes on the addition of $n$.
 \xxx[as]{This needs a bit more explanation. What is the reader looking for? Also, the caption of this figure could probably be a little clearer.}
@@
 \fill [black] (28,-41.9) -- (28.5,-41.1) -- (27.5,-41.1);
-\draw (27.5,-32.5) node [left] {$l$};
-\draw [black] (19.6,-34.06) -- (53.8,-33.64);
+\draw (27.5,-29.5) node [left] {$l$};
+\draw [dashed] (19.6,-34.06) -- (53.8,-33.64);
 \fill [black] (53.8,-33.64) -- (52.99,-33.15) -- (53.01,-34.15);
+\draw (36.25,-32.5) node [left] {$n$};
 \draw [black] (44.5,-41.9) -- (44.5,-23.1);
 \fill [black] (44.5,-23.1) -- (44,-23.9) -- (45,-23.9);
-\draw (45,-32.5) node [right] {$m$};
+\draw (42.5,-37.5) node [right] {$m$};
 \end{tikzpicture}
-Each arrow represents a path, where $(S,T)$ is the new edge being added.
+Each arrow represents a path, where $n$ is the new edge being added.
+While ~\ref{algo_online_polynomial} returns two pairs for verification, one from $P_1$ to $P_2$ and the other from $Q_1$ to $Q_2$, it actually suffices to just check a pair from $Q_1$ to $Q_2$ as demonstrated in theorem ~\ref{reductionRule}.
 \end{center}
@@
 \label{reductionRule}
-If parallel paths $g_2 = f_2; (S,T); h_2$ then it must be that $g_1 = f_1; (S,T); h_1$.
+If parallel paths $g_2 = f_2; n; h_2$ then it must be that $g_1 = f_1; n; h_1$.
 \end{theorem} 
@@
 We use the fact that $f_1$=$l; f_2$ and $h_1$=$h_2; m$.
-\[g_2 = f_2; (S,T); h_2 \Rightarrow l; g_2 = l; f_2; (S,T) ; h_2 \]
-\[\Rightarrow l ; g_2 ; m = l ; f_2 ; (S,T) ; h_2 ; m \Rightarrow g_1 = f_1 ; (S,T) ; h_1\]
+\[g_2 = f_2; n; h_2 \Rightarrow l; g_2 = l; f_2; n ; h_2 \]
+\[\Rightarrow l ; g_2 ; m = l ; f_2 ; n ; h_2 ; m \Rightarrow g_1 = f_1 ; n ; h_1\]
 
@@
 \xxx[as]{At some point, we have switched to capital letters for both paths and node variables... maybe make this consistent?} \xxx[dg]{I think these are all nodes, so it should be good}
+The rule effectively places an ordering on the informativeness of path pairs based on their terminal nodes.
+% If it is possible to draw a path from node $P_1$ to $P_2$ that passes through $Q_1$ and $Q_2$ then verifying a parallel pair from $Q_1$ to $Q_2$ implies that parallel pairs from $P_1$ to $P_2$ commute as well.
 
@@
 Which makes it hard to see why ``operations'' are relevant in this context.}
-Under the assumption that the only path operations allowed are composition and replacement of one path by a different, equal path, as would be true when edges are generic functions, and no other information is available, so that $F$ is a semi-group, this ``reduction'' rule is also the only rule to reduce the set of path pairs to check by finding implications.
+ Given that a set of path pairs are equal, suppose that one attempts to derive the proposition that a different parallel pair of paths is equal with a step-by-step application of inference rules.
+ Under the assumption that edges are generic functions, and no other information is available, $F$ is a semi-group.
+ The only inference rules allowed are composition (given that $f_1=f_2$, it must be that $g \circ f_1 = g \circ f_2$) and replacement of one path by a different, equal path (given $f_1=f_2$ and $g \circ f_1 = h \circ f_1 $, it must be true that $g \circ f_1 = h \circ f_2$).
+Any permutation of the repeated application of these two rules results in the ``reduction rule'' already described; it is therefore the only rule that can be used to reduce the set of path pairs to check.
 
@@
 %
-To summarize this algorithm, we start by constructing a graph with a node for each possible (source, sink) pair in the graph: each node then represents a possible choice for parallel pair endpoint pairs, and we greedily search for the smallest set of nodes from which the entire graph would be reachable.
-The idea is to look for ``roots'' in the graph that have to be included in the ultimate verification set because they have no predecessor an cannot be verified ``through'' the verification of some other pair.
+To summarize this algorithm conceptually, we start by constructing a graph with a node for each possible (source, sink) pair in the graph: each node then represents a possible choice for parallel pair endpoint pairs.
+Edges are drawn from node ($P_1$, $P_2$) to ($Q_1$, $Q_2$) if $Q_1$ is a successor of $P_1$ and $P_2$ is a successor of $Q_2$.
+We greedily search for the smallest set of nodes from which the entire graph would be reachable.
+The idea is to look for ``roots'' in the graph that have to be included in the ultimate verification set because they have no predecessor in the graph and cannot be verified ``through'' the verification of some other pair.
 Then all the successors whose verification is implied by the roots are eliminated.
@@
 
-The result of the algorithm cannot be further reduced.
-\xxx[as]{Do we need an argument for why?} \xxx[dg]{Yeah, we probably do}
+At the end of the greedy graph reduction we are left with the unique set of root nodes.
+The only way to reduce the set of parallel pairs is to apply the reduction rule of theorem ~\ref{reductionRule}, but all the ways in which the rule is applicable was already captured in the edges of the graph.
+The leftover set has no edges and no scope for further reduction.
+
 Also, the verification of the parallel pairs returned in the algorithm implies that the output of the previous algorithm must commute, and transitively with Theorem~\ref{verifyingSet} that the entire diagram must commute.
@@
 \subsection{Verification}
-Ultimately verification is performed by calling an equality oracle (that is beyond the scope of this work) for every comparable pair returned by the previous stage, as in \ref{algo_verification}\xxx[as]{This algorithm can probably be summarized in one sentence and doesn't seem that helpful} .
-
-\begin{algorithm}
-\DontPrintSemicolon
-\For{(path1, path2) $\in$ parallelPairs}{
-    \If{path1 $!$= path2}{
-        \KwRet{False}\;
-    }
-}
-\KwRet{True}\;
-\caption{Verification algorithm}\label{algo_verification}
-\end{algorithm}
+Ultimately verification is performed by calling an equality oracle (that is beyond the scope of this work) for every comparable pair returned by the previous stage.
 

### 2020-09-08 acmPaper/paper.tex

@@
 There are several examples of reasonably complicated transformation graphs that we can pick from.
-Gator includes several graphics examples as part of its examples package; for this evaluation, we looked at the \textit{phong}, \textit{reflection}, and \textit{shadow map} examples, sample images of which can be seen in Appendix~\ref{sec:gator-appendix}\xxx[as]{Either add text to appendix B or just reference a figure}.
+Gator includes graphics examples as part of its examples package; for this evaluation, we looked at the \textit{phong}, \textit{reflection}, and \textit{shadow map} examples, sample images of which can be seen in Appendix~\ref{sec:gator-appendix}\xxx[as]{Either add text to appendix B or just reference a figure}.
 
 We implemented a system for interfacing between the optimal set path checker (Algorithm 4) in the open-source implementation of Gator.
-Using this algorithm, we found a bug in the reflection model, where a matrix was not being divided correctly by its determinant \xxx[dg]{List other things I find, if anything}\xxx[ak]{I think it may be nice to add more detail for at least one bug (code, what was the program doing, what the graph/conflict was, time taken, etc.)}.
-We also inserted previous bugs to check the algorithm was working correctly, and it identified all inserted bugs\xxx[as]{This general observation is less interesting than specific bugs, and should be omitted to just use the specifics}.
+The system was tested with intentional bugs, of which it found them all, although no "real" bugs were found.
+The graphs used were of size 5 or less; for graphs of this size, the checker was able to run in real time with no noticeable loss of frames.
+Since the program is running at 60 frames per second, the checker was running at a rate faster than .01 seconds.
 

### 2020-09-08 acmPaper/paper.tex
Updates to sections 5-7

--- 2020-09-08 acmPaper/paper.tex
@@
 
+%Cruft for Gator
+\lstdefinelanguage{Gator}{
+	keywords={void, float, int, with, auto,
+		if, abs, \#ifdef, \#else, \#endif, in, out, uniform,
+		varying, attribute, return, void, space, is, coord, canon, type,
+		declare, vec, mat, as, has, frame, object, dimension, coord},
+	comment=[l]{//},
+}
+\lstset{
+	language=Gator,
+	columns=fullflexible,
+	keepspaces=true,
+	showspaces=false,
+	showstringspaces=false,
+	keywordstyle=\bfseries,
+	basicstyle=\ttfamily\small,
+	basewidth=0.53em,
+	aboveskip=0.9\medskipamount,
+	belowskip=0.6\medskipamount,
+	mathescape=true,
+	lineskip={-2pt},
+	commentstyle=\color{gray},
+	% xleftmargin=\parindent,
+}
+
+% Shorthand for inline code
 \newcommand{\code}{\lstinline[keywordstyle=]}
@@
 
-\xxx[as]{Just a brief overview here to explain to the reader why these applications are important and appropriate (and what we hope to learn from the case studies).}
-
-To demonstrate our algorithms applied to a real world situation,
-we search for inconsistencies in diagrams of geometry transformations, and in a diagram of the exchange rate between currencies.
+To demonstrate our algorithms applied to a real world situation, we search for inconsistencies in diagrams of geometry transformations, and in a diagram of the exchange rate between currencies.
+Each of these applications use commutative diagrams, and the commutative nature of each is necessary to reason about some form of correctness.
+We explore these examples with the intent of showing that the algorithms discussed apply to realistic settings and potentially identify real-world examples of incorrect behavior.
 
@@
 A key feature of Gator is \code{in} expressions, which insert code to automatically transform between two geometry types.
-For example, given a point \code{p} represented in 2-dimensional Cartesian coordinates (which has type \code{cart2}), we can transform this point into polar coordinates using the expression \underline{\code{p in polar}}.
+For example, given a point \code{p} represented in 2-dimensional Cartesian coordinates (which has type \code{cart2}), we can transform this point into polar coordinates using the expression \code{p in polar}.
 These \code{in} expressions create a structure of commutative diagrams, allowing use as introduced in Section~\ref{introduction}.
@@
 There are several examples of reasonably complicated transformation graphs that we can pick from.
-Gator includes graphics examples as part of its examples package; for this evaluation, we looked at the \textit{phong}, \textit{reflection}, and \textit{shadow map} examples, sample images of which can be seen in Appendix~\ref{sec:gator-appendix}\xxx[as]{Either add text to appendix B or just reference a figure}.
+Gator includes graphics examples as part of its examples package, all of which are in the Gator paper; for this evaluation, we looked at the \textit{phong}, \textit{reflection}, and \textit{shadow map} examples.
 
-We implemented a system for interfacing between the optimal set path checker (Algorithm 4) in the open-source implementation of Gator.
-The system was tested with intentional bugs, of which it found them all, although no "real" bugs were found.
+We implemented a system for interfacing between the optimal set path checker (\ref{algo_online_minimal}) in the open-source implementation of Gator.
+The system was tested with intentional bugs, of which it found them all, although no ``real'' bugs were found.
 The graphs used were of size 5 or less; for graphs of this size, the checker was able to run in real time with no noticeable loss of frames.
@@
 To ensure that it indeed commuted, we started with an empty diagram, and added in edges one by one.
-Before the addition of each edge, we used the algorithms (\ref{algo_online_polynomial} and ~\ref{algo_online_minimal}) \xxx[as]{remind the reader what these are: our algorithms?
-baselines?
-batch?
-incremental?
-etc.} to ensure the addition of a new edge did not introduce inconsistencies in the existing diagram.
+Before the addition of each edge, we used the algorithms (\ref{algo_online_polynomial} and ~\ref{algo_online_minimal}, the online polynomial and online minimal set algorithms, respectively) to ensure the addition of a new edge did not introduce inconsistencies in the existing diagram.
 If a new edge was problematic, the algorithms returned an example inconsistent pair that would arise from the addition of the edge.
@@
 Averaging over evaluation for the first 30 days of 2020, building and verifying a diagram to completion (inclusive of the time required by network calls) took 243$\pm$19 seconds using \ref{algo_online_minimal}, and in 133$\pm$13 seconds with \ref{algo_online_polynomial}.
-\xxx[as]{What is the conclusion the reader should take away?}
+For this large of a graph and data set, these times are reasonable and show these algorithms can be used in a realistic setting.
+Finding actual inconsistencies further shows the value of using these algorithms and commutative diagrams in the real world.
 
@@
 (5) \ref{algo_online_minimal}, the minimal set finding algorithm.
-\xxx[as]{Again, the reader probably needs a reminder about what the latter ones are.}
 The two metrics we evaluate are time for response and size of response set (smaller sets---tighter output results---would mean less calls to the oracle).
@@
 The average time taken by each algorithm over the course of 10 runs over randomly generated graphs with 9 nodes and 32 edges is listed in Table~\ref{tab:times}.
-\xxx[as]{Perhaps the previous paragraph summarizing the findings from this table should be moved here?}
 
@@
 
-\xxx[as]{In general, I recommend the syntax ``Figure X shows the time results.
-They are interesting.'' over ``The time results indicate something interesting, as Figure X shows.'' Putting the reference up front helps the reader immediately see what figure they should be cross-referencing.
-(Occurs in both paragraphs below.)}
-
 Figure~\ref{fig:timeVsSize} shows that the algorithms time scales with size, as expected.
 \xxx[as]{The plots are weirdly tiny on my machine... any chance your plotting tool can output PDF images?}
-Both \ref{algo_online_minimal} and \ref{algo_minimal_spanning_set} exhibit graphs that are polynomial in appearance.
-\xxx[as]{\ref{algo_minimal_spanning_set} does not appear in the figure.}
+Both \ref{algo_online_minimal} and \ref{algo_online_polynomial} exhibit graphs that are polynomial in appearance.
 The na\"{i}ve baseline as well as the two flip tolerant baseline display quick growth.
@@
 Table ~\ref{tab:sizes}, summarizes the number of output pairs that the algorithms returned on average over 10 runs, for graphs with 9 nodes and 32 edges.
-\xxx[as]{Conclusions. Are they as expected? If so, it's OK to just say why they were expected.}
+These results are essentially as expected, although it is interesting to note that Algorithm 3 produces around triple the number of pairs compared to the batch algorithm.
+Also note that Algorithm 4 produces the minimal number of paths, showing why it is the minimal set algorithm.
 
@@
 deterministic automatic type conversions and can catch inconsistencies of definition in a program with user defined conversions.
-In this paper, we have presented verification algorithms that given a commuting diagram, a new edge, efficiently compute the set of paths that would equal to each other if and only if the diagram would still commute after addition of the edge.
+In this paper, we have presented verification algorithms that efficiently compute the set of paths that would equal to each other if and only if the diagram would still commute after addition of a new edge.
 
 Integrating conversion consistency checks into widely used languages such as Scala could provide a lot of value to the program.
-Beyond implementation work, this requires an answer to the question of how to write function equality oracles.\xxx[as]{Could we say a sentence or two about why this would be useful?
-Also maybe use Scala as an example of a broader languages with this property}
+Since Scala and several other languages provide automatic conversions between types, it seems important to ensure that the choice of which path to take (an apparently arbitrary choice) does not effect the behavior of the program.
+Having an algorithm to ensure the commutativity of the resulting diagrams can ensure that behavior is correct and help prevent semantic confusion or errors when using such features.
+More engineering work still remains to implement this feature in a language such as Scala, but this paper provides the algorithms necessary to explore a solution.
 
@@
 
-\section{Example Gator Images}
-\label{sec:gator-appendix}
-\begin{figure}
-	\begin{subfigure}{\linewidth}
-		\includegraphics[width=\linewidth]{phong.png}
-		\caption{Phong Lighting Model}
-	\end{subfigure}
-
-	\begin{subfigure}{\linewidth}
-		\includegraphics[width=\linewidth]{reflection.png}
-		\caption{Reflection Model}
-	\end{subfigure}
-
-	\begin{subfigure}{\linewidth}
-		\includegraphics[width=\linewidth]{shadowmap.png}
-		\caption{Shadow Map Model}
-	\end{subfigure}
-
-	% \begin{subfigure}{\linewidth}
-	%   \includegraphics[width=\linewidth]{variance_10_TwoFlip.png}
-	%   \caption{Two flip tolerant baseline}
-	%   \label{fig:sfigTwoFlipSpread}
-	% \end{subfigure}
-
-	\caption{Example Images from Gator Programs}
-	\label{fig:gator-images}
-\end{figure}
-
 \end{document}

### 2020-09-09 acmPaper/paper.tex
Updates to sections 1-3

@@
 Figure~\ref{fig:typeExample} illustrates an example in a simple language with units-of-measure types.
-\xxx[as]{Refer here to some citations of real type systems for units of measure.}
+Such a type system is similar to existing type systems, such as that implemented in F\#~\cite{unitsOfMeasure} and a variety of C-like languages.
 In such a system, an important correctness criterion is that the diagram \emph{commutes:} when traversing the graph from any start node to any end node, applying every transformation along the path to any input value, the result is the same output value \emph{independent of the path chosen between the two nodes}.
 With our type coercion example, it is a problem if casting to a supposedly equivalent type as an intermediate step resulted in a different answer than a direct cast.
-\xxx[as]{Next, let's make the example specific: say that the expression ``(wugs) a'', for example, can produce value X or value Y, depending on whether the compiler uses path A or path B to do the conversion.}
+Specifically, given a variable \code{x} of type \code{meters}, applying the cast \code{(wugs) x} can be done in two ways: either \code{(wugs)(feet) x} or \code{(wugs)(miles) x}.
+Which path is taken depends on the compiler; we would like the choice of paths to be semantically equivalent so the compiler is free to make a choice.
 
@@
 Efficient commutativity checking is not trivial.
-\xxx[as]{Let's flip these next two thoughts, because the first thing you have to deal with is resolving cycles---then you can worry about an efficient algorithm for acyclic graphs.}
-Na\"{i}vely checking if all path pairs that begin and end at the same node in a given diagram commute could require a number of function equality checks that grow as factorial in the number of nodes, because a path consists of an ordering of nodes.
-Further, the presence of cycles implies a potentially infinite number of paths.
+The presence of cycles implies a potentially infinite number of paths..
+Further, na\"{i}vely checking if all path pairs that begin and end at the same node in a given diagram commute could require a number of function equality checks that grow as factorial in the number of nodes, because a path consists of an ordering of nodes.
 Previous work ~\cite{commutative} has identified an $O(|E|^2|V|^4)$ algorithm to verify that a complete acyclic diagram commutes; however, it addresses neither online addition nor cyclic diagrams.
 
-\xxx[as]{Let's stick to present tense, even if it sounds a little weird.
-And we can state up front what the paper accomplishes.
-Something like this: ``This paper presents an algorithm for online checking of diagram commutativity and applies it to verifying systems of value conversions.
-We identify two key insights:''}
-In the course of efficiently verifying commutativity over online addition, we identified two key insights.
-The first was that in a commutative diagram, when a new edge is to be added, at most one path needs to be checked against \xxx[as]{can we rephrase this passive voice, which makes the sentence somewhat tricky to process?} for a given source and sink pair.
+In the course of efficiently verifying commutativity over online addition, we identify two key insights.
+First, for a given source and sink pair, at most one path need be checked against in a commutative diagram when a new edge is added.
 Because the diagram commutes, all the paths between a given source and sink are equal and a representative to check against can arbitrarily be chosen.
-It lead to an $O(|V|^2(|E|+|V|))$ algorithm to verify a diagram remains commutative over the course of online addition, assuming an oracle to check the equality of functions.
-The algorithm makes an asymptotically optimal number of calls to the oracle.
+This leads to an $O(|V|^2(|E|+|V|))$ algorithm to verify a diagram remains commutative over the course of online addition, assuming an oracle to check the equality of functions.
+The algorithm makes an \emph{asymptotically} optimal number of calls to the oracle.
 
-\xxx[as]{This paragraph is a little confusing to me.
-First, what is ``informativeness''?
-Maybe we can be a little more specific.
-Second, I thought the previous paragraph said we already had an optimal number of calls to the oracle, but the second half of this paragraphs says that we need an even more expensive algorithm to obtain that minimum.}
-The second insight was that there is a single rule that places a partial, transitive ordering on the informativeness of paths.
-It allowed for the creation of a greedy, $O(|V|^4)$ optimization step that results in the number of calls to the oracle being minimal.
+The second insight is  that there is a single rule that places a partial, transitive ordering on the amount of information about the paths in the graph contained in each path.
+It allowed for the creation of a greedy, $O(|V|^4)$ optimization step that results in the number of calls to the oracle being exactly minimal.
 The optimization is very useful when the equality checking oracle is expensive.
@@
 We evaluate our algorithms against random graphs and use them in two case studies.
-\xxx[as]{Let's avoid passive voice in these next two sentences by saying ``we use the algorithm to do this'' instead of ``the algorithm is used to do this.''}
-First, our algorithm is used in the domain specific geometry type language \textit{Gator} ~\cite{gator} to ensure that user defined transformations between spaces stay consistent.
-Second, our algorithm is used to identify inefficiencies in a currency conversion graph.
+First, we use our algorithm in the domain specific geometry type language \textit{Gator} ~\cite{gator} to ensure that user defined transformations between spaces stay consistent.
+Second, we use our algorithm to identify inefficiencies in a currency conversion graph.
 
@@
 
-A pair of paths $p$ and $q$ is called \textit{parallel} iff their terminal nodes are the same, i.e., $\partial(p)=\partial(q)$.
+A pair of paths $p_1$ and $p_2$ is said to \textit{parallel} iff their terminal nodes are the same, i.e., $\partial(p_1)=\partial(p_2)$.
 $\partial$, $\partial^{+}$ and $\partial^{-}$ are extended to apply to parallel pairs.
-\xxx[as]{Let's use a name other than $p$, which we have already used for paths, for pairs.
-(Or use something else for paths above.) Also, since we just used $p$ and $q$ for two paths, maybe that would be better than $p_1$ and $p_2$.}
-For parallel pair $p$ = ($p_1$, $p_2$), $\partial(p)=\partial(p_1)=\partial(p_2)=(\partial(p)^{+}, \partial(p)^{-})$.
+For parallel pair $\phi$ = ($p_1$, $p_2$), $\partial(\phi)=\partial(p_1)=\partial(p_2)=(\partial(\phi)^{+}, \partial(\phi)^{-})$.
 
@@
 The diagram commutes iff
-$\forall (p,q)\in R_{all}$,
-$f(p)=f(q)$; that is, the composition of maps along any path connecting any pair $u$ to $v$ is independent of path choice.
+$\forall (p_1,p_2)\in R_{all}$,
+$f(p_1)=f(p_2)$; that is, the composition of maps along any path connecting any pair $u$ to $v$ is independent of path choice.
 
@@
 %
-We therefore reduce the \textsc{Online addition problem} to the \textsc{Verification set problem}.
-\xxx[as]{Although it's not technically a ``reduction,'' is it? Maybe better to say that the latter is what we solve, and we assume an oracle that uses those results to solve the former.}
-The latter, when given a diagram and a new edge, returns the set of parallel pairs of paths, such that if and only if the members in each pair are equal, then the new graph must commute.
-\xxx[as]{Let's be precise around the ``equal'' terminology in the previous and next sentences.
-It's not that the pairs are equal---it's that we check function equivalence for the pairs' corresponding functions.}
-The output to the \textsc{Online addition problem} can then be obtained as whether equality checking for all pairs succeeds.
+We therefore collapse the \textsc{Online addition problem} to the \textsc{Verification set problem}; we solve the latter and assume an oracle with the results to produce the former.
+The latter, when given a diagram and a new edge, returns the set of parallel pairs of paths, such that if and only if the members in each pair have function equivalence, then the new graph must commute.
+The output to the \textsc{Online addition problem} can then be obtained as whether function equivalence checking for all pairs succeeds.
 
-The algorithms in the rest of the paper assume that the function equivalence oracle is reflexive, symmetric, and transitive.
-\xxx[as]{The following is a little tricky to understand here... maybe this is best left to the currency section?}
-For example, the oracle for currency graph (Section~\ref{sec:currency}) only approximately preserves transitivity because of floating point error magnification, so while the algorithm catches mistakes, it does not guarantee that the diagram commutes.
+The formal algorithms in the rest of the paper assume that the function equivalence oracle is reflexive, symmetric, and transitive.
 
@@
 \subsection{Na\"{i}ve Baseline Algorithm}
+Our first goal is to develop a baseline (exponential) algorithm that can reason about cycles without producing an infinite set of paths.
+This algorithm will first pare the structure of the graph down to remove cycles, extract the pairs of paths in the graph, and finally reason about each pair to check commutativity.
+This algorithm produces the two components $C$ and $Q$: the cycle verification pairs and acyclic parallel pairs, respectively.
 
-\xxx[as]{Start with an overview of the steps.
-Something like: deal with cycles, then deal with the remaining paths in an acyclic graph.
-Maybe also give overall context to say that this algorithm is going to be exponential but at least it will be finite, unlike trying to check ``all paths.'' Then break up the discussion so there is one paragraph per step instead of one long paragraph.}
+We start with the set of all parallel pairs in the diagram.
+We pare it down to be finite by handling cycles: using a procedure like Johnson's algorithm~\cite{johnson}, we find all simple cycles in the diagram.
+Specifically, we create a cycle verification set, $C$, and verify for each cycle that a single traversal is equal to the identity function by adding $(v \rightarrow v, 1)$ for each node v in the cycle to $C$.
+Here, $v \rightarrow v$ is a simple cycle starting and ending at $v$, 1 is the identity function, and these must be verified to be equal to each other.
 
-We start with the set of all parallel pairs in the diagram (with the new edge added in) \xxx[as]{clarify: is this a batch or incremental algorithm?
-If it's a batch algorithm, then there is no need to discuss a ``new edge''}, and pare it down to be finite by handling cycles.
-Then, using a procedure like Johnson's algorithm~\cite{johnson}, we find all simple cycles in the diagram.
-\xxx[as]{I'm a tad confused here because this sentence says ``then...'' as if we had moved on from the previous step, but the previous step says it handled cycles already?}
-We create a cycle verification set, $C$, and verify for each cycle that a single traversal is equal to the identity function by adding $(v \rightarrow v, 1)$ for each node v in the cycle to $C$.
-\xxx[as]{It seems important to have already explained the ``type'' of $C$ (i.e., the output format of the algorithm?}
-Here, $v \rightarrow v$ is a simple cycle starting and ending at $v$, 1 is the identity function, and these must be verified to be equal to each other.
 We then create a set $P$ of all the paths in the diagram with no cycles, and filter the set $P \times P$, excluding pairs where the paths begin or end on different nodes, or are identical, to get the set of all cycle-free parallel pairs $Q$.
-The output of the algorithm is $C \cup Q$, the cycle verification pairs and acyclic parallel pairs.
-After verifying $C$, it is sufficient to verify only $Q$ (Lemma~\ref{one_occurence_lemma}) because cycles must now be the identity, so for any pair in the set of all parallel paths, any instance of a cycle can be removed to obtain an equivalent pair with shorter, cycle-free paths.
-\xxx[as]{What does it mean to ``verify only $Q$''? This could be a bit more precise...}
+After verifying $C$, it is sufficient to verify only $Q$ (as opposed to all pairs, per Lemma~\ref{one_occurence_lemma}) because cycles must now be the identity, so for any pair in the set of all parallel paths, any instance of a cycle can be removed to obtain an equivalent pair with shorter, cycle-free paths.
+
 If the shorter pair has equal paths then the paths in the original pair must also be equal to each other.
@@
 
-Our second baseline refines the output of the na\"{i}ve baseline.
-\xxx[as]{But even before saying that, let's give some overall context: remind the reader what the incremental setup looks like.
-Then we can explain that we are going to use something like the previous section.
-Perhaps we can also be a little more precise than ``refining'' the output, which could mean many things.}
-Like before, it start by creating a cycle verification set $C'$, but includes only the simple cycles that pass through the new edge.
+For an incremental algorithm, we explore how the addition of an edge can change the baseline to looking at a subset of the graph rather than every pair of paths.
+In achieving this, this second baseline essentially refines the results of the na\"{i}ve; a similar structure, but with a substantially reduced set of paths to examine.
+
+Like before, we start by creating a cycle verification set $C'$, but includes only the simple cycles that pass through the new edge.
 Then, instead of $Q$, the set of all non-cyclic parallel pairs, the algorithm obtains its subset $Q'$ consisting of all non-cyclic parallel pairs such that exactly one path in each pair passes through the new edge.
-To this end, the algorithm performs a \textit{two-flip tolerant path search} whose output is passed into a \textit{path extraction algorithm}.
-\xxx[as]{Maybe we can give a one-phrase explanation of what TFTPS does: it finds all the parallel pairs of which only one path includes the new edge, right?}
+To this end, the algorithm performs a \textit{two-flip tolerant path search} whose output is passed into a \textit{path extraction algorithm}; this search finds all parallel pairs for which only one path includes the new edge.
 The result of the path extraction algorithm to get the final output $Q' \cup C'$.
@@
 
-\xxx[as]{Maybe a quick paragraph here to remind the reader what is still coming in the rest of this section?}
+To resolve this algorithm fully, we will need to define the specifics of the two-flip tolerant path search and how to narrow down the results of this search into an actual set of paths to verify.
 
@@
 We use a ``two-flip tolerant'' path search from the source ($S$) to the sink ($T$) of the new edge to identify the pairs of paths where exactly one path includes the new edge.
-The idea is that every such pair corresponds to a two flip tolerant path from $S$ to $T$.
-\xxx[as]{However, ``two flip tolerant path'' has not yet been defined, so this remark sounds a little strange.}
 
@@
 The new edge is represented with a dashed arrow.
-Here, $f_1 \circ f_2 \circ f_3$ is a two flip path, and $f_1 \circ (S, T) \circ f_2$ \xxx[as]{should that be $f_2$ at the end?} is a new path created because of the addition of ($S$,$T$) that forms a parallel pair with ($f_2$).
+Here, $f_1 \circ f_2 \circ f_3$ is a two flip path, and $f_1 \circ (S, T) \circ f_3$ is a new path created because of the addition of ($S$,$T$) that forms a parallel pair with $f_2$.
 
@@
 The algorithm processes every two flip tolerant path $p$ in $P$ case-wise to obtain pairs to add to the output set.
-\xxx[as]{Bullet points (itemize) are often a good way to lay out cases?}
-In the case where p has two flips, $t_1(p) \circ (S, T) \circ t_3(p)$ and $t_2(p)$ form a parallel pair.
-When p has only the first flip (which is to say, the third phase of the path is missing), the parallel paths are $t_1(p) \circ (S, T)$ and $t_2(p)$.
-Similarly when only the second flipping point is present (so that there is no first phase), then the parallel pair is  $(S, T) \circ t_3(p)$ and $t_2(p)$.
-Finally when no flipping points are present, there are two possibilities.
+\begin{itemize}
+\item In the case where p has two flips, $t_1(p) \circ (S, T) \circ t_3(p)$ and $t_2(p)$ form a parallel pair.
+\item When p has only the first flip (which is to say, the third phase of the path is missing), the parallel paths are $t_1(p) \circ (S, T)$ and $t_2(p)$.
+\item Similarly when only the second flipping point is present (so that there is no first phase), then the parallel pair is  $(S, T) \circ t_3(p)$ and $t_2(p)$.
+\item Finally when no flipping points are present, there are two possibilities:
 Either $p$ is a path from $S$ to $T$, in which case the parallel paths are simply the edge $(S, T)$ and $p$, or $p$ is a path from $T$ to $S$.
@@
 Like with the na\"{i}ve algorithm, for every node $v$ in the cycle, we add the pair $(v \rightarrow v, 1)$, where $v \rightarrow v$ is the cycle $p \circ (S,T)$ written to start and end at $v$.
+\end{itemize}
 
-\xxx[as]{A little signposting here might be helpful... maybe a new {\textbackslash}paragraph and a brief reminder to the reader about why we need a theorem to tie things up?}
+\paragraph{Resolving the Incremental Algorithm}
+We conclude our discussion of this incremental algorithm by proving that the result is the same as if we were running the na\"{i}ve baseline algorithm.
+This in turn shows that we have found a more efficient algorithm to achieve the same result of providing a set of paths which can be used to check commutativity.
 
@@
 
-For bilinking $s$, let $F(s)$ be the vector in $\text{GF}(2)^{|E|}$ representing the edges present in s (the $n^{\text{th}}$ dimension of $F(s)$ is 1 if the corresponding edge is in $s$, and 0 otherwise).
-\xxx[AK]{GF(2) or $\mathbb{F}_2$ is standard notation for the Galois field of two elements. Do I need to write more about it here?}
-\xxx[as]{Probably just to briefly tell the reader what it is, not to fully explain it.}
+For bilinking $s$, let $F(s)$ be the vector in $\text{GF}(2)^{|E|}$ (where $\text{GF}(2)$ is the Galois field, that is, finite field of two elements) representing the edges present in s (the $n^{\text{th}}$ dimension of $F(s)$ is 1 if the corresponding edge is in $s$, and 0 otherwise).
 Let this function be extended to sets, so that for some set of bilinkings $S, F(S) = \{ F(s) | s\in S \}$. A notion of linear independence in this vector field exists.
@@
 %
-$\sigma$ is used \xxx[as]{passive voice, and a somewhat awkward start to the sentence} to iteratively check if a given pair is redundant.
-Bilinkings are eliminated until a minimum ``spanning'' subset is reached.
+We use $\sigma$ to iteratively check if a given pair is redundant.
+We eliminate Bilinkings until a minimum ``spanning'' subset is reached.
 

### 2020-09-09 acmPaper/paper.tex
typo

@@
 Frink ~\cite{frink} is a language that supports automatic conversion between units and infinite precision floating point numbers.
-It does not appear to support the implicit definition of conversion between unites \xxx[AK]{Need to spend some time confirming this} but if extended to do so, would need to contend with the problem of commuting graphs.
+It does not appear to support the implicit definition of conversion between units \xxx[AK]{Need to spend some time confirming this} but if extended to do so, would need to contend with the problem of commuting graphs.
 The same is true for F\# which has support for units of measure ~\cite{unitsOfMeasure} and Ada's GNAT compiler~\cite{gnat}.

### 2020-09-09 acmPaper/paper.tex
Updates to section 4

--- 2020-09-09 acmPaper/paper.tex
@@
 
-A path is a sequence of edges.
-The edge-to-function mapping $f$ can be naturally extended to paths: if path $p=e_1\circ e_2 \circ \cdots \circ e_n$ then $f(p)=f(e_1) \circ f(e_2) \circ \cdots \circ f(e_n)$.
+A path is a sequence of edges. The edge-to-function mapping $f$ can be naturally extended to paths: if path $p=e_1 ; e_2 ; \cdots ; e_n$ then $f(p)=f(e_1) ; f(e_2) ; \cdots ; f(e_n)$.
 %
@@
 The new edge is represented with a dashed arrow.
-Here, $f_1 \circ f_2 \circ f_3$ is a two flip path, and $f_1 \circ (S, T) \circ f_3$ is a new path created because of the addition of ($S$,$T$) that forms a parallel pair with $f_2$.
+Here, $f_1 ; f_2 ; f_3$ is a two flip path, and $f_1 ; (S, T) ; f_3$ is a new path created because of the addition of ($S$,$T$) that forms a parallel pair with $f_2$.
 
@@
 \begin{itemize}
-\item In the case where p has two flips, $t_1(p) \circ (S, T) \circ t_3(p)$ and $t_2(p)$ form a parallel pair.
-\item When p has only the first flip (which is to say, the third phase of the path is missing), the parallel paths are $t_1(p) \circ (S, T)$ and $t_2(p)$.
-\item Similarly when only the second flipping point is present (so that there is no first phase), then the parallel pair is  $(S, T) \circ t_3(p)$ and $t_2(p)$.
+\item In the case where p has two flips, $t_1(p) ; (S, T) ; t_3(p)$ and $t_2(p)$ form a parallel pair.
+\item When p has only the first flip (which is to say, the third phase of the path is missing), the parallel paths are $t_1(p) ; (S, T)$ and $t_2(p)$.
+\item Similarly when only the second flipping point is present (so that there is no first phase), then the parallel pair is  $(S, T) ; t_3(p)$ and $t_2(p)$.
 \item Finally when no flipping points are present, there are two possibilities:
 Either $p$ is a path from $S$ to $T$, in which case the parallel paths are simply the edge $(S, T)$ and $p$, or $p$ is a path from $T$ to $S$.
-In this case, we have found a cycle, $p \circ (S,T)$, to be paired with the identity function.
-Like with the na\"{i}ve algorithm, for every node $v$ in the cycle, we add the pair $(v \rightarrow v, 1)$, where $v \rightarrow v$ is the cycle $p \circ (S,T)$ written to start and end at $v$.
+In this case, we have found a cycle, $p ; (S,T)$, to be paired with the identity function.
+Like with the na\"{i}ve algorithm, for every node $v$ in the cycle, we add the pair $(v \rightarrow v, 1)$, where $v \rightarrow v$ is the cycle $p ; (S,T)$ written to start and end at $v$.
 \end{itemize}
@@
 
-    It is also clear that every cycle in $C'$ can be expressed as $(S, T) \circ p$, and corresponds to the input two flip tolerant path p.
+    It is also clear that every cycle in $C'$ can be expressed as $(S, T) ; p$, and corresponds to the input two flip tolerant path p.
 
@@
 
-\xxx[dg]{Probably put this in the appendix}
 \begin{algorithm}
@@
 If \ref{algo_online_polynomial} were applied to the diagram shown in Figure~\ref{figure_reduction_rule} there would be redundancies in the output.
-It turns out that verifying $g_2=f_2 \circ n \circ h_2$ is sufficient to ensure the diagram still commutes on the addition of $n$.
+It turns out that verifying $g_2=f_2 ; n ; h_2$ is sufficient to ensure the diagram still commutes on the addition of $n$.
 \xxx[as]{This needs a bit more explanation. What is the reader looking for? Also, the caption of this figure could probably be a little clearer.}
@@
 Nodes are also their own successors and predecessors.
-\xxx[as]{At some point, we have switched to capital letters for both paths and node variables... maybe make this consistent?} \xxx[dg]{I think these are all nodes, so it should be good}
 The rule effectively places an ordering on the informativeness of path pairs based on their terminal nodes.
@@
 
-\xxx[as]{I couldn't quite follow the following, perhaps because the figure is not quite enough for me to understand what the ``rule'' is here.
-Which makes it hard to see why ``operations'' are relevant in this context.}
  Given that a set of path pairs are equal, suppose that one attempts to derive the proposition that a different parallel pair of paths is equal with a step-by-step application of inference rules.
  Under the assumption that edges are generic functions, and no other information is available, $F$ is a semi-group.
- The only inference rules allowed are composition (given that $f_1=f_2$, it must be that $g \circ f_1 = g \circ f_2$) and replacement of one path by a different, equal path (given $f_1=f_2$ and $g \circ f_1 = h \circ f_1 $, it must be true that $g \circ f_1 = h \circ f_2$).
+ The only inference rules allowed are composition (given that $f_1=f_2$, it must be that $g ; f_1 = g ; f_2$) and replacement of one path by a different, equal path (given $f_1=f_2$ and $g ; f_1 = h ; f_1 $, it must be true that $g ; f_1 = h ; f_2$).
 Any permutation of the repeated application of these two rules results in the ``reduction rule'' already described; it is therefore the only rule that can be used to reduce the set of path pairs to check.

### 2020-09-09 acmPaper/paper.tex
Added authors, removed anonymous
Removed 2 no-longer-applicable comments
Hacks to eliminate components
Move caption to caption

@@
 \newcount\draft\draft=1 % set to 0 for submission or publication
-\documentclass[sigplan,review,anonymous]{acmart}
+% \documentclass[sigplan,review,anonymous]{acmart}
+\documentclass[sigplan,review]{acmart}
 \setcopyright{none}
@@
 %
-\author{First Author}
-\email{email}
-\author{Second Author}
-\authornotemark[1]
-\email{email 2}
+\author{Aditi Kabra \quad Dietrich Geisler \quad Adrian Sampson}
 \affiliation{%
   \institution{Cornell University}
-  \streetaddress{address}
-  \city{Ithaca}
-  \state{New York}
-  \postcode{14850}
 }
-
-\author{Third Author}
-\email{email 3}
+% ACM publication cruft.
+\acmJournal{PACMPL}
+\acmVolume{1}
+\acmNumber{TAPAS}
+\acmArticle{1}
+\acmYear{2020}
+\acmMonth{1}
+\acmDOI{} % \acmDOI{10.1145/nnnnnnn.nnnnnnn}
+\startPage{1}
 
@@
 $(\partial(p)^{+}, \partial(p)^{-})$.
-\xxx[as]{Simplified this explanation; I hope it's OK.}
 
@@
 It turns out that verifying $g_2=f_2 ; n ; h_2$ is sufficient to ensure the diagram still commutes on the addition of $n$.
-\xxx[as]{This needs a bit more explanation. What is the reader looking for? Also, the caption of this figure could probably be a little clearer.}
 
@@
 \end{tikzpicture}
+\end{center}
+\caption{Reduction rule.
 Each arrow represents a path, where $n$ is the new edge being added.
-While ~\ref{algo_online_polynomial} returns two pairs for verification, one from $P_1$ to $P_2$ and the other from $Q_1$ to $Q_2$, it actually suffices to just check a pair from $Q_1$ to $Q_2$ as demonstrated in theorem ~\ref{reductionRule}.
-\end{center}
-\caption{Reduction rule.}
+While ~\ref{algo_online_polynomial} returns two pairs for verification, one from $P_1$ to $P_2$ and the other from $Q_1$ to $Q_2$, it actually suffices to just check a pair from $Q_1$ to $Q_2$ as demonstrated in theorem ~\ref{reductionRule}.}
 \label{figure_reduction_rule}

### 2020-09-09 acmPaper/paper.tex
Kill an orphan
Little fixes

@@
     Every element in the output of the path extraction algorithm was by construction an element of $O$.
-
-    It is also clear that every cycle in $C'$ can be expressed as $(S, T) ; p$, and corresponds to the input two flip tolerant path p.
+Every cycle in $C'$ can be expressed as $(S, T) ; p$, and corresponds to the input two flip tolerant path $p$.
 
@@
 
-\begin{algorithm}
+\begin{algorithm}[t]
 \label{alg_spanning_set}
@@
 
-\caption{finding spanning set}\label{algo_spanning_set}
+\caption{Finding a spanning set.}\label{algo_spanning_set}
 \end{algorithm}
@@
 
-\caption{Minimal spanning set}\label{algo_minimal_spanning_set}
+\caption{Finding a minimal spanning set.}\label{algo_minimal_spanning_set}
 \end{algorithm}
@@
 We use randomly generated graphs of varying size: given a graph and a new edge, we time how long it takes for an algorithm to return the set of pairs that need to be verified.
-All computations were performed on a Macbook Pro 2015, 2.9 GHz Dual-Core Intel Core i5.
+All computations were performed on a MacBook Pro 2015, 2.9 GHz dual-core Intel Core i5.
 
@@
 \end{tabular}
-\caption{Computation time for 9 node graph of density 0.4, averaged over ten runs}
+\caption{Computation time for 9-node graph of density 0.4, averaged over ten runs.}
 \label{tab:times} %TODO debug and fix label not working.

### 2020-09-09 acmPaper/paper.tex
Remove footnote on the first page
Space hacks on the first page
Trim the first page

@@
 % \documentclass[sigplan,review,anonymous]{acmart}
-\documentclass[sigplan,review]{acmart}
+\documentclass[sigplan,review,nonacm=true]{acmart}
 \setcopyright{none}
@@
 \label{introduction}
-Many systems can be understood as diagrams: graphs where nodes represent domains and edges represent directed transformation functions between nodes.
+Many systems use diagrams: graphs where nodes are domains and edges are transformation functions.
 A type system with coercions, for example, corresponds to a graph whose nodes are types and whose edges are coercions.
-Figure~\ref{fig:typeExample} illustrates an example in a simple language with units-of-measure types.
-Such a type system is similar to existing type systems, such as that implemented in F\#~\cite{unitsOfMeasure} and a variety of C-like languages.
+Figure~\ref{fig:typeExample} illustrates an example in a simple language with units-of-measure types~\cite{unitsOfMeasure}.
 In such a system, an important correctness criterion is that the diagram \emph{commutes:} when traversing the graph from any start node to any end node, applying every transformation along the path to any input value, the result is the same output value \emph{independent of the path chosen between the two nodes}.
-With our type coercion example, it is a problem if casting to a supposedly equivalent type as an intermediate step resulted in a different answer than a direct cast.
+With our coercion example, it is a problem if casting to a supposedly equivalent type as an intermediate step resulted in a different answer than a direct cast.
 Specifically, given a variable \code{x} of type \code{meters}, applying the cast \code{(wugs) x} can be done in two ways: either \code{(wugs)(feet) x} or \code{(wugs)(miles) x}.
@@
         \end{verbatim}
+        \vspace{-3ex}
         \caption{A sample program with user defined type conversion.}
@@
     \end{subfigure}
-
+    \vspace{-2ex}
     \caption{In this sample program, the user implicitly defines two ways to cast variable a from meters to the new unit wugs.

### 2020-09-10 acmPaper/paper.tex
One more top-algorithm
Try to center code
Intro tweaks

@@
     \begin{subfigure}{0.4\textwidth}
+        \centering
         \begin{verbatim}
@@
 
-In the course of efficiently verifying commutativity over online addition, we identify two key insights.
-First, for a given source and sink pair, at most one path need be checked against in a commutative diagram when a new edge is added.
+For verifying commutativity over online addition, we identify two key insights.
+First, when a new edge is added,
+only one path per source and sink pair needs to be checked against the existing commutative diagram.
 Because the diagram commutes, all the paths between a given source and sink are equal and a representative to check against can arbitrarily be chosen.
@@
 
-The second insight is  that there is a single rule that places a partial, transitive ordering on the amount of information about the paths in the graph contained in each path.
-It allowed for the creation of a greedy, $O(|V|^4)$ optimization step that results in the number of calls to the oracle being exactly minimal.
-The optimization is very useful when the equality checking oracle is expensive.
+The second insight is that there is a single rule that places a partial, transitive ordering on paths indicating the amount of information they contain about other paths.
+This insight yields a greedy $O(|V|^4)$ optimization step that results in the number of oracle calls being exactly minimal.
+The optimization is critical when equality checking is expensive.
 
@@
 Second, we use our algorithm to identify inefficiencies in a currency conversion graph.
-
-We empirically compare our solution to three baseline implementations: a na\"{i}ve checking of all path pairs with only special handling for cycles, a check for all path pairs that involve the new edge, and an algorithm suggested by previous work to solve the batch version of the problem for acyclic diagrams.
+%
+We empirically compare our solution to three baseline implementations: a na\"{i}ve cycle-sensitive all-pairs check, a check for all path pairs that involve the new edge, and an algorithm suggested by previous work to solve the batch version of the problem for acyclic diagrams.
 Our proposed algorithms run orders of magnitude faster than the baselines.
@@
 
-\begin{algorithm}
+\begin{algorithm}[t]
 \DontPrintSemicolon

### 2020-09-10 acmPaper/paper.tex
Minor changes

--- 2020-09-09 acmPaper/paper.tex
@@
 We then create a set $P$ of all the paths in the diagram with no cycles, and filter the set $P \times P$, excluding pairs where the paths begin or end on different nodes, or are identical, to get the set of all cycle-free parallel pairs $Q$.
-After verifying $C$, it is sufficient to verify only $Q$ (as opposed to all pairs, per Lemma~\ref{one_occurence_lemma}) because cycles must now be the identity, so for any pair in the set of all parallel paths, any instance of a cycle can be removed to obtain an equivalent pair with shorter, cycle-free paths.
+After verifying $C$, it is sufficient to verify only $Q$ (as opposed to all pairs) because cycles must now be the identity, so for any pair in the set of all parallel paths, any instance of a cycle can be removed to obtain an equivalent pair with shorter, cycle-free paths.
 
@@
 For path pair $p$, and new edge from node $S$ to node $T$, the first segment extends from $\partial(p)^+$ to the $S$, the second, the new edge $(S, T)$ itself, and the third, from $T$ to $\partial(p)^-$.
-The new edge could only appear once because cycles have already been dealt with so only pairs where the path includes the new edge once need be checked (Lemma~\ref{one_occurence_lemma}).
+The new edge could only appear once because cycles have already been dealt with so only pairs where the path includes the new edge once need be checked.
 The first segment of both pairs would have to be equal because they existed as parallel pairs in the original diagram, and similarly the third segments would also have to be equal.
@@
 
-\begin{lemma}
-\label{one_occurence_lemma}
-A path that involves the new edge more than once is equal to a path without multiple occurrences of the edge, under the assumption that the pairs in C are verified.
-\end{lemma}
-\begin{proof}
-Consider the part of the path between the first occurrence of the new edge and the second. This forms a cycle, which has a corresponding check in $C$ and must be verified to be equal to the identity.
-This means that the loops containing the multiple occurrences of the new edge may be ignored and the path with a single occurrence of the new edge with the cycles removed is equal to this path.
-\end{proof}
-
 \begin{theorem}

### 2020-09-10 acmPaper/paper.tex
Little tweaks in 4

@@
 
-We present a polynomial time solution to the \textsc{verification set problem} .
+We present a polynomial time solution to the \textsc{verification set problem}.
 As in the online baseline algorithm, we do not concern ourselves with parallel pairs where neither or both paths pass through the new edge.
-The key observation allowing us to improve on the online baseline is a result of Theorem~\ref{reductionRule} (which will be expanded on later): for a given source and sink pair, only a single parallel pair needs to be verified.
+The key observation allowing us to improve on the online baseline is a result of Theorem~\ref{reductionRule} (which we expand on later): for a given source and sink pair, only a single parallel pair needs to be verified.
 Theorem~\ref{verifyingSet} shows that, should our selected set of pairs and cycles passing through the new edge be verified commutative, the entire diagram must commute.
-Algorithm \ref{algo_online_polynomial} uses this strategy of identifying a parallel pair with exactly one path through the edge for each (source, sink) pair.
+\ref{algo_online_polynomial} uses this strategy of identifying a parallel pair with exactly one path through the edge for each (source, sink) pair.
 
-\begin{algorithm}
+\begin{algorithm}[t]
 \DontPrintSemicolon
@@
 
-The try block is executed at most $O(|V|^2)$ times, which is also the bound on the number of pairs verified.
+The \code{try} block is executed at most $O(|V|^2)$ times, which is also the bound on the number of pairs verified.
 %
 This bound is asymptotically tight, as can be seen in the case where the graph contains $2N$ nodes along $S$ and $T$.
-We imagine dividing the nodes into two groups of $N$ nodes each.
-Every node in group 1 has a forward node to every node in group 2 as well as to $S$.
+Imagine dividing the nodes into two groups of $N$ nodes each.
+Every node in group 1 has a forward edge to every node in group 2 and to $S$.
 $T$ has a forward edge to every node in group 2.
-In this diagram, on the addition of edge $(S, T)$, $N^2$ paths need to be verified which is polynomial in the total number of nodes, $2N+2$.
+In this diagram, when adding edge $(S, T)$, $N^2$ paths need to be verified which is polynomial in the total number of nodes, $2N+2$.
 
@@
 \[\Rightarrow l ; g_2 ; m = l ; f_2 ; n ; h_2 ; m \Rightarrow g_1 = f_1 ; n ; h_1\]
-
-The proof is not affected if any of the paths used are the identity, e.g., if $f_1$ is the identity and $S$ and $P_1$ are the same node.
+%
+The proof holds if any of the paths used are the identity, e.g., if $f_1$ is the identity so $S$ and $P_1$ are the same node.
 \end{proof}
@@
 
- Given that a set of path pairs are equal, suppose that one attempts to derive the proposition that a different parallel pair of paths is equal with a step-by-step application of inference rules.
+Given that a set of path pairs are equal, suppose we attempt to derive the proposition that a different parallel pair of paths is equal with a step-by-step application of inference rules.
  Under the assumption that edges are generic functions, and no other information is available, $F$ is a semi-group.

### 2020-09-10 acmPaper/paper.tex
Removed the appendix

--- 2020-09-10 acmPaper/paper.tex
@@
 The key observation allowing us to improve on the online baseline is a result of Theorem~\ref{reductionRule} (which will be expanded on later): for a given source and sink pair, only a single parallel pair needs to be verified.
-Theorem~\ref{verifyingSet} shows that, should our selected set of pairs and cycles passing through the new edge be verified commutative, the entire diagram must commute.
+It is straightforward to see that, should our selected set of pairs and cycles passing through the new edge be verified commutative, the entire diagram must commute.
 Algorithm \ref{algo_online_polynomial} uses this strategy of identifying a parallel pair with exactly one path through the edge for each (source, sink) pair.
@@
 
-Also, the verification of the parallel pairs returned in the algorithm implies that the output of the previous algorithm must commute, and transitively with Theorem~\ref{verifyingSet} that the entire diagram must commute.
+Also, the verification of the parallel pairs returned in the algorithm implies that the output of the previous algorithm must commute and that the entire diagram must commute.
 
@@
 
-\appendix
-
-\section{Theorems and Additional Cases}
-
-Consider all the cycles in the graph that pass through the new edge, unique to their terminal point.
-Define a cycle-pair corresponding to cycle c to mean the pair of paths (c, identity on terminal node of c).
-Let $C$ be the set of all cycle-pairs (unique to their terminal point) corresponding to cycles that pass through the new edge.
-
-Let $V$, the set of pairs to verify, be $C \cup R_0$.
-
-\begin{theorem}
-\label{verifyingSet}
-Verifying that all pairs in V commute implies that all pairs in $R_{all}$ commute.
-\end{theorem}
-\begin{proof}
-
-Each pair in $R_{all}$ that is not in V would fall into one of these categories:
-
-Case A: pairs that do not involve the new edge: By assumption that the original diagram commutes, these pairs must commute.
-
-Case B: Cases where there is a different pair corresponding to this choice of (source, sink) in V already.
-There are four paths to be considered, two from each pair. Those paths which do not include the new edge must already be equal by the assumption that the original diagram commutes.
-Those paths which do include the new edge (S,T) can be reduced, by the case for multiple occurrences of a new edge, to a path with a single occurrence of the new edge as described in the preceding lemma \ref{one_occurence_lemma}.
-These new paths can be divided into three segments: the segment from the source to S, the new edge, and the segment from the last occurrence of T to the sink.
-The first segments of all paths are equal by assumption that the original diagram commutes.
-This is also true of the third segment of all the paths.
-The second segment consists of only the new edge and is the same for all paths.
-The composition of equal functions is equal, so all the paths passing through the new edge must be equal.
-Therefore it is sufficient to check one path that passes through the new edge and one that doesn't. Such a pair belongs to V by construction.
-\end{proof}
-
 \end{document}

### 2020-09-10 acmPaper/paper.tex

@@
 
-We present a polynomial time solution to the \textsc{verification set problem} .
+We present a polynomial time solution to the \textsc{verification set problem}.
 As in the online baseline algorithm, we do not concern ourselves with parallel pairs where neither or both paths pass through the new edge.
-The key observation allowing us to improve on the online baseline is a result of Theorem~\ref{reductionRule} (which will be expanded on later): for a given source and sink pair, only a single parallel pair needs to be verified.
+
+The key observation allowing us to improve on the online baseline is a result of Theorem~\ref{reductionRule} (which we expand on later): for a given source and sink pair, only a single parallel pair needs to be verified.
 It is straightforward to see that, should our selected set of pairs and cycles passing through the new edge be verified commutative, the entire diagram must commute.
-Algorithm \ref{algo_online_polynomial} uses this strategy of identifying a parallel pair with exactly one path through the edge for each (source, sink) pair.
+\ref{algo_online_polynomial} uses this strategy of identifying a parallel pair with exactly one path through the edge for each (source, sink) pair.
 
-\begin{algorithm}
+\begin{algorithm}[t]
 \DontPrintSemicolon

### 2020-09-10 acmPaper/paper.tex
small change

--- 2020-09-10 acmPaper/paper.tex
@@
 We then create a set $P$ of all the paths in the diagram with no cycles, and filter the set $P \times P$, excluding pairs where the paths begin or end on different nodes, or are identical, to get the set of all cycle-free parallel pairs $Q$.
-After verifying $C$, it is sufficient to verify only $Q$ (as opposed to all pairs) because cycles must now be the identity, so for any pair in the set of all parallel paths, any instance of a cycle can be removed to obtain an equivalent pair with shorter, cycle-free paths.
+After verifying $C$, it is sufficient to verify only $Q$ (as opposed to all pairs, per Lemma~\ref{one_occurence_lemma}) because cycles must now be the identity, so for any pair in the set of all parallel paths, any instance of a cycle can be removed to obtain an equivalent pair with shorter, cycle-free paths.
 
@@
 For path pair $p$, and new edge from node $S$ to node $T$, the first segment extends from $\partial(p)^+$ to the $S$, the second, the new edge $(S, T)$ itself, and the third, from $T$ to $\partial(p)^-$.
-The new edge could only appear once because cycles have already been dealt with so only pairs where the path includes the new edge once need be checked.
+The new edge could only appear once because cycles have already been dealt with so only pairs where the path includes the new edge once need be checked (Lemma~\ref{one_occurence_lemma}).
 The first segment of both pairs would have to be equal because they existed as parallel pairs in the original diagram, and similarly the third segments would also have to be equal.
@@
 
-We present a polynomial time solution to the \textsc{verification set problem}.
+We present a polynomial time solution to the \textsc{verification set problem} .
 As in the online baseline algorithm, we do not concern ourselves with parallel pairs where neither or both paths pass through the new edge.
-The key observation allowing us to improve on the online baseline is a result of Theorem~\ref{reductionRule} (which we expand on later): for a given source and sink pair, only a single parallel pair needs to be verified.
+The key observation allowing us to improve on the online baseline is a result of Theorem~\ref{reductionRule} (which will be expanded on later): for a given source and sink pair, only a single parallel pair needs to be verified.
 Theorem~\ref{verifyingSet} shows that, should our selected set of pairs and cycles passing through the new edge be verified commutative, the entire diagram must commute.
-\ref{algo_online_polynomial} uses this strategy of identifying a parallel pair with exactly one path through the edge for each (source, sink) pair.
+Algorithm \ref{algo_online_polynomial} uses this strategy of identifying a parallel pair with exactly one path through the edge for each (source, sink) pair.
 
-\begin{algorithm}[t]
+\begin{algorithm}
 \DontPrintSemicolon
@@
 
-The \code{try} block is executed at most $O(|V|^2)$ times, which is also the bound on the number of pairs verified.
+The try block is executed at most $O(|V|^2)$ times, which is also the bound on the number of pairs verified.
 %
 This bound is asymptotically tight, as can be seen in the case where the graph contains $2N$ nodes along $S$ and $T$.
-Imagine dividing the nodes into two groups of $N$ nodes each.
-Every node in group 1 has a forward edge to every node in group 2 and to $S$.
+We imagine dividing the nodes into two groups of $N$ nodes each.
+Every node in group 1 has a forward node to every node in group 2 as well as to $S$.
 $T$ has a forward edge to every node in group 2.
-In this diagram, when adding edge $(S, T)$, $N^2$ paths need to be verified which is polynomial in the total number of nodes, $2N+2$.
+In this diagram, on the addition of edge $(S, T)$, $N^2$ paths need to be verified which is polynomial in the total number of nodes, $2N+2$.
 
@@
 \[\Rightarrow l ; g_2 ; m = l ; f_2 ; n ; h_2 ; m \Rightarrow g_1 = f_1 ; n ; h_1\]
-%
-The proof holds if any of the paths used are the identity, e.g., if $f_1$ is the identity so $S$ and $P_1$ are the same node.
+
+The proof is not affected if any of the paths used are the identity, e.g., if $f_1$ is the identity and $S$ and $P_1$ are the same node.
 \end{proof}
@@
 
-Given that a set of path pairs are equal, suppose we attempt to derive the proposition that a different parallel pair of paths is equal with a step-by-step application of inference rules.
+ Given that a set of path pairs are equal, suppose that one attempts to derive the proposition that a different parallel pair of paths is equal with a step-by-step application of inference rules.
  Under the assumption that edges are generic functions, and no other information is available, $F$ is a semi-group.
@@
 While the batch algorithm improves on this, it still does not scale very well, with computation for a graph with 14 nodes and 0.4 density taking hours.
-Our implementation does not memoize the construction of the vector and matrix representation of paths in GF(2); profiling indicates that this construction is a major factor in the high time cost for these algorithms.
+Our implementation does not memoize the construction of the vector and matrix representation of paths in GF(2); profiling indicates that this construction is a major factor in the high time cost for this algorithm.
 \ref{algo_online_polynomial} performs only slightly better than the batch algorithm.
@@
 
+\begin{lemma}
+\label{one_occurence_lemma}
+A path that involves the new edge more than once is equal to a path without multiple occurrences of the edge, under the assumption that the pairs in C are verified.
+\end{lemma}
+\begin{proof}
+Consider the part of the path between the first occurrence of the new edge and the second. This forms a cycle, which has a corresponding check in $C$ and must be verified to be equal to the identity.
+This means that the loops containing the multiple occurrences of the new edge may be ignored and the path with a single occurrence of the new edge with the cycles removed is equal to this path.
+\end{proof}
+
 \begin{theorem}

### 2020-09-10 acmPaper/paper.tex
Under the page limit

--- 2020-09-10 acmPaper/paper.tex
@@
 Figure~\ref{fig:typeExample} illustrates an example in a simple language with units-of-measure types~\cite{unitsOfMeasure}.
-In such a system, an important correctness criterion is that the diagram \emph{commutes:} when traversing the graph from any start node to any end node, applying every transformation along the path to any input value, the result is the same output value \emph{independent of the path chosen between the two nodes}.
+In such a system, an important correctness criterion is that the diagram \emph{commutes}: when traversing the graph from any start node to any end node, applying every transformation along the path to any input value, the result is the same output value \emph{independent of the path chosen between the two nodes}.
 With our coercion example, it is a problem if casting to a supposedly equivalent type as an intermediate step resulted in a different answer than a direct cast.
@@
 
-This paper is about efficiently checking diagrams that arise in real systems for commutativity.
+This paper is about efficiently checking commutativity in diagrams that arise in real systems.
 We assume a simple equivalence checker for individual transformation functions: in our type system example, for instance, it is possible to check transformation equivalence by comparing the conversion factors.
@@
 
-The formal algorithms in the rest of the paper assume that the function equivalence oracle is reflexive, symmetric, and transitive.
+The algorithms in this paper assume that the function equivalence oracle is reflexive, symmetric, and transitive.
 
@@
 This algorithm will first pare the structure of the graph down to remove cycles, extract the pairs of paths in the graph, and finally reason about each pair to check commutativity.
-This algorithm produces the two components $C$ and $Q$: the cycle verification pairs and acyclic parallel pairs, respectively.
+This results in two components $C$ and $Q$: the cycle verification pairs and acyclic parallel pairs, respectively.
 
@@
 
-The two flip tolerant path search returns the set of all paths between a given source and sink that have up to two flips (paths that omit one or more of the three phases are also accepted).
+The two flip tolerant path search returns the set of all paths between a given source and sink that have up to two flips (paths with zero or one flip are also accepted).
 
@@
 
-Murota's main result~\cite{commutative} is a solution to a batch version of \textsc{verification set}; given an acyclic diagram, it returns the minimal set of equality checks that would succeed if and only if the diagram commutes.
-The paper describes an algorithm to find the ($|V|^2|E|$ bounded) minimal set of pairs that needs to be checked.
+Murota's main result~\cite{commutative} solves the batch version of \textsc{verification set}: given an acyclic diagram, it returns the minimal set of equality checks that succeed if and only if the diagram commutes.
+Murota describes an algorithm to find the ($|V|^2|E|$ bounded) minimal set of pairs that need be checked.
 
@@
 For a set of bilinkings $R$, the closure function $cl$ is defined as:
-$cl(R) = \{ s\in R_0| s$ is linearly {dependent on $F(R) \}$.
+$cl(R) = \{ s\in R_0| s$ is linearly dependent on $F(R) \}$.
 %
-The closure function on $R$ basically captures all the pairs that can be made by made by composing or ``gluing together'' the bilinkings in $R$. 
+The closure function on $R$ captures all the pairs that can be made by made by composing or ``gluing together'' the bilinkings in $R$. 
 %
-Using these two functions, we finally define the function $\sigma$ on a set of bilinkings $R$ as
+Using these two functions, we define the function $\sigma$ on a set of bilinkings $R$ as
 $\sigma(R) = \{s \in R_0 | s\in cl(R\cap \langle s \rangle) \}$.
@@
 We use $\sigma$ to iteratively check if a given pair is redundant.
-We eliminate Bilinkings until a minimum ``spanning'' subset is reached.
+We eliminate Bilinkings until we reach a minimum ``spanning'' subset.
 

### 2020-09-10 acmPaper/paper.tex
Final grammar pass

--- 2020-09-10 acmPaper/paper.tex
@@
-\newcount\draft\draft=1 % set to 0 for submission or publication
+\newcount\draft\draft=0 % set to 0 for submission or publication
 % \documentclass[sigplan,review,anonymous]{acmart}
@@
 For verifying commutativity over online addition, we identify two key insights.
-First, when a new edge is added,
-only one path per source and sink pair needs to be checked against the existing commutative diagram.
+First, when a new edge is added, only one path per source and sink pair needs to be checked against the existing commutative diagram.
 Because the diagram commutes, all the paths between a given source and sink are equal and a representative to check against can arbitrarily be chosen.
@@
 
-The second insight is that there is a single rule that places a partial, transitive ordering on paths indicating the amount of information they contain about other paths.
+Second, there is a single rule that places a partial, transitive ordering on paths indicating the amount of information they contain about other paths.
 This insight yields a greedy $O(|V|^4)$ optimization step that results in the number of oracle calls being exactly minimal.
@@
 We empirically compare our solution to three baseline implementations: a na\"{i}ve cycle-sensitive all-pairs check, a check for all path pairs that involve the new edge, and an algorithm suggested by previous work to solve the batch version of the problem for acyclic diagrams.
-Our proposed algorithms run orders of magnitude faster than the baselines.
+Our proposed algorithms run orders of magnitude faster than the baseline implementations.
 
@@
 We start with a directed graph $G=(V,E)$, where $V$ corresponds to sets of elements and edges $(u, v)$ in $E$ correspond to functions that maps elements of $u$ to elements in $v$.
-All these functions form a semigroup $F$, where multiplication is function composition.
+These functions form a semigroup $F$, where multiplication is function composition.
 A semigroup consists of a set and an associative binary operation, which we use to capture function composition.
@@
 Checking function equality is a domain specific, potentially hard problem, dependent on the nature of the graph.
-For our case study in graphics programming (see Section~\ref{sec:gator}), edges are matrices and nodes are vector spaces, so function composition uses matrix multiplication and equivalence checking simply compares matrix values.
+For example, in our case study in graphics programming (see Section~\ref{sec:gator}), edges are matrices and nodes are vector spaces, so function composition uses matrix multiplication and equivalence checking simply compares matrix values.
 We therefore assume some oracle for checking transformation function equivalence that will vary by domain.
@@
 To examine the efficacy of our proposed solution to the \textsc{Verification set problem}, we compare it to some potential alternatives.
-Specifically, we examine a na\"{i}ve factorial algorithm, a slightly less na\"{i}ve factorial algorithm which we identify to be a two-flip tolerant path search, and Murota's historical batch solution~\cite{commutative}.
+Specifically, we examine a na\"{i}ve factorial algorithm, a slightly less na\"{i}ve factorial algorithm which we identify to be a two-flip tolerant path search, and Murota's previous batch solution~\cite{commutative}.
 
@@
 We pare it down to be finite by handling cycles: using a procedure like Johnson's algorithm~\cite{johnson}, we find all simple cycles in the diagram.
-Specifically, we create a cycle verification set, $C$, and verify for each cycle that a single traversal is equal to the identity function by adding $(v \rightarrow v, 1)$ for each node v in the cycle to $C$.
+We create a cycle verification set, $C$, and verify for each cycle that a single traversal is equal to the identity function by adding $(v \rightarrow v, 1)$ for each node v in the cycle to $C$.
 Here, $v \rightarrow v$ is a simple cycle starting and ending at $v$, 1 is the identity function, and these must be verified to be equal to each other.
@@
 To see why this is true, each path could be thought of as consisting of the composition of three segments.
-For path pair $p$, and new edge from node $S$ to node $T$, the first segment extends from $\partial(p)^+$ to the $S$, the second, the new edge $(S, T)$ itself, and the third, from $T$ to $\partial(p)^-$.
+For path pair $p$, and new edge from node $S$ to node $T$, the first segment extends from $\partial(p)^+$ to $S$, the second, the new edge $(S, T)$ itself, and the third, from $T$ to $\partial(p)^-$.
 The new edge could only appear once because cycles have already been dealt with so only pairs where the path includes the new edge once need be checked.
-The first segment of both pairs would have to be equal because they existed as parallel pairs in the original diagram, and similarly the third segments would also have to be equal.
-The second segments, consisting of the same edge, would also have to be equal because the equivalence oracle is reflexive.
-Therefore, the two paths of the pair, a composition of these three equal components, would be equal, since the oracle would preserve transitivity of equality.
+The first segment of both pairs would have to be equal because they existed as parallel pairs in the original diagram, and similarly the third segment would also have to be equal.
+The second segment, consisting of the same edge, would also have to be equal because the equivalence oracle is reflexive.
+A composition of these three equal components would be then be equal, since the oracle would preserve transitivity of equality.
 We are left only with parallel pairs where exactly one of the paths passes through the new edge.
@@
 \paragraph{Path extraction algorithm}
-Next, the \textit{path extraction algorithm} then transforms the output of the two flip path search to the verification set, $Q' \cup C'$.
+Next, the \textit{path extraction algorithm} then transforms the output of the two flip path search into the verification set, $Q' \cup C'$.
 Given a set of two flip tolerant paths from the new edge source to sink, the algorithm outputs a set of pairs to verify.
@@
 A bilinking is defined to be a parallel pair that is disjoint but for their terminal nodes. The set of all bilinkings is $R_0$.
-In an acyclic diagram, if all bilinkings are equal, all parallel pairs must also be equal since they any given pair can be expressed as a composition of bilinkings.
+In an acyclic diagram, if all bilinkings are equal, all parallel pairs must also be equal since any given pair can be expressed as a composition of bilinkings.
 
@@
 Roughly, the algorithm proceeds by first efficiently finding a \textit{spanning} set of bilinkings (a subset whose verification implies the verification of all bilinkings in the graph).
-It does this by, starting at every node, finding the reachable subsection of the graph, and a spanning tree for the subsection.
+It does this, starting at every node, by finding the reachable subsection of the graph, and a spanning tree for the subsection.
 From each edges in the reachable section that is not a part of the tree, it generates a bilinking using the edge and a path in the tree that is parallel to the edge (\ref{algo_spanning_set}).
@@
 
-\subsection{Verification}
-Ultimately verification is performed by calling an equality oracle (that is beyond the scope of this work) for every comparable pair returned by the previous stage.
-
 \section{Case Studies}
@@
 
-Gator is a domain specific language designed around Geometry types, which are used to describe properties and transformations of geometric objects~\cite{gator}.
+Gator is a domain specific language designed around geometry types, which are used to describe properties and transformations of geometric objects~\cite{gator}.
 A key feature of Gator is \code{in} expressions, which insert code to automatically transform between two geometry types.
@@
 Specifically, Gator introduces transformations between \textit{reference frames}, which are the geometry equivalent of transforming between linear algebra basis vectors.
-Each edge on our transformation graph is a thus a matrix, with composition of edges as matrix multiplication and an oracle checking matrix equality (up to a rounding error $\epsilon$).
+Each edge on our transformation graph is thus a matrix, with composition of edges as matrix multiplication and an oracle checking matrix equality (up to a rounding error $\epsilon$).
 
@@
 Consider a diagram with nodes as currencies and a directed edge being the conversion rate from its source node's currency to its sink node's currency.
-Since the transformation exchange rate of money from any given base currency to a target currency can be expected to be the same regardless of which intermediate currency transformations are used, this diagram should commute.
+Since the exchange rate of money from any given base currency to a target currency can be expected to be the same regardless of which intermediate currency transformations are used, this diagram should commute.
 
@@
 
-Figure~\ref{fig:timeVsSize} shows that the algorithms time scales with size, as expected.
+Figure~\ref{fig:timeVsSize} shows that the algorithms' time scales with size, as expected.
 \xxx[as]{The plots are weirdly tiny on my machine... any chance your plotting tool can output PDF images?}

### 2020-10-04 acmPaper/paper.tex
tweak in figure 1

@@
         \begin{verbatim}
-        var a : meters = 1;
-        var b : miles = (miles) a;
-        var c : feet = (feet) a;
+        var x : meters = 1;
+        define foot:
+            1 meter = 3.28 feet;
+        define miles:
+            1 meter = 0.000621 miles;
         define wugs:
@@
             1 foot = 10 wugs;
-        var d : wugs = (wugs) a;
+        var y : wugs = (wugs) x;
         \end{verbatim}

### 2020-10-04 acmPaper/paper.tex
changed sets to use mathcal

@@
 We start with a directed graph $G=(V,E)$, where $V$ corresponds to sets of elements and edges $(u, v)$ in $E$ correspond to functions that maps elements of $u$ to elements in $v$.
-These functions form a semigroup $F$, where multiplication is function composition.
+These functions form a semigroup $\mathcal{F}$, where multiplication is function composition.
 A semigroup consists of a set and an associative binary operation, which we use to capture function composition.
@@
 
-Let $R_{all}$ be the set of all parallel pairs of paths in a given diagram.
+Let $\mathcal{R}_{all}$ be the set of all parallel pairs of paths in a given diagram.
 The diagram commutes iff
-$\forall (p_1,p_2)\in R_{all}$,
+$\forall (p_1,p_2)\in \mathcal{R}_{all}$,
 $f(p_1)=f(p_2)$; that is, the composition of maps along any path connecting any pair $u$ to $v$ is independent of path choice.
@@
 This algorithm will first pare the structure of the graph down to remove cycles, extract the pairs of paths in the graph, and finally reason about each pair to check commutativity.
-This results in two components $C$ and $Q$: the cycle verification pairs and acyclic parallel pairs, respectively.
+This results in two components $\mathcal{C}$ and $\mathcal{Q}$: the cycle verification pairs and acyclic parallel pairs, respectively.
 
@@
 We pare it down to be finite by handling cycles: using a procedure like Johnson's algorithm~\cite{johnson}, we find all simple cycles in the diagram.
-We create a cycle verification set, $C$, and verify for each cycle that a single traversal is equal to the identity function by adding $(v \rightarrow v, 1)$ for each node v in the cycle to $C$.
+We create a cycle verification set, $\mathcal{C}$, and verify for each cycle that a single traversal is equal to the identity function by adding $(v \rightarrow v, 1)$ for each node v in the cycle to $\mathcal{C}$.
 Here, $v \rightarrow v$ is a simple cycle starting and ending at $v$, 1 is the identity function, and these must be verified to be equal to each other.
 
-We then create a set $P$ of all the paths in the diagram with no cycles, and filter the set $P \times P$, excluding pairs where the paths begin or end on different nodes, or are identical, to get the set of all cycle-free parallel pairs $Q$.
-After verifying $C$, it is sufficient to verify only $Q$ (as opposed to all pairs) because cycles must now be the identity, so for any pair in the set of all parallel paths, any instance of a cycle can be removed to obtain an equivalent pair with shorter, cycle-free paths.
+We then create a set $\mathcal{P}$ of all the paths in the diagram with no cycles, and filter the set $\mathcal{P} \times \mathcal{P}$, excluding pairs where the paths begin or end on different nodes, or are identical, to get the set of all cycle-free parallel pairs $\mathcal{Q}$.
+After verifying $\mathcal{C}$, it is sufficient to verify only $\mathcal{Q}$ (as opposed to all pairs) because cycles must now be the identity, so for any pair in the set of all parallel paths, any instance of a cycle can be removed to obtain an equivalent pair with shorter, cycle-free paths.
 
@@
 It is therefore safe to remove all pairs of paths with cycles, leaving only parallel pairs where neither path has a cycle.
-$P$ is finite, bounded by $2^{|V|}$, as a path without cycles is an ordering on nodes, each node occurring at most once.
-$|P \times P|$, and consequently, $|Q|$, are also finite, bounded by $2^{2|V|}$.
+$\mathcal{P}$ is finite, bounded by $2^{|V|}$, as a path without cycles is an ordering on nodes, each node occurring at most once.
+$|\mathcal{P} \times \mathcal{P}|$, and consequently, $|\mathcal{Q}|$, are also finite, bounded by $2^{2|V|}$.
 Thus the algorithm terminates and returns a finite (if large) set.
@@
 
-Like before, we start by creating a cycle verification set $C'$, but includes only the simple cycles that pass through the new edge.
-Then, instead of $Q$, the set of all non-cyclic parallel pairs, the algorithm obtains its subset $Q'$ consisting of all non-cyclic parallel pairs such that exactly one path in each pair passes through the new edge.
+Like before, we start by creating a cycle verification set $\mathcal{C}'$, but includes only the simple cycles that pass through the new edge.
+Then, instead of $\mathcal{Q}$, the set of all non-cyclic parallel pairs, the algorithm obtains its subset $\mathcal{Q}'$ consisting of all non-cyclic parallel pairs such that exactly one path in each pair passes through the new edge.
 To this end, the algorithm performs a \textit{two-flip tolerant path search} whose output is passed into a \textit{path extraction algorithm}; this search finds all parallel pairs for which only one path includes the new edge.
-The result of the path extraction algorithm to get the final output $Q' \cup C'$.
+The result of the path extraction algorithm to get the final output $\mathcal{Q}' \cup \mathcal{C}'$.
 
@@
 \paragraph{Path extraction algorithm}
-Next, the \textit{path extraction algorithm} then transforms the output of the two flip path search into the verification set, $Q' \cup C'$.
+Next, the \textit{path extraction algorithm} then transforms the output of the two flip path search into the verification set, $\mathcal{Q}' \cup \mathcal{C}'$.
 Given a set of two flip tolerant paths from the new edge source to sink, the algorithm outputs a set of pairs to verify.
 
-Let the new edge added to the diagram be $(S, T)$ and the input set of paths, $P$.
-The algorithm processes every two flip tolerant path $p$ in $P$ case-wise to obtain pairs to add to the output set.
+Let the new edge added to the diagram be $(S, T)$ and the input set of paths, $\mathcal{P}$.
+The algorithm processes every two flip tolerant path $p$ in $\mathcal{P}$ case-wise to obtain pairs to add to the output set.
 \begin{itemize}
@@
     Perform the two-flip tolerant path search from the source to sink node of the edge that is to be added followed, and on the output, apply the path extraction algorithm. 
-    The result is the set $O = Q'\cup C'$ of new parallel pairs with exactly one path passing through the new edge and neither paths containing any cycles, and the set of simple cycles passing through the new edge.
+    The result is the set $\mathcal{O} = \mathcal{Q}'\cup \mathcal{C}'$ of new parallel pairs with exactly one path passing through the new edge and neither paths containing any cycles, and the set of simple cycles passing through the new edge.
 \end{theorem}
@@
 \begin{proof}
-    Every element in the output of the path extraction algorithm was by construction an element of $O$.
-Every cycle in $C'$ can be expressed as $(S, T) ; p$, and corresponds to the input two flip tolerant path $p$.
+    Every element in the output of the path extraction algorithm was by construction an element of $\mathcal{O}$.
+Every cycle in $\mathcal{C}'$ can be expressed as $(S, T) ; p$, and corresponds to the input two flip tolerant path $p$.
 
-    It remains to show that every new parallel pair $p$ in $Q'$ corresponds to a two flip tolerant path.
+    It remains to show that every new parallel pair $p$ in $\mathcal{Q}'$ corresponds to a two flip tolerant path.
     Let $\partial(p)^+ = F_1$ and $\partial(p)^- = F_2$. 
@@
 
-A bilinking is defined to be a parallel pair that is disjoint but for their terminal nodes. The set of all bilinkings is $R_0$.
+A bilinking is defined to be a parallel pair that is disjoint but for their terminal nodes. The set of all bilinkings is $\mathcal{R}_0$.
 In an acyclic diagram, if all bilinkings are equal, all parallel pairs must also be equal since any given pair can be expressed as a composition of bilinkings.
 
-Define $r_1>r_2$ for bilinkings $r_1$ = $\{p_1,q_1\}$, $r_2$= $\{p_2, q_2)$ $\in$ $R_0$, if there exists a path $p$ such that $\partial(p) = \partial(r_1)$ and $p$ contains $p_2$.
+Define $r_1>r_2$ for bilinkings $r_1$ = $\{p_1,q_1\}$, $r_2$= $\{p_2, q_2)$ $\in$ $\mathcal{R}_0$, if there exists a path $p$ such that $\partial(p) = \partial(r_1)$ and $p$ contains $p_2$.
 Define $\langle\rangle$ as:
-$\langle r \rangle = \{ s\in R_0| r>s\}$.
+$\langle r \rangle = \{ s\in \mathcal{R}_0| r>s\}$.
 
 For bilinking $s$, let $F(s)$ be the vector in $\text{GF}(2)^{|E|}$ (where $\text{GF}(2)$ is the Galois field, that is, finite field of two elements) representing the edges present in s (the $n^{\text{th}}$ dimension of $F(s)$ is 1 if the corresponding edge is in $s$, and 0 otherwise).
-Let this function be extended to sets, so that for some set of bilinkings $S, F(S) = \{ F(s) | s\in S \}$. A notion of linear independence in this vector field exists.
+Let this function be extended to sets, so that for some set of bilinkings $\mathcal{S}$, $F(\mathcal{S}) = \{ F(s) | s\in \mathcal{S} \}$. A notion of linear independence in this vector field exists.
 
-For a set of bilinkings $R$, the closure function $cl$ is defined as:
-$cl(R) = \{ s\in R_0| s$ is linearly dependent on $F(R) \}$.
+For a set of bilinkings $\mathcal{R}$, the closure function $cl$ is defined as:
+$cl(\mathcal{R}) = \{ s\in \mathcal{R}_0| s$ is linearly dependent on $F(\mathcal{R}) \}$.
 %
-The closure function on $R$ captures all the pairs that can be made by made by composing or ``gluing together'' the bilinkings in $R$. 
+The closure function on $\mathcal{R}$ captures all the pairs that can be made by made by composing or ``gluing together'' the bilinkings in $\mathcal{R}$. 
 %
-Using these two functions, we define the function $\sigma$ on a set of bilinkings $R$ as
-$\sigma(R) = \{s \in R_0 | s\in cl(R\cap \langle s \rangle) \}$.
-This is the function used to capture all the pairs whose equivalence is implied by the equivalence of pairs in $R$.
+Using these two functions, we define the function $\sigma$ on a set of bilinkings $\mathcal{R}$ as
+$\sigma(\mathcal{R}) = \{s \in \mathcal{R}_0 | s\in cl(\mathcal{R}\cap \langle s \rangle) \}$.
+This is the function used to capture all the pairs whose equivalence is implied by the equivalence of pairs in $\mathcal{R}$.
 %
@@
 Given that a set of path pairs are equal, suppose we attempt to derive the proposition that a different parallel pair of paths is equal with a step-by-step application of inference rules.
- Under the assumption that edges are generic functions, and no other information is available, $F$ is a semi-group.
+ Under the assumption that edges are generic functions, and no other information is available, $\mathcal{F}$ is a semi-group.
  The only inference rules allowed are composition (given that $f_1=f_2$, it must be that $g ; f_1 = g ; f_2$) and replacement of one path by a different, equal path (given $f_1=f_2$ and $g ; f_1 = h ; f_1 $, it must be true that $g ; f_1 = h ; f_2$).

### 2020-10-05 acmPaper/paper.tex
Tweaking algorithm listings

@@
     T $\gets$ createMinimumSpanningTree(S)\;
-    excludedEdges = S.edges - T.edges\;
+    excludedEdges = edges in subgraph S - edges in T\;
     \ForEach{edge e $\in$ excludedEdges}{
         firstPath = T.findPath(source: e.source, sink: e.sink)\;
-        $R_s$.addElement(new Bilinking(firstPath, e)\;
+        $R_s$.addElement((firstPath, e))\;
     } 
@@
 
-\caption{Finding a spanning set.}\label{algo_spanning_set}
+\caption{Finding a spanning set of path pairs.}\label{algo_spanning_set}
 \end{algorithm}
@@
 \DontPrintSemicolon
-\KwResult{Find a minimal spanning set R.}
+\KwResult{Find a minimal spanning set of path pairs (bilinkings) R.}
 
@@
 \SetKwProg{Fn}{Function}{:}{}
-\Fn{\FSigma{inputSet S, codomain Rs}}{
+\Fn{\FSigma{input set S, spanning set Rs}}{
     output $\gets$ \{\}\;
     \For{bilinking $\in$ Rs}{
-        \tcc{Get fragments that could build up to the bilinking.}
-        smallerPairs $\gets$ getSmallerPairs(bilinking)\;
+        \tcc{Get fragments that could build up to the bilinking. Corresponds to applying <> function.}
+        smallerPairs $\gets$ allShorterPieces(bilinking)\;
         consideredPieces $\gets$ smallerPairs $\cap$ S \;
-        \tcc{See if bilinking can be built from these pieces.}
+        \tcc{Now see if bilinking can be built from these pieces.}
         \If{linearlyDependent(consideredPieces, bilinking)}{
@@
 parallelPairs $\gets$ \{\}\;
-\For{S in existingGraph.Nodes}{
-    \For{T in existingGraph}{
+\For{src in existingGraph.Nodes}{
+    \For{snk in existingGraph}{
         \try{}{
             Path pathWithNewEdge $\gets$ FindPath(
-                graph: existingGraph, 
-                sourceNode: S,
-                sinkNode: newEdge.Source()) +\;
+                % graph: existingGraph, 
+                sourceNode: src,
+                sinkNode: newEdge.Source +\;
             newEdge as Path +\;
             FindPath(
-                graph: existingGraph, 
-                sourceNode: newEdge.Sink(), 
-                sinkNode: T)\;
-            \If{S == T}{
-                pathInOldGraph $\gets$ identity(S)\;
+                % graph: exisingGraph, 
+                sourceNode: newEdge.Sink, 
+                sinkNode: snk)\;
+            \If{src == snk}{
+                pathInOldGraph $\gets$ identity(src)\;
             }
@@
                 Path pathInOldGraph $\gets$ FindPath(
-                    graph: existingGraph, 
-                    sourceNode: S, 
-                    sinkNode: T)\;
+                    % graph: existingGraph, 
+                    sourceNode: src, 
+                    sinkNode: snk)\;
             }
@@
         \catch{PathFindingFailedException}{
-           \tcc{No comparable pairs from node S to node T that need to be checked}
+           \tcc{No comparable pairs from node src to node snk that need to be checked}
            continue\;
@@
 
-predecessors $\gets$ S.predecessors(existingGraph)\;
-successors $\gets$ T.successors(existingGraph)\;
+predecessors $\gets$ predecessors of S in existingGraph\;
+successors $\gets$ successors of T in existingGraph\;
 
-Graph terminalPairGraph\;
+Graph terminalPairGraph $\gets$ empty \;
 \For{q $\in$ successors}{
     \For{p $\in$ predecessors}{
-        terminalPairGraph.addNode(q, p)\;
-        \For{qPred $\in$ q.predecessors(existingGraph)}{
-            \For{pSucc $\in$ p.successors(existingGraph)}{
-                terminalPairGraph.addEdge((qPred, pSucc))\;
+        terminalPairGraph.addNode((q, p))\;
+        \For{predecessor $\in$ predecessors of q in existingGraph}{
+            \For{successor $\in$ predecessors of p in existingGraph}{
+                terminalPairGraph.addEdge((predecessor, successor))\;
             }
@@
 verificationSet $\gets$ \{\}\;
-\While{len(terminalPairGraph.nodes) $>$ 0}{
-    currentNode $\gets$ terminalPairGraph.nodes[0]\;
-    visitedNodes $\gets$ \{\}\;
-    \While{len(currentNode.parents()) $>$ 0}{
-        visitedNodes.add(currentNode)\;
-        currentNode $\gets$ currentNode.parents()[0]\;
-        \If{currentNode $\in$ visitedNodes}{
-            edges $\gets$ getAllEdges(visitedNodes, terminalPairGraph)\;
-            terminalPairGraph.removeNodes(visitedNodes)\;
-            terminalPairGraph.addNode(currentNode, edges)\; 
-        }
+\While{terminalPairGraph.nodes not empty}{
+    currentNode $\gets$ terminalPairGraph.node\;
+    \tcp*{an arbitrarily chosen node of terminalPairGraph}
+    \While{currentNode has predecessors}{
+        currentNode $\gets$ predecessor\;
+        \tcp*{an arbitrarily chosen predecessor of current Node}
     }
     verificationSet.add(currentNode)\;
-    terminalPairGraph.removeNodes( currentNode.successors(terminalPairGraph))\;
+    terminalPairGraph.removeAllSuccessors(currentNode)\;
 }

### 2020-10-06 implementation/evaluator2.py
Messily create new evaluator copies for each class of graphs

@@ -15,15 +15,15 @@ import traceback
 from typing import List
 
 NUM_TRIES = 10
-DO_ACYCLIC = False
+DO_ACYCLIC = True
 OUTFILE = ''
 
 densityStep = 0.5
 sizeStep = 1
-maxSize = 250
+maxSize = 12
 
 # densities = [i*densityStep for i in range(1, int(1.0/densityStep))] 
-densities = [0.1, 0.5, 0.9]
+densities = [0.1, 0.25, 0.4]
 sizes = [i*sizeStep for i in range(1, int(maxSize/sizeStep))]
 # sizes = [1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12]
 
@@ -33,11 +33,11 @@ evaluationList = [(density, size) for size in sizes for density in densities]
 colors = ["black", "purple", "orange", "blue", "green", "red"]
 
 checkers : List[IPathChecker] = [
-    PolynomialPathChecker(),
-    OptimalSetPathChecker(),
+    # PolynomialPathChecker(),
+    # OptimalSetPathChecker(),
     # NaiveChecker(),
     # TwoFlipPathChecker(),
-    # BatchChecker()
+    BatchChecker()
 ]
 
 class attemptInfo:

### 2020-10-06 implementation/evaluator3.py

@@ -20,7 +20,7 @@ OUTFILE = ''
 
 densityStep = 0.5
 sizeStep = 1
-maxSize = 250
+maxSize = 10
 
 # densities = [i*densityStep for i in range(1, int(1.0/densityStep))] 
 densities = [0.1, 0.5, 0.9]
@@ -33,10 +33,10 @@ evaluationList = [(density, size) for size in sizes for density in densities]
 colors = ["black", "purple", "orange", "blue", "green", "red"]
 
 checkers : List[IPathChecker] = [
-    PolynomialPathChecker(),
-    OptimalSetPathChecker(),
-    # NaiveChecker(),
-    # TwoFlipPathChecker(),
+    # PolynomialPathChecker(),
+    # OptimalSetPathChecker(),
+    NaiveChecker(),
+    TwoFlipPathChecker()
     # BatchChecker()
 ]
 

### 2020-10-06 implementation/evaluator4.py

@@ -23,7 +23,7 @@ sizeStep = 1
 maxSize = 250
 
 # densities = [i*densityStep for i in range(1, int(1.0/densityStep))] 
-densities = [0.1, 0.5, 0.9]
+densities = [0.5]
 sizes = [i*sizeStep for i in range(1, int(maxSize/sizeStep))]
 # sizes = [1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12]
 
@@ -34,7 +34,7 @@ colors = ["black", "purple", "orange", "blue", "green", "red"]
 
 checkers : List[IPathChecker] = [
     PolynomialPathChecker(),
-    OptimalSetPathChecker(),
+    OptimalSetPathChecker()
     # NaiveChecker(),
     # TwoFlipPathChecker(),
     # BatchChecker()
@@ -169,8 +169,8 @@ errorBars=False, median=False):
                 for x in sizes:
                     result = results[(density, x)]
                     points = result.getTimes()
-                    ax1.plot([x for point in points], points, '.', color=colors[i]) 
-                    # ax1.plot([x for point in points], points, '.', color="grey") 
+                    # ax1.plot([x for point in points], points, '.', color=colors[i]) 
+                    ax1.plot([x for point in points], points, '.', color="grey") 
         except:
             print("Exception while plotting for density {}".format(density))
             print(traceback.print_stack())
@@ -218,8 +218,8 @@ errorBars=False):
 def plot(checkerName, results):
     # plot3d(checkerName, results)
     # Modify densities to plot only a subset of those computed.
-    # plotTimeVsSize(checkerName, results, densities, True, True)
-    plotTimeVsSize(checkerName, results, densities, False, False)
+    plotTimeVsSize(checkerName, results, densities, True, True)
+    # plotTimeVsSize(checkerName, results, densities, False, False)
 
 def dumpResult(results: dict, runDetails: runDetails):
     """Write results to file."""

### 2020-10-06 implementation/evaluator2.py
revert batch checker optimizations

@@ -20,12 +20,12 @@ OUTFILE = ''
 
 densityStep = 0.5
 sizeStep = 1
-maxSize = 12
+maxSize = 10
 
 # densities = [i*densityStep for i in range(1, int(1.0/densityStep))] 
 densities = [0.1, 0.25, 0.4]
-sizes = [i*sizeStep for i in range(1, int(maxSize/sizeStep))]
-# sizes = [1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12]
+# sizes = [i*sizeStep for i in range(1, int(maxSize/sizeStep))]
+sizes = [2, 3, 4, 5, 6, 7, 8, 9, 10]
 
 evaluationList = [(density, size) for size in sizes for density in densities]
 

### 2020-10-06 implementation/pathCheckers/batchChecker.py

@@ -28,9 +28,6 @@ class BatchChecker(NaiveChecker, BaseOnlineChecker):
     def __init__(self):
         super().__init__()
         BaseOnlineChecker.__init__(self)
-        self.vectors = dict()
-        self.matrices = dict()
-        self.gaussianReduceResults = dict()
 
     #### Utility functions ####
 
@@ -41,12 +38,9 @@ class BatchChecker(NaiveChecker, BaseOnlineChecker):
 
     def vectorize(self, path):
         """Get a vector corresponding to a path."""
-        if tuple(path) in self.vectors:
-            return self.vectors[path]
         vector = [0 for i in range(len(self.graph)**2)]
         for i in range(len(path)-1):
             vector[self.getIndex(path[i], path[i+1])] = (vector[self.getIndex(i, i+1)] + 1) % 2
-        self.vectors[tuple(path)] = vector
         return vector
 
     def vectorizePair(self, pair):
@@ -109,9 +103,6 @@ class BatchChecker(NaiveChecker, BaseOnlineChecker):
         column vector is linearly dependent on all others."""
         if len(matrix) == 0:
             return True
-        key = (tuple(row) for row in matrix)
-        if key in self.gaussianReduceResults:
-            return self.gaussianReduceResults[key]
         pivot = 0
         for column in range(len(matrix[0])-1):
             if matrix[pivot][column] == 0:
@@ -128,9 +119,7 @@ class BatchChecker(NaiveChecker, BaseOnlineChecker):
                 pivot += 1
         for row in range(pivot, len(matrix)):
             if matrix[row][-1] == 1:
-                self.gaussianReduceResults[key] = False
                 return False
-        self.gaussianReduceResults[key] = True
         return True
 
     def sigma(self, subst: List[Tuple[List, List]], allPairs: List[Tuple[List, List]]) -> List:
@@ -159,10 +148,6 @@ class BatchChecker(NaiveChecker, BaseOnlineChecker):
     def matrixify(self, pieces, lastPair):
         """Given a list of pairs and the last pair in the problem, create the
         corresponding matrix in order to Gaussian reduce."""
-        piecesTuple = (tuple(piece) for piece in pieces)
-        key = (piecesTuple, tuple(lastPair))
-        if key in self.matrices:
-            return self.matrices[key]
         matrix = []
         pieces = [self.vectorizePair(piece) for piece in pieces]
         pair = self.vectorizePair(lastPair)
[…]

### 2020-10-06 implementation/evaluator2_batch.py
new evalutor scripts for each graph

@@ -20,7 +20,7 @@ OUTFILE = ''
 
 densityStep = 0.5
 sizeStep = 1
-maxSize = 10
+maxSize = 13
 
 # densities = [i*densityStep for i in range(1, int(1.0/densityStep))] 
 densities = [0.1, 0.25, 0.4]

### 2020-10-06 implementation/evaluator3_twoFlip.py

@@ -1,0 +1,265 @@
+from datetime import datetime
+import cProfile
+import functools
+import json
+import matplotlib.pyplot as plt
+import numpy as np
+from pathCheckers.batchChecker import BatchChecker
+from pathCheckers.iPathChecker import IPathChecker
+from pathCheckers.optimalSetPathChecker import OptimalSetPathChecker
+from pathCheckers.polynomialPathChecker import PolynomialPathChecker
+from pathCheckers.naiveChecker import NaiveChecker
+from pathCheckers.twoFlipChecker import TwoFlipPathChecker
+from randomGraphGenerator import generateGraph, generateAcyclicGraph
+import traceback
+from typing import List
+
+NUM_TRIES = 10
+DO_ACYCLIC = False
+OUTFILE = ''
+
+densityStep = 0.5
+sizeStep = 1
+maxSize = 12 
+
+# densities = [i*densityStep for i in range(1, int(1.0/densityStep))] 
+densities = [0.1, 0.5, 0.9]
+sizes = [i*sizeStep for i in range(1, int(maxSize/sizeStep))]
+# sizes = [1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12]
+
+evaluationList = [(density, size) for size in sizes for density in densities]
+
+# First three colors look good in black and white.
+colors = ["black", "purple", "orange", "blue", "green", "red"]
+
+checkers : List[IPathChecker] = [
+    # PolynomialPathChecker(),
+    # OptimalSetPathChecker(),
+    # NaiveChecker(),
+    TwoFlipPathChecker()
+    # BatchChecker()
+]
+
+class attemptInfo:
+    """Stores information for a given density, size"""
+
+    def __init__(self):
+        self.times = []
+        self.sizes = []
+    
+    def addTime(self,time):
+        self.times += [time]
+    
+    def addSize(self, sze):
+        self.sizes += [sze]
+    
+    def avgList(self, lst):
+        self.avg = functools.reduce(lambda  acc, n: acc + n, lst, 0)/len(lst)
+        return self.avg
+    
+    def getAvgTime(self, median=False): 
+        if median:
+            return self.getMedianTime()
+        return self.avgList(self.times)
+
+    def getAvgSizes(self):
+        return self.avgList(self.sizes)
+    
+    def getTimes(self):
+        return self.times
+
+    def getMedianTime(self):
+        copy = self.times
+        copy.sort()
+        if len(copy)%2 == 1:
[…]

### 2020-10-06 implementation/evaluator5_naive.py

@@ -1,0 +1,265 @@
+from datetime import datetime
+import cProfile
+import functools
+import json
+import matplotlib.pyplot as plt
+import numpy as np
+from pathCheckers.batchChecker import BatchChecker
+from pathCheckers.iPathChecker import IPathChecker
+from pathCheckers.optimalSetPathChecker import OptimalSetPathChecker
+from pathCheckers.polynomialPathChecker import PolynomialPathChecker
+from pathCheckers.naiveChecker import NaiveChecker
+from pathCheckers.twoFlipChecker import TwoFlipPathChecker
+from randomGraphGenerator import generateGraph, generateAcyclicGraph
+import traceback
+from typing import List
+
+NUM_TRIES = 10
+DO_ACYCLIC = False
+OUTFILE = ''
+
+densityStep = 0.5
+sizeStep = 1
+maxSize = 10
+
+# densities = [i*densityStep for i in range(1, int(1.0/densityStep))] 
+densities = [0.1, 0.5, 0.9]
+sizes = [i*sizeStep for i in range(1, int(maxSize/sizeStep))]
+# sizes = [1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12]
+
+evaluationList = [(density, size) for size in sizes for density in densities]
+
+# First three colors look good in black and white.
+colors = ["black", "purple", "orange", "blue", "green", "red"]
+
+checkers : List[IPathChecker] = [
+    # PolynomialPathChecker(),
+    # OptimalSetPathChecker(),
+    NaiveChecker()
+    # TwoFlipPathChecker()
+    # BatchChecker()
+]
+
+class attemptInfo:
+    """Stores information for a given density, size"""
+
+    def __init__(self):
+        self.times = []
+        self.sizes = []
+    
+    def addTime(self,time):
+        self.times += [time]
+    
+    def addSize(self, sze):
+        self.sizes += [sze]
+    
+    def avgList(self, lst):
+        self.avg = functools.reduce(lambda  acc, n: acc + n, lst, 0)/len(lst)
+        return self.avg
+    
+    def getAvgTime(self, median=False): 
+        if median:
+            return self.getMedianTime()
+        return self.avgList(self.times)
+
+    def getAvgSizes(self):
+        return self.avgList(self.sizes)
+    
+    def getTimes(self):
+        return self.times
+
+    def getMedianTime(self):
+        copy = self.times
+        copy.sort()
+        if len(copy)%2 == 1:
[…]

### 2020-10-06 implementation/evaluator6_oneSize.py

@@ -20,12 +20,12 @@ OUTFILE = ''
 
 densityStep = 0.5
 sizeStep = 1
-maxSize = 10
+maxSize = 250
 
 # densities = [i*densityStep for i in range(1, int(1.0/densityStep))] 
 densities = [0.1, 0.5, 0.9]
-sizes = [i*sizeStep for i in range(1, int(maxSize/sizeStep))]
-# sizes = [1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12]
+# sizes = [i*sizeStep for i in range(1, int(maxSize/sizeStep))]
+sizes = [5000]
 
 evaluationList = [(density, size) for size in sizes for density in densities]
 
@@ -33,10 +33,10 @@ evaluationList = [(density, size) for size in sizes for density in densities]
 colors = ["black", "purple", "orange", "blue", "green", "red"]
 
 checkers : List[IPathChecker] = [
-    # PolynomialPathChecker(),
-    # OptimalSetPathChecker(),
-    NaiveChecker(),
-    TwoFlipPathChecker()
+    PolynomialPathChecker(),
+    OptimalSetPathChecker(),
+    # NaiveChecker(),
+    # TwoFlipPathChecker(),
     # BatchChecker()
 ]
 

### 2020-10-07 acmPaper/paper.tex
ACM stuff

@@
 % \documentclass[sigplan,review,anonymous]{acmart}
-\documentclass[sigplan,review,nonacm=true]{acmart}
+% \documentclass[sigplan,nonacm=true]{acmart}
+\documentclass[sigplan]{acmart}
 \setcopyright{none}
@@
 
+%%% The following is specific to TAPAS '20 and the paper
+%%% 'Online Verification of Commutativity'
+%%% by Aditi Kabra, Dietrich Geisler, and Adrian Sampson.
+%%%
+\setcopyright{acmlicensed}
+\acmPrice{15.00}
+\acmDOI{10.1145/3427764.3428322}
+\acmYear{2020}
+\copyrightyear{2020}
+\acmSubmissionID{splashws20tapasmain-id5-p}
+\acmISBN{978-1-4503-8189-5/20/11}
+\acmConference[TAPAS '20]{Proceedings of the 11th ACM SIGPLAN International Workshop on Tools for Automatic Program Analysis}{November 17, 2020}{Virtual, USA}
+\acmBooktitle{Proceedings of the 11th ACM SIGPLAN International Workshop on Tools for Automatic Program Analysis (TAPAS '20), November 17, 2020, Virtual, USA}
+
+\begin{CCSXML}
+    <ccs2012>
+       <concept>
+           <concept_id>10003752.10003766.10003776</concept_id>
+           <concept_desc>Theory of computation~Regular languages</concept_desc>
+           <concept_significance>500</concept_significance>
+           </concept>
+       <concept>
+           <concept_id>10003752.10003809.10003635.10010038</concept_id>
+           <concept_desc>Theory of computation~Dynamic graph algorithms</concept_desc>
+           <concept_significance>500</concept_significance>
+           </concept>
+     </ccs2012>
+\end{CCSXML}
+
+\ccsdesc[500]{Theory of computation~Regular languages}
+\ccsdesc[500]{Theory of computation~Dynamic graph algorithms}
+
 % Shorthand for inline code
@@
 }
-% ACM publication cruft.
-\acmJournal{PACMPL}
-\acmVolume{1}
-\acmNumber{TAPAS}
-\acmArticle{1}
-\acmYear{2020}
-\acmMonth{1}
-\acmDOI{} % \acmDOI{10.1145/nnnnnnn.nnnnnnn}
+
 \startPage{1}
@@
 \end{abstract}
+
+\keywords{commuting diagrams, automatic type conversion, online verification}
 

### 2020-10-07 acmPaper/paper.tex
algorithm captions

@@
 \usepackage{fncylab}
+\usepackage{caption}
 \usepackage{subcaption}
 \usepackage{listings}
+\usepackage{microtype}
 %
@@
 \ForEach{node v in V}{
-    S $\gets$ existingGraph.extractReachableSection(v)\;
-    T $\gets$ createMinimumSpanningTree(S)\;
-    excludedEdges = edges in subgraph S - edges in T\;
+    Subgraph $\gets$ existingGraph.extractReachableSection(v)\;
+    Tree $\gets$ createMinimumSpanningTree(Subgraph)\;
+    excludedEdges = edges in Subgraph - edges in Tree\;
     \ForEach{edge e $\in$ excludedEdges}{
-        firstPath = T.findPath(source: e.source, sink: e.sink)\;
+        firstPath = Tree.findPath(source: e.source, sink: e.sink)\;
         $R_s$.addElement((firstPath, e))\;
@@
 
-\caption{Finding a spanning set of path pairs.}\label{algo_spanning_set}
+\caption{Finding a spanning set of path pairs, as in section ~\ref{sec:batchBaseline}}\label{algo_spanning_set}
 \end{algorithm}
@@
 
-\caption{Finding a minimal spanning set.}\label{algo_minimal_spanning_set}
+\caption{Finding a minimal spanning set, as described in section ~\ref{sec:batchBaseline}}\label{algo_minimal_spanning_set}
 \end{algorithm}
@@
 \begin{table}
+    \caption{Computation time for 9-node graph of density 0.4, averaged over ten runs.}
 \begin{tabular}{|c|c|}
@@
 \end{tabular}
-\caption{Computation time for 9-node graph of density 0.4, averaged over ten runs.}
 \label{tab:times} %TODO debug and fix label not working.
@@
 \begin{table}
+    \caption{Output size for 9 node graph of density 0.4, averaged over ten runs}
 \begin{tabular}{|c|c|}
@@
 \end{tabular}
-\caption{Output size for 9 node graph of density 0.4, averaged over ten runs}
 \label{tab:sizes}

### 2020-10-08 acmPaper/paper.tex
author info
replaces all but one of the pdfs, hacks to acm class
final pdf image
add pdf's
fixes

@@
 \documentclass[sigplan]{acmart}
-\setcopyright{none}
 %
+
 \usepackage{tikz}
@@
        <concept>
-           <concept_id>10003752.10003766.10003776</concept_id>
-           <concept_desc>Theory of computation~Regular languages</concept_desc>
-           <concept_significance>500</concept_significance>
-           </concept>
-       <concept>
            <concept_id>10003752.10003809.10003635.10010038</concept_id>
@@
 
-\ccsdesc[500]{Theory of computation~Regular languages}
 \ccsdesc[500]{Theory of computation~Dynamic graph algorithms}
@@
 \begin{document}
-\settopmatter{printacmref=false}
+\settopmatter{printacmref=true}
 \title{Online Verification of Commutativity}
@@
 %
-\author{Aditi Kabra \quad Dietrich Geisler \quad Adrian Sampson}
+\author{Aditi Kabra}
+\affiliation{%
+  \institution{Carnegie Mellon University}
+  \city{Pittsburgh}
+  \state{PA}
+  \country{USA}
+}
+\email{akabra@cs.cmu.edu}
+
+\author{Dietrich Geisler}
 \affiliation{%
   \institution{Cornell University}
+  \city{Ithaca}
+  \state{NY}
+  \country{USA}
 }
+\email{dag368@cornell.edu}
+
+\author{Adrian Sampson}
+\affiliation{%
+  \institution{Cornell University}
+  \city{Ithaca}
+  \state{NY}
+  \country{USA}
+}
+\email{asampson@cs.cornell.edu}
 
@@
     \begin{subfigure}{\linewidth}
-      \includegraphics[width=\linewidth]{timeVsSize_10_OptimalSet.png}
+      \includegraphics[width=\linewidth]{timeVsSize_10_OptimalSet.pdf}
       \caption{~\ref{algo_online_minimal}}
@@
     \begin{subfigure}{\linewidth}
-      \includegraphics[width=\linewidth]{timeVsSize_10_Polynomial.png}
+      \includegraphics[width=\linewidth]{timeVsSize_10_Polynomial.pdf}
       \caption{~\ref{algo_online_polynomial}}
@@
     \begin{subfigure}{\linewidth}
-      \includegraphics[width=\linewidth]{timeVsSize_10_NaiveChecker.png}
+      \includegraphics[width=\linewidth]{timeVsSize_10_NaiveChecker.pdf}
       \caption{Naive baseline}
@@
     \begin{subfigure}{\linewidth}
-      \includegraphics[width=\linewidth]{timeVsSize_10_TwoFlipPathChecker.png}
+      \includegraphics[width=\linewidth]{timeVsSize_10_TwoFlipPathChecker.pdf}
       \caption{Two flip tolerant baseline}
@@
     \begin{subfigure}{\linewidth}
-      \includegraphics[width=\linewidth]{timeVsSize_10_BatchChecker.png}
+      \includegraphics[width=\linewidth]{timeVsSize_10_BatchChecker.pdf}
       \caption{Batch algorithm baseline}
@@
     \begin{subfigure}{\linewidth}
-      \includegraphics[width=\linewidth]{variance_10_OptimalSet.png}
+      \includegraphics[width=\linewidth]{variance_10_OptimalSet.pdf}
       \caption{~\ref{algo_online_minimal}}
@@
     \begin{subfigure}{\linewidth}
-      \includegraphics[width=\linewidth]{variance_10_Polynomial.png}
+      \includegraphics[width=\linewidth]{variance_10_Polynomial.pdf}
       \caption{~\ref{algo_online_polynomial}}
@@
     % \begin{subfigure}{\linewidth}
-    %   \includegraphics[width=\linewidth]{variance_10_TwoFlip.png}
+    %   \includegraphics[width=\linewidth]{variance_10_TwoFlip.pdf}
     %   \caption{Two flip tolerant baseline}

### 2020-10-09 acmPaper/paper.tex
Abstract wording tweaks

@@
 \begin{abstract}
-Systems of transformations arise in many programming systems, such as in type graphs of implicit type conversion functions.
-It is important to ensure that these diagrams commute: that any composing path of transformations from the same source to the same destination yields the same result.
+Systems of transformations arise in many programming systems, such as in graphs of implicit type conversion functions.
+It is important to ensure that these diagrams commute: that composing any path of transformations from the same source to the same destination yields the same result.
 However, a straightforward approach to verifying commutativity must contend with cycles, and even so it runs in exponential time.

### 2020-10-13 acmPaper/paper.tex
period after captions, added verification as relevent topic

--- 2020-10-08 acmPaper/paper.tex
@@
            </concept>
+           <concept>
+           <concept_id>10003752.10010124.10010138.10010142</concept_id>
+           <concept_desc>Theory of computation~Program verification</concept_desc>
+           <concept_significance>500</concept_significance>
+           </concept>
      </ccs2012>
@@
 \ccsdesc[500]{Theory of computation~Dynamic graph algorithms}
+\ccsdesc[500]{Theory of computation~Program verification}
 
@@
             \end{tikzpicture}
-        \caption{Diagram for the type conversions in the program}
+        \caption{Diagram for the type conversions in the program.}
     \end{subfigure}
@@
 
-\caption{Finding a spanning set of path pairs, as in section ~\ref{sec:batchBaseline}}\label{algo_spanning_set}
+\caption{Finding a spanning set of path pairs, as in section ~\ref{sec:batchBaseline}.}\label{algo_spanning_set}
 \end{algorithm}
@@
 
-\caption{Finding a minimal spanning set, as described in section ~\ref{sec:batchBaseline}}\label{algo_minimal_spanning_set}
+\caption{Finding a minimal spanning set, as described in section ~\ref{sec:batchBaseline}.}\label{algo_minimal_spanning_set}
 \end{algorithm}
@@
 \KwRet{parallelPairs}\;
-\caption{Online polynomial time algorithm to find parallel pair set}
+\caption{Online polynomial time algorithm to find parallel pair set.}
 \label{algo_online_polynomial}
@@
 \KwRet{verificationSet}\;
-\caption{Minimal set finding algorithm}
+\caption{Minimal set finding algorithm.}
 \label{algo_online_minimal}
@@
       \includegraphics[width=\linewidth]{timeVsSize_10_OptimalSet.pdf}
-      \caption{~\ref{algo_online_minimal}}
+      \caption{~\ref{algo_online_minimal}.}
       \label{fig:sfigOptimalTvsS}
@@
       \includegraphics[width=\linewidth]{timeVsSize_10_Polynomial.pdf}
-      \caption{~\ref{algo_online_polynomial}}
+      \caption{~\ref{algo_online_polynomial}.}
       \label{fig:sfigPolynomialTvsS}
@@
       \includegraphics[width=\linewidth]{timeVsSize_10_NaiveChecker.pdf}
-      \caption{Naive baseline}
+      \caption{Naive baseline.}
       \label{fig:sfigNa\"{i}veTvsS}
@@
       \includegraphics[width=\linewidth]{timeVsSize_10_TwoFlipPathChecker.pdf}
-      \caption{Two flip tolerant baseline}
+      \caption{Two flip tolerant baseline.}
       \label{fig:sfigTwoFlipTvsS}
@@
       \includegraphics[width=\linewidth]{timeVsSize_10_BatchChecker.pdf}
-      \caption{Batch algorithm baseline}
+      \caption{Batch algorithm baseline.}
       \label{fig:sfigBatchTvsS}
     \end{subfigure}
-    \caption{Plots of scaling of time with input size and density}
+    \caption{Plots of scaling of time with input size and density.}
     \label{fig:timeVsSize}
@@
       \includegraphics[width=\linewidth]{variance_10_OptimalSet.pdf}
-      \caption{~\ref{algo_online_minimal}}
+      \caption{~\ref{algo_online_minimal}.}
       \label{fig:sfigOptimalSpread}
@@
       \includegraphics[width=\linewidth]{variance_10_Polynomial.pdf}
-      \caption{~\ref{algo_online_polynomial}}
+      \caption{~\ref{algo_online_polynomial}.}
       \label{fig:sfigPolynomialSpread}
@@
 
-    \caption{Plots of spread of results}
+    \caption{Plots of spread of results.}
     \label{fig:variance}
@@
 \begin{table}
-    \caption{Output size for 9 node graph of density 0.4, averaged over ten runs}
+    \caption{Output size for 9 node graph of density 0.4, averaged over ten runs.}
 \begin{tabular}{|c|c|}

### 2020-10-14 acmPaper/paper.tex
Addressed algorithms stuff

@@
 \usepackage{microtype}
+\usepackage{float}
+% \usepackage{endfloat}
+% \usepackage{afterpage}
 %
@@
 
-\begin{algorithm}[t]
+\begin{algorithm}[tb]
 \label{alg_spanning_set}
@@
 \ForEach{node v in V}{
-    Subgraph $\gets$ existingGraph.extractReachableSection(v)\;
-    Tree $\gets$ createMinimumSpanningTree(Subgraph)\;
-    excludedEdges = edges in Subgraph - edges in Tree\;
+    subgraph $\gets$ existingGraph.extractReachableSection(v)\;
+    \tcc{Get the portion of the graph that can be reached starting from v.}
+    tree $\gets$ createMinimumSpanningTree(subgraph)\;
+    excludedEdges = edges in subgraph - edges in tree\;
     \ForEach{edge e $\in$ excludedEdges}{
-        firstPath = Tree.findPath(source: e.source, sink: e.sink)\;
+        firstPath = tree.findPath(source: e.source, sink: e.sink)\;
         $R_s$.addElement((firstPath, e))\;
@@
 
-\begin{algorithm}[t]
+\begin{algorithm}[tb]
 \DontPrintSemicolon
@@
     \For{bilinking $\in$ Rs}{
+        smallerPairs $\gets$ allShorterPieces(bilinking)\;
         \tcc{Get fragments that could build up to the bilinking. Corresponds to applying <> function.}
-        smallerPairs $\gets$ allShorterPieces(bilinking)\;
         consideredPieces $\gets$ smallerPairs $\cap$ S \;
         \tcc{Now see if bilinking can be built from these pieces.}
+        \tcc{Linear independence is in GF(2) as in algorithm description.}
         \If{linearlyDependent(consideredPieces, bilinking)}{
@@
         \try{}{
+            \tcc{Use any standard path finding algorithm such as BFS to find a path in the existing graph from the specified source to sink.}
             Path pathWithNewEdge $\gets$ FindPath(
@@
                 sourceNode: src,
-                sinkNode: newEdge.Source +\;
-            newEdge as Path +\;
+                sinkNode: newEdge.Source) +\;
+            newEdge +\;
             FindPath(
@@
             \If{src == snk}{
-                pathInOldGraph $\gets$ identity(src)\;
+                \tcc{Assign the nullary path from src to snk.}
+                pathInOldGraph $\gets$ src\;
             }
@@
 
-\begin{algorithm}
+\begin{algorithm}[tb]
 \DontPrintSemicolon
@@
     \While{currentNode has predecessors}{
-        currentNode $\gets$ predecessor\;
+        currentNode $\gets$ predecessorOfCurrentNode\;
         \tcp*{an arbitrarily chosen predecessor of current Node}
@@
 
-\begin{figure}
+\begin{figure*}
     % TODO format nicely.
-    \begin{subfigure}{\linewidth}
+    \begin{subfigure}{0.3\linewidth}
       \includegraphics[width=\linewidth]{timeVsSize_10_OptimalSet.pdf}
@@
     \end{subfigure}
-
-    \begin{subfigure}{\linewidth}
+    ~
+    \begin{subfigure}{0.3\linewidth}
       \includegraphics[width=\linewidth]{timeVsSize_10_Polynomial.pdf}
@@
     \end{subfigure}
-
-    \begin{subfigure}{\linewidth}
+    ~
+    \begin{subfigure}{0.3\linewidth}
       \includegraphics[width=\linewidth]{timeVsSize_10_NaiveChecker.pdf}
@@
 
-\end{figure}
-\begin{figure}\ContinuedFloat
+    ~
+% \end{figure}
+% \begin{figure}\ContinuedFloat
 
-    \begin{subfigure}{\linewidth}
+    \begin{subfigure}{0.3\linewidth}
       \includegraphics[width=\linewidth]{timeVsSize_10_TwoFlipPathChecker.pdf}
@@
     \end{subfigure}
-
-    \begin{subfigure}{\linewidth}
+    ~
+    \begin{subfigure}{0.3\linewidth}
       \includegraphics[width=\linewidth]{timeVsSize_10_BatchChecker.pdf}
@@
     \label{fig:timeVsSize}
-\end{figure}
+\end{figure*}
 

### 2020-10-15 acmPaper/paper.tex
Caption tweaks
Figure positioning
Try to center figure
Try to make figure fit in column
Switch tables to booktabs

@@
     \begin{subfigure}{0.4\textwidth}
+        \hspace{-20pt}
         \begin{tikzpicture}[scale=0.2]
@@
 \begin{center}
-\begin{tikzpicture}[scale=0.2]
+\begin{tikzpicture}[scale=0.18]
 \tikzstyle{every node}+=[inner sep=0pt]
@@
     \caption{Computation time for 9-node graph of density 0.4, averaged over ten runs.}
-\begin{tabular}{|c|c|}
-    \hline
-    Algorithm & average seconds of computation \\
-    \hline
+\begin{tabular}{cc}
+    \toprule
+    Algorithm & Average seconds of computation \\
+    \midrule
     Na\"{i}ve baseline & 0.77 \\
@@
     ~\ref{algo_online_minimal} & 0.00086 \\
-    \hline
+    \bottomrule
 \end{tabular}
@@
 \end{table}
-
-The na\"{i}ve baseline performs poorly, taking well over a thousand seconds for even small graphs of 10 nodes.
-While the batch algorithm improves on this, it still does not scale very well, with computation for a graph with 14 nodes and 0.4 density taking hours.
-Our implementation does not memoize the construction of the vector and matrix representation of paths in GF(2); profiling indicates that this construction is a major factor in the high time cost for this algorithm.
-\ref{algo_online_polynomial} performs only slightly better than the batch algorithm.
-Surprisingly, the optimal set algorithm cuts time cost by several orders of magnitude, and runs in milliseconds for small graphs.
-All implementations are sensitive to density, performing better when density is low.
-
-\subsection{Scaling of Time with Input Size}
-
-Figure~\ref{fig:timeVsSize} shows that the algorithms' time scales with size, as expected.
-\xxx[as]{The plots are weirdly tiny on my machine... any chance your plotting tool can output PDF images?}
-Both \ref{algo_online_minimal} and \ref{algo_online_polynomial} exhibit graphs that are polynomial in appearance.
-The na\"{i}ve baseline as well as the two flip tolerant baseline display quick growth.
-The batch algorithm also grows fast, though not as much as the online checking baselines.
-
-We define density to be the ratio of the number of edges in the graph to the total possible number of edges (which is $|V|^2$, where $|V|$ is the number of nodes).
-Run time relates to the density of edges in the input graph.
-The degree of the effect differs with the algorithms, as Figure~\ref{fig:timeVsSize} shows.
-Generally, denser graphs entail longer computation time.
-For the batch algorithm we use lower densities since the input graph must be acyclic. This puts an upper bound on density that approaches 0.5 in large graphs.
-% Our proposed algorithms, ~\ref{algo_minimal_spanning_set} and ~\ref{algo_online_polynomial} are less affected because they aggressively remove redundancies.
 
@@
     \end{subfigure}
-    \caption{Plots of scaling of time with input size and density.}
+    \caption{Running time for various algorithms as input graph size and density scale.}
     \label{fig:timeVsSize}
 \end{figure*}
+
+The na\"{i}ve baseline performs poorly, taking well over a thousand seconds for even small graphs of 10 nodes.
+While the batch algorithm improves on this, it still does not scale very well, with computation for a graph with 14 nodes and 0.4 density taking hours.
+Our implementation does not memoize the construction of the vector and matrix representation of paths in GF(2); profiling indicates that this construction is a major factor in the high time cost for this algorithm.
+\ref{algo_online_polynomial} performs only slightly better than the batch algorithm.
+Surprisingly, the optimal set algorithm cuts time cost by several orders of magnitude, and runs in milliseconds for small graphs.
+All implementations are sensitive to density, performing better when density is low.
+
+\subsection{Scaling of Time with Input Size}
+
+Figure~\ref{fig:timeVsSize} shows that the algorithms' time scales with size, as expected.
+\xxx[as]{The plots are weirdly tiny on my machine... any chance your plotting tool can output PDF images?}
+Both \ref{algo_online_minimal} and \ref{algo_online_polynomial} exhibit graphs that are polynomial in appearance.
+The na\"{i}ve baseline as well as the two flip tolerant baseline display quick growth.
+The batch algorithm also grows fast, though not as much as the online checking baselines.
+
+We define density to be the ratio of the number of edges in the graph to the total possible number of edges (which is $|V|^2$, where $|V|$ is the number of nodes).
+Run time relates to the density of edges in the input graph.
+The degree of the effect differs with the algorithms, as Figure~\ref{fig:timeVsSize} shows.
+Generally, denser graphs entail longer computation time.
+For the batch algorithm we use lower densities since the input graph must be acyclic. This puts an upper bound on density that approaches 0.5 in large graphs.
+% Our proposed algorithms, ~\ref{algo_minimal_spanning_set} and ~\ref{algo_online_polynomial} are less affected because they aggressively remove redundancies.
 
@@
 
-    \caption{Plots of spread of results.}
+    \caption{Spreads of algorithm running times.}
     \label{fig:variance}
@@
     \caption{Output size for 9 node graph of density 0.4, averaged over ten runs.}
-\begin{tabular}{|c|c|}
-    \hline
-    Algorithm & average number of output pairs \\
-    \hline
+\begin{tabular}{cc}
+    \toprule
+    Algorithm & Average number of output pairs \\
+    \midrule
     Na\"{i}ve baseline & 39754.9 \\
@@
     ~\ref{algo_online_minimal} & 1 \\
-    \hline
+    \bottomrule
 \end{tabular}

### 2020-10-15 acmPaper/paper.tex
Slightly larger Figure 4
Trim Figure 4 for better wrap

@@
     % TODO format nicely.
-    \begin{subfigure}{0.3\linewidth}
-      \includegraphics[width=\linewidth]{timeVsSize_10_OptimalSet.pdf}
+    \begin{subfigure}{0.35\linewidth}
+      \includegraphics[width=\linewidth,trim=0 0 5 30,clip]{timeVsSize_10_OptimalSet.pdf}
       \caption{~\ref{algo_online_minimal}.}
@@
     \end{subfigure}
-    ~
-    \begin{subfigure}{0.3\linewidth}
-      \includegraphics[width=\linewidth]{timeVsSize_10_Polynomial.pdf}
+    %
+    \hspace{-20pt}
+    %
+    \begin{subfigure}{0.35\linewidth}
+      \includegraphics[width=\linewidth,trim=0 0 5 30,clip]{timeVsSize_10_Polynomial.pdf}
       \caption{~\ref{algo_online_polynomial}.}
@@
     \end{subfigure}
-    ~
-    \begin{subfigure}{0.3\linewidth}
-      \includegraphics[width=\linewidth]{timeVsSize_10_NaiveChecker.pdf}
+    %
+    \hspace{-20pt}
+    %
+    \begin{subfigure}{0.35\linewidth}
+      \includegraphics[width=\linewidth,trim=0 0 5 30,clip]{timeVsSize_10_NaiveChecker.pdf}
       \caption{Naive baseline.}
-      \label{fig:sfigNa\"{i}veTvsS}
+      \label{fig:sfigNaiveTvsS}
     \end{subfigure}
 
-    ~
 % \end{figure}
@@
 
-    \begin{subfigure}{0.3\linewidth}
-      \includegraphics[width=\linewidth]{timeVsSize_10_TwoFlipPathChecker.pdf}
+    \begin{subfigure}{0.35\linewidth}
+      \includegraphics[width=\linewidth,trim=0 0 5 30,clip]{timeVsSize_10_TwoFlipPathChecker.pdf}
       \caption{Two flip tolerant baseline.}
@@
     \end{subfigure}
-    ~
-    \begin{subfigure}{0.3\linewidth}
-      \includegraphics[width=\linewidth]{timeVsSize_10_BatchChecker.pdf}
+    %
+    \hspace{-20pt}
+    %
+    \begin{subfigure}{0.35\linewidth}
+      \includegraphics[width=\linewidth,trim=0 0 5 30,clip]{timeVsSize_10_BatchChecker.pdf}
       \caption{Batch algorithm baseline.}

### 2020-10-16 acmPaper/paper.tex
tex cleanup; everythong passing except font type 3

@@
 \newcount\draft\draft=0 % set to 0 for submission or publication
-% \documentclass[sigplan,review,anonymous]{acmart}
-% \documentclass[sigplan,nonacm=true]{acmart}
-\documentclass[sigplan]{acmart}
+\documentclass[sigplan, scree, balance]{acmart}
 %
@@
 \usepackage{float}
-% \usepackage{endfloat}
-% \usepackage{afterpage}
+\usepackage{balance}
 %
 \usepackage{graphicx}
-% Used for displaying a sample figure. If possible, figure files should
-% be included in EPS format.
-%
-% If you use the hyperref package, please uncomment the following line
-% to display URLs in blue roman font according to Springer's eBook style:
-% \renewcommand\UrlFont{\color{blue}\rmfamily}
+
 \AtBeginDocument{%
@@
 \title{Online Verification of Commutativity}
-%
-%\titlerunning{Abbreviated paper title}
-% If the paper title is too long for the running head, you can set
-% an abbreviated paper title here
-%
+
 \author{Aditi Kabra}
@@
 Previous work has shown how to verify commutativity in the special case of acyclic diagrams in $O(|V|^4|E|^2)$ time,
-% I think this is clear enough (at least for the abstract). --AS
-% where V is the set of vertices in the diagram, and E, the set of edges.
 but this is a \emph{batch} algorithm: the entire diagram must be known ahead of time.
@@
 
-A path is a sequence of edges. The edge-to-function mapping $f$ can be naturally extended to paths: if path $p=e_1 ; e_2 ; \cdots ; e_n$ then $f(p)=f(e_1) ; f(e_2) ; \cdots ; f(e_n)$.
+A path is a sequence of edges. The edge-to-function mapping $f$ can be naturally extended to paths: if path $p$=$e_1 ; \cdots ; e_n$ then $f(p)=f(e_1) ; f(e_2) ; \cdots ; f(e_n)$.
 %
@@
             Path pathWithNewEdge $\gets$ FindPath(
-                % graph: existingGraph, 
                 sourceNode: src,
@@
             FindPath(
-                % graph: exisingGraph, 
                 sourceNode: newEdge.Sink, 
@@
                 Path pathInOldGraph $\gets$ FindPath(
-                    % graph: existingGraph, 
                     sourceNode: src, 
@@
 The rule effectively places an ordering on the informativeness of path pairs based on their terminal nodes.
-% If it is possible to draw a path from node $P_1$ to $P_2$ that passes through $Q_1$ and $Q_2$ then verifying a parallel pair from $Q_1$ to $Q_2$ implies that parallel pairs from $P_1$ to $P_2$ commute as well.
 
@@
 \end{tabular}
-\label{tab:times} %TODO debug and fix label not working.
+\label{tab:times}
+\end{table}
+
+\begin{table}
+    \caption{Output size for 9 node graph of density 0.4, averaged over ten runs.}
+\begin{tabular}{cc}
+    \toprule
+    Algorithm & Average number of output pairs \\
+    \midrule
+    Na\"{i}ve baseline & 39754.9 \\
+    Two Flip tolerant & 748.9 \\
+    Batch algorithm & 23 \\
+    ~\ref{algo_online_polynomial} & 78.3 \\
+    ~\ref{algo_online_minimal} & 1 \\
+    \bottomrule
+\end{tabular}
+\label{tab:sizes}
 \end{table}
@@
 \begin{figure*}
-    % TODO format nicely.
     \begin{subfigure}{0.35\linewidth}
@@
 
-% \end{figure}
-% \begin{figure}\ContinuedFloat
-
     \begin{subfigure}{0.35\linewidth}
@@
 \end{figure*}
+
+\begin{figure}
+    \begin{subfigure}{\linewidth}
+      \includegraphics[width=\linewidth]{variance_10_OptimalSet.pdf}
+      \caption{~\ref{algo_online_minimal}.}
+      \label{fig:sfigOptimalSpread}
+    \end{subfigure}
+
+    \begin{subfigure}{\linewidth}
+      \includegraphics[width=\linewidth]{variance_10_Polynomial.pdf}
+      \caption{~\ref{algo_online_polynomial}.}
+      \label{fig:sfigPolynomialSpread}
+    \end{subfigure}
+
+    \caption{Spreads of algorithm running times.}
+    \label{fig:variance}
+\end{figure}
 
@@
 For the batch algorithm we use lower densities since the input graph must be acyclic. This puts an upper bound on density that approaches 0.5 in large graphs.
-% Our proposed algorithms, ~\ref{algo_minimal_spanning_set} and ~\ref{algo_online_polynomial} are less affected because they aggressively remove redundancies.
 
@@
 
-\begin{figure}
-    % TODO format nicely.
-    \begin{subfigure}{\linewidth}
-      \includegraphics[width=\linewidth]{variance_10_OptimalSet.pdf}
-      \caption{~\ref{algo_online_minimal}.}
-      \label{fig:sfigOptimalSpread}
-    \end{subfigure}
-
-    \begin{subfigure}{\linewidth}
-      \includegraphics[width=\linewidth]{variance_10_Polynomial.pdf}
-      \caption{~\ref{algo_online_polynomial}.}
-      \label{fig:sfigPolynomialSpread}
-    \end{subfigure}
-
-    % \begin{subfigure}{\linewidth}
-    %   \includegraphics[width=\linewidth]{variance_10_TwoFlip.pdf}
-    %   \caption{Two flip tolerant baseline}
-    %   \label{fig:sfigTwoFlipSpread}
-    % \end{subfigure}
-
-    \caption{Spreads of algorithm running times.}
-    \label{fig:variance}
-\end{figure}
-
 \subsection{Size of Output}
@@
 
-\begin{table}
-    \caption{Output size for 9 node graph of density 0.4, averaged over ten runs.}
-\begin{tabular}{cc}
-    \toprule
-    Algorithm & Average number of output pairs \\
-    \midrule
-    Na\"{i}ve baseline & 39754.9 \\
-    Two Flip tolerant & 748.9 \\
-    Batch algorithm & 23 \\
-    ~\ref{algo_online_polynomial} & 78.3 \\
-    ~\ref{algo_online_minimal} & 1 \\
-    \bottomrule
-\end{tabular}
-\label{tab:sizes}
-\end{table}
-
 \section{Related Work}
@@
 We did not find any other work that solves the question of verifying that diagrams commute.
-
+\balance
 However, the question of commuting does come up in programming languages with implicit type conversion.
@@
 
-% Units of measure work resource: https://gmpreussner.com/research/dimensional-analysis-in-programming-languages
 \section{Conclusion}
